### 2018-12-22 tim-lucas-wsc.tex
Working doc.

\documentclass[11pt]{article}
\fontfamily{times}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{geometry}
\usepackage{subcaption}

\geometry{verbose,tmargin=30mm,bmargin=25mm,lmargin=25mm,rmargin=25mm}
\input{figures.tex}
\pagestyle{empty}
\begin{document}
\templatefigures{}

\small{

\begin{center}
%The title should be centred and in bold letters. It should be informative but not too long (preferably no more than two lines).
\textbf{Model ensembles with different response variables for base and meta models: malaria disaggregation regression combining prevalence and incidence data}
\end{center}

% Model ensembles with different response variables for base and meta models:

\begin{center}
{Tim C. D. Lucas*}\\
{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - timcdlucas@gmail.com}\\ 
\vspace{0.2cm}

{Anita Nandi}\\
{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - anita.nandi@bdi.ox.ac.uk }\\

\vspace{0.1cm}

{Michele Nguyen}\\
{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - michele.nguyen@bdi.ox.ac.uk }\\

\vspace{0.1cm}

{Susan Rumisha}\\
{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - susan.rumisha@bdi.ox.ac.uk }\\

\vspace{0.1cm}

{Ewan Cameron}\\
{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - dr.ewan.cameron@gmail.com }\\

\vspace{0.1cm}
{Pete Gething}\\
{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - peter.gething@bdi.ox.ac.uk }\\

\vspace{0.1cm}
{Daniel J. Weiss}\\
{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - daniel.weiss@bdi.ox.ac.uk }\\

\end{center}

\begin{center}
{\bf Abstract}
\end{center}

\setlength{\parindent}{0pt}

This is where the abstract is placed. It should include a statement about the problem being addressed in the presentation (and paper, if submitted). Continue with a discussion of why it is important to address this problem. This may be followed by some summary information about the models and methods developed and/or used to address the problem. Conclude with a description of the key results and contributions that will be covered in the presentation (and paper).\\

{\bf Keywords}: first keyword; second keyword; third keyword; fourth keyword.
}\\

\setlength{\parindent}{0pt}

{\bf 1. Introduction}

%We need to switch to downscaling models
%But they have issues of low power.
%While prevalence points should not be primary data source, they should still be used.
%One way is ML.

High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \cite{gething2011new, bhatt2017improved, gething2012long} are often not effective due to the large sample sizes needed to for accurate estimates at low prevalence rates and because of the lack of nationally representative prevalence surveys in low burden countries \cite{sturrock2016mapping, sturrock2014fine}. 
Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more widely available and of better quality and recent work has focussed on methods for estimating high-resolution malaria risk from these data \cite{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}. 
However, the aggregating of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogenous areas. 
Given these restrictions, modelling any of the non linearities known to be important in malaria estimation is difficult. 
A model that combines point surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.

One approach for combining these data is to use prevalence point-surveys to train a suite of machine learning models, and then use predictions from these models as covariates in a model trained on polygon-level incidence data. 
This process of stacking models has proven effective in many realms however typical stacking uses a single dataset on a consistent scale \cite{Sill2009, bhatt2017improved}. 
Here we propose training the level zero machine learning models on point-level, binomial prevalence data and stacking these models with a polygon-level, Poisson incidence model. 
As case studies we use data from Colombia and Madagascar as these countries have  good routine surveillance data.

{\bf 2. Methods}

We used two data sources that reflect malaria burden; point prevalence surveys and polygon-level, aggregated incidence data. 
We selected Colombia and Madagascar as case examples as they  both have good surveillance data.

The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from, 1990 onwards \cite{bhatt2015effect}. 
As the prevalence surveys cover different age ranges they were standardised to an age range of 2-10 using the model from \cite{smith2007standardizing}. 
To allow for differing relationships between environment and malaria prevalence we fitted separate models for Colombia and Indonesia. 
For Colombia we used all points from South America while for Madagascar we used only Madagascan data.

The polygon level surveillance data were collated from various government reports. To account for underreporting of clinical cases due to lack of treatment seeking, missed case reports and cases that sought medical attention outside the public health systems, the methods in \cite{cibulskis2011worldwide} were used. Where species specific reports were given, these were used. For reports where no species specific, national estimates of the ratio between \emph{P. falciparum} and \emph{P. vivax} cases were used to calculate \emph{P. falciparum} only cases. To minimise temporal effects we selected, for each country, one year of surveillance data. 
We used surveillance data from 2015 for Colombia and data from 2013 for Indonesia.

Raster surfaces of population for the years 2005, 2010 and 2015, were created using a hybrid of data from GPWv4 \cite{gpw4} and WorldPop \cite{tatem2017worldpop}, with the latter taking priority for those pixels where both had population data. 
Final population rasters were created by linear interpolation of the surrounding five-yearly rasters. 
Finally, for each year, national population estimates from the UN were raked over the interpolated population surfaces.

We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included land surface temperature annual mean and standard deviation, enhanced vegetation index, mosquito temperature suitability index, elevation, tassel cap brightness, tassel cap wetness, accessibility to cities, night lights and proportion of urban land cover. 
Elevation, land surface temperature standard deviation, accessibility to cities and night lights were all log transformed to reduce skewness. 
After preliminary analyses, tassel cap brightness and urban land cover were removed as they were highly correlated with other variables. 
The covariates were standardised to have a mean of zero and a standard deviation of one. 
These covariates were used for both the machine learning models and the polygon-level models.

For each region we fitted five models via caret \cite{caret}: elastic net \cite{enet}, Random Forest \cite{ranger}, projection pursuit regression \cite{PPR}, neural networks \cite{nnet} and boosted regression trees \cite{gbm}.
Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
For each model we ran 5 fold cross validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
The parameters tried can be seen in Table S1.
Predictions from these models were then made across Colombia and Madagascar.

This model is defined by a likelihood at the level of the polygon (Figure 1A) with covariates and a spatial random field at the pixel-level. Values at the polygon-level are given the subscript $a$ while pixel level values are indexed with $b$.

The polygon case count data, $y_j$ is given a Poisson likelihood

$$y_a \sim \operatorname{Pois}(i_a\mathrm{pop_a})$$

where $i_a$ is the estimated polygon incidence rate and $\mathrm{pop_a}$ is the observed polygon population-at-risk. This polygon-level likelihood is linked to the pixel level linear predictor by

$$i_a = \frac{ \sum(i_b \mathrm{pop}_b)}{\sum  \mathrm{pop}_b} $$
$$i_b = \mathrm{p2i}(p_b)$$

where $\mathrm{p2i}$ is from a model that was published previously. \cite{cameron2015defining} which defines a function
$${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$

The linear predictor of the model, $\eta_b$, is related to prevalence by a typical logit link function.

$$p_b = \operatorname{logit}^{-1}(\eta_b)$$

The form of the link function means we get predictions of prevalence and incidence simultaneously whether we used both data types or just one.
The linear predictor is composed of an intercept, covariates a spatial, Gaussian random field two iid random effects.

$$\eta_b = \beta_0 + \beta X  + u(s, \rho, \sigma_u) + v_j(\sigma_v) + w_i(\sigma_w)$$

The Gaussian spatial effect $u(s, \rho, \sigma_u)$ has a Mat\'ern covariance function and two hyper parameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
The first iid random effect, $v_j \sim \operatorname{Norm}(0, \sigma_v)$,  was grouped by polygon, with all pixels and point surveys within polygon $j$ being grouped together.
This random effect modelled both missing covariates and extra-Poisson sampling error. 
The second iid random effect, $w_i \sim \operatorname{Norm}(0, \sigma_w)$, was applied to each point survey.
This effect modelled extra-binomial sampling noise.
As such, this random effect is not included in the predicted uncertainty in the incidence or prevalence layers.
However, it is included in predictions of out-of-sample point surveys.

Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$, $\sigma_v$ and $\sigma_w$. We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \cite{fuglstad2018constructing} such that $P(\rho < \zeta) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$. We used different $\zeta$ values for each country: Colombia $\zeta = 1$, Indonesia $\zeta = 3$. We believe that a large proportion of the variance of malaria prevalence and incidence cannot be explained by a linear combination of the covariates selected \cite{bhatt2017improved}, so we set this prior such that the random field could explain most of the range of the data.

We assigned $\sigma_v$ a penalised complexity prior \cite{simpson2017penalising} such that $P(\sigma_v > 0.05) = 0.0000001$. This was based on a comparison of the variance of Poisson random variables, with rates given by the number of cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \cite{cibulskis2011worldwide}. We found that an iid effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
We assigned $\sigma_w$ a penalised complexity prior such that $P(\sigma_w > 0.3) = 0.0000001$. 
This was chosen by finding the maximum difference in prevalence between survey points (with a sample size greater than 500) within the same raster pixel.
The differences between points within the same pixel can only be accounted for by the binomial error and this iid effect.
Given that the error on a prevalence estimate with sample size greater than 500 is quite small, the iid effect needs to be able to explain this difference.

Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.04)$. 
Given the standardised covariates, a regression coefficient from the 95\% IQR of this distribution, and an intercept of 3, each covariate would be able to predict prevalences between 0.004 and 0.27. 
This prior encodes our belief that the full range of malaria transmission can not be explained by a single covariate and our desire to regularised our model.

The models were implemented and fitted using Template Model Builder \cite{TMB} in R \cite{R}

To compare the three models we used two cross validation schemes. 
In the first, polygon incidence data was randomly split into 10 cross-validation folds.
In the second, polygon incidence data was split spatially (5 folds in Colombia, 7 folds in Indonesia).
This cross-validation experiment is testing the models’ ability to make predictions far from data where the spatial random field is not informative.
Our primary performance metric was mean absolute error.

{\bf 3. Results}

\begin{figure}
\centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs/cv1_scatter.png}
        \caption{}
        \label{fig:gull}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs/cv2_scatter.png}
        \caption{}
        \label{fig:gull}
    \end{subfigure}
\caption{{\bf Observed-predicted plots for random cross-validation experiments.}
A) Indonesia, B) Senegal, C) Madagascar. Square root prevalence.
Results from all three models are included with the points only model being shown in orange, the downscaling model being shown in blue and the combined model being shown in green.
}
\label{randompredobspointfacet}
\end{figure}

\begin{figure}
\centering
\includegraphics[trim={0 40mm 0 100mm}, width = 0.6\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
\caption{{\bf Observed-predicted plots for random cross-validation experiments.}
A) Indonesia, B) Senegal, C) Madagascar. Square root prevalence.
Results from all three models are included with the points only model being shown in orange, the downscaling model being shown in blue and the combined model being shown in green.
}
\label{randompredobspointfacet}
\end{figure}

xxx  \\

{\bf 5. Conclusions}

xxx\\

{\bf References}\\

\bibliography{Malaria} 
\bibliographystyle{plain}

\end{document}

### 2018-12-22 tim-lucas-wsc.tex
First pass edits to methods.

@@
 As the prevalence surveys cover different age ranges they were standardised to an age range of 2-10 using the model from \cite{smith2007standardizing}. 
-To allow for differing relationships between environment and malaria prevalence we fitted separate models for Colombia and Indonesia. 
 For Colombia we used all points from South America while for Madagascar we used only Madagascan data.
@@
 For each model we ran 5 fold cross validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
-The parameters tried can be seen in Table S1.
+%The parameters tried can be seen in Table S1.
 Predictions from these models were then made across Colombia and Madagascar.
+These predictions were finally inverse logit transformed so that they are on the correct scale.
 
-This model is defined by a likelihood at the level of the polygon (Figure 1A) with covariates and a spatial random field at the pixel-level.
+The top level model is a disaggregation regression model \cite{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
+This model is defined by a likelihood at the level of the polygon with covariates and a spatial random field at the pixel-level. 
 Values at the polygon-level are given the subscript $a$ while pixel level values are indexed with $b$.
@@
 
-where $i_a$ is the estimated polygon incidence rate and $\mathrm{pop_a}$ is the observed polygon population-at-risk. This polygon-level likelihood is linked to the pixel level linear predictor by
+where $i_a$ is the estimated polygon incidence rate and $\mathrm{pop_a}$ is the observed polygon population-at-risk. 
+This polygon-level likelihood is linked to the pixel level linear predictor by
 
@@
 
-where $\mathrm{p2i}$ is from a model that was published previously. \cite{cameron2015defining} which defines a function
+where $\mathrm{p2i}$ is from a model that was published previously \cite{cameron2015defining} which defines a function
 $${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$
 
+The fact that the model passes through prevalence space allows predictions of prevalence to be made at the same time as predictions of incidence rate.
+Furthermore, it means that the predictions from the machine learning models are on an appropriate scale.
 The linear predictor of the model, $\eta_b$, is related to prevalence by a typical logit link function.
@@
 
-The form of the link function means we get predictions of prevalence and incidence simultaneously whether we used both data types or just one.
-The linear predictor is composed of an intercept, covariates a spatial, Gaussian random field two iid random effects.
+The linear predictor is composed of an intercept, covariates, a spatial, Gaussian random and an iid random effects.
 
-$$\eta_b = \beta_0 + \beta X  + u(s, \rho, \sigma_u) + v_j(\sigma_v) + w_i(\sigma_w)$$
+$$\eta_b = \beta_0 + \beta X  + u(s, \rho, \sigma_u) + v_j(\sigma_v)$$
 
 The Gaussian spatial effect $u(s, \rho, \sigma_u)$ has a Mat\'ern covariance function and two hyper parameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
-The first iid random effect, $v_j \sim \operatorname{Norm}(0, \sigma_v)$,  was grouped by polygon, with all pixels and point surveys within polygon $j$ being grouped together.
-This random effect modelled both missing covariates and extra-Poisson sampling error. 
-The second iid random effect, $w_i \sim \operatorname{Norm}(0, \sigma_w)$, was applied to each point survey.
-This effect modelled extra-binomial sampling noise.
-As such, this random effect is not included in the predicted uncertainty in the incidence or prevalence layers.
-However, it is included in predictions of out-of-sample point surveys.
+The iid random effect, $v_j \sim \operatorname{Norm}(0, \sigma_v)$, modelled both missing covariates and extra-Poisson sampling error
 
-Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$, $\sigma_v$ and $\sigma_w$.
-We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \cite{fuglstad2018constructing} such that $P(\rho < \zeta) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$.
-We used different $\zeta$ values for each country: Colombia $\zeta = 1$, Indonesia $\zeta = 3$.
+Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$ and $\sigma_v$.
+We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \cite{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$.
 We believe that a large proportion of the variance of malaria prevalence and incidence cannot be explained by a linear combination of the covariates selected \cite{bhatt2017improved}, so we set this prior such that the random field could explain most of the range of the data.
@@
 We found that an iid effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
-We assigned $\sigma_w$ a penalised complexity prior such that $P(\sigma_w > 0.3) = 0.0000001$. 
-This was chosen by finding the maximum difference in prevalence between survey points (with a sample size greater than 500) within the same raster pixel.
-The differences between points within the same pixel can only be accounted for by the binomial error and this iid effect.
-Given that the error on a prevalence estimate with sample size greater than 500 is quite small, the iid effect needs to be able to explain this difference.
-
-Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.04)$. 
-Given the standardised covariates, a regression coefficient from the 95\% IQR of this distribution, and an intercept of 3, each covariate would be able to predict prevalences between 0.004 and 0.27. 
-This prior encodes our belief that the full range of malaria transmission can not be explained by a single covariate and our desire to regularised our model.
-
+Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.4)$. 
 The models were implemented and fitted using Template Model Builder \cite{TMB} in R \cite{R}
 
+We compared the performance of the models with three sets of covariates.
+Firstly, we used the environmental and anthropogenic covariates, centered and standardised.
+Secondly, we used just the predictions from the machine learning models.
+Finally we combined these two sets of covariates.
+
 To compare the three models we used two cross validation schemes. 
-In the first, polygon incidence data was randomly split into 10 cross-validation folds.
-In the second, polygon incidence data was split spatially (5 folds in Colombia, 7 folds in Indonesia).
+In the first, polygon incidence data was randomly split into six cross-validation folds.
+In the second, polygon incidence data was split spatially into three folds (via k-means clustering on the polygon centroids).
 This cross-validation experiment is testing the models’ ability to make predictions far from data where the spatial random field is not informative.
-Our primary performance metric was mean absolute error.
+Our primary performance metric was correlation between observed and predicted data.
 
@@
 \includegraphics[trim={0 40mm 0 100mm}, width = 0.6\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
-\caption{{\bf Observed-predicted plots for random cross-validation experiments.}
-A) Indonesia, B) Senegal, C) Madagascar. Square root prevalence.
-Results from all three models are included with the points only model being shown in orange, the downscaling model being shown in blue and the combined model being shown in green.
+\caption{
+  Left: Observed polygon data for Colombia.
+Grey indicates incidence of zero.
+Right: Out-of-sample predictions for the random cross-validation, machine learning model only model.
+For each cross-validation fold, predictions are made for the held out data.
+These predictions are all plotted which together covers the entire study area.
 }

### 2018-12-22 tim-lucas-wsc.tex
Start writing results.

@@
 \begin{center}
-{Tim C. D. Lucas*}\\
-{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - timcdlucas@gmail.com}\\ 
-\vspace{0.2cm}
-
-{Anita Nandi}\\
-{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - anita.nandi@bdi.ox.ac.uk }\\
-
-\vspace{0.1cm}
-
-{Michele Nguyen}\\
-{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - michele.nguyen@bdi.ox.ac.uk }\\
-
-\vspace{0.1cm}
-
-{Susan Rumisha}\\
-{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - susan.rumisha@bdi.ox.ac.uk }\\
-
-\vspace{0.1cm}
-
-{Ewan Cameron}\\
-{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - dr.ewan.cameron@gmail.com }\\
-
-\vspace{0.1cm}
-{Pete Gething}\\
-{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - peter.gething@bdi.ox.ac.uk }\\
-
-\vspace{0.1cm}
-{Daniel J. Weiss}\\
-{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - daniel.weiss@bdi.ox.ac.uk }\\
+{Tim C. D. Lucas\textsuperscript{1}; Anita Nandi\textsuperscript{1}; Michele Nguyen\textsuperscript{1}; 
+Susan Rumisha\textsuperscript{1}; Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel J. Weiss\textsuperscript{1};}\\
+{1. Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - timcdlucas@gmail.com}\\
 
@@
 
-{\bf 2. Methods}
+{\bf 2. Methodology}
 
@@
 
+In both Colombia and Madagascar, and in both the random and spatial cross-validation scheme, the model using only machine learning predictions as covariates was the best performing model (Figure~\ref{f:scatter}, Table~\ref{t:results}).
+As expected models performed better in the random cross-validation scheme than the spatial cross-validation scheme in all cases.
+Furthermore the predictive ability of the all models in Colombia in the spatial cross-validation scheme was poor.
+This shows that spatial random field is an important component of the model even when using machine learning models.
+
 \begin{figure}
@@
         \includegraphics[width=\textwidth]{figs/cv1_scatter.png}
-        \caption{}
-        \label{fig:gull}
+        \caption{Random cross-validation}
+        \label{fig:cv1}
     \end{subfigure}
@@
         \includegraphics[width=\textwidth]{figs/cv2_scatter.png}
-        \caption{}
-        \label{fig:gull}
+        \caption{Spatial cross-validation}
+        \label{fig:cv1}
     \end{subfigure}
-\caption{{\bf Observed-predicted plots for random cross-validation experiments.}
-A) Indonesia, B) Senegal, C) Madagascar. Square root prevalence.
-Results from all three models are included with the points only model being shown in orange, the downscaling model being shown in blue and the combined model being shown in green.
+\caption{
+  Observed data against predictions for cross-validation hold-out samples on a square root transformed scale.
+  a) Seven-fold random cross-validation.
+  b) Three-fold spatial cross-validation with folds indicated by colour.
 }
-\label{randompredobspointfacet}
+\label{f:scatter}
 \end{figure}
@@
 
-xxx  \\
+\begin{table}
+\caption{Pearson correlations between observed and predicted values. }
+\centering
+\begin{tabular}{lllll}
+Cross-validation scheme & Country &  Covariates &  ML &  Covs + ML \\
+\hline 
+ Random &  Colombia &  0.45 &  \textbf{0.55} &  0.54 \\
+ Random &  Madagascar &  0.70 &  \textbf{0.76} &  0.75 \\
+ Spatial &  Colombia &  0.05 &  \textbf{0.18} &  0.10 \\
+ Spatial &  Madagascar &  0.22 &  \textbf{0.63} &  0.61 
+\end{tabular}
+\label{t:results}
+\end{table}
 
 {\bf 5. Conclusions}
-
-xxx\\
 

### 2018-12-23 tim-lucas-wsc.tex
Some tidying up.

@@
 
-{\bf Keywords}: first keyword; second keyword; third keyword; fourth keyword.
+{\bf Keywords}: Spatial statistics; Ensemble; Stacking; Epidemiology.
 }\\
@@
 One approach for combining these data is to use prevalence point-surveys to train a suite of machine learning models, and then use predictions from these models as covariates in a model trained on polygon-level incidence data.
-This process of stacking models has proven effective in many realms however typical stacking uses a single dataset on a consistent scale \cite{Sill2009, bhatt2017improved}. 
+This process of stacking models has proven effective in many realms however typical stacking uses a single dataset on a consistent scale \cite{sill2009feature, bhatt2017improved}. 
 Here we propose training the level zero machine learning models on point-level, binomial prevalence data and stacking these models with a polygon-level, Poisson incidence model. 
@@
 
-For each region we fitted five models via caret \cite{caret}: elastic net \cite{enet}, Random Forest \cite{ranger}, projection pursuit regression \cite{PPR}, neural networks \cite{nnet} and boosted regression trees \cite{gbm}.
+For each region we fitted five models via caret \cite{caret}: elastic net \cite{enet}, Random Forest \cite{wright2015ranger}, projection pursuit regression \cite{friedman1981projection}, neural networks \cite{nnet} and boosted regression trees \cite{gbm}.
 Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
@@
 
-In both Colombia and Madagascar, and in both the random and spatial cross-validation scheme, the model using only machine learning predictions as covariates was the best performing model (Figure~\ref{f:scatter}, Table~\ref{t:results}).
+Overall, model performance was good in Madagascar under both random and spatial cross-validation (Figure~\ref{f:scatter}).
+Model performance was reasonable in Colombia under random cross-validation but poor under spatial cross-validation (Figure~\ref{f:scatter}).
+This indicates that the covariates alone cannot explain malaria incidence in this area.
+Mapped predictions of the best performing model look reasonable and are not unduly dominated by the random field.
+
+The model using only machine learning predictions as covariates was the best performing model in both countries and both cross-validation schemes (Table~\ref{t:results}).
 As expected models performed better in the random cross-validation scheme than the spatial cross-validation scheme in all cases.
-Furthermore the predictive ability of the all models in Colombia in the spatial cross-validation scheme was poor.
-This shows that spatial random field is an important component of the model even when using machine learning models.
+The difference in performance between the best and worst models was always smaller than the difference between the random and spatial cross-validation schemes.
+
+% models were quite good except spatial colombia
+% which model performed best
+% which ml model was best Vs parameters.
+% parameters are negative and don't sum to one
+%
+% differences between models was less than difference between data.
 
@@
   Observed data against predictions for cross-validation hold-out samples on a square root transformed scale.
-  a) Seven-fold random cross-validation.
+  a) Six-fold random cross-validation.
   b) Three-fold spatial cross-validation with folds indicated by colour.
@@
 \centering
-\begin{tabular}{lllll}
+\begin{tabular}{llrrr}
 Cross-validation scheme & Country &  Covariates &  ML &  Covs + ML \\
@@
 
+Overall, we have shown that using predictions from machine learning models trained on prevalence points is better than using environmental covariates.
+This increased performance comes despite the data being on different scales, the data being measurements of different aspects of malaria transmission and despite the imperfect model we have used to translate the between the two scales.
+However, the poor performance in the spatial cross-validation schemes highlights that better spatial coverage of data would improve predictions more than the improved model we have suggested.
+
+While we have used a small suite if covariates, the fact that models such as Random Forest and elastic net can have many covariates means we could include a much larger set of environmental variables.
+This is in strong contrast to previous downscaling models which had to use a very small number of covariates due to the relatively small number of polygons available and because data associated with polygons is under-powered for learning many parameters.
+
+% this method is an improvement
+% still reliant on spatial field
+
+% allows broad suite of covariates
+
+% sum to one stacking
+% global ml models, local experts
+
+Our top-level polygon model is not technically stacking as we have not contrained the regression parameters to be positive now or added a sum-to-constraint.
+We did not include these constraints because the base models and the meta model are trained on response data on different scales.
+However, future work could examine whether using a positive constraint on the regression parameters improves performance.
+
+Another area of potential improvement is ranking varying the data used to train the base level learners.
+Here we only used data from the region of interest (South America fit the Colombia model and Madagascar for the Madagascar model).
+However, the global dataset is much larger than these subsets.
+Training some base level models on local s data and some on the global dataset and then combining predictions from all these models has potential to further improve model performance.
+
 {\bf References}\\
@@
 
+\begin{table}
+\caption{ML model results and fitted parameters. }
+\centering
+\begin{tabular}{l|rrr|rrr}
+               &&      Madagascar &&      &                                     Colombia&    \\
+Model          & PR RMSE & Poly beta 1 & Poly beta 2 & PR RMSE & Poly beta 1  & Poly beta 2\\
+\hline
+nnet           & 0.113      &  0.031         &  0.025         & \textbf{0.058}& -0.250          & -0.246          \\
+RF   &  \textbf{0.100}      &  0.337         &  0.350         & \textbf{0.058}&  0.782          &  0.742          \\
+gbm            & 0.109      &  \textbf{0.450}&  \textbf{0.402}& 0.066         &  \textbf{0.835} &  \textbf{0.775} \\
+enet           & 0.116      &  0.326         &  0.307         & \textbf{0.058}& -0.563          & -0.369          \\
+ppr            & 0.110      & -0.233         & -0.204         & 0.059         &  0.210          &  0.166          \\
+\end{tabular}
+\label{t:mlresults}
+\end{table}
+
 \end{document}

### 2018-12-23 tim-lucas-wsc.tex
Sort refs out I think.

@@
 \usepackage{subcaption}
+
+\usepackage{natbib}
+
+%\setcitestyle{authoryear}
 
@@
 \begin{center}
-{Tim C. D. Lucas\textsuperscript{1}; Anita Nandi\textsuperscript{1}; Michele Nguyen\textsuperscript{1}; 
+{Tim C. D. Lucas*\textsuperscript{1}; Anita Nandi\textsuperscript{1}; Michele Nguyen\textsuperscript{1}; 
 Susan Rumisha\textsuperscript{1}; Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel J. Weiss\textsuperscript{1};}\\
@@
 
-High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \cite{gething2011new, bhatt2017improved, gething2012long} are often not effective due to the large sample sizes needed to for accurate estimates at low prevalence rates and because of the lack of nationally representative prevalence surveys in low burden countries \cite{sturrock2016mapping, sturrock2014fine}.
-Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more widely available and of better quality and recent work has focussed on methods for estimating high-resolution malaria risk from these data \cite{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
+High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{gething2011new, bhatt2017improved, gething2012long} are often not effective due to the large sample sizes needed to for accurate estimates at low prevalence rates and because of the lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}.
+Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more widely available and of better quality and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
 However, the aggregating of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogenous areas. 
@@
 One approach for combining these data is to use prevalence point-surveys to train a suite of machine learning models, and then use predictions from these models as covariates in a model trained on polygon-level incidence data.
-This process of stacking models has proven effective in many realms however typical stacking uses a single dataset on a consistent scale \cite{sill2009feature, bhatt2017improved}. 
+This process of stacking models has proven effective in many realms however typical stacking uses a single dataset on a consistent scale \citep{sill2009feature, bhatt2017improved}. 
 Here we propose training the level zero machine learning models on point-level, binomial prevalence data and stacking these models with a polygon-level, Poisson incidence model. 
@@
 
-The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from, 1990 onwards \cite{bhatt2015effect}. 
-As the prevalence surveys cover different age ranges they were standardised to an age range of 2-10 using the model from \cite{smith2007standardizing}. 
+The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from, 1990 onwards \citep{bhatt2015effect}. 
+As the prevalence surveys cover different age ranges they were standardised to an age range of 2-10 using the model from \citep{smith2007standardizing}. 
 For Colombia we used all points from South America while for Madagascar we used only Madagascan data.
@@
 The polygon level surveillance data were collated from various government reports.
-To account for underreporting of clinical cases due to lack of treatment seeking, missed case reports and cases that sought medical attention outside the public health systems, the methods in \cite{cibulskis2011worldwide} were used.
+To account for underreporting of clinical cases due to lack of treatment seeking, missed case reports and cases that sought medical attention outside the public health systems, the methods in \citep{cibulskis2011worldwide} were used.
 Where species specific reports were given, these were used.
@@
 
-Raster surfaces of population for the years 2005, 2010 and 2015, were created using a hybrid of data from GPWv4 \cite{gpw4} and WorldPop \cite{tatem2017worldpop}, with the latter taking priority for those pixels where both had population data.
+Raster surfaces of population for the years 2005, 2010 and 2015, were created using a hybrid of data from GPWv4 \citep{gpw4} and WorldPop \citep{tatem2017worldpop}, with the latter taking priority for those pixels where both had population data.
 Final population rasters were created by linear interpolation of the surrounding five-yearly rasters. 
@@
 
-For each region we fitted five models via caret \cite{caret}: elastic net \cite{enet}, Random Forest \cite{wright2015ranger}, projection pursuit regression \cite{friedman1981projection}, neural networks \cite{nnet} and boosted regression trees \cite{gbm}.
+For each region we fitted five models via caret \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
 Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
@@
 
-The top level model is a disaggregation regression model \cite{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
+The top level model is a disaggregation regression model \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
 This model is defined by a likelihood at the level of the polygon with covariates and a spatial random field at the pixel-level. 
@@
 
-where $\mathrm{p2i}$ is from a model that was published previously \cite{cameron2015defining} which defines a function
+where $\mathrm{p2i}$ is from a model that was published previously \citep{cameron2015defining} which defines a function
 $${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$
@@
 Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$ and $\sigma_v$.
-We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \cite{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$.
-We believe that a large proportion of the variance of malaria prevalence and incidence cannot be explained by a linear combination of the covariates selected \cite{bhatt2017improved}, so we set this prior such that the random field could explain most of the range of the data.
+We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \citep{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$.
+We believe that a large proportion of the variance of malaria prevalence and incidence cannot be explained by a linear combination of the covariates selected \citep{bhatt2017improved}, so we set this prior such that the random field could explain most of the range of the data.
 
-We assigned $\sigma_v$ a penalised complexity prior \cite{simpson2017penalising} such that $P(\sigma_v > 0.05) = 0.0000001$.
-This was based on a comparison of the variance of Poisson random variables, with rates given by the number of cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \cite{cibulskis2011worldwide}.
+We assigned $\sigma_v$ a penalised complexity prior \citep{simpson2017penalising} such that $P(\sigma_v > 0.05) = 0.0000001$.
+This was based on a comparison of the variance of Poisson random variables, with rates given by the number of cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \citep{cibulskis2011worldwide}.
 We found that an iid effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
 Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.4)$. 
-The models were implemented and fitted using Template Model Builder \cite{TMB} in R \cite{R}
+The models were implemented and fitted using Template Model Builder \citep{TMB} in R \citep{R}
 
@@
         \caption{Spatial cross-validation}
-        \label{fig:cv1}
+        \label{fig:cv2}
     \end{subfigure}
@@
 
-{\bf References}\\
-
 \bibliography{Malaria} 
-\bibliographystyle{plain}
+\bibliographystyle{apalike}
 

### 2018-12-28 tim-lucas-wsc.tex
More edits.

@@
 This indicates that the covariates alone cannot explain malaria incidence in this area.
-Mapped predictions of the best performing model look reasonable and are not unduly dominated by the random field.
+Mapped predictions of the best performing model look reasonable and are not unduly dominated by the random field (Figure~\ref{f:map}).
 
@@
 
+The cross-validation results for the machine learning models are shown in Table~\ref{t:ml results}.
+Also in Table~\ref{t:ml results} is the means (across folds) of the  regression coefficients for these models from the polygon-level models that used only predictions from machine learning models as covariates.
+Predictive performance was similar across different models, with Random Forest performing best in Madagascar and neural networks, Random Forests and elastic net performing equally well on Colombia.
+The estimated regression parameters are similar between the random and spatial cross-validation schemes.
+However, the best performing machibe learning models do not have the largest estimated regression coefficients.
+Also of note is that some models were estimated to have a negative relationship (conditional on the presence of the other models) with polygon incidence.
+
 % models were quite good except spatial colombia
 % which model performed best
+% Random field parameters?
+% coverage
 % which ml model was best Vs parameters.
@@
 }
-\label{randompredobspointfacet}
+\label{f:map}
 \end{figure}

### 2018-12-28 template_paper_v1.tex
Add template.

--- 2018-12-23 tim-lucas-wsc.tex
@@
 \usepackage{graphicx}
-\usepackage{amsmath}
 \usepackage{geometry}
-\usepackage{subcaption}
-
 \geometry{verbose,tmargin=30mm,bmargin=25mm,lmargin=25mm,rmargin=25mm}
@@
 %The title should be centred and in bold letters. It should be informative but not too long (preferably no more than two lines).
-\textbf{Model ensembles with different response variables for base and meta models: malaria disaggregation regression combining prevalence and incidence data}
+\textbf{Title of the paper (preferably up to two lines)}
 \end{center}
 
-% Model ensembles with different response variables for base and meta models:
+\begin{center}
+{Name Surname*}\\
+{Institution, City, Country - e-mail address}\\ 
+\vspace{0.5cm}
 
-\begin{center}
-{Tim C. D. Lucas\textsuperscript{1}; Anita Nandi\textsuperscript{1}; Michele Nguyen\textsuperscript{1}; 
-Susan Rumisha\textsuperscript{1}; Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel J. Weiss\textsuperscript{1};}\\
-{1. Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - timcdlucas@gmail.com}\\
+{Name Surname}\\
+{Institution, City, Country - e-mail address}\\
 
@@
 
-{\bf Keywords}: Spatial statistics; Ensemble; Stacking; Epidemiology.
+{\bf Keywords}: first keyword; second keyword; third keyword; fourth keyword.
 }\\
@@
 
-%We need to switch to downscaling models
-%But they have issues of low power.
-%While prevalence points should not be primary data source, they should still be used.
-%One way is ML.
+xxx\\
 
-High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \cite{gething2011new, bhatt2017improved, gething2012long} are often not effective due to the large sample sizes needed to for accurate estimates at low prevalence rates and because of the lack of nationally representative prevalence surveys in low burden countries \cite{sturrock2016mapping, sturrock2014fine}.
-Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more widely available and of better quality and recent work has focussed on methods for estimating high-resolution malaria risk from these data \cite{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
-However, the aggregating of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogenous areas. 
-Given these restrictions, modelling any of the non linearities known to be important in malaria estimation is difficult. 
-A model that combines point surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.
+{\bf 2. Section 2}
 
-One approach for combining these data is to use prevalence point-surveys to train a suite of machine learning models, and then use predictions from these models as covariates in a model trained on polygon-level incidence data.
-This process of stacking models has proven effective in many realms however typical stacking uses a single dataset on a consistent scale \cite{sill2009feature, bhatt2017improved}. 
-Here we propose training the level zero machine learning models on point-level, binomial prevalence data and stacking these models with a polygon-level, Poisson incidence model. 
-As case studies we use data from Colombia and Madagascar as these countries have  good routine surveillance data.
+xxx\\
 
-{\bf 2. Methodology}
+{\bf 3. Section 3}
 
-We used two data sources that reflect malaria burden; point prevalence surveys and polygon-level, aggregated incidence data. 
-We selected Colombia and Madagascar as case examples as they both have good surveillance data.
+xxx  \\
 
-The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from, 1990 onwards \cite{bhatt2015effect}. 
-As the prevalence surveys cover different age ranges they were standardised to an age range of 2-10 using the model from \cite{smith2007standardizing}. 
-For Colombia we used all points from South America while for Madagascar we used only Madagascan data.
+{\bf 4. Section 4}
 
-The polygon level surveillance data were collated from various government reports.
-To account for underreporting of clinical cases due to lack of treatment seeking, missed case reports and cases that sought medical attention outside the public health systems, the methods in \cite{cibulskis2011worldwide} were used.
-Where species specific reports were given, these were used.
-For reports where no species specific, national estimates of the ratio between \emph{P.
-falciparum} and \emph{P.
-vivax} cases were used to calculate \emph{P.
-falciparum} only cases.
-To minimise temporal effects we selected, for each country, one year of surveillance data.
-We used surveillance data from 2015 for Colombia and data from 2013 for Indonesia.
-
-Raster surfaces of population for the years 2005, 2010 and 2015, were created using a hybrid of data from GPWv4 \cite{gpw4} and WorldPop \cite{tatem2017worldpop}, with the latter taking priority for those pixels where both had population data.
-Final population rasters were created by linear interpolation of the surrounding five-yearly rasters. 
-Finally, for each year, national population estimates from the UN were raked over the interpolated population surfaces.
-
-We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included land surface temperature annual mean and standard deviation, enhanced vegetation index, mosquito temperature suitability index, elevation, tassel cap brightness, tassel cap wetness, accessibility to cities, night lights and proportion of urban land cover.
-Elevation, land surface temperature standard deviation, accessibility to cities and night lights were all log transformed to reduce skewness. 
-After preliminary analyses, tassel cap brightness and urban land cover were removed as they were highly correlated with other variables. 
-The covariates were standardised to have a mean of zero and a standard deviation of one. 
-These covariates were used for both the machine learning models and the polygon-level models.
-
-For each region we fitted five models via caret \cite{caret}: elastic net \cite{enet}, Random Forest \cite{wright2015ranger}, projection pursuit regression \cite{friedman1981projection}, neural networks \cite{nnet} and boosted regression trees \cite{gbm}.
-Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
-For each model we ran 5 fold cross validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
-%The parameters tried can be seen in Table S1.
-Predictions from these models were then made across Colombia and Madagascar.
-These predictions were finally inverse logit transformed so that they are on the correct scale.
-
-The top level model is a disaggregation regression model \cite{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
-This model is defined by a likelihood at the level of the polygon with covariates and a spatial random field at the pixel-level. 
-Values at the polygon-level are given the subscript $a$ while pixel level values are indexed with $b$.
-
-The polygon case count data, $y_j$ is given a Poisson likelihood
-
-$$y_a \sim \operatorname{Pois}(i_a\mathrm{pop_a})$$
-
-where $i_a$ is the estimated polygon incidence rate and $\mathrm{pop_a}$ is the observed polygon population-at-risk. 
-This polygon-level likelihood is linked to the pixel level linear predictor by
-
-$$i_a = \frac{ \sum(i_b \mathrm{pop}_b)}{\sum  \mathrm{pop}_b} $$
-$$i_b = \mathrm{p2i}(p_b)$$
-
-where $\mathrm{p2i}$ is from a model that was published previously \cite{cameron2015defining} which defines a function
-$${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$
-
-The fact that the model passes through prevalence space allows predictions of prevalence to be made at the same time as predictions of incidence rate.
-Furthermore, it means that the predictions from the machine learning models are on an appropriate scale.
-The linear predictor of the model, $\eta_b$, is related to prevalence by a typical logit link function.
-
-$$p_b = \operatorname{logit}^{-1}(\eta_b)$$
-
-The linear predictor is composed of an intercept, covariates, a spatial, Gaussian random and an iid random effects.
-
-$$\eta_b = \beta_0 + \beta X  + u(s, \rho, \sigma_u) + v_j(\sigma_v)$$
-
-The Gaussian spatial effect $u(s, \rho, \sigma_u)$ has a Mat\'ern covariance function and two hyper parameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
-The iid random effect, $v_j \sim \operatorname{Norm}(0, \sigma_v)$, modelled both missing covariates and extra-Poisson sampling error
-
-Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$ and $\sigma_v$.
-We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \cite{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$.
-We believe that a large proportion of the variance of malaria prevalence and incidence cannot be explained by a linear combination of the covariates selected \cite{bhatt2017improved}, so we set this prior such that the random field could explain most of the range of the data.
-
-We assigned $\sigma_v$ a penalised complexity prior \cite{simpson2017penalising} such that $P(\sigma_v > 0.05) = 0.0000001$.
-This was based on a comparison of the variance of Poisson random variables, with rates given by the number of cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \cite{cibulskis2011worldwide}.
-We found that an iid effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
-Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.4)$. 
-The models were implemented and fitted using Template Model Builder \cite{TMB} in R \cite{R}
-
-We compared the performance of the models with three sets of covariates.
-Firstly, we used the environmental and anthropogenic covariates, centered and standardised.
-Secondly, we used just the predictions from the machine learning models.
-Finally we combined these two sets of covariates.
-
-To compare the three models we used two cross validation schemes. 
-In the first, polygon incidence data was randomly split into six cross-validation folds.
-In the second, polygon incidence data was split spatially into three folds (via k-means clustering on the polygon centroids).
-This cross-validation experiment is testing the models’ ability to make predictions far from data where the spatial random field is not informative.
-Our primary performance metric was correlation between observed and predicted data.
-
-{\bf 3. Results}
-
-Overall, model performance was good in Madagascar under both random and spatial cross-validation (Figure~\ref{f:scatter}).
-Model performance was reasonable in Colombia under random cross-validation but poor under spatial cross-validation (Figure~\ref{f:scatter}).
-This indicates that the covariates alone cannot explain malaria incidence in this area.
-Mapped predictions of the best performing model look reasonable and are not unduly dominated by the random field.
-
-The model using only machine learning predictions as covariates was the best performing model in both countries and both cross-validation schemes (Table~\ref{t:results}).
-As expected models performed better in the random cross-validation scheme than the spatial cross-validation scheme in all cases.
-The difference in performance between the best and worst models was always smaller than the difference between the random and spatial cross-validation schemes.
-
-% models were quite good except spatial colombia
-% which model performed best
-% which ml model was best Vs parameters.
-% parameters are negative and don't sum to one
-%
-% differences between models was less than difference between data.
-
-\begin{figure}
-\centering
-    \begin{subfigure}[b]{0.49\textwidth}
-        \includegraphics[width=\textwidth]{figs/cv1_scatter.png}
-        \caption{Random cross-validation}
-        \label{fig:cv1}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.49\textwidth}
-        \includegraphics[width=\textwidth]{figs/cv2_scatter.png}
-        \caption{Spatial cross-validation}
-        \label{fig:cv1}
-    \end{subfigure}
-\caption{
-  Observed data against predictions for cross-validation hold-out samples on a square root transformed scale.
-  a) Six-fold random cross-validation.
-  b) Three-fold spatial cross-validation with folds indicated by colour.
-}
-\label{f:scatter}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[trim={0 40mm 0 100mm}, width = 0.6\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
-\caption{
-  Left: Observed polygon data for Colombia.
-Grey indicates incidence of zero.
-Right: Out-of-sample predictions for the random cross-validation, machine learning model only model.
-For each cross-validation fold, predictions are made for the held out data.
-These predictions are all plotted which together covers the entire study area.
-}
-\label{randompredobspointfacet}
-\end{figure}
-
-\begin{table}
-\caption{Pearson correlations between observed and predicted values. }
-\centering
-\begin{tabular}{llrrr}
-Cross-validation scheme & Country &  Covariates &  ML &  Covs + ML \\
-\hline 
- Random &  Colombia &  0.45 &  \textbf{0.55} &  0.54 \\
- Random &  Madagascar &  0.70 &  \textbf{0.76} &  0.75 \\
- Spatial &  Colombia &  0.05 &  \textbf{0.18} &  0.10 \\
- Spatial &  Madagascar &  0.22 &  \textbf{0.63} &  0.61 
-\end{tabular}
-\label{t:results}
-\end{table}
+xxx\\
 
@@
 
-Overall, we have shown that using predictions from machine learning models trained on prevalence points is better than using environmental covariates.
-This increased performance comes despite the data being on different scales, the data being measurements of different aspects of malaria transmission and despite the imperfect model we have used to translate the between the two scales.
-However, the poor performance in the spatial cross-validation schemes highlights that better spatial coverage of data would improve predictions more than the improved model we have suggested.
-
-While we have used a small suite if covariates, the fact that models such as Random Forest and elastic net can have many covariates means we could include a much larger set of environmental variables.
-This is in strong contrast to previous downscaling models which had to use a very small number of covariates due to the relatively small number of polygons available and because data associated with polygons is under-powered for learning many parameters.
-
-% this method is an improvement
-% still reliant on spatial field
-
-% allows broad suite of covariates
-
-% sum to one stacking
-% global ml models, local experts
-
-Our top-level polygon model is not technically stacking as we have not contrained the regression parameters to be positive now or added a sum-to-constraint.
-We did not include these constraints because the base models and the meta model are trained on response data on different scales.
-However, future work could examine whether using a positive constraint on the regression parameters improves performance.
-
-Another area of potential improvement is ranking varying the data used to train the base level learners.
-Here we only used data from the region of interest (South America fit the Colombia model and Madagascar for the Madagascar model).
-However, the global dataset is much larger than these subsets.
-Training some base level models on local s data and some on the global dataset and then combining predictions from all these models has potential to further improve model performance.
+xxx\\
 
@@
 
-\bibliography{Malaria} 
-\bibliographystyle{plain}
+Fisher, R. A. (1925). Statistical methods for research workers. Genesis Publishing Pvt Ltd.\\
 
-\begin{table}
-\caption{ML model results and fitted parameters. }
-\centering
-\begin{tabular}{l|rrr|rrr}
-               &&      Madagascar &&      &                                     Colombia&    \\
-Model          & PR RMSE & Poly beta 1 & Poly beta 2 & PR RMSE & Poly beta 1  & Poly beta 2\\
-\hline
-nnet           & 0.113      &  0.031         &  0.025         & \textbf{0.058}& -0.250          & -0.246          \\
-RF   &  \textbf{0.100}      &  0.337         &  0.350         & \textbf{0.058}&  0.782          &  0.742          \\
-gbm            & 0.109      &  \textbf{0.450}&  \textbf{0.402}& 0.066         &  \textbf{0.835} &  \textbf{0.775} \\
-enet           & 0.116      &  0.326         &  0.307         & \textbf{0.058}& -0.563          & -0.369          \\
-ppr            & 0.110      & -0.233         & -0.204         & 0.059         &  0.210          &  0.166          \\
-\end{tabular}
-\label{t:mlresults}
-\end{table}
+McCullagh, P., \& Nelder, J. A. (1983). Generalized linear models. London, England, Chapman and Hall.\\
+
+Box, G. E., Jenkins, G. M., \& Reinsel, G. C. (2013). Time series analysis: forecasting and control. John Wiley \& Sons.
 

### 2019-01-07 tim-lucas-wsc.tex
add abstrct and edits.

@@
 {Tim C. D. Lucas*\textsuperscript{1}; Anita Nandi\textsuperscript{1}; Michele Nguyen\textsuperscript{1}; 
-Susan Rumisha\textsuperscript{1}; Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel J. Weiss\textsuperscript{1};}\\
+Susan Rumisha\textsuperscript{1}; Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel J. Weiss\textsuperscript{1}.}\\
 {1. Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - timcdlucas@gmail.com}\\
@@
 
-This is where the abstract is placed.
+% This is where the abstract is placed.
 It should include a statement about the problem being addressed in the presentation (and paper, if submitted).
@@
 Conclude with a description of the key results and contributions that will be covered in the presentation (and paper).\\
+
+Maps of infection risk are a vital tool for the elimination of malaria.
+Routine surveillance data of malaria case counts, often aggregated over administrative regions, is becoming more widely available and can better measure low malaria risk than prevalence surveys.
+However, aggregation of case counts over large, heterogenous areas means that these data are often underpowered for learning relationships between the environment and malaria risk.
+A model that combines point surveys and aggregated surveillance data could have the benefits of both but must be able to account for the fact that these two data types are on different scales.
+Here, we train multiple machine learning models on point surveys and then combine the predictions from these with a geostatistical disaggregation model that uses routine surveillance data.
+We find that, in tests using data from Colombia and Madagascar, combining predictions from machine learning models out performs models using environmental covariates.
 
@@
 
-High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{gething2011new, bhatt2017improved, gething2012long} are often not effective due to the large sample sizes needed to for accurate estimates at low prevalence rates and because of the lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}.
+High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{gething2011new, bhatt2017improved, gething2012long, bhatt2015effect} are often not effective due to the large sample sizes needed for accurate estimates at low prevalence rates and because of the lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}.
 Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more widely available and of better quality and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
-However, the aggregating of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogenous areas. 
-Given these restrictions, modelling any of the non linearities known to be important in malaria estimation is difficult. 
+However, the aggregation of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogenous areas. 
+Given these restrictions, modelling the non-linearities known to be important in malaria mapping is difficult \cite{bhatt2017improved, bhatt2015effect}. 
 A model that combines point surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.
@@
 Here we propose training the level zero machine learning models on point-level, binomial prevalence data and stacking these models with a polygon-level, Poisson incidence model.
-As case studies we use data from Colombia and Madagascar as these countries have  good routine surveillance data.
 
@@
 
-We used two data sources that reflect malaria burden; point prevalence surveys and polygon-level, aggregated incidence data. 
+We used two data sources that reflect malaria burden; point-prevalence surveys and polygon-level, aggregated incidence data. 
 We selected Colombia and Madagascar as case examples as they both have good surveillance data.
@@
 
-The cross-validation results for the machine learning models are shown in Table~\ref{t:ml results}.
-Also in Table~\ref{t:ml results} is the means (across folds) of the  regression coefficients for these models from the polygon-level models that used only predictions from machine learning models as covariates.
+The cross-validation results for the machine learning models are shown in Table~\ref{t:mlresults}.
+Also in Table~\ref{t:mlresults} is the means (across folds) of the  regression coefficients for these models from the polygon-level models that used only predictions from machine learning models as covariates.
 Predictive performance was similar across different models, with Random Forest performing best in Madagascar and neural networks, Random Forests and elastic net performing equally well on Colombia.
@@
 \centering
-\includegraphics[trim={0 40mm 0 100mm}, width = 0.6\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
+\includegraphics[trim={0 40mm 0 40mm}, width = 0.55\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
 \caption{
@@
 \label{t:results}
+\end{table}
+
+\begin{table}
+\caption{ML model results and fitted parameters. }
+\centering
+\begin{tabular}{l|rrr|rrr}
+               &&      Madagascar &&      &                                     Colombia&    \\
+Model          & PR RMSE & Poly beta 1 & Poly beta 2 & PR RMSE & Poly beta 1  & Poly beta 2\\
+\hline
+nnet           & 0.113      &  0.031         &  0.025         & \textbf{0.058}& -0.250          & -0.246          \\
+RF   &  \textbf{0.100}      &  0.337         &  0.350         & \textbf{0.058}&  0.782          &  0.742          \\
+gbm            & 0.109      &  \textbf{0.450}&  \textbf{0.402}& 0.066         &  \textbf{0.835} &  \textbf{0.775} \\
+enet           & 0.116      &  0.326         &  0.307         & \textbf{0.058}& -0.563          & -0.369          \\
+ppr            & 0.110      & -0.233         & -0.204         & 0.059         &  0.210          &  0.166          \\
+\end{tabular}
+\label{t:mlresults}
 \end{table}
@@
 
-\begin{table}
-\caption{ML model results and fitted parameters. }
-\centering
-\begin{tabular}{l|rrr|rrr}
-               &&      Madagascar &&      &                                     Colombia&    \\
-Model          & PR RMSE & Poly beta 1 & Poly beta 2 & PR RMSE & Poly beta 1  & Poly beta 2\\
-\hline
-nnet           & 0.113      &  0.031         &  0.025         & \textbf{0.058}& -0.250          & -0.246          \\
-RF   &  \textbf{0.100}      &  0.337         &  0.350         & \textbf{0.058}&  0.782          &  0.742          \\
-gbm            & 0.109      &  \textbf{0.450}&  \textbf{0.402}& 0.066         &  \textbf{0.835} &  \textbf{0.775} \\
-enet           & 0.116      &  0.326         &  0.307         & \textbf{0.058}& -0.563          & -0.369          \\
-ppr            & 0.110      & -0.233         & -0.204         & 0.059         &  0.210          &  0.166          \\
-\end{tabular}
-\label{t:mlresults}
-\end{table}
-
 \end{document}

### 2019-01-07 tim-lucas-wsc.tex
More edits.

@@
 We selected Colombia and Madagascar as case examples as they both have good surveillance data.
-
-The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from, 1990 onwards \citep{bhatt2015effect}. 
+The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect}.  % ceck todo
 As the prevalence surveys cover different age ranges they were standardised to an age range of 2-10 using the model from \citep{smith2007standardizing}. 
-For Colombia we used all points from South America while for Madagascar we used only Madagascan data.
-
+For Colombia we used all points from South America (n = todo) while for Madagascar we used only Madagascan data (n = todo).
 The polygon level surveillance data were collated from various government reports. 
@@
 To minimise temporal effects we selected, for each country, one year of surveillance data.
-We used surveillance data from 2015 for Colombia and data from 2013 for Indonesia.
+We used surveillance data from 2015 for Colombia (n = todo) and data from todo for Madagascar (n = todo).
 
@@
 Final population rasters were created by linear interpolation of the surrounding five-yearly rasters.
-Finally, for each year, national population estimates from the UN were raked over the interpolated population surfaces.
 
-We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included land surface temperature annual mean and standard deviation, enhanced vegetation index, mosquito temperature suitability index, elevation, tassel cap brightness, tassel cap wetness, accessibility to cities, night lights and proportion of urban land cover.
-Elevation, land surface temperature standard deviation, accessibility to cities and night lights were all log transformed to reduce skewness. 
+We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, mosquito temperature suitability index, elevation, tassel cap brightness, tassel cap wetness, log accessibility to cities, log night lights and proportion of urban land cover.
 After preliminary analyses, tassel cap brightness and urban land cover were removed as they were highly correlated with other variables. 
@@
 Predictions from these models were then made across Colombia and Madagascar.
-These predictions were finally inverse logit transformed so that they are on the correct scale.
+These predictions were finally inverse logit transformed so that they are on the correct scale for the top level model.
 
@@
 
-The polygon case count data, $y_j$ is given a Poisson likelihood
+The polygon case count data, $y_a$ is given a Poisson likelihood
 
@@
 where $i_a$ is the estimated polygon incidence rate and $\mathrm{pop_a}$ is the observed polygon population-at-risk. 
-This polygon-level likelihood is linked to the pixel level linear predictor by
+This polygon-level likelihood is linked to the pixel level prevealence
 
@@
 
-The fact that the model passes through prevalence space allows predictions of prevalence to be made at the same time as predictions of incidence rate.
-Furthermore, it means that the predictions from the machine learning models are on an appropriate scale.
+The fact that the model passes through prevalence space allows predictions of prevalence to be made at the same time as predictions of incidence rate and that the predictions from the machine learning models are on an appropriate scale.
 The linear predictor of the model, $\eta_b$, is related to prevalence by a typical logit link function.
@@
 
-The linear predictor is composed of an intercept, covariates, a spatial, Gaussian random and an iid random effects.
+The linear predictor is composed of an intercept, covariates, a spatial, Gaussian random field and an iid random effect.
 
@@
 
-The Gaussian spatial effect $u(s, \rho, \sigma_u)$ has a Mat\'ern covariance function and two hyper parameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
+The Gaussian spatial effect $u(s, \rho, \sigma_u)$ has a Mat\'ern covariance function and two hyperparameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
 The iid random effect, $v_j \sim \operatorname{Norm}(0, \sigma_v)$, modelled both missing covariates and extra-Poisson sampling error
@@
 We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \citep{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$. 
-We believe that a large proportion of the variance of malaria prevalence and incidence cannot be explained by a linear combination of the covariates selected \citep{bhatt2017improved}, so we set this prior such that the random field could explain most of the range of the data.
+We set this prior such that the random field could explain most of the range of the data if required.
 
@@
 This indicates that the covariates alone cannot explain malaria incidence in this area.
-Mapped predictions of the best performing model look reasonable and are not unduly dominated by the random field (Figure~\ref{f:map}).
+Mapped predictions of the best performing model in Colombia look reasonable and are not unduly dominated by the random field (Figure~\ref{f:map}).
 
@@
 As expected models performed better in the random cross-validation scheme than the spatial cross-validation scheme in all cases.
+The difference between the covariate only model and the machine learning predictions only model was greater in the spatial cross-validation scheme than the random cross-validation.
 The difference in performance between the best and worst models was always smaller than the difference between the random and spatial cross-validation schemes.
@@
 The cross-validation results for the machine learning models are shown in Table~\ref{t:mlresults}.
-Also in Table~\ref{t:mlresults} is the means (across folds) of the  regression coefficients for these models from the polygon-level models that used only predictions from machine learning models as covariates.
-Predictive performance was similar across different models, with Random Forest performing best in Madagascar and neural networks, Random Forests and elastic net performing equally well on Colombia.
+Also in Table~\ref{t:mlresults} is the means (across folds) of the  regression coefficients (i.e.
+the weights of the machine learning models in the level zero model) from the polygon-level models that used only predictions from machine learning models as covariates.
+Predictive performance was similar across different models, with Random Forest performing best in Madagascar and neural networks, Random Forests and elastic net performing equally well in Colombia.
 The estimated regression parameters are similar between the random and spatial cross-validation schemes.
-However, the best performing machibe learning models do not have the largest estimated regression coefficients.
-Also of note is that some models were estimated to have a negative relationship (conditional on the presence of the other models) with polygon incidence.
+However, the best performing machine learning models do not have the largest estimated regression coefficients as would be expected if prevalence and incidence were completely correlated.
+Also of note is that some models were estimated to have a negative relationship (conditional on the inclusion of predictions from other machine learning models) with polygon incidence.
 
@@
 \begin{table}
-\caption{ML model results and fitted parameters. }
+\caption{Machine learning model results and means of fitted parameters across $k$ model. }
 \centering
+\small
 \begin{tabular}{l|rrr|rrr}
                &&      Madagascar &&      &                                     Colombia&    \\
-Model          & PR RMSE & Poly beta 1 & Poly beta 2 & PR RMSE & Poly beta 1  & Poly beta 2\\
+Model          & ML RMSE & Random CV $\bar{\beta}_i$ & Spatial CV $\bar{\beta}_i$  & ML RMSE & Random CV $\bar{\beta}_i$  & Spatial CV $\bar{\beta}_i$  \\
 \hline
@@
 
-Overall, we have shown that using predictions from machine learning models trained on prevalence points is better than using environmental covariates.
-This increased performance comes despite the data being on different scales, the data being measurements of different aspects of malaria transmission and despite the imperfect model we have used to translate the between the two scales.
+Overall, our experiments suggest that using predictions from machine learning models trained on prevalence points is better than using environmental covariates when fitting disaggregation models of malaria incidence.
+This increased performance comes despite the data being on different scales, the data being measurements of different aspects of malaria transmission and despite the imperfect model we have used to translate between the two scales.
 However, the poor performance in the spatial cross-validation schemes highlights that better spatial coverage of data would improve predictions more than the improved model we have suggested.
 
-While we have used a small suite if covariates, the fact that models such as Random Forest and elastic net can have many covariates means we could include a much larger set of environmental variables.
-This is in strong contrast to previous downscaling models which had to use a very small number of covariates due to the relatively small number of polygons available and because data associated with polygons is under-powered for learning many parameters.
+While we have used a small suite of covariates, the fact that models such as Random Forest and elastic net can have many covariates means we could include a much larger set of environmental variables.
+This is in strong contrast to previous downscaling models which had to use a very small number of covariates due to the relatively small number of polygons available and because data associated with polygons is underpowered for learning many parameters \citep{sturrock2014fine}.
 
@@
 
-Our top-level polygon model is not technically stacking as we have not contrained the regression parameters to be positive now or added a sum-to-constraint.
+Our top-level polygon model is not technically stacking as we have not contrained the regression parameters to be positive nor included a sum-to-one constraint.
 We did not include these constraints because the base models and the meta model are trained on response data on different scales.
@@
 
-Another area of potential improvement is ranking varying the data used to train the base level learners.
-Here we only used data from the region of interest (South America fit the Colombia model and Madagascar for the Madagascar model).
+Another area of potential improvement is varying the data used to train the base level learners.
+Here we only used data from the region of interest (South America for the Colombia model and Madagascar for the Madagascar model).
 However, the global dataset is much larger than these subsets.
-Training some base level models on local s data and some on the global dataset and then combining predictions from all these models has potential to further improve model performance.
+Training some base level models on local data and some on the global dataset and then combining predictions from all these models has potential to further improve model performance.
 

### 2019-01-07 tim-lucas-wsc.tex
Add ns

@@
 {Tim C. D. Lucas*\textsuperscript{1}; Anita Nandi\textsuperscript{1}; Michele Nguyen\textsuperscript{1}; 
-Susan Rumisha\textsuperscript{1}; Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel J. Weiss\textsuperscript{1}.}\\
+Susan Rumisha\textsuperscript{1}; 
+Katherine E. Battle\textsuperscript{1}; Rosalind E. Howes\textsuperscript{1}; 
+Chantal Hendriks\textsuperscript{1}; Andre Python\textsuperscript{1}; Penny Hancock\textsuperscript{1}; 
+Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel~J.~Weiss\textsuperscript{1}.}\\
 {1. Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - timcdlucas@gmail.com}\\
@@
 The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect}.  % ceck todo
-As the prevalence surveys cover different age ranges they were standardised to an age range of 2-10 using the model from \citep{smith2007standardizing}. 
-For Colombia we used all points from South America (n = todo) while for Madagascar we used only Madagascan data (n = todo).
+As the prevalence surveys cover different age ranges they were standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}. 
+For Colombia we used all points from South America (n = 522) while for Madagascar we used only Madagascan data (n = 1505).
 The polygon level surveillance data were collated from various government reports. 
@@
 To minimise temporal effects we selected, for each country, one year of surveillance data.
-We used surveillance data from 2015 for Colombia (n = todo) and data from todo for Madagascar (n = todo).
+We used surveillance data from 2015 for Colombia (n = 952) and data from todo for Madagascar (n = 110).
 

### 2019-01-10 tim-lucas-wsc.tex
Andre edits.

@@
 Here, we train multiple machine learning models on point surveys and then combine the predictions from these with a geostatistical disaggregation model that uses routine surveillance data.
-We find that, in tests using data from Colombia and Madagascar, combining predictions from machine learning models out performs models using environmental covariates.
+We find that, in tests using data from Colombia and Madagascar, using a disaggregation regression model to combine predictions from machine learning models trained in point surveys improves model accuracy.
 
@@
 
-High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{gething2011new, bhatt2017improved, gething2012long, bhatt2015effect} are often not effective due to the large sample sizes needed for accurate estimates at low prevalence rates and because of the lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}.
-Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more widely available and of better quality and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
-However, the aggregation of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogenous areas. 
-Given these restrictions, modelling the non-linearities known to be important in malaria mapping is difficult \cite{bhatt2017improved, bhatt2015effect}. 
+High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{gething2011new, bhatt2017improved, gething2012long, bhatt2015effect} is often not effective  because, firstly, so few individuals are infected that most surveys will detect zero cases, and secondly, because of the lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}.
+Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more widely available and more reliable and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
+However, the aggregation of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogenous areas, because it is unclear where within the polygon, and in which environments, the cases occurred.
+This data is therefore often under-powered for fitting flexible, non-linear models known to be required for accurate malaria maps \cite{bhatt2017improved, bhatt2015effect}. 
 A model that combines point surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.
@@
 We selected Colombia and Madagascar as case examples as they both have good surveillance data.
-The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect}.  % ceck todo
+The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling}.  % ceck todo
 As the prevalence surveys cover different age ranges they were standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}. 
 For Colombia we used all points from South America (n = 522) while for Madagascar we used only Madagascan data (n = 1505).
-The polygon level surveillance data were collated from various government reports. 
-To account for underreporting of clinical cases due to lack of treatment seeking, missed case reports and cases that sought medical attention outside the public health systems, the methods in \citep{cibulskis2011worldwide} were used.
-Where species specific reports were given, these were used.
-For reports where no species specific, national estimates of the ratio between \emph{P.
-falciparum} and \emph{P.
+The polygon incidence data were collected from government reports and standardised using methods defined in \cite{cibulskis2011worldwide}.
+This standardisation step accounts for missed cases due to lack of treatment seeking, missing case reports, and cases that sought medical attention outside the public health systems \citep{battle2016treatment}.
+For reports where cases were not reported at the species level, national estimates of the ratio between \emph{Plasmodium falciparum} and \emph{P.
 vivax} cases were used to calculate \emph{P.
@@
 To minimise temporal effects we selected, for each country, one year of surveillance data. 
-We used surveillance data from 2015 for Colombia (n = 952) and data from todo for Madagascar (n = 110).
+We used surveillance data from 2015 for Colombia (n = 952) and data from 2013 for Madagascar (n = 110).
 
-Raster surfaces of population for the years 2005, 2010 and 2015, were created using a hybrid of data from GPWv4 \citep{gpw4} and WorldPop \citep{tatem2017worldpop}, with the latter taking priority for those pixels where both had population data.
-Final population rasters were created by linear interpolation of the surrounding five-yearly rasters.
+Raster surfaces of population for the years 2005, 2010 and 2015, were created using data from GPWv4 \citep{gpw4} and WorldPop \citep{tatem2017worldpop}, with the latter taking priority for those pixels where both had population data.
+Population rasters for the remaining years were created by linear interpolation.
 
-We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, mosquito temperature suitability index, elevation, tassel cap brightness, tassel cap wetness, log accessibility to cities, log night lights and proportion of urban land cover.
+We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, mosquito temperature suitability index, elevation, tassel cap brightness, tassel cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \cite{weiss2015re}.
 After preliminary analyses, tassel cap brightness and urban land cover were removed as they were highly correlated with other variables. 
@@
 Predictions from these models were then made across Colombia and Madagascar.
-These predictions were finally inverse logit transformed so that they are on the correct scale for the top level model.
+These predictions were finally inverse logit transformed so that they are on the linear predictor scale of the top level model.
 
@@
 The polygon case count data, $y_a$ is given a Poisson likelihood
-
 $$y_a \sim \operatorname{Pois}(i_a\mathrm{pop_a})$$
-
 where $i_a$ is the estimated polygon incidence rate and $\mathrm{pop_a}$ is the observed polygon population-at-risk. 
 This polygon-level likelihood is linked to the pixel level prevealence 
-
 $$i_a = \frac{ \sum(i_b \mathrm{pop}_b)}{\sum  \mathrm{pop}_b} $$
 $$i_b = \mathrm{p2i}(p_b)$$
-
 where $\mathrm{p2i}$ is from a model that was published previously \citep{cameron2015defining} which defines a function
 $${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$
-
-The fact that the model passes through prevalence space allows predictions of prevalence to be made at the same time as predictions of incidence rate and that the predictions from the machine learning models are on an appropriate scale.
-The linear predictor of the model, $\eta_b$, is related to prevalence by a typical logit link function.
-
-$$p_b = \operatorname{logit}^{-1}(\eta_b)$$
-
-The linear predictor is composed of an intercept, covariates, a spatial, Gaussian random field and an iid random effect.
-
-$$\eta_b = \beta_0 + \beta X  + u(s, \rho, \sigma_u) + v_j(\sigma_v)$$
-
+The fact that the model passes through prevalence space means that the predictions from the machine learning models are on an appropriate scale.
+The linear predictor of the model is related to prevalence by a typical logit link function and includes an intercept, covariates, a spatial, Gaussian, random field and an iid random effect.
+$$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u(s, \rho, \sigma_u) + v_j(\sigma_v)\right)$$
 The Gaussian spatial effect $u(s, \rho, \sigma_u)$ has a Mat\'ern covariance function and two hyperparameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
@@
 We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \citep{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$. 
+This prior encodes our a priori preference for a simpler, smoother random field.
 We set this prior such that the random field could explain most of the range of the data if required.
@@
 
-Overall, model performance was good in Madagascar under both random and spatial cross-validation (Figure~\ref{f:scatter}).
-Model performance was reasonable in Colombia under random cross-validation but poor under spatial cross-validation (Figure~\ref{f:scatter}).
+Overall, model performance was good in Madagascar under both random and spatial cross-validation (Figure~\ref{f:scatter}, Table~\ref{t:results}).
+Model performance was reasonable in Colombia under random cross-validation but poor under spatial cross-validation (Figure~\ref{f:scatter}, Table~\ref{t:results}).
 This indicates that the covariates alone cannot explain malaria incidence in this area.
@@
 \begin{table}
-\caption{Machine learning model results and means of fitted parameters across $k$ model. }
+\caption{Machine learning model results and means of fitted parameters (i.e.
+model weights) across $k$ cross-validation folds of the model which only contained machine learning model predictions as covariates.
+}
 \centering

### 2019-01-10 tim-lucas-wsc.tex
Finis andre edits.

@@
 Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
-For each model we ran 5 fold cross validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
+For each model we ran 5 fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
 %The parameters tried can be seen in Table S1.
@@
 $${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$
-The fact that the model passes through prevalence space means that the predictions from the machine learning models are on an appropriate scale.
+The fact that the model passes through prevalence space means that the predictions from the machine learning models can be appropriately scaled.
 The linear predictor of the model is related to prevalence by a typical logit link function and includes an intercept, covariates, a spatial, Gaussian, random field and an iid random effect.
@@
 We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \citep{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$. 
-This prior encodes our a priori preference for a simpler, smoother random field.
+This prior encodes our \emph{a priori} preference for a simpler, smoother random field.
 We set this prior such that the random field could explain most of the range of the data if required.
@@
 
-To compare the three models we used two cross validation schemes. 
+To compare the three models we used two cross-validation schemes. 
 In the first, polygon incidence data was randomly split into six cross-validation folds.
@@
 
-Overall, model performance was good in Madagascar under both random and spatial cross-validation (Figure~\ref{f:scatter}, Table~\ref{t:results}).
-Model performance was reasonable in Colombia under random cross-validation but poor under spatial cross-validation (Figure~\ref{f:scatter}, Table~\ref{t:results}).
-This indicates that the covariates alone cannot explain malaria incidence in this area.
-Mapped predictions of the best performing model in Colombia look reasonable and are not unduly dominated by the random field (Figure~\ref{f:map}).
+Figure~\ref{f:scatter} shows the model performance under random and spatial cross-validation for both Madagascar and Colombia. 
+The poor model performance in Colombia under spatial cross-validation indicates that the covariates alone cannot explain malaria incidence in this area. 
+For all other models that use the machine learning predictions as covariates, correlations of 0.54 -- 0.76, between observed and predicted data, are obtained (Table~\ref{t:results}).
+Input data and mapped out-of-sample predictions in Colombia, of the best performing model, are shown in in Figure~\ref{f:map}.
 
@@
 
-Overall, our experiments suggest that using predictions from machine learning models trained on prevalence points is better than using environmental covariates when fitting disaggregation models of malaria incidence.
+Overall, our experiments suggest that using predictions from machine learning models trained on prevalence points provides more accurate predictions than using environmental covariates when fitting disaggregation models of malaria incidence.
 This increased performance comes despite the data being on different scales, the data being measurements of different aspects of malaria transmission and despite the imperfect model we have used to translate between the two scales.
-However, the poor performance in the spatial cross-validation schemes highlights that better spatial coverage of data would improve predictions more than the improved model we have suggested.
+However, the reduced model accuracy in the spatial cross-validation schemes, relative to the random cross-validation scheme, highlights that better spatial coverage of data would improve predictions more than the improved model we have suggested.
 
-While we have used a small suite of covariates, the fact that models such as Random Forest and elastic net can have many covariates means we could include a much larger set of environmental variables.
-This is in strong contrast to previous downscaling models which had to use a very small number of covariates due to the relatively small number of polygons available and because data associated with polygons is underpowered for learning many parameters \citep{sturrock2014fine}.
+Due to the low power of typical aggregated incidence datasets, previous analysis using disaggregation regression had to use a small number of covariates \citep{sturrock2014fine}.
+However, the increased amount of data provided by using prevalence points, and the the fact that model such as Random Forest and elastic net can use many coviariates means future work could include many more covariates, potentially increasing predictive performance further.
 

### 2019-01-10 tim-lucas-wsc.tex
Add anita edits.

@@
 The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling}.  % ceck todo
+For Colombia we used all points from South America (n = 522) while for Madagascar we used only Malagasy data (n = 1505).
+We chose these geographic regions based on a trade-off between wanting a large sample size but wanting data from geographically similar areas.
 As the prevalence surveys cover different age ranges they were standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
-For Colombia we used all points from South America (n = 522) while for Madagascar we used only Madagascan data (n = 1505).
 The polygon incidence data were collected from government reports and standardised using methods defined in \cite{cibulskis2011worldwide}.
@@
 To minimise temporal effects we selected, for each country, one year of surveillance data. 
-We used surveillance data from 2015 for Colombia (n = 952) and data from 2013 for Madagascar (n = 110).
+We used surveillance data from 2015 for Colombia (n = 952) and data from 2013 for Madagascar (n = 110) as these years had the most data in each case.
 
@@
 The fact that the model passes through prevalence space means that the predictions from the machine learning models can be appropriately scaled.
-The linear predictor of the model is related to prevalence by a typical logit link function and includes an intercept, covariates, a spatial, Gaussian, random field and an iid random effect.
+The linear predictor of the model is related to prevalence by a typical logit link function and includes an intercept, $\beta_0$, covariates, $X$ with regression parameters $\beta$, a spatial, Gaussian, random field, $u(s, \rho, \sigma_u)$, and an iid random effect, $v_j(\sigma_v)$.
 $$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u(s, \rho, \sigma_u) + v_j(\sigma_v)\right)$$
-The Gaussian spatial effect $u(s, \rho, \sigma_u)$ has a Mat\'ern covariance function and two hyperparameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
-The iid random effect, $v_j \sim \operatorname{Norm}(0, \sigma_v)$, modelled both missing covariates and extra-Poisson sampling error
+The Gaussian spatial effect has a Mat\'ern covariance function and two hyperparameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
+The iid random effect modelled both missing covariates and extra-Poisson sampling error.
 
@@
 Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.4)$. 
-The models were implemented and fitted using Template Model Builder \citep{TMB} in R \citep{R}
+The models were implemented and fitted using Template Model Builder \citep{TMB} in R \citep{R}.
 
-We compared the performance of the models with three sets of covariates.
+We compared the performance of the models with three sets of covariates, $X$.
 Firstly, we used the environmental and anthropogenic covariates, centered and standardised.
-Secondly, we used just the predictions from the machine learning models.
+Secondly, we used the predictions from the machine learning models.
 Finally we combined these two sets of covariates.

### 2019-01-10 tim-lucas-wsc.tex
Kate B edits.

@@
 
-We used two data sources that reflect malaria burden; point-prevalence surveys and polygon-level, aggregated incidence data. 
-We selected Colombia and Madagascar as case examples as they both have good surveillance data.
+We used two data sources that reflect malaria transmission; point-prevalence surveys and polygon-level, aggregated incidence data. 
+We selected Colombia and Madagascar as case examples as they both have fairly complete surveillance data at a finer geographical resolution than admin 1.
 The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling}.  % ceck todo
@@
 
-Raster surfaces of population for the years 2005, 2010 and 2015, were created using data from GPWv4 \citep{gpw4} and WorldPop \citep{tatem2017worldpop}, with the latter taking priority for those pixels where both had population data.
+Raster surfaces of population for the years 2005, 2010 and 2015, were created using data from WorldPop \citep{tatem2017worldpop} and from GPWv4 \citep{gpw4} where WorldPop did not have values. 
 Population rasters for the remaining years were created by linear interpolation.
 
-We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, mosquito temperature suitability index, elevation, tassel cap brightness, tassel cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \cite{weiss2015re}.
+We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tassel cap brightness, tassel cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \cite{weiss2015re}.
 After preliminary analyses, tassel cap brightness and urban land cover were removed as they were highly correlated with other variables. 
@@
 
-For each region we fitted five models via caret \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
+For each region we fitted five models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
 Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
-For each model we ran 5 fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
+For each model we ran  five-fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
 %The parameters tried can be seen in Table S1.
@@
 Input data and mapped out-of-sample predictions in Colombia, of the best performing model, are shown in in Figure~\ref{f:map}.
+
+\begin{table}[h!]
+\caption{Pearson correlations between observed and predicted values. }
+\centering
+\begin{tabular}{llrrr}
+Cross-validation scheme & Country &  Covariates &  ML &  Covs + ML \\
+\hline 
+ Random &  Colombia &  0.45 &  \textbf{0.55} &  0.54 \\
+ Random &  Madagascar &  0.70 &  \textbf{0.76} &  0.75 \\
+ Spatial &  Colombia &  0.05 &  \textbf{0.18} &  0.10 \\
+ Spatial &  Madagascar &  0.22 &  \textbf{0.63} &  0.61 
+\end{tabular}
+\label{t:results}
+\end{table}
 
@@
 \begin{table}
-\caption{Pearson correlations between observed and predicted values. }
-\centering
-\begin{tabular}{llrrr}
-Cross-validation scheme & Country &  Covariates &  ML &  Covs + ML \\
-\hline 
- Random &  Colombia &  0.45 &  \textbf{0.55} &  0.54 \\
- Random &  Madagascar &  0.70 &  \textbf{0.76} &  0.75 \\
- Spatial &  Colombia &  0.05 &  \textbf{0.18} &  0.10 \\
- Spatial &  Madagascar &  0.22 &  \textbf{0.63} &  0.61 
-\end{tabular}
-\label{t:results}
-\end{table}
-
-\begin{table}
 \caption{Machine learning model results and means of fitted parameters (i.e.
@@
 Due to the low power of typical aggregated incidence datasets, previous analysis using disaggregation regression had to use a small number of covariates \citep{sturrock2014fine}.
-However, the increased amount of data provided by using prevalence points, and the the fact that model such as Random Forest and elastic net can use many coviariates means future work could include many more covariates, potentially increasing predictive performance further.
+However, as models such as Random Forest and elastic net are robust in the presence of high dimensional data, future work could include many more covariates, potentially increasing predictive performance further.
 
@@
 
-Our top-level polygon model is not technically stacking as we have not contrained the regression parameters to be positive nor included a sum-to-one constraint.
+While the approach presented here is related to stacking, it differs in that we have not constrained the regression parameters to be positive nor included a sum-to-one constraint i.e.\thinspace the result is not simply a weighted average of the level zero model predictions.
 We did not include these constraints because the base models and the meta model are trained on response data on different scales.

### 2019-01-10 tim-lucas-wsc.tex
Ros edits.

@@
 However, aggregation of case counts over large, heterogenous areas means that these data are often underpowered for learning relationships between the environment and malaria risk.
-A model that combines point surveys and aggregated surveillance data could have the benefits of both but must be able to account for the fact that these two data types are on different scales.
+A model that combines point surveys and aggregated surveillance data could have the benefits of both but must be able to account for the fact that these two data types are different malariometric units.
 Here, we train multiple machine learning models on point surveys and then combine the predictions from these with a geostatistical disaggregation model that uses routine surveillance data.
@@
 
-We used two data sources that reflect malaria transmission; point-prevalence surveys and polygon-level, aggregated incidence data. 
-We selected Colombia and Madagascar as case examples as they both have fairly complete surveillance data at a finer geographical resolution than admin 1.
+We used two data sources that reflect \emph{Plasmodium falciparum} malaria transmission; point-prevalence surveys and polygon-level, aggregated incidence data. 
+We selected Colombia and Madagascar as case examples as they both have fairly complete, publically available, surveillance data at a finer geographical resolution than admin 1.
 The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling}.  % ceck todo
@@
 This standardisation step accounts for missed cases due to lack of treatment seeking, missing case reports, and cases that sought medical attention outside the public health systems \citep{battle2016treatment}.
-For reports where cases were not reported at the species level, national estimates of the ratio between \emph{Plasmodium falciparum} and \emph{P.
+For reports where cases were not reported at the species level, national estimates of the ratio between \emph{P.
+falciparum} and \emph{P.
 vivax} cases were used to calculate \emph{P.
@@
 To minimise temporal effects we selected, for each country, one year of surveillance data. 
-We used surveillance data from 2015 for Colombia (n = 952) and data from 2013 for Madagascar (n = 110) as these years had the most data in each case.
+We used annual surveillance data from 2015 for Colombia (952 municipalities) and data from 2013 for Madagascar (110 districts) as these years had the most data in each case.
 
@@
 
-We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tassel cap brightness, tassel cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \cite{weiss2015re}.
-After preliminary analyses, tassel cap brightness and urban land cover were removed as they were highly correlated with other variables. 
+We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \cite{weiss2015re}.
+After preliminary analyses, tasseled cap brightness and urban land cover were removed as they were highly correlated with other variables. 
 The covariates were standardised to have a mean of zero and a standard deviation of one. 
@@
 
-For each region we fitted five models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
+For each country we fitted five models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
 Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
@@
 $${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$
-The fact that the model passes through prevalence space means that the predictions from the machine learning models can be appropriately scaled.
+The fact that the model passes through prevalence space ensures that the predictions from the machine learning models can be appropriately scaled.
 The linear predictor of the model is related to prevalence by a typical logit link function and includes an intercept, $\beta_0$, covariates, $X$ with regression parameters $\beta$, a spatial, Gaussian, random field, $u(s, \rho, \sigma_u)$, and an iid random effect, $v_j(\sigma_v)$.
@@
 We assigned $\sigma_v$ a penalised complexity prior \citep{simpson2017penalising} such that $P(\sigma_v > 0.05) = 0.0000001$. 
-This was based on a comparison of the variance of Poisson random variables, with rates given by the number of cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \citep{cibulskis2011worldwide}.
+This was based on a comparison of the variance of Poisson random variables, with rates given by the number of polygon-level cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \citep{cibulskis2011worldwide}.
 We found that an iid effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
@@
 In the second, polygon incidence data was split spatially into three folds (via k-means clustering on the polygon centroids).
-This cross-validation experiment is testing the models’ ability to make predictions far from data where the spatial random field is not informative.
+This cross-validation experiment is testing the models' ability to make predictions far from data where the spatial random field is not informative.
 Our primary performance metric was correlation between observed and predicted data.

### 2019-01-21 tim-lucas-wsc.tex
Final edits before submisison.

@@
 Here, we train multiple machine learning models on point surveys and then combine the predictions from these with a geostatistical disaggregation model that uses routine surveillance data.
-We find that, in tests using data from Colombia and Madagascar, using a disaggregation regression model to combine predictions from machine learning models trained in point surveys improves model accuracy.
+We find that, in tests using data from Colombia and Madagascar, using a disaggregation regression model to combine predictions from machine learning models trained on point surveys improves model accuracy relative to using the environmental covariates directly.
 
@@
 High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{gething2011new, bhatt2017improved, gething2012long, bhatt2015effect} is often not effective  because, firstly, so few individuals are infected that most surveys will detect zero cases, and secondly, because of the lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}.
-Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more widely available and more reliable and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
+Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more reliable and more widely available \citep{sturrock2016mapping} and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
 However, the aggregation of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogenous areas, because it is unclear where within the polygon, and in which environments, the cases occurred.
-This data is therefore often under-powered for fitting flexible, non-linear models known to be required for accurate malaria maps \cite{bhatt2017improved, bhatt2015effect}. 
+This data is therefore often under-powered for fitting flexible, non-linear models as is required for accurate malaria maps \citep{bhatt2017improved, bhatt2015effect}. 
 A model that combines point surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.
@@
 We chose these geographic regions based on a trade-off between wanting a large sample size but wanting data from geographically similar areas.
-As the prevalence surveys cover different age ranges they were standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
+The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
 The polygon incidence data were collected from government reports and standardised using methods defined in \cite{cibulskis2011worldwide}.
@@
 
+We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \citep{weiss2015re}.
+Tasseled cap brightness and urban land cover were subsequently removed as they were highly correlated with other variables. 
+The covariates were standardised to have a mean of zero and a standard deviation of one. 
+These covariates were used for both the machine learning models and the polygon-level models.
 Raster surfaces of population for the years 2005, 2010 and 2015, were created using data from WorldPop \citep{tatem2017worldpop} and from GPWv4 \citep{gpw4} where WorldPop did not have values. 
@@
 
-We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \cite{weiss2015re}.
-After preliminary analyses, tasseled cap brightness and urban land cover were removed as they were highly correlated with other variables. 
-The covariates were standardised to have a mean of zero and a standard deviation of one. 
-These covariates were used for both the machine learning models and the polygon-level models.
-
 For each country we fitted five models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
-Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
+Our response variable was prevalence and we weighted the data by sample size (i.e.\thinspace the number of people tested for malaria in each survey.
 For each model we ran  five-fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
 %The parameters tried can be seen in Table S1.
-Predictions from these models were then made across Colombia and Madagascar.
+Predictions from these models were then made across Colombia and Madagascar respectively.
 These predictions were finally inverse logit transformed so that they are on the linear predictor scale of the top level model.
 
-The top level model is a disaggregation regression model \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
+The top level model was a disaggregation regression model \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
 This model is defined by a likelihood at the level of the polygon with covariates and a spatial random field at the pixel-level. 
@@
 where $\mathrm{p2i}$ is from a model that was published previously \citep{cameron2015defining} which defines a function
-$${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$
+$$\mathrm{p2i}: f(P) = 2.616P - 3.596P^2 + 1.594P^3.$$
 The fact that the model passes through prevalence space ensures that the predictions from the machine learning models can be appropriately scaled.
-The linear predictor of the model is related to prevalence by a typical logit link function and includes an intercept, $\beta_0$, covariates, $X$ with regression parameters $\beta$, a spatial, Gaussian, random field, $u(s, \rho, \sigma_u)$, and an iid random effect, $v_j(\sigma_v)$.
+The linear predictor of the model is related to prevalence by a typical logit link function and includes an intercept, $\beta_0$, covariates, $X$ with regression parameters $\beta$, a spatial, Gaussian, random field, $u(s, \rho, \sigma_u)$, and an \emph{iid} random effect, $v_j(\sigma_v)$.
 $$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u(s, \rho, \sigma_u) + v_j(\sigma_v)\right)$$
 The Gaussian spatial effect has a Mat\'ern covariance function and two hyperparameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
-The iid random effect modelled both missing covariates and extra-Poisson sampling error.
+The \emph{iid} random effect models both missing covariates and extra-Poisson sampling error.
 
@@
 We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \citep{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$. 
-This prior encodes our \emph{a priori} preference for a simpler, smoother random field.
+This prior encoded our \emph{a priori} preference for a simpler, smoother random field.
 We set this prior such that the random field could explain most of the range of the data if required.
@@
 This was based on a comparison of the variance of Poisson random variables, with rates given by the number of polygon-level cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \citep{cibulskis2011worldwide}.
-We found that an iid effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
+We found that an \emph{iid} effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
 Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.4)$. 
@@
 In the second, polygon incidence data was split spatially into three folds (via k-means clustering on the polygon centroids).
-This cross-validation experiment is testing the models' ability to make predictions far from data where the spatial random field is not informative.
+This spatial cross-validation scheme is testing the models' ability to make predictions far from data where the spatial random field is not informative.
 Our primary performance metric was correlation between observed and predicted data.
@@
 The poor model performance in Colombia under spatial cross-validation indicates that the covariates alone cannot explain malaria incidence in this area. 
-For all other models that use the machine learning predictions as covariates, correlations of 0.54 -- 0.76, between observed and predicted data, are obtained (Table~\ref{t:results}).
-Input data and mapped out-of-sample predictions in Colombia, of the best performing model, are shown in in Figure~\ref{f:map}.
+For all other models that use the machine learning predictions as covariates, correlations between observed and predicted data of 0.54 -- 0.76 were achieved (Table~\ref{t:results}).
+Input data and mapped out-of-sample predictions of the best performing model, in Colombia, are shown in in Figure~\ref{f:map}.
 
@@
 The model using only machine learning predictions as covariates was the best performing model in both countries and both cross-validation schemes (Table~\ref{t:results}).
-As expected models performed better in the random cross-validation scheme than the spatial cross-validation scheme in all cases.
-The difference between the covariate only model and the machine learning predictions only model was greater in the spatial cross-validation scheme than the random cross-validation.
-The difference in performance between the best and worst models was always smaller than the difference between the random and spatial cross-validation schemes.
+As expected, models performed better in the random cross-validation scheme than the spatial cross-validation scheme.
+The difference between the covariate only model and the machine learning predictions only model was greater in the spatial cross-validation scheme than in the random cross-validation.
+The improvement in performance between the worst and best models was always smaller than the difference between the random and spatial cross-validation schemes.
 
-The cross-validation results for the machine learning models are shown in Table~\ref{t:mlresults}.
-Also in Table~\ref{t:mlresults} is the means (across folds) of the  regression coefficients (i.e.
-the weights of the machine learning models in the level zero model) from the polygon-level models that used only predictions from machine learning models as covariates.
-Predictive performance was similar across different models, with Random Forest performing best in Madagascar and neural networks, Random Forests and elastic net performing equally well in Colombia.
+Predictive performance of machine learning models was similar, with Random Forest performing best in Madagascar and neural networks, Random Forests and elastic net performing equally well in Colombia (Table~\ref{t:mlresults}).
+The means (across folds) of the  regression coefficients (i.e.\thinspace the weights of the machine learning models in the level zero model) from the polygon-level models that used only predictions from machine learning models as covariates can also be seen in Table~\ref{t:mlresults}.
 The estimated regression parameters are similar between the random and spatial cross-validation schemes.
 However, the best performing machine learning models do not have the largest estimated regression coefficients as would be expected if prevalence and incidence were completely correlated.
-Also of note is that some models were estimated to have a negative relationship (conditional on the inclusion of predictions from other machine learning models) with polygon incidence.
+Also of note is that some models were estimated to have a negative relationship with incidence (conditional on the inclusion of predictions from other machine learning models).
 
@@
 \centering
-    \begin{subfigure}[b]{0.49\textwidth}
+    \begin{subfigure}[b]{0.45\textwidth}
         \includegraphics[width=\textwidth]{figs/cv1_scatter.png}
@@
     \end{subfigure}
-    \begin{subfigure}[b]{0.49\textwidth}
+    \begin{subfigure}[b]{0.45\textwidth}
         \includegraphics[width=\textwidth]{figs/cv2_scatter.png}
@@
 \centering
-\includegraphics[trim={0 40mm 0 40mm}, width = 0.55\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
+\includegraphics[trim={0 30mm 0 40mm}, width = 0.55\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
 \caption{
-  Left: Observed polygon data for Colombia.
-Grey indicates incidence of zero.
-Right: Out-of-sample predictions for the random cross-validation, machine learning model only model.
-For each cross-validation fold, predictions are made for the held out data.
-These predictions are all plotted which together covers the entire study area.
+  Left: Observed data for Colombia (grey for zero incidence).
+Right: Out-of-sample predictions for the random cross-validation, machine learning only model.
+For each cross-validation fold, predictions are made for the held out data which are then combined to make a single surface.
 }
@@
 \begin{table}
-\caption{Machine learning model results and means of fitted parameters (i.e.
-model weights) across $k$ cross-validation folds of the model which only contained machine learning model predictions as covariates.
-}
+\caption{Machine learning model results and means of fitted parameters (i.e.\thinspace model weights) across cross-validation folds of the machine learning predictions only model. }
 \centering
@@
 
-Due to the low power of typical aggregated incidence datasets, previous analysis using disaggregation regression had to use a small number of covariates \citep{sturrock2014fine}.
-However, as models such as Random Forest and elastic net are robust in the presence of high dimensional data, future work could include many more covariates, potentially increasing predictive performance further.
+Due to the low power of typical aggregated incidence datasets, previous analyses using disaggregation regression used a small number of covariates \citep{sturrock2014fine}.
+However, as models such as Random Forest and elastic net can robustly handle high dimensional data, future work could include many more covariates, potentially increasing predictive performance.
 
@@
 Another area of potential improvement is varying the data used to train the base level learners.
-Here we only used data from the region of interest (South America for the Colombia model and Madagascar for the Madagascar model).
+Here we only used data from the region of interest.
 However, the global dataset is much larger than these subsets.

### 2019-01-21 tim-lucas-wsc.tex
Spell check.

@@
 Routine surveillance data of malaria case counts, often aggregated over administrative regions, is becoming more widely available and can better measure low malaria risk than prevalence surveys.
-However, aggregation of case counts over large, heterogenous areas means that these data are often underpowered for learning relationships between the environment and malaria risk.
+However, aggregation of case counts over large, heterogeneous areas means that these data are often underpowered for learning relationships between the environment and malaria risk.
 A model that combines point surveys and aggregated surveillance data could have the benefits of both but must be able to account for the fact that these two data types are different malariometric units.
@@
 Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more reliable and more widely available \citep{sturrock2016mapping} and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
-However, the aggregation of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogenous areas, because it is unclear where within the polygon, and in which environments, the cases occurred.
+However, the aggregation of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogeneous areas, because it is unclear where within the polygon, and in which environments, the cases occurred.
 This data is therefore often under-powered for fitting flexible, non-linear models as is required for accurate malaria maps \citep{bhatt2017improved, bhatt2015effect}. 
@@
 We used two data sources that reflect \emph{Plasmodium falciparum} malaria transmission; point-prevalence surveys and polygon-level, aggregated incidence data. 
-We selected Colombia and Madagascar as case examples as they both have fairly complete, publically available, surveillance data at a finer geographical resolution than admin 1.
+We selected Colombia and Madagascar as case examples as they both have fairly complete, publicly available, surveillance data at a finer geographical resolution than admin 1.
 The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling}.  % ceck todo
@@
 where $i_a$ is the estimated polygon incidence rate and $\mathrm{pop_a}$ is the observed polygon population-at-risk. 
-This polygon-level likelihood is linked to the pixel level prevealence 
+This polygon-level likelihood is linked to the pixel level prevalence 
 $$i_a = \frac{ \sum(i_b \mathrm{pop}_b)}{\sum  \mathrm{pop}_b} $$

### 2019-07-30 tim-lucas-model-ensembles.tex
Switch to non wsc ms.
Add newer bigger table comparing RMSE and beta.

--- 2019-01-21 tim-lucas-wsc.tex
@@
 \geometry{verbose,tmargin=30mm,bmargin=25mm,lmargin=25mm,rmargin=25mm}
-\input{figures.tex}
+%\input{figures.tex}
 \pagestyle{empty}
 \begin{document}
-\templatefigures{}
+%\templatefigures{}
 
@@
 This process of stacking models has proven effective in many realms however typical stacking uses a single dataset on a consistent scale \citep{sill2009feature, bhatt2017improved}. 
+% more about stacking
+% good for bias variance
+% SDM examples. any others?
+
+Diversity of predictions is good.
+Diversity of models
+Diversity of data. global local.
+
+Different scales. is it transfer learning?
+
 Here we propose training the level zero machine learning models on point-level, binomial prevalence data and stacking these models with a polygon-level, Poisson incidence model.
@@
 The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling}.  % ceck todo
-For Colombia we used all points from South America (n = 522) while for Madagascar we used only Malagasy data (n = 1505).
-We chose these geographic regions based on a trade-off between wanting a large sample size but wanting data from geographically similar areas.
 The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
+
 The polygon incidence data were collected from government reports and standardised using methods defined in \cite{cibulskis2011worldwide}.
@@
 
-For each country we fitted five models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
+For each country we fitted 5 models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
+These models were fitted to both the full malaria prevalence dataset and to more regional subsets of the data.
+For Colombia we used all points from South America (n = 522) while for Madagascar we used only Malagasy data (n = 1505).
+We chose these geographic regions based on a trade-off between wanting a large sample size but wanting data from geographically similar areas.
 Our response variable was prevalence and we weighted the data by sample size (i.e.\thinspace the number of people tested for malaria in each survey.
 For each model we ran  five-fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
-%The parameters tried can be seen in Table S1.
+The parameters tried can be seen in Table S1.
 Predictions from these models were then made across Colombia and Madagascar respectively.
@@
 
+This is not true stacking because we don't constrain betas to be positive.
+preliminary analyses indicated that this didn't improve performance.
+
 Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$ and $\sigma_v$. 
@@
 
-We compared the performance of the models with three sets of covariates, $X$.
-Firstly, we used the environmental and anthropogenic covariates, centered and standardised.
-Secondly, we used the predictions from the machine learning models.
-Finally we combined these two sets of covariates.
+We compared the performance of the models with four sets of covariates, $X$.
+Firstly, we used the environmental and anthropogenic covariates, centered and standardised as a baseline model.
+Secondly, we used the predictions from the machine learning models trained on regional data.
+Thirdly, we used predictions from the machine learning models trained on the global data.
+Finally we combined predictions from the machine learning models trained on regional data and global data.
+Preliminary analyses showed that combining environmental covariates with machine learning predictions was not effective so we did not pursue this further.
 
-To compare the three models we used two cross-validation schemes. 
+To compare the four models we used two cross-validation schemes. 
 In the first, polygon incidence data was randomly split into six cross-validation folds.
@@
 Our primary performance metric was correlation between observed and predicted data.
+We also examined the calibration of models by calculating the proportion of held out data that were within their 80\% credible intervals.
 
@@
 
+% to do table of coverage.
+
 The model using only machine learning predictions as covariates was the best performing model in both countries and both cross-validation schemes (Table~\ref{t:results}).
@@
 Predictive performance of machine learning models was similar, with Random Forest performing best in Madagascar and neural networks, Random Forests and elastic net performing equally well in Colombia (Table~\ref{t:mlresults}).
-The means (across folds) of the  regression coefficients (i.e.\thinspace the weights of the machine learning models in the level zero model) from the polygon-level models that used only predictions from machine learning models as covariates can also be seen in Table~\ref{t:mlresults}.
+The regression coefficients (i.e.\thinspace the weights of the machine learning models in the level zero model) from the polygon-level models that used only predictions from machine learning models as covariates can also be seen in Table~\ref{t:mlresults}.
 The estimated regression parameters are similar between the random and spatial cross-validation schemes.
@@
 Also of note is that some models were estimated to have a negative relationship with incidence (conditional on the inclusion of predictions from other machine learning models).
+% Compare plots of global vs local. eg amazon predicted low by global.
+% Look at \betas for global and local.
 
@@
 
+% copy Sam's plot. predictions from ml and stacked.
+
 \begin{table}
@@
 ppr            & 0.110      & -0.233         & -0.204         & 0.059         &  0.210          &  0.166          \\
+\end{tabular}
+\label{t:mlresults}
+\end{table}
+
+% reconsider. scatter plot?
+% rmse ~ beta. facet as country, colour as global vs local. do on full model not CV. Need to include global vs local.
+%
+% Or table:
+% Country Model       RMSE beta_local beta_global beta_local+global
+% MDG     RF local    12   0.1        -           0.2    
+% MDG     RF global   14   -          0.4         0.3
+% COL     nnet local  200  0.01       -           0.0001
+
+\begin{table}
+\caption{Machine learning model results and means of fitted parameters (i.e.\thinspace model weights) across cross-validation folds of the machine learning predictions only model. }
+\centering
+\small
+\begin{tabular}{ll|rrrr}
+     Country          & Model &      RMSE & $\beta_l $ &   $\beta_g$   &  $\beta_{g+l}$                                 \\ \hline
+Madagascar & RF_l & 0.100 & 0.337 &-&\\
+Madagascar & RF_g &  & - &&\\
+Madagascar & GBM_l & 0.109 & 0.450 & - &\\
+Madagascar & GBM_g &  & - &&\\
+Madagascar & enet_l & 0.100 & 0.326 &-&\\
+Madagascar & enet_g &  & - &&\\
+Madagascar & nnet_l & 0.113 & 0.031 &-&\\
+Madagascar & nnet_g &  & - &&\\
+Madagascar & ppr_l & 0.110 & -0.233 &-&\\
+Madagascar & ppr_g &  &  &&\vspace{0.3cm}\\ 
+Colombia & RF_l & 0.100 & 0.337 &-&\\
+Colombia & RF_g &  & - &&\\
+Colombia & GBM_l & 0.109 & 0.450 & - &\\
+Colombia & GBM_g &  & - &&\\
+Colombia & enet_l & 0.100 & 0.326 &-&\\
+Colombia & enet_g &  & - &&\\
+Colombia & nnet_l & 0.113 & 0.031 &-&\\
+Colombia & nnet_g &  & - &&\\
+Colombia & ppr_l & 0.110 & -0.233 &-&\\
+Colombia & ppr_g &  &  &&\vspace{0.3cm}\\ 
+Senegal & RF_l & 0.100 & 0.337 &-&\\
+Senegal & RF_g &  & - &&\\
+Senegal & GBM_l & 0.109 & 0.450 & - &\\
+Senegal & GBM_g &  & - &&\\
+Senegal & enet_l & 0.100 & 0.326 &-&\\
+Senegal & enet_g &  & - &&\\
+Senegal & nnet_l & 0.113 & 0.031 &-&\\
+Senegal & nnet_g &  & - &&\\
+Senegal & ppr_l & 0.110 & -0.233 &-&\\
+Senegal & ppr_g &  &  &&\vspace{0.3cm}\\ 
+Indonesia & RF_l & 0.100 & 0.337 &-&\\
+Indonesia & RF_g &  & - &&\\
+Indonesia & GBM_l & 0.109 & 0.450 & - &\\
+Indonesia & GBM_g &  & - &&\\
+Indonesia & enet_l & 0.100 & 0.326 &-&\\
+Indonesia & enet_g &  & - &&\\
+Indonesia & nnet_l & 0.113 & 0.031 &-&\\
+Indonesia & nnet_g &  & - &&\\
+Indonesia & ppr_l & 0.110 & -0.233 &-&\\
+Indonesia & ppr_g &  &  &&\\
 \end{tabular}

### 2019-07-30 tim-lucas-model-ensembles.tex
Remove old RMSE vs beta table.
Some intro edits.

@@
 Susan Rumisha\textsuperscript{1}; 
-Katherine E. Battle\textsuperscript{1}; Rosalind E. Howes\textsuperscript{1}; 
-Chantal Hendriks\textsuperscript{1}; Andre Python\textsuperscript{1}; Penny Hancock\textsuperscript{1}; 
+Katherine E. Battle\textsuperscript{1}; Rosalind E. Howes\textsuperscript{1}; Andre Python\textsuperscript{1}; Penny Hancock\textsuperscript{1}; Punam Someone\textsuperscript{1};
 Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel~J.~Weiss\textsuperscript{1}.}\\
@@
 
-Maps of infection risk are a vital tool for the elimination of malaria.
-Routine surveillance data of malaria case counts, often aggregated over administrative regions, is becoming more widely available and can better measure low malaria risk than prevalence surveys.
-However, aggregation of case counts over large, heterogeneous areas means that these data are often underpowered for learning relationships between the environment and malaria risk.
+Maps of infection rate are one of the core tools needed for the elimination of malaria.
+Reliable routine surveillance data of malaria incidence is becoming more widely available and can better measure low malaria risk than prevalence surveys.
+However, these data are often aggregated over large, heterogeneous areas which means that they are often underpowered for learning complex relationships between the environment and malaria risk.
+In contrast, point prevalence surveys are directly linked to loval environmental conditions but are not common in many areas of the world that are working towards elimination.
 A model that combines point surveys and aggregated surveillance data could have the benefits of both but must be able to account for the fact that these two data types are different malariometric units.
 Here, we train multiple machine learning models on point surveys and then combine the predictions from these with a geostatistical disaggregation model that uses routine surveillance data.
-We find that, in tests using data from Colombia and Madagascar, using a disaggregation regression model to combine predictions from machine learning models trained on point surveys improves model accuracy relative to using the environmental covariates directly.
+We find that, in four national case studies, using a disaggregation regression model to combine predictions from machine learning models trained on point surveys improves model accuracy relative to using the environmental covariates directly.
 
@@
 This process of stacking models has proven effective in many realms however typical stacking uses a single dataset on a consistent scale \citep{sill2009feature, bhatt2017improved}. 
+Stacking improves predictions by controlling bias and variance; as long as suitably diverse models are averaged, they will have different biases while high variance in models should be averaged out even in relatively similar models.
+This understanding of how stacking improves model performance indicates that diversity in models is important for stacking to be effective.
+Diversity in models is typically created in two ways: by using diverse training datasets (as in Random Forests for example) and by using functionally different models (for example by averaging tree based models and neural networks).
+
 % more about stacking
@@
 \small
-\begin{tabular}{l|rrr|rrr}
-               &&      Madagascar &&      &                                     Colombia&    \\
-Model          & ML RMSE & Random CV $\bar{\beta}_i$ & Spatial CV $\bar{\beta}_i$  & ML RMSE & Random CV $\bar{\beta}_i$  & Spatial CV $\bar{\beta}_i$  \\
-\hline
-nnet           & 0.113      &  0.031         &  0.025         & \textbf{0.058}& -0.250          & -0.246          \\
-RF   &  \textbf{0.100}      &  0.337         &  0.350         & \textbf{0.058}&  0.782          &  0.742          \\
-gbm            & 0.109      &  \textbf{0.450}&  \textbf{0.402}& 0.066         &  \textbf{0.835} &  \textbf{0.775} \\
-enet           & 0.116      &  0.326         &  0.307         & \textbf{0.058}& -0.563          & -0.369          \\
-ppr            & 0.110      & -0.233         & -0.204         & 0.059         &  0.210          &  0.166          \\
-\end{tabular}
-\label{t:mlresults}
-\end{table}
-
-% reconsider. scatter plot?
-% rmse ~ beta. facet as country, colour as global vs local. do on full model not CV. Need to include global vs local.
-%
-% Or table:
-% Country Model       RMSE beta_local beta_global beta_local+global
-% MDG     RF local    12   0.1        -           0.2    
-% MDG     RF global   14   -          0.4         0.3
-% COL     nnet local  200  0.01       -           0.0001
-
-\begin{table}
-\caption{Machine learning model results and means of fitted parameters (i.e.\thinspace model weights) across cross-validation folds of the machine learning predictions only model. }
-\centering
-\small
 \begin{tabular}{ll|rrrr}
      Country          & Model &      RMSE & $\beta_l $ &   $\beta_g$   &  $\beta_{g+l}$                                 \\ \hline
-Madagascar & RF_l & 0.100 & 0.337 &-&\\
-Madagascar & RF_g &  & - &&\\
-Madagascar & GBM_l & 0.109 & 0.450 & - &\\
-Madagascar & GBM_g &  & - &&\\
-Madagascar & enet_l & 0.100 & 0.326 &-&\\
-Madagascar & enet_g &  & - &&\\
-Madagascar & nnet_l & 0.113 & 0.031 &-&\\
-Madagascar & nnet_g &  & - &&\\
-Madagascar & ppr_l & 0.110 & -0.233 &-&\\
-Madagascar & ppr_g &  &  &&\vspace{0.3cm}\\ 
-Colombia & RF_l & 0.100 & 0.337 &-&\\
-Colombia & RF_g &  & - &&\\
-Colombia & GBM_l & 0.109 & 0.450 & - &\\
-Colombia & GBM_g &  & - &&\\
-Colombia & enet_l & 0.100 & 0.326 &-&\\
-Colombia & enet_g &  & - &&\\
-Colombia & nnet_l & 0.113 & 0.031 &-&\\
-Colombia & nnet_g &  & - &&\\
-Colombia & ppr_l & 0.110 & -0.233 &-&\\
-Colombia & ppr_g &  &  &&\vspace{0.3cm}\\ 
-Senegal & RF_l & 0.100 & 0.337 &-&\\
-Senegal & RF_g &  & - &&\\
-Senegal & GBM_l & 0.109 & 0.450 & - &\\
-Senegal & GBM_g &  & - &&\\
-Senegal & enet_l & 0.100 & 0.326 &-&\\
-Senegal & enet_g &  & - &&\\
-Senegal & nnet_l & 0.113 & 0.031 &-&\\
-Senegal & nnet_g &  & - &&\\
-Senegal & ppr_l & 0.110 & -0.233 &-&\\
-Senegal & ppr_g &  &  &&\vspace{0.3cm}\\ 
-Indonesia & RF_l & 0.100 & 0.337 &-&\\
-Indonesia & RF_g &  & - &&\\
-Indonesia & GBM_l & 0.109 & 0.450 & - &\\
-Indonesia & GBM_g &  & - &&\\
-Indonesia & enet_l & 0.100 & 0.326 &-&\\
-Indonesia & enet_g &  & - &&\\
-Indonesia & nnet_l & 0.113 & 0.031 &-&\\
-Indonesia & nnet_g &  & - &&\\
-Indonesia & ppr_l & 0.110 & -0.233 &-&\\
-Indonesia & ppr_g &  &  &&\\
+Madagascar & RF\textsubscript{l} & 0.100 & 0.337 &-&\\
+Madagascar & RF\textsubscript{g} &  & - &&\\
+Madagascar & GBM\textsubscript{l} & 0.109 & 0.450 & - &\\
+Madagascar & GBM\textsubscript{g} &  & - &&\\
+Madagascar & enet\textsubscript{l} & 0.100 & 0.326 &-&\\
+Madagascar & enet\textsubscript{g} &  & - &&\\
+Madagascar & nnet\textsubscript{l} & 0.113 & 0.031 &-&\\
+Madagascar & nnet\textsubscript{g} &  & - &&\\
+Madagascar & ppr\textsubscript{l} & 0.110 & -0.233 &-&\\
+Madagascar & ppr\textsubscript{g} &  &  &&\vspace{0.3cm}\\ 
+Colombia & RF\textsubscript{l} & 0.100 & 0.337 &-&\\
+Colombia & RF\textsubscript{g} &  & - &&\\
+Colombia & GBM\textsubscript{l} & 0.109 & 0.450 & - &\\
+Colombia & GBM\textsubscript{g} &  & - &&\\
+Colombia & enet\textsubscript{l} & 0.100 & 0.326 &-&\\
+Colombia & enet\textsubscript{g} &  & - &&\\
+Colombia & nnet\textsubscript{l} & 0.113 & 0.031 &-&\\
+Colombia & nnet\textsubscript{g} &  & - &&\\
+Colombia & ppr\textsubscript{l} & 0.110 & -0.233 &-&\\
+Colombia & ppr\textsubscript{g} &  &  &&\vspace{0.3cm}\\ 
+Senegal & RF\textsubscript{l} & 0.100 & 0.337 &-&\\
+Senegal & RF\textsubscript{g} &  & - &&\\
+Senegal & GBM\textsubscript{l} & 0.109 & 0.450 & - &\\
+Senegal & GBM\textsubscript{g} &  & - &&\\
+Senegal & enet\textsubscript{l} & 0.100 & 0.326 &-&\\
+Senegal & enet\textsubscript{g} &  & - &&\\
+Senegal & nnet\textsubscript{l} & 0.113 & 0.031 &-&\\
+Senegal & nnet\textsubscript{g} &  & - &&\\
+Senegal & ppr\textsubscript{l} & 0.110 & -0.233 &-&\\
+Senegal & ppr\textsubscript{g} &  &  &&\vspace{0.3cm}\\ 
+Indonesia & RF\textsubscript{l} & 0.100 & 0.337 &-&\\
+Indonesia & RF\textsubscript{g} &  & - &&\\
+Indonesia & GBM\textsubscript{l} & 0.109 & 0.450 & - &\\
+Indonesia & GBM\textsubscript{g} &  & - &&\\
+Indonesia & enet\textsubscript{l} & 0.100 & 0.326 &-&\\
+Indonesia & enet\textsubscript{g} &  & - &&\\
+Indonesia & nnet\textsubscript{l} & 0.113 & 0.031 &-&\\
+Indonesia & nnet\textsubscript{g} &  & - &&\\
+Indonesia & ppr\textsubscript{l} & 0.110 & -0.233 &-&\\
+Indonesia & ppr\textsubscript{g} &  &  &&\\
 \end{tabular}

### 2019-09-12 tim-lucas-model-ensembles.tex
Add new results after fixing ml predictions at 0. Still weird though.

@@
 
+\begin{table}[h!]
+\caption{Pearson correlations between observed and predicted values. }
+\centering
+\begin{tabular}{llrrrr}
+Cross-validation scheme & Country &  Covariates &  ML\textsubscript{l} &  ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
+\hline 
+ Random &  Colombia & 0.58& 0.51 &   & 0.53 \\
+ Random &  Madagascar &  0.70 &  \textbf{0.73} &   0.70 & 0.71 \\
+ Random &  Senegal &  0.58 &  0.11 &  0.30 & 0.08 \\
+ Random &  Indonesia &   &   &   &  \\
+ Spatial &  Colombia &  0.12 &  0.09 &   & 0.10\\
+ Spatial &  Madagascar &  0.22 &  \textbf{0.66} &  0.63 & 0.60\\
+ Spatial &  Senegal &  0.63 &  0.27 &  0.36 & 0.35 \\
+ Spatial &  Indonesia &   &   &   &  \\
+\end{tabular}
+\label{t:results}
+\end{table}
+
 % to do table of coverage.
@@
 Madagascar & RF\textsubscript{l} & 0.100 & 0.337 &-&\\
-Madagascar & RF\textsubscript{g} &  & - &&\\
+Madagascar & RF\textsubscript{g} &  0.169 & - &&\\
 Madagascar & GBM\textsubscript{l} & 0.109 & 0.450 & - &\\
-Madagascar & GBM\textsubscript{g} &  & - &&\\
+Madagascar & GBM\textsubscript{g} &  0.178& - &&\\
 Madagascar & enet\textsubscript{l} & 0.100 & 0.326 &-&\\
-Madagascar & enet\textsubscript{g} &  & - &&\\
+Madagascar & enet\textsubscript{g} & 0.233 & - &&\\
 Madagascar & nnet\textsubscript{l} & 0.113 & 0.031 &-&\\
-Madagascar & nnet\textsubscript{g} &  & - &&\\
+Madagascar & nnet\textsubscript{g} &  0.220 & - &&\\
 Madagascar & ppr\textsubscript{l} & 0.110 & -0.233 &-&\\
-Madagascar & ppr\textsubscript{g} &  &  &&\vspace{0.3cm}\\ 
+Madagascar & ppr\textsubscript{g} & 0.205 &  &&\vspace{0.3cm}\\ 
 Colombia & RF\textsubscript{l} & 0.100 & 0.337 &-&\\

### 2019-10-09 tim-lucas-model-ensembles.tex
Full main results table

@@
 
+%\begin{table}[h!]
+%\caption{Pearson correlations between observed and predicted values. }
+%\centering
+%\begin{tabular}{llrrr}
+%Cross-validation scheme & Country &  Covariates &  ML &  Covs + ML \\
+%\hline 
+% Random &  Colombia &  0.45 &  \textbf{0.55} &  0.54 \\
+% Random &  Madagascar &  0.70 &  \textbf{0.76} &  0.75 \\
+% Spatial &  Colombia &  0.05 &  \textbf{0.18} &  0.10 \\
+% Spatial &  Madagascar &  0.22 &  \textbf{0.63} &  0.61 
+%\end{tabular}
+%\label{t:results}
+%\end{table}
+
 \begin{table}[h!]
@@
 \centering
-\begin{tabular}{llrrr}
-Cross-validation scheme & Country &  Covariates &  ML &  Covs + ML \\
+\begin{tabular}{llrrrr}
+Cross-validation scheme & Country &  Enviro &  ML\textsubscript{l} &  Enviro + ML\textsubscript{l} \\
 \hline 
- Random &  Colombia &  0.45 &  \textbf{0.55} &  0.54 \\
- Random &  Madagascar &  0.70 &  \textbf{0.76} &  0.75 \\
- Spatial &  Colombia &  0.05 &  \textbf{0.18} &  0.10 \\
- Spatial &  Madagascar &  0.22 &  \textbf{0.63} &  0.61 
+ Random &  Colombia & 0.59 & 0.59 & \textbf{0.61} \\
+ Random &  Madagascar &  0.70 &  \textbf{0.73} & 0.69 \\
+ Random &  Senegal &  \textbf{0.58} &  0.51 & 0.57 \\
+ Random &  Indonesia &  0.52 &  0.48 &  \textbf{0.59} \\
+ Spatial &  Colombia &  0.12 &  \textbf{0.33} &  0.25\\
+ Spatial &  Madagascar &  0.22 &  \textbf{0.66} & 0.18\\
+ Spatial &  Senegal &  \textbf{0.63} &  0.27 &  \textbf{0.63} \\
+ Spatial &  Indonesia &  0.45 &  0.44 &  \textbf{0.51} \\
 \end{tabular}
@@
 \begin{tabular}{llrrrr}
-Cross-validation scheme & Country &  Covariates &  ML\textsubscript{l} &  ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
+Cross-validation scheme & Country &  Covariates &   ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
 \hline 
- Random &  Colombia & 0.58& 0.51 &   & 0.53 \\
- Random &  Madagascar &  0.70 &  \textbf{0.73} &   0.70 & 0.71 \\
- Random &  Senegal &  0.58 &  0.11 &  0.30 & 0.08 \\
- Random &  Indonesia &   &   &   &  \\
- Spatial &  Colombia &  0.12 &  0.09 &   & 0.10\\
- Spatial &  Madagascar &  0.22 &  \textbf{0.66} &  0.63 & 0.60\\
- Spatial &  Senegal &  0.63 &  0.27 &  0.36 & 0.35 \\
- Spatial &  Indonesia &   &   &   &  \\
+ Random &  Colombia & \textbf{0.59} &0.58 & 0.56 \\
+ Random &  Madagascar &  0.70 &  0.70 & \textbf{0.71} \\
+ Random &  Senegal &  \textbf{0.58} &  0.30 & 0.08 \\
+ Random &  Indonesia &  \textbf{0.52} & 0.46 & 0.32 \\
+ Spatial &  Colombia &  0.12 &  0.18 & \textbf{0.34}\\
+ Spatial &  Madagascar &  0.22 &  \textbf{0.63} & 0.60\\
+ Spatial &  Senegal &  \textbf{0.63} & 0.36 & 0.35 \\
+ Spatial &  Indonesia &  \textbf{0.45} & 0.41 & \textbf{0.45} \\
 \end{tabular}
@@
 \end{table}
+
+%\begin{table}[h!]
+%\caption{Pearson correlations between observed and predicted values. }
+%\centering
+%\begin{tabular}{llrrrr}
+%Cross-validation scheme & Country &  Covariates &  ML\textsubscript{l} &  ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
+%\hline 
+% Random &  Colombia & \textbf{0.59} & \textbf{0.59} &  0.58 & 0.56 \\
+% Random &  Madagascar &  0.70 &  \textbf{0.73} &   0.70 & 0.71 \\
+% Random &  Senegal &  \textbf{0.58} &  0.51 &  0.30 & 0.08 \\
+% Random &  Indonesia &  \textbf{0.52} &  0.48 &  0.46 & 0.32 \\
+% Spatial &  Colombia &  0.12 &  0.33 &  0.18 & \textbf{0.34}\\
+% Spatial &  Madagascar &  0.22 &  \textbf{0.66} &  0.63 & 0.60\\
+% Spatial &  Senegal &  \textbf{0.63} &  0.27 &  0.36 & 0.35 \\
+% Spatial &  Indonesia &  \textbf{0.45} &  0.44 &  0.41 & \textbf{0.45} \\
+%\end{tabular}
+%\label{t:results}
+%\end{table}
 

### 2019-10-09 tim-lucas-model-ensembles.tex
Switch tables to sort by country then CV.

@@
  Random &  Colombia & 0.59 & 0.59 & \textbf{0.61} \\
+ Spatial &  Colombia &  0.12 &  \textbf{0.33} &  0.25\\
  Random &  Madagascar &  0.70 &  \textbf{0.73} & 0.69 \\
+ Spatial &  Madagascar &  0.22 &  \textbf{0.66} & 0.18\\
  Random &  Senegal &  \textbf{0.58} &  0.51 & 0.57 \\
+ Spatial &  Senegal &  \textbf{0.63} &  0.27 &  \textbf{0.63} \\
  Random &  Indonesia &  0.52 &  0.48 &  \textbf{0.59} \\
- Spatial &  Colombia &  0.12 &  \textbf{0.33} &  0.25\\
- Spatial &  Madagascar &  0.22 &  \textbf{0.66} & 0.18\\
- Spatial &  Senegal &  \textbf{0.63} &  0.27 &  \textbf{0.63} \\
  Spatial &  Indonesia &  0.45 &  0.44 &  \textbf{0.51} \\
@@
  Random &  Colombia & \textbf{0.59} &0.58 & 0.56 \\
+ Spatial &  Colombia &  0.12 &  0.18 & \textbf{0.34}\\
  Random &  Madagascar &  0.70 &  0.70 & \textbf{0.71} \\
+ Spatial &  Madagascar &  0.22 &  \textbf{0.63} & 0.60\\
  Random &  Senegal &  \textbf{0.58} &  0.30 & 0.08 \\
+ Spatial &  Senegal &  \textbf{0.63} & 0.36 & 0.35 \\
  Random &  Indonesia &  \textbf{0.52} & 0.46 & 0.32 \\
- Spatial &  Colombia &  0.12 &  0.18 & \textbf{0.34}\\
- Spatial &  Madagascar &  0.22 &  \textbf{0.63} & 0.60\\
- Spatial &  Senegal &  \textbf{0.63} & 0.36 & 0.35 \\
  Spatial &  Indonesia &  \textbf{0.45} & 0.41 & \textbf{0.45} \\

### 2019-10-09 tim-lucas-model-ensembles.tex
Continue messing with table.

@@
 \small
-\begin{tabular}{ll|rrrr}
-     Country          & Model &      RMSE & $\beta_l $ &   $\beta_g$   &  $\beta_{g+l}$                                 \\ \hline
-Madagascar & RF\textsubscript{l} & 0.100 & 0.337 &-&\\
-Madagascar & RF\textsubscript{g} &  0.169 & - &&\\
-Madagascar & GBM\textsubscript{l} & 0.109 & 0.450 & - &\\
-Madagascar & GBM\textsubscript{g} &  0.178& - &&\\
-Madagascar & enet\textsubscript{l} & 0.100 & 0.326 &-&\\
-Madagascar & enet\textsubscript{g} & 0.233 & - &&\\
-Madagascar & nnet\textsubscript{l} & 0.113 & 0.031 &-&\\
-Madagascar & nnet\textsubscript{g} &  0.220 & - &&\\
-Madagascar & ppr\textsubscript{l} & 0.110 & -0.233 &-&\\
-Madagascar & ppr\textsubscript{g} & 0.205 &  &&\vspace{0.3cm}\\ 
-Colombia & RF\textsubscript{l} & 0.100 & 0.337 &-&\\
-Colombia & RF\textsubscript{g} &  & - &&\\
-Colombia & GBM\textsubscript{l} & 0.109 & 0.450 & - &\\
-Colombia & GBM\textsubscript{g} &  & - &&\\
-Colombia & enet\textsubscript{l} & 0.100 & 0.326 &-&\\
-Colombia & enet\textsubscript{g} &  & - &&\\
-Colombia & nnet\textsubscript{l} & 0.113 & 0.031 &-&\\
-Colombia & nnet\textsubscript{g} &  & - &&\\
-Colombia & ppr\textsubscript{l} & 0.110 & -0.233 &-&\\
-Colombia & ppr\textsubscript{g} &  &  &&\vspace{0.3cm}\\ 
-Senegal & RF\textsubscript{l} & 0.100 & 0.337 &-&\\
-Senegal & RF\textsubscript{g} &  & - &&\\
-Senegal & GBM\textsubscript{l} & 0.109 & 0.450 & - &\\
-Senegal & GBM\textsubscript{g} &  & - &&\\
-Senegal & enet\textsubscript{l} & 0.100 & 0.326 &-&\\
-Senegal & enet\textsubscript{g} &  & - &&\\
-Senegal & nnet\textsubscript{l} & 0.113 & 0.031 &-&\\
-Senegal & nnet\textsubscript{g} &  & - &&\\
-Senegal & ppr\textsubscript{l} & 0.110 & -0.233 &-&\\
-Senegal & ppr\textsubscript{g} &  &  &&\vspace{0.3cm}\\ 
-Indonesia & RF\textsubscript{l} & 0.100 & 0.337 &-&\\
-Indonesia & RF\textsubscript{g} &  & - &&\\
-Indonesia & GBM\textsubscript{l} & 0.109 & 0.450 & - &\\
-Indonesia & GBM\textsubscript{g} &  & - &&\\
-Indonesia & enet\textsubscript{l} & 0.100 & 0.326 &-&\\
-Indonesia & enet\textsubscript{g} &  & - &&\\
-Indonesia & nnet\textsubscript{l} & 0.113 & 0.031 &-&\\
-Indonesia & nnet\textsubscript{g} &  & - &&\\
-Indonesia & ppr\textsubscript{l} & 0.110 & -0.233 &-&\\
-Indonesia & ppr\textsubscript{g} &  &  &&\\
+\begin{tabular}{ll|rr}
+     Country          & Model &      RMSE & $\beta$ \\ \hline
+Madagascar & RF\textsubscript{l} & 0.100 & 0.337 \\
+Madagascar & GBM\textsubscript{l} & 0.105 & 0.450 \\
+Madagascar & enet\textsubscript{l} & 0.116 & 0.326 \\
+Madagascar & nnet\textsubscript{l} & 0.113 & 0.031 \\
+Madagascar & ppr\textsubscript{l} & 0.109 & -0.233\vspace{0.3cm}\\ 
+Madagascar & RF\textsubscript{g} &  0.169 & - \\
+Madagascar & GBM\textsubscript{g} &  0.178& - \\
+Madagascar & enet\textsubscript{g} & 0.233 & - \\
+Madagascar & nnet\textsubscript{g} &  0.220 & - \\
+Madagascar & ppr\textsubscript{g} & 0.205 &   \vspace{0.3cm}\\
+Colombia & RF\textsubscript{l} & 0.100 & 0.337 \\
+Colombia & GBM\textsubscript{l} & 0.109 & 0.450  \\
+Colombia & enet\textsubscript{l} & 0.100 & 0.326 \\
+Colombia & nnet\textsubscript{l} & 0.113 & 0.031 \\
+Colombia & ppr\textsubscript{l} & 0.110 & -0.233\vspace{0.3cm}\\
+Colombia & RF\textsubscript{g} & 0.169 & -\\
+Colombia & GBM\textsubscript{g} & 0.178 & - \\
+Colombia & enet\textsubscript{g} & 0.233 & - \\
+Colombia & nnet\textsubscript{g} & 0.220 & - \\
+Colombia & ppr\textsubscript{g} & 0.205 & \vspace{0.3cm}\\ 
+Senegal & RF\textsubscript{l} & 0.100 & 0.337 \\
+Senegal & GBM\textsubscript{l} & 0.109 & 0.450 \\
+Senegal & enet\textsubscript{l} & 0.100 & 0.326 \\
+Senegal & nnet\textsubscript{l} & 0.113 & 0.031 \\
+Senegal & ppr\textsubscript{l} & 0.110 & -0.233\vspace{0.3cm}\\
+Senegal & RF\textsubscript{g} & 0.169 & \\
+Senegal & GBM\textsubscript{g} & 0.178 & - \\
+Senegal & enet\textsubscript{g} & 0.233 & - \\
+Senegal & nnet\textsubscript{g} & 0.220 & - \\
+Senegal & ppr\textsubscript{g} & 0.205 & \vspace{0.3cm} \\ 
+Indonesia & RF\textsubscript{l} & 0.100 & 0.337 \\
+Indonesia & GBM\textsubscript{l} & 0.109 & 0.450 \\
+Indonesia & enet\textsubscript{l} & 0.100 & 0.326\\
+Indonesia & nnet\textsubscript{l} & 0.113 & 0.031 \\
+Indonesia & ppr\textsubscript{l} & 0.110 & -0.233\vspace{0.3cm}\\
+Indonesia & RF\textsubscript{g} & 0.169 & \\
+Indonesia & GBM\textsubscript{g} & 0.178 & \\
+Indonesia & enet\textsubscript{g} &0.233  &\\
+Indonesia & nnet\textsubscript{g} &0.220  &\\
+Indonesia & ppr\textsubscript{g} & 0.205 &  \\
 \end{tabular}

### 2019-10-09 tim-lucas-model-ensembles.tex
More edits bringing in new figs.

@@
 
-Figure~\ref{f:scatter} shows the model performance under random and spatial cross-validation for both Madagascar and Colombia. 
+Figures~\ref{f:scatter1} and \ref{f:scatter2} show the model performance under random and spatial cross-validation for all countries. 
 The poor model performance in Colombia under spatial cross-validation indicates that the covariates alone cannot explain malaria incidence in this area. 
@@
 \begin{figure}
-\centering
-    \begin{subfigure}[b]{0.45\textwidth}
   \includegraphics[width=\textwidth]{figs/cv1_scatter.png}
-        \caption{Random cross-validation}
-        \label{fig:cv1}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[width=\textwidth]{figs/cv2_scatter.png}
-        \caption{Spatial cross-validation}
-        \label{fig:cv2}
-    \end{subfigure}
 \caption{
   Observed data against predictions for cross-validation hold-out samples on a square root transformed scale.
-  a) Six-fold random cross-validation.
-  b) Three-fold spatial cross-validation with folds indicated by colour.
 }
-\label{f:scatter}
+\label{f:scatter1}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{figs/cv2_scatter.png}
+\caption{
+  Observed data against predictions for cross-validation hold-out samples on a square root transformed scale.
+}
+\label{f:scatter1}
 \end{figure}
@@
 \small
-\begin{tabular}{ll|rr}
-     Country          & Model &      RMSE & $\beta$ \\ \hline
-Madagascar & RF\textsubscript{l} & 0.100 & 0.337 \\
-Madagascar & GBM\textsubscript{l} & 0.105 & 0.450 \\
-Madagascar & enet\textsubscript{l} & 0.116 & 0.326 \\
-Madagascar & nnet\textsubscript{l} & 0.113 & 0.031 \\
-Madagascar & ppr\textsubscript{l} & 0.109 & -0.233\vspace{0.3cm}\\ 
-Madagascar & RF\textsubscript{g} &  0.169 & - \\
-Madagascar & GBM\textsubscript{g} &  0.178& - \\
-Madagascar & enet\textsubscript{g} & 0.233 & - \\
-Madagascar & nnet\textsubscript{g} &  0.220 & - \\
-Madagascar & ppr\textsubscript{g} & 0.205 &   \vspace{0.3cm}\\
-Colombia & RF\textsubscript{l} & 0.100 & 0.337 \\
-Colombia & GBM\textsubscript{l} & 0.109 & 0.450  \\
-Colombia & enet\textsubscript{l} & 0.100 & 0.326 \\
-Colombia & nnet\textsubscript{l} & 0.113 & 0.031 \\
-Colombia & ppr\textsubscript{l} & 0.110 & -0.233\vspace{0.3cm}\\
-Colombia & RF\textsubscript{g} & 0.169 & -\\
-Colombia & GBM\textsubscript{g} & 0.178 & - \\
-Colombia & enet\textsubscript{g} & 0.233 & - \\
-Colombia & nnet\textsubscript{g} & 0.220 & - \\
-Colombia & ppr\textsubscript{g} & 0.205 & \vspace{0.3cm}\\ 
-Senegal & RF\textsubscript{l} & 0.100 & 0.337 \\
-Senegal & GBM\textsubscript{l} & 0.109 & 0.450 \\
-Senegal & enet\textsubscript{l} & 0.100 & 0.326 \\
-Senegal & nnet\textsubscript{l} & 0.113 & 0.031 \\
-Senegal & ppr\textsubscript{l} & 0.110 & -0.233\vspace{0.3cm}\\
-Senegal & RF\textsubscript{g} & 0.169 & \\
-Senegal & GBM\textsubscript{g} & 0.178 & - \\
-Senegal & enet\textsubscript{g} & 0.233 & - \\
-Senegal & nnet\textsubscript{g} & 0.220 & - \\
-Senegal & ppr\textsubscript{g} & 0.205 & \vspace{0.3cm} \\ 
-Indonesia & RF\textsubscript{l} & 0.100 & 0.337 \\
-Indonesia & GBM\textsubscript{l} & 0.109 & 0.450 \\
-Indonesia & enet\textsubscript{l} & 0.100 & 0.326\\
-Indonesia & nnet\textsubscript{l} & 0.113 & 0.031 \\
-Indonesia & ppr\textsubscript{l} & 0.110 & -0.233\vspace{0.3cm}\\
-Indonesia & RF\textsubscript{g} & 0.169 & \\
-Indonesia & GBM\textsubscript{g} & 0.178 & \\
-Indonesia & enet\textsubscript{g} &0.233  &\\
-Indonesia & nnet\textsubscript{g} &0.220  &\\
-Indonesia & ppr\textsubscript{g} & 0.205 &  \\
+\begin{tabular}{ll|rr|rr}
+     Country          & Model &      RMSE\textsubscript{l} & $\beta_l$ & RMSE\textsubscript{g} & $\beta_g$ \\ \hline
+Madagascar & RF & \textbf{0.100} & 0.538 &  \textbf{0.169} & \textbf{0.529}\\
+Madagascar & GBM & 0.105 & \textbf{0.570} & 0.178& 0.432 \\
+Madagascar & enet & 0.116 & 0.301 &0.233 & 0.262 \\
+Madagascar & nnet & 0.113 & 0.033 &0.220 & 0.364 \\
+Madagascar & ppr & 0.109 & 0.469 & 0.205 &  0.403\vspace{0.3cm}\\ 
+Colombia & RF & \textbf{0.068} & 0.505 &  \textbf{0.169} & \textbf{0.522}\\
+Colombia & GBM & 0.073 & \textbf{0.879} & 0.178& -0.231  \\
+Colombia & enet & 0.070 & 0.079 &0.233 & 0.080 \\
+Colombia & nnet & 0.070 & -0.008 &0.220 & 0.144 \\
+Colombia & ppr & 0.070 & 0.530 & 0.205 &  0.468\vspace{0.3cm}\\
+Senegal & RF & \textbf{0.092} & \textbf{0.339} & \textbf{0.169} & \textbf{0.425} \\
+Senegal & GBM & 0.099 & 0.261& 0.178& 0.408 \\
+Senegal & enet& 0.103 & 0.344  &0.233 & 0.205 \\
+Senegal & nnet & 0.099 & 0.254 &0.220 & 0.190 \\
+Senegal & ppr & 0.098 & 0.268& 0.205 &  0.126\vspace{0.3cm}\\
+Indonesia & RF& \textbf{0.081} & 0.447 & \textbf{0.169} & 0.178\\
+Indonesia & GBM & 0.085 & 0.357 & 0.178& 0.289 \\
+Indonesia & enet & 0.091 & 0.303 &0.233 & \textbf{0.526} \\
+Indonesia & nnet & 0.089 & \textbf{0.506} &0.220 & 0.316 \\
+Indonesia & ppr & 0.089 & 0.364 & 0.205 &  0.089\vspace{0.3cm}\\
+
 \end{tabular}

### 2019-10-10 tim-lucas-model-ensembles.tex
Redo intro.

@@
 \end{center}
-
-\begin{center}
-{\bf Abstract}
-\end{center}
-
-\setlength{\parindent}{0pt}
+\section{Abstract}
 
@@
 Maps of infection rate are one of the core tools needed for the elimination of malaria.
-Reliable routine surveillance data of malaria incidence is becoming more widely available and can better measure low malaria risk than prevalence surveys.
+Reliable routine surveillance data of malaria incidence is becoming more widely available and can better measure low malaria risk than prevalence point-surveys.
 However, these data are often aggregated over large, heterogeneous areas which means that they are often underpowered for learning complex relationships between the environment and malaria risk.
-In contrast, point prevalence surveys are directly linked to loval environmental conditions but are not common in many areas of the world that are working towards elimination.
+In contrast, prevalence point-surveys are directly linked to loval environmental conditions but are not common in many areas of the world.
 A model that combines point surveys and aggregated surveillance data could have the benefits of both but must be able to account for the fact that these two data types are different malariometric units.
-Here, we train multiple machine learning models on point surveys and then combine the predictions from these with a geostatistical disaggregation model that uses routine surveillance data.
-We find that, in four national case studies, using a disaggregation regression model to combine predictions from machine learning models trained on point surveys improves model accuracy relative to using the environmental covariates directly.
+Here, we train multiple machine learning models on prevalence point-surveys and then combine the predictions from these with a geostatistical disaggregation model that uses routine surveillance data.
+We find that, in four national case studies, using a disaggregation regression model to combine predictions from machine learning models trained on point surveys improves model accuracy relative to a baseline disaggregation regression model that uses the environmental covariates only.
 
@@
 
-\setlength{\parindent}{0pt}
-
-{\bf 1. Introduction}
+\section{Introduction}
 
@@
 
-High-resolution maps of malaria risk are vital for elimination but mapping malaria in low burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{gething2011new, bhatt2017improved, gething2012long, bhatt2015effect} is often not effective  because, firstly, so few individuals are infected that most surveys will detect zero cases, and secondly, because of the lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}.
+High-resolution maps of malaria risk are vital for control and elimination \citep{weiss2019mapping, battle2019mapping}.
+However, mapping malaria in lower burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{weiss2019mapping, bhatt2017improved, battle2019mapping, bhatt2015effect} is often not effective for two reasons.
+Firstly, so few individuals are infected that most surveys will detect zero cases \citep{sturrock2016mapping}.
+Secondly, there is a lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}. 
 Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more reliable and more widely available \citep{sturrock2016mapping} and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
@@
 This data is therefore often under-powered for fitting flexible, non-linear models as is required for accurate malaria maps \citep{bhatt2017improved, bhatt2015effect}. 
-A model that combines point surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.
+A model that combines prevalence point-surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.
 
-One approach for combining these data is to use prevalence point-surveys to train a suite of machine learning models, and then use predictions from these models as covariates in a model trained on polygon-level incidence data.
-This process of stacking models has proven effective in many realms however typical stacking uses a single dataset on a consistent scale \citep{sill2009feature, bhatt2017improved}. 
+Model stacking has proven effective in many realms \citep{sill2009feature, bhatt2017improved, hao2019review}. 
 Stacking improves predictions by controlling bias and variance; as long as suitably diverse models are averaged, they will have different biases while high variance in models should be averaged out even in relatively similar models.
@@
 Diversity in models is typically created in two ways: by using diverse training datasets (as in Random Forests for example) and by using functionally different models (for example by averaging tree based models and neural networks).
+One important trade-off in spatial modelling is whether to use local data (with a smaller sample size but that is likely to be representative of the area of study) or global data that have a larger sample size but a less close association with the areas of study.
+In the context of malaria mapping, we can think about diversity of training data in this context and expect that stacking seperate models trained on local and global data will also increase the diversity of predictions in a useful way.
 
-% more about stacking
-% good for bias variance
-% SDM examples. any others?
+Here we propose training a suite of level-zero machine learning models on point-level, binomial prevalence data and stacking these models with a polygon-level, Poisson incidence model.
+This modelling scheme has similarities to previous stacking methods used for malaria mapping \citep{bhatt2017improved}.
+However, as the response data in the level-zero models and the stacker are on different scales (prevalence is a proportion while incidence is a rate) there are a number of important differences.
+Notably, we cannot simply take a weighted average of the predictions from the level-zero models; they need to be transformed to the incidence scale and some additional flexibility to allow for mispecification of this transform needs to be allowed.
+To test the effectiveness of this approach we use data from four countries: Madagascar, Colombia, Indonesia and Senegal.
+We train the machine learning models on both local and global data and examine a number of different combinations of covariates: just local or global machine learning predictions, or combining the two or combining local machine learning predictions with the raw environmental covariates.
+While there isn't one consistently best model, we find that in nearly all cases, using some information from local machine learning models improves model performance.
 
-Diversity of predictions is good.
-Diversity of models
-Diversity of data. global local.
-
-Different scales. is it transfer learning?
-
-Here we propose training the level zero machine learning models on point-level, binomial prevalence data and stacking these models with a polygon-level, Poisson incidence model.
-
-{\bf 2. Methodology}
+\section{Methods}
 
@@
 We selected Colombia and Madagascar as case examples as they both have fairly complete, publicly available, surveillance data at a finer geographical resolution than admin 1.
-The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling}.  % ceck todo
+The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling, pfeffer2018ma}.
+% ceck todo
 The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
@@
 
-{\bf 3. Results}
+\section{Results}
 
@@
 }
-\label{f:scatter1}
+\label{f:scatter2}
 \end{figure}
@@
 \centering
-\includegraphics[trim={0 30mm 0 40mm}, width = 0.55\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
+\includegraphics[trim={0 30mm 0 40mm}, width = 0.8\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
 \caption{
@@
 \begin{table}
-\caption{Machine learning model results and means of fitted parameters (i.e.\thinspace model weights) across cross-validation folds of the machine learning predictions only model. }
+\caption{Machine learning model results and fitted parameters (i.e.\thinspace model weights) of the machine learning predictions only models (local only and global only). }
 \centering
@@
 
-{\bf 5. Conclusions}
+\section{Discussion}
+
+\section{Conclusions}
 

### 2019-10-10 tim-lucas-model-ensembles.tex
Revise methods.

@@
 We used two data sources that reflect \emph{Plasmodium falciparum} malaria transmission; point-prevalence surveys and polygon-level, aggregated incidence data. 
-We selected Colombia and Madagascar as case examples as they both have fairly complete, publicly available, surveillance data at a finer geographical resolution than admin 1.
+We selected Madagascar, Colombia, Indonesia and Senegal as case examples as they all have fairly complete, publicly available, surveillance data at a finer geographical resolution than admin 1.
 The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling, pfeffer2018ma}.
-% ceck todo
 The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
@@
 To minimise temporal effects we selected, for each country, one year of surveillance data. 
-We used annual surveillance data from 2015 for Colombia (952 municipalities) and data from 2013 for Madagascar (110 districts) as these years had the most data in each case.
+We used annual surveillance data from data from 2013 for Madagascar (110 districts), 2015 for Colombia (952 municipalities), 2013 for Indonesia (244 regencies and cities) and 2009 for Senegal (34 departments).
+These years were selected as they had the most data in each case.
 
-We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \citep{weiss2015re}.
+We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \citep{weiss2015re, weiss2018global}.
 Tasseled cap brightness and urban land cover were subsequently removed as they were highly correlated with other variables. 
@@
 These models were fitted to both the full malaria prevalence dataset and to more regional subsets of the data.
-For Colombia we used all points from South America (n = 522) while for Madagascar we used only Malagasy data (n = 1505).
+For Colombia we used all points from South America (locations = 522, individuals = 7,719) while for the other countries we used only data from that country (Madagascar: n = 1,505, individuals = 89,381.
+Indonesia: n = 4,778, individuals = 1,512,888.
+Senegal: n = 1,762, individuals = 80,896).
 We chose these geographic regions based on a trade-off between wanting a large sample size but wanting data from geographically similar areas.
@@
 For each model we ran  five-fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
-The parameters tried can be seen in Table S1.
-Predictions from these models were then made across Colombia and Madagascar respectively.
-These predictions were finally inverse logit transformed so that they are on the linear predictor scale of the top level model.
+Predictions from these models were then made across each country respectively.
+These predictions were finally empirical logit transformed so that they are on the linear predictor scale of the top level model.
+An empirical logit was used rather than a standard logit as there were many predictions of exactly zero.
 
 The top level model was a disaggregation regression model \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
+The models were implemented and fitted using Template Model Builder \citep{TMB} in R \citep{R}.
 This model is defined by a likelihood at the level of the polygon with covariates and a spatial random field at the pixel-level. 
@@
 The fact that the model passes through prevalence space ensures that the predictions from the machine learning models can be appropriately scaled.
-The linear predictor of the model is related to prevalence by a typical logit link function and includes an intercept, $\beta_0$, covariates, $X$ with regression parameters $\beta$, a spatial, Gaussian, random field, $u(s, \rho, \sigma_u)$, and an \emph{iid} random effect, $v_j(\sigma_v)$.
-$$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u(s, \rho, \sigma_u) + v_j(\sigma_v)\right)$$
+The linear predictor of the model is related to prevalence by a typical logit link function and includes an intercept, $\beta_0$, covariates, $X$, with regression parameters $\beta$, a spatial, Gaussian, random field, $u_s(\rho, \sigma_u)$, and an \emph{iid} random effect, $v_j(\sigma_v)$.
+$$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u_s(\rho, \sigma_u) + v_j(\sigma_v)\right)$$
 The Gaussian spatial effect has a Mat\'ern covariance function and two hyperparameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
@@
 
-This is not true stacking because we don't constrain betas to be positive.
-preliminary analyses indicated that this didn't improve performance.
-
 Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$ and $\sigma_v$. 
-We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \citep{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$. 
+We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \citep{fuglstad2018constructing} such that $P(\rho < 1) = 0.0001$ (except for Indonesia where we set $P(\rho < 3) = 0.0001$ due to it's much larger size) and $P(\sigma_u > 1) = 0.0001$.
 This prior encoded our \emph{a priori} preference for a simpler, smoother random field.
@@
 
-We assigned $\sigma_v$ a penalised complexity prior \citep{simpson2017penalising} such that $P(\sigma_v > 0.05) = 0.0000001$. 
+We assigned $\sigma_v$ a penalised complexity prior \citep{simpson2017penalising} such that $P(\sigma_v > 0.05) =  0.0001$. 
 This was based on a comparison of the variance of Poisson random variables, with rates given by the number of polygon-level cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \citep{cibulskis2011worldwide}.
 We found that an \emph{iid} effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
-Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.4)$. 
-The models were implemented and fitted using Template Model Builder \citep{TMB} in R \citep{R}.
 
-We compared the performance of the models with four sets of covariates, $X$.
+Finally, we set different priors on the regression coefficients depending on which covariates were used.
+When environmental covariates or a mix of environmental covariates and predictions from machine learning models were used we set the prior to be weakly regularising, $\beta_i \sim \operatorname{ Norm}(0, 0.4)$, such that it is unlikely that any single covariate could explain the full range of the response data.
+When only machine learning model predictions were used we set $\beta_i \sim \operatorname{ Norm}(\frac{1}{M}, 0.4)$ where $M$ is the number of machine learning models being used. 
+This prior  sets our \emph{a priori} expectation that all the machine learning prediction models are positively and equally correlated with incidence i.e. this prior encodes standard model averaging.
+It is important to note that this setup does not constitute true stacking in which we would enforce $\sum_i \beta_i = 1$ \citep{bhatt2017improved}.
+In a preliminary analysis we tested the case where we force $\beta_i > 0$ which in practice is largely the same as the $\sum_i \beta_i = 1$  \citep{breiman1996stacked} but allows a small amount of flexibility to handle mispecification in $\mathrm{p2i}$.
+This analysis did not find any benefits to this approach so it was not considered further.
+
+We compared the performance of the models with five sets of covariates, $X$.
 Firstly, we used the environmental and anthropogenic covariates, centered and standardised as a baseline model.
-Secondly, we used the predictions from the machine learning models trained on regional data.
-Thirdly, we used predictions from the machine learning models trained on the global data.
-Finally we combined predictions from the machine learning models trained on regional data and global data.
-Preliminary analyses showed that combining environmental covariates with machine learning predictions was not effective so we did not pursue this further.
+Secondly, we used the predictions from the machine learning models trained on local data.
+Thirdly, we combined the predictions from the machine learning models and the environmental coviarates in one model.
+Fourthly, we used predictions from the machine learning models trained on the global data.
+Finally, we combined predictions from the machine learning models trained on regional data and global data.
 
-To compare the four models we used two cross-validation schemes. 
+To compare the five models we used two cross-validation schemes. 
 In the first, polygon incidence data was randomly split into six cross-validation folds.
-In the second, polygon incidence data was split spatially into three folds (via k-means clustering on the polygon centroids).
+In the second, polygon incidence data was split spatially into k folds (via k-means clustering on the polygon centroids).
+We set k as 3 for Madagascar and Colombia.
+Due to its large size we set k as 7 for Indonesia.
+Due to the small sample size, we set k as 5 for Senegal.
 This spatial cross-validation scheme is testing the models' ability to make predictions far from data where the spatial random field is not informative.
@@
 Figures~\ref{f:scatter1} and \ref{f:scatter2} show the model performance under random and spatial cross-validation for all countries.
-The poor model performance in Colombia under spatial cross-validation indicates that the covariates alone cannot explain malaria incidence in this area. 
-For all other models that use the machine learning predictions as covariates, correlations between observed and predicted data of 0.54 -- 0.76 were achieved (Table~\ref{t:results}).
-Input data and mapped out-of-sample predictions of the best performing model, in Colombia, are shown in in Figure~\ref{f:map}.
 
@@
 \end{tabular}
-\label{t:results}
+\label{t:results2}
 \end{table}
@@
 % to do table of coverage.
-
-The model using only machine learning predictions as covariates was the best performing model in both countries and both cross-validation schemes (Table~\ref{t:results}).
-As expected, models performed better in the random cross-validation scheme than the spatial cross-validation scheme.
-The difference between the covariate only model and the machine learning predictions only model was greater in the spatial cross-validation scheme than in the random cross-validation.
-The improvement in performance between the worst and best models was always smaller than the difference between the random and spatial cross-validation schemes.
-
-Predictive performance of machine learning models was similar, with Random Forest performing best in Madagascar and neural networks, Random Forests and elastic net performing equally well in Colombia (Table~\ref{t:mlresults}).
-The regression coefficients (i.e.\thinspace the weights of the machine learning models in the level zero model) from the polygon-level models that used only predictions from machine learning models as covariates can also be seen in Table~\ref{t:mlresults}.
-The estimated regression parameters are similar between the random and spatial cross-validation schemes.
-However, the best performing machine learning models do not have the largest estimated regression coefficients as would be expected if prevalence and incidence were completely correlated.
-Also of note is that some models were estimated to have a negative relationship with incidence (conditional on the inclusion of predictions from other machine learning models).
-% Compare plots of global vs local. eg amazon predicted low by global.
-% Look at \betas for global and local.
-
-% models were quite good except spatial colombia
-% which model performed best
-% Random field parameters?
-% coverage
-% which ml model was best Vs parameters.
-% parameters are negative and don't sum to one
-%
-% differences between models was less than difference between data.
 
@@
 \end{figure}
-
-% copy Sam's plot. predictions from ml and stacked.
 

### 2019-10-10 tim-lucas-model-ensembles.tex
Add experimental condition names and Malagasy acknowledgement.

@@
 We compared the performance of the models with five sets of covariates, $X$.
-Firstly, we used the environmental and anthropogenic covariates, centered and standardised as a baseline model.
-Secondly, we used the predictions from the machine learning models trained on local data.
-Thirdly, we combined the predictions from the machine learning models and the environmental coviarates in one model.
-Fourthly, we used predictions from the machine learning models trained on the global data.
-Finally, we combined predictions from the machine learning models trained on regional data and global data.
+Firstly, we used the environmental and anthropogenic covariates, centered and standardised as a baseline model (subsequantly called Enviro).
+Secondly, we used the predictions from the machine learning models trained on local data (ML\textsubscript{l}).
+Thirdly, we combined the predictions from the machine learning models and the environmental coviarates in one model (Enviro + ML\textsubscript{l}).
+Fourthly, we used predictions from the machine learning models trained on the global data (ML\textsubscript{g}).
+Finally, we combined predictions from the machine learning models trained on regional data and global data (ML\textsubscript{l} + ML\textsubscript{g}).
 
@@
 
+\section*{Acknowledgments}
+The authors acknowledge the National Malaria Control Programme of Madagascar for sharing their routine case data for this analysis.
+
 \bibliography{Malaria} 

### 2019-10-10 tim-lucas-model-ensembles.tex
Switch table columns.

@@
 The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling, pfeffer2018ma}.
+While the data cover a large time period, we did not model time explicitely but future work could include this component.
 The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
@@
 \begin{tabular}{llrrrr}
-Cross-validation scheme & Country &  Enviro &  ML\textsubscript{l} &  Enviro + ML\textsubscript{l} \\
+Country &  Cross-validation & Enviro &  ML\textsubscript{l} &  Enviro + ML\textsubscript{l} \\
 \hline 
- Random &  Colombia & 0.59 & 0.59 & \textbf{0.61} \\
- Spatial &  Colombia &  0.12 &  \textbf{0.33} &  0.25\\
- Random &  Madagascar &  0.70 &  \textbf{0.73} & 0.69 \\
- Spatial &  Madagascar &  0.22 &  \textbf{0.66} & 0.18\\
- Random &  Senegal &  \textbf{0.58} &  0.51 & 0.57 \\
- Spatial &  Senegal &  \textbf{0.63} &  0.27 &  \textbf{0.63} \\
- Random &  Indonesia &  0.52 &  0.48 &  \textbf{0.59} \\
- Spatial &  Indonesia &  0.45 &  0.44 &  \textbf{0.51} \\
+ Colombia & Random &  0.59 & 0.59 & \textbf{0.61} \\
+ Colombia &  Spatial &  0.12 &  \textbf{0.33} &  0.25\\
+ Madagascar & Random &   0.70 &  \textbf{0.73} & 0.69 \\
+ Madagascar &  Spatial &  0.22 &  \textbf{0.66} & 0.18\\
+ Senegal &  Random &  \textbf{0.58} &  0.51 & 0.57 \\
+ Senegal &  Spatial &  \textbf{0.63} &  0.27 &  \textbf{0.63} \\
+ Indonesia &  Random &  0.52 &  0.48 &  \textbf{0.59} \\
+ Indonesia &  Spatial &  0.45 &  0.44 &  \textbf{0.51} \\
 \end{tabular}
@@
 \begin{tabular}{llrrrr}
-Cross-validation scheme & Country &  Covariates &   ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
+Country &  Cross-validation & Enviro &   ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
 \hline 
- Random &  Colombia & \textbf{0.59} &0.58 & 0.56 \\
- Spatial &  Colombia &  0.12 &  0.18 & \textbf{0.34}\\
- Random &  Madagascar &  0.70 &  0.70 & \textbf{0.71} \\
- Spatial &  Madagascar &  0.22 &  \textbf{0.63} & 0.60\\
- Random &  Senegal &  \textbf{0.58} &  0.30 & 0.08 \\
- Spatial &  Senegal &  \textbf{0.63} & 0.36 & 0.35 \\
- Random &  Indonesia &  \textbf{0.52} & 0.46 & 0.32 \\
- Spatial &  Indonesia &  \textbf{0.45} & 0.41 & \textbf{0.45} \\
+ Colombia & Random &  \textbf{0.59} &0.58 & 0.56 \\
+ Colombia &  Spatial &  0.12 &  0.18 & \textbf{0.34}\\
+ Madagascar &  Random &  0.70 &  0.70 & \textbf{0.71} \\
+ Madagascar &  Spatial &  0.22 &  \textbf{0.63} & 0.60\\
+ Senegal &  Random &  \textbf{0.58} &  0.30 & 0.08 \\
+ Senegal &  Spatial &  \textbf{0.63} & 0.36 & 0.35 \\
+ Indonesia &  Random &  \textbf{0.52} & 0.46 & 0.32 \\
+ Indonesia &  Spatial &  \textbf{0.45} & 0.41 & \textbf{0.45} \\
 \end{tabular}
@@
 
+\begin{table}[h!]
+\caption{Covarage of 80\% credible intervals. }
+\centering
+\begin{tabular}{llrrrr}
+CV & Country &  Enviro & ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
+\hline 
+ Random &  Colombia &  \\
+ Spatial &  Colombia &  \\
+ Random &  Madagascar &  \\
+ Spatial &  Madagascar & \\
+ Random &  Senegal & \\
+ Spatial &  Senegal &  \\
+ Random &  Indonesia &  \\
+ Spatial &  Indonesia &   \\
+\end{tabular}
+\label{t:results2}
+\end{table}
+
 \section{Discussion}

### 2019-10-10 tim-lucas-model-ensembles.tex
Add coverage table.

@@
 \begin{table}[h!]
-\caption{Covarage of 80\% credible intervals. }
+\caption{Covarage of 80\% credible intervals. Values outside 0.7--0.9 are shown in bold.}
 \centering
-\begin{tabular}{llrrrr}
-CV & Country &  Enviro & ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
+\begin{tabular}{llrrrrr}
+CV & Country &  Enviro & ML\textsubscript{l} &  Enviro + ML\textsubscript{l} & ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
 \hline 
- Random &  Colombia &  \\
- Spatial &  Colombia &  \\
- Random &  Madagascar &  \\
- Spatial &  Madagascar & \\
- Random &  Senegal & \\
- Spatial &  Senegal &  \\
- Random &  Indonesia &  \\
- Spatial &  Indonesia &   \\
+ Colombia & Random & \textbf{0.28} & \textbf{0.28} & \textbf{0.29} & \textbf{0.29} & \textbf{0.29} \\
+ Madagascar &  Random & 0.80 & 0.77& 0.75& 0.77& 0.76 \\
+ Senegal & Random &0.79 & 0.79& 0.79& 0.79& 0.82 \\
+ Indonesia & Random &0.80 & 0.81& 0.78& 0.79& 0.77  \\
+ Colombia &  Spatial & \textbf{0.31} & \textbf{0.34}  & \textbf{0.32} & \textbf{0.34} & \textbf{0.34}  \\
+ Madagascar & Spatial & \textbf{0.65} & 0.74& 0.70& 0.75 & 0.76   \\
+ Senegal & Spatial & 0.85 & \textbf{0.91}& 0.85& \textbf{0.94} & 0.85  \\
+  Indonesia & Spatial & 0.80 & 0.78& 0.78& 0.76& 0.75  \\
 \end{tabular}

### 2019-10-10 tim-lucas-model-ensembles.tex
Add commented out para table.

@@
 \begin{tabular}{llrrrrr}
-CV & Country &  Enviro & ML\textsubscript{l} &  Enviro + ML\textsubscript{l} & ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
+Country &  CV & Enviro & ML\textsubscript{l} &  Enviro + ML\textsubscript{l} & ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
 \hline 
@@
 \end{table}
+
+%\begin{table}
+%\caption{Fitted regression parameters of the environmental and machine learning covariates for thee models: just environmental variables, just local machine learning models and the combination of the two.}
+%\centering
+%\tiny
+%\begin{tabular}{ll|rrr}
+%Country    & Variable &  $\beta_{enviro}$   & $\beta_l$ & $\beta_{enviro+l}$\\ \hline
+%Madagascar & LST mean & 0.100 & 0.538 &  0.169\\
+%Madagascar & EVI & 0.105 & 0.570 & 0.178 \\
+%Madagascar & TSI & 0.116 & 0.301 &0.233  \\
+%Madagascar & access & 0.113 & 0.033 &0.220  \\
+%Madagascar & Elev & 0.109 & 0.469 & 0.205\\ 
+%Madagascar & LST sd & 0.116 & 0.301 &0.233  \\
+%Madagascar & VIIRS & 0.113 & 0.033 &0.220  \\
+%Madagascar & TCW & 0.109 & 0.469 & 0.205\\ 
+%Madagascar & RF & 0.100 & 0.538 &  0.169\\
+%Madagascar & GBM & 0.105 & 0.570 & 0.178 \\
+%Madagascar & enet & 0.116 & 0.301 &0.233  \\
+%Madagascar & nnet & 0.113 & 0.033 &0.220  \\
+%Madagascar & ppr & 0.109 & 0.469 & 0.205\vspace{0.2cm}\\ 
+%Colombia & LST mean & 0.100 & 0.538 &  0.169\\
+%Colombia & EVI & 0.105 & 0.570 & 0.178 \\
+%Colombia & TSI & 0.116 & 0.301 &0.233  \\
+%Colombia & access & 0.113 & 0.033 &0.220  \\
+%Colombia & Elev & 0.109 & 0.469 & 0.205\\ 
+%Colombia & LST sd & 0.116 & 0.301 &0.233  \\
+%Colombia & VIIRS & 0.113 & 0.033 &0.220  \\
+%Colombia & TCW & 0.109 & 0.469 & 0.205\\ 
+%Colombia & RF & 0.068 & 0.505 &  0.169 \\
+%Colombia & GBM & 0.073 & 0.879 & 0.178 \\
+%Colombia & enet & 0.070 & 0.079 &0.233 \\
+%Colombia & nnet & 0.070 & -0.008 &0.220  \\
+%Colombia & ppr & 0.070 & 0.530 & 0.205 \vspace{0.2cm}\\
+%Senegal & LST mean & 0.100 & 0.538 &  0.169\\
+%Senegal & EVI & 0.105 & 0.570 & 0.178 \\
+%Senegal & TSI & 0.116 & 0.301 &0.233  \\
+%Senegal & access & 0.113 & 0.033 &0.220  \\
+%Senegal & Elev & 0.109 & 0.469 & 0.205\\ 
+%Senegal & LST sd & 0.116 & 0.301 &0.233  \\
+%Senegal & VIIRS & 0.113 & 0.033 &0.220  \\
+%Senegal & TCW & 0.109 & 0.469 & 0.205\\ 
+%Senegal & RF & 0.092 & 0.339 & 0.169  \\
+%Senegal & GBM & 0.099 & 0.261& 0.178 \\
+%Senegal & enet& 0.103 & 0.344  &0.233  \\
+%Senegal & nnet & 0.099 & 0.254 &0.220 \\
+%Senegal & ppr & 0.098 & 0.268& 0.205 \vspace{0.2cm}\\
+%Indonesia & LST mean & 0.100 & 0.538 &  0.169\\
+%Indonesia & EVI & 0.105 & 0.570 & 0.178 \\
+%Indonesia & TSI & 0.116 & 0.301 &0.233  \\
+%Indonesia & access & 0.113 & 0.033 &0.220  \\
+%Indonesia & Elev & 0.109 & 0.469 & 0.205\\ 
+%Indonesia & LST sd & 0.116 & 0.301 &0.233  \\
+%Indonesia & VIIRS & 0.113 & 0.033 &0.220  \\
+%Indonesia & TCW & 0.109 & 0.469 & 0.205\\ 
+%Indonesia & RF& 0.081 & 0.447 & 0.169 \\
+%Indonesia & GBM & 0.085 & 0.357 & 0.178\\
+%Indonesia & enet & 0.091 & 0.303 &0.233  \\
+%Indonesia & nnet & 0.089 & 0.506 &0.220  \\
+%Indonesia & ppr & 0.089 & 0.364 & 0.205 \\
+%
+%
+%
+%\end{tabular}
+%\label{t:mlresults}
+%\end{table}
+%
+%
 

### 2019-10-10 tim-lucas-model-ensembles.tex
Fiull in table of mle. Using means over cv though.

--- 2019-10-10 tim-lucas-model-ensembles.tex
@@
 \label{t:results2}
+\end{table}
+
+\begin{table}
+\caption{Fitted regression parameters of the environmental and machine learning covariates for thee models: just environmental variables, just local machine learning models and the combination of the two.}
+\centering
+\tiny
+\begin{tabular}{ll|rrr}
+Country    & Variable &  $\beta_{enviro}$   & $\beta_l$ & $\beta_{enviro+l}$\\ \hline
+Madagascar & LST mean & 0.17 &  &  -0.01\\
+Madagascar & EVI & 0.32 &  & -0.04 \\
+Madagascar & TSI & 0.42 &  &0.06  \\
+Madagascar & access & 0.34 &  &0.00  \\
+Madagascar & Elev & -0.14 &  & 0.00\\ 
+Madagascar & LST sd & -0.78 &  &-0.04  \\
+Madagascar & VIIRS & -0.1 &  &-0.52  \\
+Madagascar & TCW & -0.22 &  & -0.06\\ 
+Madagascar & RF &  & 0.29 &  0.01\\
+Madagascar & GBM &  & 0.52 & 0.25 \\
+Madagascar & enet &  & 0.31 &0.09  \\
+Madagascar & nnet &  & 0.43 &0.14  \\
+Madagascar & ppr &  & 0.49 & 0.22\vspace{0.2cm}\\ 
+Colombia & LST mean &-0.12  &  & -0.22\\
+Colombia & EVI &  -0.08&  & 0.17\\
+Colombia & TSI &  0.54&  & -0.06\\
+Colombia & access &  0.35&  &0.00 \\
+Colombia & Elev &  0.47&  & 0.00\\
+Colombia & LST sd & -0.11 &  & -0.10\\
+Colombia & VIIRS & 0.24 &  & 0.01\\
+Colombia & TCW &  0.21&  & 0.00\\
+Colombia & RF &  & 0.36 & -0.18\\
+Colombia & GBM &  & 0.54 & 0.26\\
+Colombia & enet &  & 0.32 & -0.34\\
+Colombia & nnet &  &0.33  & -0.09\\
+Colombia & ppr &  & 0.54 & 0.01\vspace{0.2cm}\\
+Senegal & LST mean &-0.04  &  &-0.02\\
+Senegal & EVI &  0.33&  & 0.30\\
+Senegal & TSI &  -0.58&  &  -0.59\\
+Senegal & access & -0.03 &  &  -0.05\\
+Senegal & Elev &  0.42&  &0.37\\ 
+Senegal & LST sd &  0.21&  & 0.10 \\
+Senegal & VIIRS &  0.13&  &  0.13\\
+Senegal & TCW & 0.08 &  &-0.01\\ 
+Senegal & RF &  & 0.34 &  0.20\\
+Senegal & GBM &  &  0.28& 0.05\\
+Senegal & enet &  &  0.27&  -0.05\\
+Senegal & nnet &  & 0.29 &-0.09 \\
+Senegal & ppr &  &  0.35& 0.25\vspace{0.2cm}\\
+Indonesia & LST mean & -0.32 &  &-0.34\\
+Indonesia & EVI &  0.49&  & 0.36\\
+Indonesia & TSI &  0.58&  &0.52\\
+Indonesia & access &0.33 &  & 0.06 \\
+Indonesia & Elev& -0.06&  &-0.13\\ 
+Indonesia & LST sd & -0.23 &  & -0.33\\
+Indonesia & VIIRS &  -0.12&  & 0.03\\
+Indonesia & TCW& 0.12 &  &0.10\\ 
+Indonesia & RF &  & 0.32 & 0.06\\
+Indonesia & GBM &  &  0.38&0.03\\
+Indonesia & enet &  &  0.49&  0.56\\
+Indonesia & nnet &  &  0.36&  0.18\\
+Indonesia & ppr &  &  0.45& 0.32\\
+
+\end{tabular}
+\label{t:mlresults}
 \end{table}

### 2019-10-10 tim-lucas-model-ensembles.tex
Start writing results.

@@
 \section{Results}
+
+Table \ref{t:results} gives the correlation between observed and held out data (under random and spatially cross-validation) for models that use environmental covariates only, local machine learning models only and a combination of the two.
+Many of the differences in performance are rather marginal.
+Overall, in only one case (random cross-validation in Senegal) is Enviro the best performing model while under spatially cross-validation in Senegal Enviro and Envir + ML\textsubscript{l} perform equally well.
+Of the remaining six cases, in three cases ML\textsubscript{l} performs best and in three cases Envir + ML\textsubscript{l} performs best.
+The greatest benefits in predictive performance occur under spatial cross validation and in the cases when the Enviro model is particularly poor.
+
+Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatially cross-validation) for models that use environmental covariates only, global machine learning models only and a combination of the local and global machine learning models.
+In five cases, Enviro is the best or tied best performing model.
+In only one case in the ML\textsubscript{g} model the best.
+Comparing across Tables \ref{t:results} and \ref{t:results2} we can see that ML\textsubscript{g} is never the best performing model.
+In only three cases (Spatial Colombia, Spatial Senegal and Spatial Indonesia) does ML\textsubscript{g} outperform ML\textsubscript{l}.
 

### 2019-10-10 tim-lucas-model-ensembles.tex
Delete commented out table from wsc
Vaguely complete 0.9 draft.

@@
 Susan Rumisha\textsuperscript{1}; 
-Katherine E. Battle\textsuperscript{1}; Rosalind E. Howes\textsuperscript{1}; Andre Python\textsuperscript{1}; Penny Hancock\textsuperscript{1}; Punam Someone\textsuperscript{1};
+Katherine E. Battle\textsuperscript{1}; Rosalind E. Howes\textsuperscript{1}; Andre Python\textsuperscript{1}; Penny Hancock\textsuperscript{1}; Punam Amratia\textsuperscript{1};
 Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel~J.~Weiss\textsuperscript{1}.}\\
@@
 For each model we ran  five-fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
+Root mean square error (RMSE) was used to select the best performing model.
 Predictions from these models were then made across each country respectively.
@@
 
-Table \ref{t:results} gives the correlation between observed and held out data (under random and spatially cross-validation) for models that use environmental covariates only, local machine learning models only and a combination of the two.
-Many of the differences in performance are rather marginal.
-Overall, in only one case (random cross-validation in Senegal) is Enviro the best performing model while under spatially cross-validation in Senegal Enviro and Envir + ML\textsubscript{l} perform equally well.
-Of the remaining six cases, in three cases ML\textsubscript{l} performs best and in three cases Envir + ML\textsubscript{l} performs best.
-The greatest benefits in predictive performance occur under spatial cross validation and in the cases when the Enviro model is particularly poor.
-
-Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatially cross-validation) for models that use environmental covariates only, global machine learning models only and a combination of the local and global machine learning models.
-In five cases, Enviro is the best or tied best performing model.
-In only one case in the ML\textsubscript{g} model the best.
-Comparing across Tables \ref{t:results} and \ref{t:results2} we can see that ML\textsubscript{g} is never the best performing model.
-In only three cases (Spatial Colombia, Spatial Senegal and Spatial Indonesia) does ML\textsubscript{g} outperform ML\textsubscript{l}.
-
-Figures~\ref{f:scatter1} and \ref{f:scatter2} show the model performance under random and spatial cross-validation for all countries.
-
-%\begin{table}[h!]
-%\caption{Pearson correlations between observed and predicted values. }
-%\centering
-%\begin{tabular}{llrrr}
-%Cross-validation scheme & Country &  Covariates &  ML &  Covs + ML \\
-%\hline 
-% Random &  Colombia &  0.45 &  \textbf{0.55} &  0.54 \\
-% Random &  Madagascar &  0.70 &  \textbf{0.76} &  0.75 \\
-% Spatial &  Colombia &  0.05 &  \textbf{0.18} &  0.10 \\
-% Spatial &  Madagascar &  0.22 &  \textbf{0.63} &  0.61 
-%\end{tabular}
-%\label{t:results}
-%\end{table}
-
-\begin{table}[h!]
+\begin{table}[t!]
 \caption{Pearson correlations between observed and predicted values. }
@@
 
-\begin{table}[h!]
+Table \ref{t:results} gives the correlation between observed and held out data (under random and spatially cross-validation) for models that use environmental covariates only, local machine learning models only and a combination of the two.
+Many of the differences in performance are rather marginal.
+Overall, in only one case (random cross-validation in Senegal) is Enviro the best performing model while under spatially cross-validation in Senegal Enviro and Envir + ML\textsubscript{l} perform equally well.
+Of the remaining six cases, in three cases ML\textsubscript{l} performs best and in three cases Envir + ML\textsubscript{l} performs best.
+The greatest benefits in predictive performance occur under spatial cross validation and in the cases when the Enviro model is particularly poor.
+
+Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatially cross-validation) for models that use environmental covariates only, global machine learning models only and a combination of the local and global machine learning models.
+In five cases, Enviro is the best or tied best performing model.
+In only one case in the ML\textsubscript{g} model the best.
+Comparing across Tables \ref{t:results} and \ref{t:results2} we can see that ML\textsubscript{g} is never the best performing model.
+In only three cases (Spatial Colombia, Spatial Senegal and Spatial Indonesia) does ML\textsubscript{g} outperform ML\textsubscript{l}.
+
+Figures~\ref{f:scatter1} shows scatter plots of the model performance under random cross-validation for all countries and all models.
+Figures~\ref{f:scatter2} shows scatter plots of the model performance under spatial cross-validation for all countries and all models.
+It can be seen that without environmental covariates, the models in Madagascar fail to predict very high or very low values correctly.
+Figure~\ref{f:map} shows the input data and out-of-sample predictions of the ML\textsubscript{l} model in Colombia.
+Table~\ref{t:coverage} shows the out-of-sample coverage of the 80\% credible intervals for all models, countries and cross-validation schemes.
+The coverage in Colombia is very poor with no models achieving a coverage above 0.4.
+Otherwise the coverage is generally acceptable with most values lying between 0.7 and 0.9.
+Over there is a general tendancy for models to be slightly over confident.
+
+We can examine the relationship between the RMSE of the machine learning models to their fitted regression coefficients (weights).
+These values are given in Table~\ref{t:mlresults} (with the RMSE of the global models being copied four times).
+In all five sets of machine learning models, Random Forest performs the best.
+While we have not forced $\beta_i > 0$, the priors for these coefficients were positive.
+In nearly all cases the fitted values are positive.
+If the prevalence data and incidence data are not biased relative to each other we would expect the models with the lowest RMSE to also have the biggest regression coefficient.
+This occurs in four cases where Random Forest has the lowest RMSE and the biggest coefficient.
+In a further two cases, Random Forest has the lowest RMSE but GBM has the highest coefficient.
+The predictions from Random Forest and GBM are highly correlated in part because they both perform well and because they are both tree based models.
+Finally, it can be seen that relationship between RMSE and regression coefficients is much weaker in Indonesia.
+For the models trained on local data a neural network has the highest fitted coefficient while for the models trained on global data an elastic net has the highest fitted coefficient.
+These model both produce very smooth estimates relative to the tree based methods.
+
+\begin{table}[t!]
 \caption{Pearson correlations between observed and predicted values. }
@@
 
+\begin{table}[h!]
+\caption{Covarage of 80\% credible intervals. Values outside 0.7--0.9 are shown in bold.}
+\centering
+\begin{tabular}{llrrrrr}
+Country &  CV & Enviro & ML\textsubscript{l} &  Enviro + ML\textsubscript{l} & ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
+\hline 
+ Colombia & Random & \textbf{0.28} & \textbf{0.28} & \textbf{0.29} & \textbf{0.29} & \textbf{0.29} \\
+ Madagascar &  Random & 0.80 & 0.77& 0.75& 0.77& 0.76 \\
+ Senegal & Random &0.79 & 0.79& 0.79& 0.79& 0.82 \\
+ Indonesia & Random &0.80 & 0.81& 0.78& 0.79& 0.77  \\
+ Colombia &  Spatial & \textbf{0.31} & \textbf{0.34}  & \textbf{0.32} & \textbf{0.34} & \textbf{0.34}  \\
+ Madagascar & Spatial & \textbf{0.65} & 0.74& 0.70& 0.75 & 0.76   \\
+ Senegal & Spatial & 0.85 & \textbf{0.91}& 0.85& \textbf{0.94} & 0.85  \\
+  Indonesia & Spatial & 0.80 & 0.78& 0.78& 0.76& 0.75  \\
+\end{tabular}
+\label{t:coverage}
+\end{table}
+
 \begin{table}
@@
 \label{t:mlresults}
-\end{table}
-
-\begin{table}[h!]
-\caption{Covarage of 80\% credible intervals. Values outside 0.7--0.9 are shown in bold.}
-\centering
-\begin{tabular}{llrrrrr}
-Country &  CV & Enviro & ML\textsubscript{l} &  Enviro + ML\textsubscript{l} & ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
-\hline 
- Colombia & Random & \textbf{0.28} & \textbf{0.28} & \textbf{0.29} & \textbf{0.29} & \textbf{0.29} \\
- Madagascar &  Random & 0.80 & 0.77& 0.75& 0.77& 0.76 \\
- Senegal & Random &0.79 & 0.79& 0.79& 0.79& 0.82 \\
- Indonesia & Random &0.80 & 0.81& 0.78& 0.79& 0.77  \\
- Colombia &  Spatial & \textbf{0.31} & \textbf{0.34}  & \textbf{0.32} & \textbf{0.34} & \textbf{0.34}  \\
- Madagascar & Spatial & \textbf{0.65} & 0.74& 0.70& 0.75 & 0.76   \\
- Senegal & Spatial & 0.85 & \textbf{0.91}& 0.85& \textbf{0.94} & 0.85  \\
-  Indonesia & Spatial & 0.80 & 0.78& 0.78& 0.76& 0.75  \\
-\end{tabular}
-\label{t:results2}
 \end{table}
@@
 
-\section{Conclusions}
-
 Overall, our experiments suggest that using predictions from machine learning models trained on prevalence points provides more accurate predictions than using environmental covariates when fitting disaggregation models of malaria incidence.
 This increased performance comes despite the data being on different scales, the data being measurements of different aspects of malaria transmission and despite the imperfect model we have used to translate between the two scales.
-However, the reduced model accuracy in the spatial cross-validation schemes, relative to the random cross-validation scheme, highlights that better spatial coverage of data would improve predictions more than the improved model we have suggested.
+However, there is no clear best model between ML\textsubscript{l} and Enviro + ML\textsubscript{l}.
+Therefore, when using these methods, both of these models should be fitted and the most predictive model selected for a given dataset.
+Furthermore, many of the performance improvements are rather marginal. 
+However, in a few cases such as Colombia and Madagascar under spatial cross-validation the performance boost is large.
+For example, using just environmental covariates in Colombia under spatial cross-validation gives a correlation between observed and predicted data of 0.12.
+Such a model would be unusable for any applied or policy work.
+In contrast, using the predictions from the machine learning models trained on local data gives a correlation value of 0.33, which while still poor is possibly useable.
 
-Due to the low power of typical aggregated incidence datasets, previous analyses using disaggregation regression used a small number of covariates \citep{sturrock2014fine}.
-However, as models such as Random Forest and elastic net can robustly handle high dimensional data, future work could include many more covariates, potentially increasing predictive performance.
+As expected, the model performance was generally worse under spatial cross-validation than under random cross-validation.
+This implies that the models are still relying heavily on the spatial Gaussian random field.
+Furthermore, the difference between random and spatial cross-validation is usually bigger than the difference between different models.
+This suggests that better data coverage is more important than which specific model is used.
 
@@
 % sum to one stacking
+
 % global ml models, local experts
+
+Overall, the models fitted using predictions from machine learning models trained on the global database of prevalence point-surveys were less good than those using predictions from machine learning models trained on local data.
+This in itself is not particularly surprising.
+However, it does indicate that continental scale models should consider using a mosaic of locally trained machine learning models for example.
+What is more surprising is that using both the ML\textsubscript{l} and ML\textsubscript{g} covariates together did not improve performance given that this approach of creating diverse predictions is a core element of stacking methodology.
 
@@
 We did not include these constraints because the base models and the meta model are trained on response data on different scales.
-However, future work could examine whether using a positive constraint on the regression parameters improves performance.
+However, given our priors, nearly all the fitted regression coefficients in models with only machine learning predictions were positive.
+Therefore, in practice these models are working in a way similar to standard stacking.
+However, the coefficients certainly do not sum to one, and fitted intercepts are very small to account for this.
 
-Another area of potential improvement is varying the data used to train the base level learners.
-Here we only used data from the region of interest.
-However, the global dataset is much larger than these subsets.
-Training some base level models on local data and some on the global dataset and then combining predictions from all these models has potential to further improve model performance.
+Due to the low power of typical aggregated incidence datasets, previous analyses using disaggregation regression used a small number of covariates \citep{sturrock2014fine}.
+However, as models such as Random Forest and elastic net can robustly handle high dimensional data, future work could include many more covariates in the machine learning models, potentially increasing predictive performance.
+However, to make comparisons more easy, here we have used the same environmental covariates in the machine learning models and in the Enviro disaggregation regression models.
+
+\section{Conclusions}
+
+Overall we find that including predictions from machine learning models trained on prevalence point-surveys typically improves the performance of disaggregation regression models for malaria incidence.
+Training the machine learning models on local data gave much better performance, but how local the data should be remains an open question.
+As more countries start providing timely and accurate routine surveillance data of malaria cases disaggregation regression may become more popular.
+We have here presented a method for improving these models by using ancillary data.
 

### 2019-10-14 tim-lucas-model-ensembles.tex
Sort authors a bit.

@@
 
-\begin{center}
-{Tim C. D. Lucas*\textsuperscript{1}; Anita Nandi\textsuperscript{1}; Michele Nguyen\textsuperscript{1}; 
-Susan Rumisha\textsuperscript{1}; 
-Katherine E. Battle\textsuperscript{1}; Rosalind E. Howes\textsuperscript{1}; Andre Python\textsuperscript{1}; Penny Hancock\textsuperscript{1}; Punam Amratia\textsuperscript{1};
-Ewan Cameron\textsuperscript{1}; Pete Gething\textsuperscript{1}; Daniel~J.~Weiss\textsuperscript{1}.}\\
-{1. Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK - timcdlucas@gmail.com}\\
+%\begin{center}
+{
+\noindent Tim C. D. Lucas*\textsuperscript{1}; 
+Anita K. Nandi\textsuperscript{1},
+Emma L. Collins\textsuperscript{1},
+Suzanne H. Keddie\textsuperscript{1},
+Katherine A. Twohig\textsuperscript{1},
+Elisabeth G. Chestnutt\textsuperscript{1},
+Rosalind E. Howes\textsuperscript{1},
+Michele Nguyen\textsuperscript{1},
+Susan F. Rumisha\textsuperscript{1},
+Andre Python\textsuperscript{1},
+Rohan Arambepola\textsuperscript{1},
+Amelia Bertozzi-Villa\textsuperscript{1, 2},
+Penelope Hancock\textsuperscript{1},
+Punam Amratia\textsuperscript{1},
+Katherine E. Battle\textsuperscript{1},
+Ewan Cameron\textsuperscript{1},
+Peter W. Gething\textsuperscript{1, 3, 4},
+Daniel J. Weiss\textsuperscript{1}}\vspace{0.4cm}\\
 
-\end{center}
+{
+\noindent 1. Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK \\
+2. Institute for Disease Modeling, Bellevue WA, USA\\
+3. Telethon Kids Institute, Perth Children’s Hospital, Perth, Australia\\
+4. Curtin University, Perth, Australia\\
+* timcdlucas@gmail.com}\\
+
+%\end{center}
 \section{Abstract}
@@
 \hline 
- Colombia & Random &  0.59 & 0.59 & \textbf{0.61} \\
+ Colombia & Random &  \textbf{0.59} & \textbf{0.59} & 0.58 \\
  Colombia &  Spatial &  0.12 &  \textbf{0.33} &  0.25\\
@@
 \hline 
- Colombia & Random &  \textbf{0.59} &0.58 & 0.56 \\
- Colombia &  Spatial &  0.12 &  0.18 & \textbf{0.34}\\
+ Colombia & Random &  \textbf{0.59} &0.58 & 0.55 \\
+ Colombia &  Spatial &  0.12 &  0.12 & \textbf{0.33}\\
  Madagascar &  Random &  0.70 &  0.70 & \textbf{0.71} \\
@@
 \hline 
- Colombia & Random & \textbf{0.28} & \textbf{0.28} & \textbf{0.29} & \textbf{0.29} & \textbf{0.29} \\
+ Colombia & Random & \textbf{0.28} & \textbf{0.28} & \textbf{0.29} & \textbf{0.28} & \textbf{0.30} \\
  Madagascar &  Random & 0.80 & 0.77& 0.75& 0.77& 0.76 \\
@@
  Indonesia & Random &0.80 & 0.81& 0.78& 0.79& 0.77  \\
- Colombia &  Spatial & \textbf{0.31} & \textbf{0.34}  & \textbf{0.32} & \textbf{0.34} & \textbf{0.34}  \\
+ Colombia &  Spatial & \textbf{0.30} & \textbf{0.33}  & \textbf{0.33} & \textbf{0.33} & \textbf{0.34}  \\
  Madagascar & Spatial & \textbf{0.65} & 0.74& 0.70& 0.75 & 0.76   \\
@@
 Madagascar & ppr & 0.109 & 0.469 & 0.205 &  0.403\vspace{0.3cm}\\ 
-Colombia & RF & \textbf{0.068} & 0.505 &  \textbf{0.169} & \textbf{0.522}\\
-Colombia & GBM & 0.073 & \textbf{0.879} & 0.178& -0.231  \\
-Colombia & enet & 0.070 & 0.079 &0.233 & 0.080 \\
-Colombia & nnet & 0.070 & -0.008 &0.220 & 0.144 \\
-Colombia & ppr & 0.070 & 0.530 & 0.205 &  0.468\vspace{0.3cm}\\
+Colombia & RF & \textbf{0.068} & 0.625 &  \textbf{0.169} & \textbf{0.522}\\
+Colombia & GBM & 0.073 & \textbf{0.952} & 0.178& -0.231  \\
+Colombia & enet & 0.070 & 0.219 &0.233 & 0.080 \\
+Colombia & nnet & 0.070 & 0.129 &0.220 & 0.144 \\
+Colombia & ppr & 0.070 & 0.0.667 & 0.205 &  0.468\vspace{0.3cm}\\
 Senegal & RF & \textbf{0.092} & \textbf{0.339} & \textbf{0.169} & \textbf{0.425} \\

### 2019-10-14 tim-lucas-model-ensembles.tex
Switch to numbered refs.

@@
 
-\usepackage{natbib}
-
+\usepackage[numbers, sort&compress]{natbib}
 %\setcitestyle{authoryear}
@@
 %The title should be centred and in bold letters. It should be informative but not too long (preferably no more than two lines).
-\textbf{Model ensembles with different response variables for base and meta models: malaria disaggregation regression combining prevalence and incidence data}
+\textbf{Improving disaggregation models of malaria incidence by ensembling non-linear models of prevalence}
 \end{center}
@@
 \noindent 1. Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK \\
-2. Institute for Disease Modeling, Bellevue WA, USA\\
+2. Institute for Disease Modeling, Bellevue, WA, USA\\
 3. Telethon Kids Institute, Perth Children’s Hospital, Perth, Australia\\
@@
 \section{Abstract}
-
-% This is where the abstract is placed.
-It should include a statement about the problem being addressed in the presentation (and paper, if submitted).
-Continue with a discussion of why it is important to address this problem.
-This may be followed by some summary information about the models and methods developed and/or used to address the problem.
-Conclude with a description of the key results and contributions that will be covered in the presentation (and paper).\\
 
@@
 Many of the differences in performance are rather marginal.
-Overall, in only one case (random cross-validation in Senegal) is Enviro the best performing model while under spatially cross-validation in Senegal Enviro and Envir + ML\textsubscript{l} perform equally well.
+Overall, in only one case (random cross-validation in Senegal) is Enviro the best performing model.
+In two cases, Enviro perfoms equally well as another model: under spatially cross-validation in Senegal Enviro and Envir + ML\textsubscript{l} perform equally well, under random cross-validation in Colombia Enviro and ML\textsubscript{l} perform equally well.
 Of the remaining six cases, in three cases ML\textsubscript{l} performs best and in three cases Envir + ML\textsubscript{l} performs best.
@@
 \bibliography{Malaria} 
-\bibliographystyle{apalike}
+\bibliographystyle{unsrt}
 

### 2019-10-15 tim-lucas-model-ensembles.tex
Add final full colombia data results.

@@
 
-\usepackage[numbers, sort&compress]{natbib}
+\usepackage{natbib}
+%\usepackage[numbers, sort&compress]{natbib}
 %\setcitestyle{authoryear}
@@
 
-Maps of infection rate are one of the core tools needed for the elimination of malaria.
-Reliable routine surveillance data of malaria incidence is becoming more widely available and can better measure low malaria risk than prevalence point-surveys.
-However, these data are often aggregated over large, heterogeneous areas which means that they are often underpowered for learning complex relationships between the environment and malaria risk.
+Maps of infection rate are one of the core tools needed for the control and elimination of malaria.
+As reliable routine surveillance data of malaria incidence becomes more widely available disaggregation regression is becoming an important model framework.
+However, the aggregation of incidence over large, heterogeneous areas means that they are often underpowered for learning complex, non-linear relationships between the environment and malaria risk.
 In contrast, prevalence point-surveys are directly linked to loval environmental conditions but are not common in many areas of the world.
-A model that combines point surveys and aggregated surveillance data could have the benefits of both but must be able to account for the fact that these two data types are different malariometric units.
-Here, we train multiple machine learning models on prevalence point-surveys and then combine the predictions from these with a geostatistical disaggregation model that uses routine surveillance data.
-We find that, in four national case studies, using a disaggregation regression model to combine predictions from machine learning models trained on point surveys improves model accuracy relative to a baseline disaggregation regression model that uses the environmental covariates only.
+A model that combines point surveys and aggregated surveillance data could have the benefits of both but must be able to account for the fact that these two data types measure different malariometric units.
+Here, we train multiple non-linear, machine learning models on prevalence point-surveys and then ensemble the predictions from these with a geostatistical disaggregation model that uses routine surveillance data.
+We find that, in four national case studies, using a disaggregation regression model to combine predictions from machine learning models trained on prevalence point-surveys improves model accuracy relative to a baseline disaggregation regression model that uses linear effects of the environmental covariates only.
 
@@
 In all five sets of machine learning models, Random Forest performs the best.
-While we have not forced $\beta_i > 0$, the priors for these coefficients were positive.
+We have not forced $\beta_i > 0$ but we have set the priors for these coefficients with a positive mean.
 In nearly all cases the fitted values are positive.
 If the prevalence data and incidence data are not biased relative to each other we would expect the models with the lowest RMSE to also have the biggest regression coefficient.
-This occurs in four cases where Random Forest has the lowest RMSE and the biggest coefficient.
+This occurs in three cases where Random Forest has the lowest RMSE and the biggest coefficient.
 In a further two cases, Random Forest has the lowest RMSE but GBM has the highest coefficient.
@@
 Madagascar & ppr & 0.109 & 0.469 & 0.205 &  0.403\vspace{0.3cm}\\ 
-Colombia & RF & \textbf{0.068} & 0.625 &  \textbf{0.169} & \textbf{0.522}\\
-Colombia & GBM & 0.073 & \textbf{0.952} & 0.178& -0.231  \\
-Colombia & enet & 0.070 & 0.219 &0.233 & 0.080 \\
-Colombia & nnet & 0.070 & 0.129 &0.220 & 0.144 \\
-Colombia & ppr & 0.070 & 0.0.667 & 0.205 &  0.468\vspace{0.3cm}\\
+Colombia & RF & \textbf{0.068} & 0.625 &  \textbf{0.169} & 0.180\\
+Colombia & GBM & 0.073 & \textbf{0.952} & 0.178& -0.218  \\
+Colombia & enet & 0.070 & 0.219 &0.233 & 0.183 \\
+Colombia & nnet & 0.070 & 0.129 &0.220 & 0.527 \\
+Colombia & ppr & 0.070 & 0.0.667 & 0.205 &  \textbf{0.546}\vspace{0.3cm}\\
 Senegal & RF & \textbf{0.092} & \textbf{0.339} & \textbf{0.169} & \textbf{0.425} \\
@@
 \end{tabular}
-\label{t:mlresults}
+\label{t:mlresults2}
 \end{table}
@@
 \bibliography{Malaria} 
-\bibliographystyle{unsrt}
+%\bibliographystyle{unsrt}
+\bibliographystyle{apalike}
 

### 2019-10-16 tim-lucas-model-ensembles.tex
Restructure intro.

@@
 
-%We need to switch to downscaling models
-%But they have issues of low power.
-%While prevalence points should not be primary data source, they should still be used.
-%One way is ML.
+% New plan
+
+% We need to do disaggregation but how do we do nonlinear
+
+% We can do machine learning on prevalence and then combine...
+
+% Stacking is a thing and has some nice theory.
+
+% We found that it worked ok.
 
@@
 
+Here we propose training a suite of level-zero machine learning models on point-level, binomial prevalence data and combining these predictions by using them as covariates in a polygon-level, Poisson incidence model.
+This modelling scheme has similarities to previous stacking methods used for malaria mapping \citep{bhatt2017improved}.
+However, as the response data in the level-zero models and the stacker are on different scales (prevalence is a proportion while incidence is a rate) there are a number of important differences.
+Notably, we cannot simply take a weighted average of the predictions from the level-zero models; they need to be transformed to the incidence scale and some additional flexibility to allow for mispecification of this transform needs to be allowed.
+
 Model stacking has proven effective in many realms \citep{sill2009feature, bhatt2017improved, hao2019review}. 
@@
 
-Here we propose training a suite of level-zero machine learning models on point-level, binomial prevalence data and stacking these models with a polygon-level, Poisson incidence model.
-This modelling scheme has similarities to previous stacking methods used for malaria mapping \citep{bhatt2017improved}.
-However, as the response data in the level-zero models and the stacker are on different scales (prevalence is a proportion while incidence is a rate) there are a number of important differences.
-Notably, we cannot simply take a weighted average of the predictions from the level-zero models; they need to be transformed to the incidence scale and some additional flexibility to allow for mispecification of this transform needs to be allowed.
-To test the effectiveness of this approach we use data from four countries: Madagascar, Colombia, Indonesia and Senegal.
+To test the effectiveness of the proposed approach we use data from four countries: Madagascar, Colombia, Indonesia and Senegal.
 We train the machine learning models on both local and global data and examine a number of different combinations of covariates: just local or global machine learning predictions, or combining the two or combining local machine learning predictions with the raw environmental covariates.
@@
 In two cases, Enviro perfoms equally well as another model: under spatially cross-validation in Senegal Enviro and Envir + ML\textsubscript{l} perform equally well, under random cross-validation in Colombia Enviro and ML\textsubscript{l} perform equally well.
-Of the remaining six cases, in three cases ML\textsubscript{l} performs best and in three cases Envir + ML\textsubscript{l} performs best.
-The greatest benefits in predictive performance occur under spatial cross validation and in the cases when the Enviro model is particularly poor.
+Of the remaining five cases, in three cases ML\textsubscript{l} performs best and in two cases (random and spatial cross-validation in Indonesia) Envir + ML\textsubscript{l} performs best.
+The greatest benefits in predictive performance occur under spatial cross-validation and in the cases when the Enviro model is particularly poor.
 

### 2019-10-16 tim-lucas-model-ensembles.tex
More work on intro.

@@
 Here we propose training a suite of level-zero machine learning models on point-level, binomial prevalence data and combining these predictions by using them as covariates in a polygon-level, Poisson incidence model.
-This modelling scheme has similarities to previous stacking methods used for malaria mapping \citep{bhatt2017improved}.
+This modelling scheme has similarities to previous stacking methods used for malaria mapping \citep{bhatt2017improved} and elsewhere \citep{wolpert1992stacked, sill2009feature, hao2019review, breiman1996stacked}.
 However, as the response data in the level-zero models and the stacker are on different scales (prevalence is a proportion while incidence is a rate) there are a number of important differences.
-Notably, we cannot simply take a weighted average of the predictions from the level-zero models; they need to be transformed to the incidence scale and some additional flexibility to allow for mispecification of this transform needs to be allowed.
+Notably, we cannot simply take a weighted average of the predictions from the level-zero models as in a standard stacking scheme \citep{sill2009feature, hao2019review}.
+Instead the predictions  need to be transformed to the incidence scale with a seperately fitted model \citep{cameron2015defining}.
+This scheme is also similar to applications in other disease contexts where data from vector or host species are used to train models, the predictions from which are then used in a final model \citep{pigott2014mapping, shearer2016estimating}.
+In those applications we would always expect to need additional covariates as well the modelled distributions of hosts or vectors while in this application, because both sets of data are direct measures of some aspect of malaria transmission rate, it is possible that we would need no further covariates.
 
-Model stacking has proven effective in many realms \citep{sill2009feature, bhatt2017improved, hao2019review}. 
-Stacking improves predictions by controlling bias and variance; as long as suitably diverse models are averaged, they will have different biases while high variance in models should be averaged out even in relatively similar models.
+Model stacking \citep{wolpert1992stacked} has proven effective in many realms \citep{breiman1996stacked, sill2009feature, bhatt2017improved, hao2019review}. 
+Stacking improves predictions by controlling bias and variance; as long as suitably diverse models are averaged, they will have different biases while high variance in models should be averaged out.
 This understanding of how stacking improves model performance indicates that diversity in models is important for stacking to be effective.
-Diversity in models is typically created in two ways: by using diverse training datasets (as in Random Forests for example) and by using functionally different models (for example by averaging tree based models and neural networks).
+Diversity in models is typically created in two ways: by using diverse training datasets\citep{breiman1996bagging} (as in Random Forests for example) and by using functionally different models (for example by averaging tree based models and neural networks) \citep{breiman1996stacked}.
 One important trade-off in spatial modelling is whether to use local data (with a smaller sample size but that is likely to be representative of the area of study) or global data that have a larger sample size but a less close association with the areas of study.
-In the context of malaria mapping, we can think about diversity of training data in this context and expect that stacking seperate models trained on local and global data will also increase the diversity of predictions in a useful way.
+For the application of malaria mapping, we can think about diversity of training data in this context and expect that stacking seperate models trained on local and global data will also increase the diversity of predictions in a useful way.
 
@@
 Finally, we set different priors on the regression coefficients depending on which covariates were used.
-When environmental covariates or a mix of environmental covariates and predictions from machine learning models were used we set the prior to be weakly regularising, $\beta_i \sim \operatorname{ Norm}(0, 0.4)$, such that it is unlikely that any single covariate could explain the full range of the response data.
+When environmental covariates or a mix of environmental covariates and predictions from machine learning models were used we set the prior to be weakly regularising, $\beta_i \sim \operatorname{Norm}(0, 0.4)$, such that it is unlikely that any single covariate could explain the full range of the response data.
 When only machine learning model predictions were used we set $\beta_i \sim \operatorname{ Norm}(\frac{1}{M}, 0.4)$ where $M$ is the number of machine learning models being used. 

### 2019-10-23 tim-lucas-model-ensembles.tex
Get elsevier template working.

@@
-\documentclass[11pt]{article}
-\fontfamily{times}
+\documentclass[review]{elsarticle}
+
 \usepackage{graphicx}
 \usepackage{amsmath}
-\usepackage{geometry}
+
 \usepackage{subcaption}
+
+\usepackage{lineno,hyperref}
+\modulolinenumbers[5]
+\bibliographystyle{elsarticle-num}
+%\usepackage{numcompress}\bibliographystyle{model3-num-names}
 
@@
 %\usepackage[numbers, sort&compress]{natbib}
-%\setcitestyle{authoryear}
+%\setcitestyle{author[oxford]year}
 
-\geometry{verbose,tmargin=30mm,bmargin=25mm,lmargin=25mm,rmargin=25mm}
+%\geometry{verbose,tmargin=30mm,bmargin=25mm,lmargin=25mm,rmargin=25mm}
 %\input{figures.tex}
-\pagestyle{empty}
+%\pagestyle{empty}
 \begin{document}
@@
 
-\small{
-
-\begin{center}
-%The title should be centred and in bold letters. It should be informative but not too long (preferably no more than two lines).
-\textbf{Improving disaggregation models of malaria incidence by ensembling non-linear models of prevalence}
-\end{center}
-
 % Model ensembles with different response variables for base and meta models:
 
-%\begin{center}
-{
-\noindent Tim C. D. Lucas*\textsuperscript{1}; 
-Anita K. Nandi\textsuperscript{1},
-Emma L. Collins\textsuperscript{1},
-Suzanne H. Keddie\textsuperscript{1},
-Katherine A. Twohig\textsuperscript{1},
-Elisabeth G. Chestnutt\textsuperscript{1},
-Rosalind E. Howes\textsuperscript{1},
-Michele Nguyen\textsuperscript{1},
-Susan F. Rumisha\textsuperscript{1},
-Andre Python\textsuperscript{1},
-Rohan Arambepola\textsuperscript{1},
-Amelia Bertozzi-Villa\textsuperscript{1, 2},
-Penelope Hancock\textsuperscript{1},
-Punam Amratia\textsuperscript{1},
-Katherine E. Battle\textsuperscript{1},
-Ewan Cameron\textsuperscript{1},
-Peter W. Gething\textsuperscript{1, 3, 4},
-Daniel J. Weiss\textsuperscript{1}}\vspace{0.4cm}\\
+% Spatial and Spatio-temporal Epidemiology
+% Special issue: Disease Surveillance and Infectious disease Modeling
 
-{
-\noindent 1. Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK \\
-2. Institute for Disease Modeling, Bellevue, WA, USA\\
-3. Telethon Kids Institute, Perth Children’s Hospital, Perth, Australia\\
-4. Curtin University, Perth, Australia\\
-* timcdlucas@gmail.com}\\
+\begin{frontmatter}
 
-%\end{center}
-\section{Abstract}
+\title{Improving disaggregation models of malaria incidence by ensembling non-linear models of prevalence}
+
+\author[oxford]{Tim C. D. Lucas\corref{mycorrespondingauthor}}
+\cortext[mycorrespondingauthor]{Corresponding author}
+\ead{timcdlucas@gmail.com}
+\author[oxford]{Anita K. Nandi}
+\author[oxford]{Emma L. Collins}
+\author[oxford]{Suzanne H. Keddie}
+\author[oxford]{Katherine A. Twohig}
+\author[oxford]{Elisabeth G. Chestnutt}
+\author[oxford]{Rosalind E. Howes}
+\author[oxford]{Michele Nguyen}
+\author[oxford]{Susan F. Rumisha}
+\author[oxford]{Andre Python}
+\author[oxford]{Rohan Arambepola}
+\author[oxford,idm]{Amelia Bertozzi-Villa}
+\author[oxford]{Penelope Hancock}
+\author[oxford]{Punam Amratia}
+\author[oxford]{Katherine E. Battle}
+\author[oxford]{Ewan Cameron}
+\author[oxford,telethon,curtin]{Peter W. Gething}
+\author[oxford]{Daniel J. Weiss}
+
+\address[oxford]{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK}
+\address[idm]{Institute for Disease Modeling, Bellevue, WA, USA}
+\address[telethon]{Telethon Kids Institute, Perth Children’s Hospital, Perth, Australia}
+\address[curtin]{Curtin University, Perth, Australia}
+
+\begin{abstract}
 
@@
 
-{\bf Keywords}: Spatial statistics; Ensemble; Stacking; Epidemiology.
-}\\
+\end{abstract}
+
+\begin{keyword}
+Spatial statistics; Disaggregation regression; Stacking; Surveillance data
+\end{keyword}
+
+\end{frontmatter}
+
+\linenumbers
 
@@
 \section*{Acknowledgments}
-The authors acknowledge the National Malaria Control Programme of Madagascar for sharing their routine case data for this analysis.
+The author[oxford]s acknowledge the National Malaria Control Programme of Madagascar for sharing their routine case data for this analysis.
 
@@
 %\bibliographystyle{unsrt}
-\bibliographystyle{apalike}
+%\bibliographystyle{apalike}
 

### 2019-10-23 tim-lucas-model-ensembles.tex
smaller intro edits.

@@
 
-Here we propose training a suite of level-zero machine learning models on point-level, binomial prevalence data and combining these predictions by using them as covariates in a polygon-level, Poisson incidence model.
+Here we propose training a suite of level-zero machine learning models on point-level, binomial prevalence data and combining these predictions by using them as covariates in a level-one, polygon-level, Poisson incidence model.
 This modelling scheme has similarities to previous stacking methods used for malaria mapping \citep{bhatt2017improved} and elsewhere \citep{wolpert1992stacked, sill2009feature, hao2019review, breiman1996stacked}.
-However, as the response data in the level-zero models and the stacker are on different scales (prevalence is a proportion while incidence is a rate) there are a number of important differences.
+However, as the response data in the level-zero models and the level-one models are on different scales (prevalence is a proportion while incidence is a rate) there are a number of important differences.
 Notably, we cannot simply take a weighted average of the predictions from the level-zero models as in a standard stacking scheme \citep{sill2009feature, hao2019review}.
 Instead the predictions  need to be transformed to the incidence scale with a seperately fitted model \citep{cameron2015defining}.
-This scheme is also similar to applications in other disease contexts where data from vector or host species are used to train models, the predictions from which are then used in a final model \citep{pigott2014mapping, shearer2016estimating}.
-In those applications we would always expect to need additional covariates as well the modelled distributions of hosts or vectors while in this application, because both sets of data are direct measures of some aspect of malaria transmission rate, it is possible that we would need no further covariates.
+Applications in other disease contexts have used a similar scheme where data from vector or wild-animal host species are used to train models, the predictions from which are then used in a final model \citep{pigott2014mapping, shearer2016estimating}.
+In such applications we would always expect to need additional covariates in additional to the modelled distributions of hosts or vectors.
+However, in the case examined here, both sets of data are direct measures of some aspect of malaria transmission rate, and therefore it is possible that we would need no further covariates.
 

### 2019-10-24 tim-lucas-model-ensembles.tex
More messing around with results.

@@
 
+\subsection{Malaria Data}
+
 We used two data sources that reflect \emph{Plasmodium falciparum} malaria transmission; point-prevalence surveys and polygon-level, aggregated incidence data. 
@@
 The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
+This data was used as both a global dataset and as regional subsets.
+The global dataset contained 55,914 surveys in 44,842 distinct locations and represented samples from 5,687,304 individuals.
+For Colombia we used all points from South America (locations = 522, individuals = 7,719) while for the other countries we used only data from that country (Madagascar: locations = 1,505, individuals = 89,381.
+Indonesia: locations = 4,778, individuals = 1,512,888.
+Senegal: locations = 1,762, individuals = 80,896).
+We chose these geographic regions based on a trade-off between wanting a large sample size but wanting data from geographically relevant areas.
 
@@
 
+\subsection{Covariates}
+
 We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \citep{weiss2015re, weiss2018global}.
@@
 
+\subsection{Machine learning models}
 For each country we fitted 5 models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
-These models were fitted to both the full malaria prevalence dataset and to more regional subsets of the data.
-For Colombia we used all points from South America (locations = 522, individuals = 7,719) while for the other countries we used only data from that country (Madagascar: n = 1,505, individuals = 89,381.
-Indonesia: n = 4,778, individuals = 1,512,888.
-Senegal: n = 1,762, individuals = 80,896).
-We chose these geographic regions based on a trade-off between wanting a large sample size but wanting data from geographically similar areas.
-Our response variable was prevalence and we weighted the data by sample size (i.e.\thinspace the number of people tested for malaria in each survey.
+These models were fitted to both the full malaria prevalence dataset and to the regional subsets of the data.
+Our response variable was prevalence and we weighted the data by sample size (i.e.\thinspace the number of people tested for malaria in each survey).
 For each model we ran  five-fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
@@
 
+\subsection{Disaggregation regression}
 The top level model was a disaggregation regression model \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
@@
 
+\subsection{Experiments}
 We compared the performance of the models with five sets of covariates, $X$.
@@
 
-\section{Results}
-
 \begin{table}[t!]
@@
 \hline 
- Colombia & Random &  \textbf{0.59} & \textbf{0.59} & 0.58 \\
+ Colombia & Random &  0.59 & 0.59 & \textbf{0.61} \\
  Colombia &  Spatial &  0.12 &  \textbf{0.33} &  0.25\\
- Madagascar & Random &   0.70 &  \textbf{0.73} & 0.69 \\
- Madagascar &  Spatial &  0.22 &  \textbf{0.66} & 0.18\\
+ Madagascar & Random &   \textbf{0.70} &  0.68 & 0.69 \\
+ Madagascar &  Spatial &  0.22 &  \textbf{0.55} & 0.18\\
  Senegal &  Random &  \textbf{0.58} &  0.51 & 0.57 \\
- Senegal &  Spatial &  \textbf{0.63} &  0.27 &  \textbf{0.63} \\
- Indonesia &  Random &  0.52 &  0.48 &  \textbf{0.59} \\
- Indonesia &  Spatial &  0.45 &  0.44 &  \textbf{0.51} \\
+ Senegal &  Spatial &  \textbf{0.63} &  0.51 &  \textbf{0.63} \\
+ Indonesia &  Random &  0.58 &  0.48 &  \textbf{0.59} \\
+ Indonesia &  Spatial &  \textbf{0.63} &  0.44 &  0.51 \\
 \end{tabular}
@@
 \end{table}
+
+\section{Results}
 
@@
 
+\begin{figure}
+  \includegraphics[width=\textwidth]{figs/cv1_l_scatter.png}
+\caption{
+  Observed data against predictions for random cross-validation hold-out samples on a square root transformed scale.
+}
+\label{f:scatter1}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{figs/cv2_l_scatter.png}
+\caption{
+  Observed data against predictions for spatial cross-validation hold-out samples on a square root transformed scale.
+}
+\label{f:scatter2}
+\end{figure}
+
 Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatially cross-validation) for models that use environmental covariates only, global machine learning models only and a combination of the local and global machine learning models.
@@
 
+\begin{table}[t!]
+\caption{Pearson correlations between observed and predicted values. }
+\centering
+\begin{tabular}{llrrrr}
+Country &  Cross-validation & Enviro &   ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
+\hline 
+ Colombia & Random &  \textbf{0.59} &0.55 & 0.58 \\
+ Colombia &  Spatial &  0.12 &  \textbf{0.33} & 0.25\\
+ Madagascar &  Random &  \textbf{0.70} &  0.67 & 0.68 \\
+ Madagascar &  Spatial &  0.22 &  \textbf{0.55} & 0.18\\
+ Senegal &  Random &  \textbf{0.58} &  0.50 & 0.49 \\
+ Senegal &  Spatial &  \textbf{0.63} & 0.51 & \textbf{0.63} \\
+ Indonesia &  Random &  \textbf{0.52} & 0.32 & 0.46 \\
+ Indonesia &  Spatial &  \textbf{0.63} & 0.44 & 0.51 \\
+\end{tabular}
+\label{t:results2}
+\end{table}
 Figures~\ref{f:scatter1} shows scatter plots of the model performance under random cross-validation for all countries and all models.
@@
 
-\begin{table}[t!]
-\caption{Pearson correlations between observed and predicted values. }
-\centering
-\begin{tabular}{llrrrr}
-Country &  Cross-validation & Enviro &   ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
-\hline 
- Colombia & Random &  \textbf{0.59} &0.58 & 0.55 \\
- Colombia &  Spatial &  0.12 &  0.12 & \textbf{0.33}\\
- Madagascar &  Random &  0.70 &  0.70 & \textbf{0.71} \\
- Madagascar &  Spatial &  0.22 &  \textbf{0.63} & 0.60\\
- Senegal &  Random &  \textbf{0.58} &  0.30 & 0.08 \\
- Senegal &  Spatial &  \textbf{0.63} & 0.36 & 0.35 \\
- Indonesia &  Random &  \textbf{0.52} & 0.46 & 0.32 \\
- Indonesia &  Spatial &  \textbf{0.45} & 0.41 & \textbf{0.45} \\
-\end{tabular}
-\label{t:results2}
-\end{table}
-
 %\begin{table}[h!]
@@
 % to do table of coverage.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{figs/cv1_scatter.png}
-\caption{
-  Observed data against predictions for cross-validation hold-out samples on a square root transformed scale.
-}
-\label{f:scatter1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{figs/cv2_scatter.png}
-\caption{
-  Observed data against predictions for cross-validation hold-out samples on a square root transformed scale.
-}
-\label{f:scatter2}
-\end{figure}
 
@@
 \section*{Acknowledgments}
-The author[oxford]s acknowledge the National Malaria Control Programme of Madagascar for sharing their routine case data for this analysis.
+The authors acknowledge the National Malaria Control Programme of Madagascar for sharing their routine case data for this analysis.
 

### 2019-10-24 tim-lucas-model-ensembles.tex
Comment out big table.

@@
 
-\begin{table}
-\caption{Fitted regression parameters of the environmental and machine learning covariates for thee models: just environmental variables, just local machine learning models and the combination of the two.}
-\centering
-\tiny
-\begin{tabular}{ll|rrr}
-Country    & Variable &  $\beta_{enviro}$   & $\beta_l$ & $\beta_{enviro+l}$\\ \hline
-Madagascar & LST mean & 0.17 &  &  -0.01\\
-Madagascar & EVI & 0.32 &  & -0.04 \\
-Madagascar & TSI & 0.42 &  &0.06  \\
-Madagascar & access & 0.34 &  &0.00  \\
-Madagascar & Elev & -0.14 &  & 0.00\\ 
-Madagascar & LST sd & -0.78 &  &-0.04  \\
-Madagascar & VIIRS & -0.1 &  &-0.52  \\
-Madagascar & TCW & -0.22 &  & -0.06\\ 
-Madagascar & RF &  & 0.29 &  0.01\\
-Madagascar & GBM &  & 0.52 & 0.25 \\
-Madagascar & enet &  & 0.31 &0.09  \\
-Madagascar & nnet &  & 0.43 &0.14  \\
-Madagascar & ppr &  & 0.49 & 0.22\vspace{0.2cm}\\ 
-Colombia & LST mean &-0.12  &  & -0.22\\
-Colombia & EVI &  -0.08&  & 0.17\\
-Colombia & TSI &  0.54&  & -0.06\\
-Colombia & access &  0.35&  &0.00 \\
-Colombia & Elev &  0.47&  & 0.00\\
-Colombia & LST sd & -0.11 &  & -0.10\\
-Colombia & VIIRS & 0.24 &  & 0.01\\
-Colombia & TCW &  0.21&  & 0.00\\
-Colombia & RF &  & 0.36 & -0.18\\
-Colombia & GBM &  & 0.54 & 0.26\\
-Colombia & enet &  & 0.32 & -0.34\\
-Colombia & nnet &  &0.33  & -0.09\\
-Colombia & ppr &  & 0.54 & 0.01\vspace{0.2cm}\\
-Senegal & LST mean &-0.04  &  &-0.02\\
-Senegal & EVI &  0.33&  & 0.30\\
-Senegal & TSI &  -0.58&  &  -0.59\\
-Senegal & access & -0.03 &  &  -0.05\\
-Senegal & Elev &  0.42&  &0.37\\ 
-Senegal & LST sd &  0.21&  & 0.10 \\
-Senegal & VIIRS &  0.13&  &  0.13\\
-Senegal & TCW & 0.08 &  &-0.01\\ 
-Senegal & RF &  & 0.34 &  0.20\\
-Senegal & GBM &  &  0.28& 0.05\\
-Senegal & enet &  &  0.27&  -0.05\\
-Senegal & nnet &  & 0.29 &-0.09 \\
-Senegal & ppr &  &  0.35& 0.25\vspace{0.2cm}\\
-Indonesia & LST mean & -0.32 &  &-0.34\\
-Indonesia & EVI &  0.49&  & 0.36\\
-Indonesia & TSI &  0.58&  &0.52\\
-Indonesia & access &0.33 &  & 0.06 \\
-Indonesia & Elev& -0.06&  &-0.13\\ 
-Indonesia & LST sd & -0.23 &  & -0.33\\
-Indonesia & VIIRS &  -0.12&  & 0.03\\
-Indonesia & TCW& 0.12 &  &0.10\\ 
-Indonesia & RF &  & 0.32 & 0.06\\
-Indonesia & GBM &  &  0.38&0.03\\
-Indonesia & enet &  &  0.49&  0.56\\
-Indonesia & nnet &  &  0.36&  0.18\\
-Indonesia & ppr &  &  0.45& 0.32\\
-
-\end{tabular}
-\label{t:mlresults2}
-\end{table}
+%\begin{table}
+%\caption{Fitted regression parameters of the environmental and machine learning covariates for thee models: just environmental variables, just local machine learning models and the combination of the two.}
+%\centering
+%\tiny
+%\begin{tabular}{ll|rrr}
+%Country    & Variable &  $\beta_{enviro}$   & $\beta_l$ & $\beta_{enviro+l}$\\ \hline
+%Madagascar & LST mean & 0.17 &  &  -0.01\\
+%Madagascar & EVI & 0.32 &  & -0.04 \\
+%Madagascar & TSI & 0.42 &  &0.06  \\
+%Madagascar & access & 0.34 &  &0.00  \\
+%Madagascar & Elev & -0.14 &  & 0.00\\ 
+%Madagascar & LST sd & -0.78 &  &-0.04  \\
+%Madagascar & VIIRS & -0.1 &  &-0.52  \\
+%Madagascar & TCW & -0.22 &  & -0.06\\ 
+%Madagascar & RF &  & 0.29 &  0.01\\
+%Madagascar & GBM &  & 0.52 & 0.25 \\
+%Madagascar & enet &  & 0.31 &0.09  \\
+%Madagascar & nnet &  & 0.43 &0.14  \\
+%Madagascar & ppr &  & 0.49 & 0.22\vspace{0.2cm}\\ 
+%Colombia & LST mean &-0.12  &  & -0.22\\
+%Colombia & EVI &  -0.08&  & 0.17\\
+%Colombia & TSI &  0.54&  & -0.06\\
+%Colombia & access &  0.35&  &0.00 \\
+%Colombia & Elev &  0.47&  & 0.00\\
+%Colombia & LST sd & -0.11 &  & -0.10\\
+%Colombia & VIIRS & 0.24 &  & 0.01\\
+%Colombia & TCW &  0.21&  & 0.00\\
+%Colombia & RF &  & 0.36 & -0.18\\
+%Colombia & GBM &  & 0.54 & 0.26\\
+%Colombia & enet &  & 0.32 & -0.34\\
+%Colombia & nnet &  &0.33  & -0.09\\
+%Colombia & ppr &  & 0.54 & 0.01\vspace{0.2cm}\\
+%Senegal & LST mean &-0.04  &  &-0.02\\
+%Senegal & EVI &  0.33&  & 0.30\\
+%Senegal & TSI &  -0.58&  &  -0.59\\
+%Senegal & access & -0.03 &  &  -0.05\\
+%Senegal & Elev &  0.42&  &0.37\\ 
+%Senegal & LST sd &  0.21&  & 0.10 \\
+%Senegal & VIIRS &  0.13&  &  0.13\\
+%Senegal & TCW & 0.08 &  &-0.01\\ 
+%Senegal & RF &  & 0.34 &  0.20\\
+%Senegal & GBM &  &  0.28& 0.05\\
+%Senegal & enet &  &  0.27&  -0.05\\
+%Senegal & nnet &  & 0.29 &-0.09 \\
+%Senegal & ppr &  &  0.35& 0.25\vspace{0.2cm}\\
+%Indonesia & LST mean & -0.32 &  &-0.34\\
+%Indonesia & EVI &  0.49&  & 0.36\\
+%Indonesia & TSI &  0.58&  &0.52\\
+%Indonesia & access &0.33 &  & 0.06 \\
+%Indonesia & Elev& -0.06&  &-0.13\\ 
+%Indonesia & LST sd & -0.23 &  & -0.33\\
+%Indonesia & VIIRS &  -0.12&  & 0.03\\
+%Indonesia & TCW& 0.12 &  &0.10\\ 
+%Indonesia & RF &  & 0.32 & 0.06\\
+%Indonesia & GBM &  &  0.38&0.03\\
+%Indonesia & enet &  &  0.49&  0.56\\
+%Indonesia & nnet &  &  0.36&  0.18\\
+%Indonesia & ppr &  &  0.45& 0.32\\
+%
+%
+%
+%\end{tabular}
+%\label{t:mlresults2}
+%\end{table}
+%
 

### 2019-10-25 tim-lucas-model-ensembles.tex
New figures and things.

@@
  Senegal &  Random &  \textbf{0.58} &  0.51 & 0.57 \\
- Senegal &  Spatial &  \textbf{0.63} &  0.51 &  \textbf{0.63} \\
- Indonesia &  Random &  0.58 &  0.48 &  \textbf{0.59} \\
- Indonesia &  Spatial &  \textbf{0.63} &  0.44 &  0.51 \\
+ Senegal &  Spatial &  0.63 &  0.51 &  0.63 \\
+ Indonesia &  Random &  0.52 &  0.48 &  \textbf{0.59} \\
+ Indonesia &  Spatial &  0.45 &  0.44 &  \textbf{0.51} \\
 \end{tabular}
@@
  Colombia & Random &  \textbf{0.59} &0.55 & 0.58 \\
- Colombia &  Spatial &  0.12 &  \textbf{0.33} & 0.25\\
+ Colombia &  Spatial &  0.12 &  0.12 & \textbf{0.33}\\
  Madagascar &  Random &  \textbf{0.70} &  0.67 & 0.68 \\
- Madagascar &  Spatial &  0.22 &  \textbf{0.55} & 0.18\\
+ Madagascar &  Spatial &  0.22 &  0.51 & \textbf{0.55}\\
  Senegal &  Random &  \textbf{0.58} &  0.50 & 0.49 \\
- Senegal &  Spatial &  \textbf{0.63} & 0.51 & \textbf{0.63} \\
+ Senegal &  Spatial &   \textbf{0.63} & 0.55 & 0.52 \\
  Indonesia &  Random &  \textbf{0.52} & 0.32 & 0.46 \\
- Indonesia &  Spatial &  \textbf{0.63} & 0.44 & 0.51 \\
+ Indonesia &  Spatial &  0.45 & 0.41 & 0.45 \\
 \end{tabular}
@@
 Otherwise the coverage is generally acceptable with most values lying between 0.7 and 0.9.
-Over there is a general tendancy for models to be slightly over confident.
+Overall there is a general tendancy for models to be slightly over confident.
 
@@
 \centering
-\includegraphics[trim={0 30mm 0 40mm}, width = 0.8\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
+\includegraphics[width = 1\textwidth]{figs/col_comparison_map.png} %\caption{Indonesia spatial crossvalidation} 
 \caption{
   Left: Observed data for Colombia (grey for zero incidence).
-Right: Out-of-sample predictions for the random cross-validation, machine learning only model.
+Middle: Out-of-sample predictions for the spatial cross-validation, environmental covariates only model.
+Right: Out-of-sample predictions for the spatial cross-validation, local machine learning only model.
 For each cross-validation fold, predictions are made for the held out data which are then combined to make a single surface.
@@
 
-\begin{table}[h!]
+\begin{table}[t!]
 \caption{Covarage of 80\% credible intervals. Values outside 0.7--0.9 are shown in bold.}
@@
 
-\begin{table}
+\begin{table}[t!]
 \caption{Machine learning model results and fitted parameters (i.e.\thinspace model weights) of the machine learning predictions only models (local only and global only). }
@@
 Colombia & nnet & 0.070 & 0.129 &0.220 & 0.527 \\
-Colombia & ppr & 0.070 & 0.0.667 & 0.205 &  \textbf{0.546}\vspace{0.3cm}\\
+Colombia & ppr & 0.070 & 0.667 & 0.205 &  \textbf{0.546}\vspace{0.3cm}\\
 Senegal & RF & \textbf{0.092} & \textbf{0.339} & \textbf{0.169} & \textbf{0.425} \\

### 2019-11-05 tim-lucas-model-ensembles.tex
Try again at experimental design.

@@
 
-To test the effectiveness of the proposed approach we use data from four countries: Madagascar, Colombia, Indonesia and Senegal.
-We train the machine learning models on both local and global data and examine a number of different combinations of covariates: just local or global machine learning predictions, or combining the two or combining local machine learning predictions with the raw environmental covariates.
-While there isn't one consistently best model, we find that in nearly all cases, using some information from local machine learning models improves model performance.
+To test the effectiveness of the proposed approach we used data from four countries: Madagascar, Colombia, Indonesia and Senegal.
+We first examined whether using predictions from machine learning models trained on local prevalence data could improve predictions.
+We then tested whether using predictions from machine learning models trained on global data could improve predictions either on their own or when combined with predictions from machine learning models trained on local data.
+While there was not one consistently best model, we found that in nearly all cases, using predictions from local machine learning models improved model performance.
+In contrast, using predictions from machine learning models trained on the global prevalence dataset rarely improved predictive performance.
 
@@
 The global dataset contained 55,914 surveys in 44,842 distinct locations and represented samples from 5,687,304 individuals.
-For Colombia we used all points from South America (locations = 522, individuals = 7,719) while for the other countries we used only data from that country (Madagascar: locations = 1,505, individuals = 89,381.
-Indonesia: locations = 4,778, individuals = 1,512,888.
-Senegal: locations = 1,762, individuals = 80,896).
+For Colombia we used all points from South America (7,719 individuals from 522 locations) while for the other countries we used only data from that country (Madagascar: 89,381 individuals from 1,505 locations.
+Indonesia: 1,512,888 individuals from 4,778 locations.
+Senegal: 80,896 individuals from 1,762 locations).
 We chose these geographic regions based on a trade-off between wanting a large sample size but wanting data from geographically relevant areas.
@@
 \subsection{Machine learning models}
-For each country we fitted 5 models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
+For each country specific dataset and for the global dataset we fitted 5 models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
 These models were fitted to both the full malaria prevalence dataset and to the regional subsets of the data.
@@
 Predictions from these models were then made across each country respectively.
-These predictions were finally empirical logit transformed so that they are on the linear predictor scale of the top level model.
+These predictions were finally empirical logit transformed so that they were on the linear predictor scale of the top level model.
 An empirical logit was used rather than a standard logit as there were many predictions of exactly zero.
@@
 \subsection{Experiments}
-We compared the performance of the models with five sets of covariates, $X$.
-Firstly, we used the environmental and anthropogenic covariates, centered and standardised as a baseline model (subsequantly called Enviro).
-Secondly, we used the predictions from the machine learning models trained on local data (ML\textsubscript{l}).
-Thirdly, we combined the predictions from the machine learning models and the environmental coviarates in one model (Enviro + ML\textsubscript{l}).
-Fourthly, we used predictions from the machine learning models trained on the global data (ML\textsubscript{g}).
-Finally, we combined predictions from the machine learning models trained on regional data and global data (ML\textsubscript{l} + ML\textsubscript{g}).
 
-To compare the five models we used two cross-validation schemes. 
+We used the environmental and anthropogenic covariates, centered and standardised, as a baseline model (subsequantly called Enviro).
+We then performed two experiments.
+In the first (\emph{i}) we tested whether using predictions from machine learning models trained on local prevalence data improved predictions.
+In the second (\emph{ii}) we tested whether using predictions from machine learning models trained on global prevalence data improved predictions.
+Our primary performance metric was correlation between observed and predicted data.
+We also examined the calibration of models by calculating the proportion of held out data that were within their 80\% credible intervals.
+
+In experiment \emph{i} we compared the baseline to two models that used locally trained machine learning models.
+First we used only the predictions from the machine learning models trained on local data (ML\textsubscript{l}).
+Secondly, we combined the predictions from the machine learning models and the environmental coviarates in one model (Enviro + ML\textsubscript{l}).
+
+In experiment \emph{ii} we compared the baseline to two models that used globally trained machine learning models.
+We used only predictions from the machine learning models trained on the global data (ML\textsubscript{g}).
+We also combined predictions from the machine learning models trained on regional data and global data (ML\textsubscript{l} + ML\textsubscript{g}).
+
+In each experiment we used two cross-validation schemes. 
 In the first, polygon incidence data was randomly split into six cross-validation folds.
@@
 This spatial cross-validation scheme is testing the models' ability to make predictions far from data where the spatial random field is not informative.
-Our primary performance metric was correlation between observed and predicted data.
-We also examined the calibration of models by calculating the proportion of held out data that were within their 80\% credible intervals.
 
@@
 In a further two cases, Random Forest has the lowest RMSE but GBM has the highest coefficient.
-The predictions from Random Forest and GBM are highly correlated in part because they both perform well and because they are both tree based models.
+The predictions from Random Forest and GBM are highly correlated in part because they both perform well and in part because they are both tree based models.
 Finally, it can be seen that relationship between RMSE and regression coefficients is much weaker in Indonesia.

### 2019-11-07 tim-lucas-model-ensembles.tex
Rework results and discussion to be experiment i and ii

@@
 In experiment \emph{i} we compared the baseline to two models that used locally trained machine learning models.
-First we used only the predictions from the machine learning models trained on local data (ML\textsubscript{l}).
+First we used only the predictions from the machine learning models trained on local data (subsequantly called ML\textsubscript{l}).
 Secondly, we combined the predictions from the machine learning models and the environmental coviarates in one model (Enviro + ML\textsubscript{l}).
@@
 \begin{tabular}{llrrrr}
-Country &  Cross-validation & Enviro &  ML\textsubscript{l} &  Enviro + ML\textsubscript{l} \\
+Country &  Cross-validation & Enviro &  Enviro + ML\textsubscript{l} &  ML\textsubscript{l} \\
 \hline 
- Colombia & Random &  0.59 & 0.59 & \textbf{0.61} \\
- Colombia &  Spatial &  0.12 &  \textbf{0.33} &  0.25\\
- Madagascar & Random &   \textbf{0.70} &  0.68 & 0.69 \\
- Madagascar &  Spatial &  0.22 &  \textbf{0.55} & 0.18\\
- Senegal &  Random &  \textbf{0.58} &  0.51 & 0.57 \\
- Senegal &  Spatial &  0.63 &  0.51 &  0.63 \\
- Indonesia &  Random &  0.52 &  0.48 &  \textbf{0.59} \\
- Indonesia &  Spatial &  0.45 &  0.44 &  \textbf{0.51} \\
+ Colombia & Random &  0.59 & \textbf{0.61}& 0.59  \\
+ Colombia &  Spatial &  0.12 &  0.25&  \textbf{0.33} \\
+ Indonesia &  Random &  0.52 &  \textbf{0.59}&  0.48  \\
+ Indonesia &  Spatial &  0.45 &  \textbf{0.51}&  0.44  \\
+ Madagascar & Random &   \textbf{0.70}& 0.69 &  0.68  \\
+ Madagascar &  Spatial &  0.22 & 0.18&  \textbf{0.55} \\
+ Senegal &  Random &  \textbf{0.58} & 0.57&  0.51  \\
+ Senegal &  Spatial &  0.63 &  0.63&  0.51  \\
 \end{tabular}
@@
 
-Table \ref{t:results} gives the correlation between observed and held out data (under random and spatially cross-validation) for models that use environmental covariates only, local machine learning models only and a combination of the two.
+Table \ref{t:results} gives the correlation between observed and held out data (under random and spatially cross-validation) for experiment \emph{i} (models Enviro, ML\textsubscript{l} and Envir + ML\textsubscript{l}).
 Many of the differences in performance are rather marginal.
-Overall, in only one case (random cross-validation in Senegal) is Enviro the best performing model.
-In two cases, Enviro perfoms equally well as another model: under spatially cross-validation in Senegal Enviro and Envir + ML\textsubscript{l} perform equally well, under random cross-validation in Colombia Enviro and ML\textsubscript{l} perform equally well.
-Of the remaining five cases, in three cases ML\textsubscript{l} performs best and in two cases (random and spatial cross-validation in Indonesia) Envir + ML\textsubscript{l} performs best.
+Enviro is the best performing model in two cases (random cross-validation in Madagascar and Senegal).
+In one case, Enviro perfoms equally well as another model; under spatial cross-validation in Senegal Enviro and Enviro + ML\textsubscript{l} perform equally well.
+Of the remaining five cases, in two cases ML\textsubscript{l} performs best and in three cases Envir + ML\textsubscript{l} performs best.
 The greatest benefits in predictive performance occur under spatial cross-validation and in the cases when the Enviro model is particularly poor.
@@
 
-Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatially cross-validation) for models that use environmental covariates only, global machine learning models only and a combination of the local and global machine learning models.
-In five cases, Enviro is the best or tied best performing model.
-In only one case in the ML\textsubscript{g} model the best.
-Comparing across Tables \ref{t:results} and \ref{t:results2} we can see that ML\textsubscript{g} is never the best performing model.
-In only three cases (Spatial Colombia, Spatial Senegal and Spatial Indonesia) does ML\textsubscript{g} outperform ML\textsubscript{l}.
+Figures~\ref{f:scatter1} shows scatter plots of the model performance under random cross-validation for experiment \emph{i}.
+Figures~\ref{f:scatter2} shows scatter plots of the model performance under spatial cross-validation for experiment \emph{i}.
+It can be seen that without environmental covariates, the models in Madagascar fail to predict very high or very low values correctly.
+Figure~\ref{f:map} shows the input data and spatially out-of-sample predictions of the Enviro model and ML\textsubscript{l} model in Colombia.
 
-\begin{table}[t!]
-\caption{Pearson correlations between observed and predicted values. }
+Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatially cross-validation) for experiment \emph{ii} (models Enviro, ML\textsubscript{g} and ML\textsubscript{l} + ML\textsubscript{g}).
+In six cases, Enviro is the best or tied best performing model.
+The ML\textsubscript{g} model is never the best performing model and only outpeforms Enviro in one case (spatial cross-validation in Madagascar).
+In two cases ML\textsubscript{l} + ML\textsubscript{g} is the best performing model (spatial Colombia and spatial Madagascar).
+Comparing across Tables \ref{t:results} and \ref{t:results2} we can see that ML\textsubscript{g} outpeforms ML\textsubscript{l} once (spatial cross-validation in Senegal).
+In only two cases (Spatial Senegal and Spatial Indonesia) does ML\textsubscript{l} + ML\textsubscript{g} outperform ML\textsubscript{l}.
+
+\begin{figure}[t!]
 \centering
-\begin{tabular}{llrrrr}
-Country &  Cross-validation & Enviro &   ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
-\hline 
- Colombia & Random &  \textbf{0.59} &0.55 & 0.58 \\
- Colombia &  Spatial &  0.12 &  0.12 & \textbf{0.33}\\
- Madagascar &  Random &  \textbf{0.70} &  0.67 & 0.68 \\
- Madagascar &  Spatial &  0.22 &  0.51 & \textbf{0.55}\\
- Senegal &  Random &  \textbf{0.58} &  0.50 & 0.49 \\
- Senegal &  Spatial &   \textbf{0.63} & 0.55 & 0.52 \\
- Indonesia &  Random &  \textbf{0.52} & 0.32 & 0.46 \\
- Indonesia &  Spatial &  0.45 & 0.41 & 0.45 \\
-\end{tabular}
-\label{t:results2}
-\end{table}
-Figures~\ref{f:scatter1} shows scatter plots of the model performance under random cross-validation for all countries and all models.
-Figures~\ref{f:scatter2} shows scatter plots of the model performance under spatial cross-validation for all countries and all models.
-It can be seen that without environmental covariates, the models in Madagascar fail to predict very high or very low values correctly.
-Figure~\ref{f:map} shows the input data and out-of-sample predictions of the ML\textsubscript{l} model in Colombia.
+\includegraphics[width = 1\textwidth]{figs/col_comparison_map.png} %\caption{Indonesia spatial crossvalidation} 
+\caption{
+  A) Observed data for Colombia (grey for zero incidence).
+B) Out-of-sample predictions for the spatial cross-validation, environmental covariates only model.
+C) Out-of-sample predictions for the spatial cross-validation, local machine learning only model.
+For each cross-validation fold, predictions are made for the held out data which are then combined to make a single surface.
+}
+\label{f:map}
+\end{figure}
+
 Table~\ref{t:coverage} shows the out-of-sample coverage of the 80\% credible intervals for all models, countries and cross-validation schemes.
@@
 We can examine the relationship between the RMSE of the machine learning models to their fitted regression coefficients (weights).
-These values are given in Table~\ref{t:mlresults} (with the RMSE of the global models being copied four times).
-In all five sets of machine learning models, Random Forest performs the best.
+These values are given in Table~\ref{t:mlresults}.
+In all five sets of machine learning models (four sets trained on local data and one set trained on local data), Random Forest performs the best.
 We have not forced $\beta_i > 0$ but we have set the priors for these coefficients with a positive mean.
@@
 The predictions from Random Forest and GBM are highly correlated in part because they both perform well and in part because they are both tree based models.
-Finally, it can be seen that relationship between RMSE and regression coefficients is much weaker in Indonesia.
+Finally, it can be seen that the relationship between RMSE and regression coefficients was much weaker in Indonesia.
 For the models trained on local data a neural network has the highest fitted coefficient while for the models trained on global data an elastic net has the highest fitted coefficient.
-These model both produce very smooth estimates relative to the tree based methods.
 
@@
 
-\begin{figure}
+\begin{table}[t!]
+\caption{Pearson correlations between observed and predicted values. }
 \centering
-\includegraphics[width = 1\textwidth]{figs/col_comparison_map.png} %\caption{Indonesia spatial crossvalidation} 
-\caption{
-  Left: Observed data for Colombia (grey for zero incidence).
-Middle: Out-of-sample predictions for the spatial cross-validation, environmental covariates only model.
-Right: Out-of-sample predictions for the spatial cross-validation, local machine learning only model.
-For each cross-validation fold, predictions are made for the held out data which are then combined to make a single surface.
-}
-\label{f:map}
-\end{figure}
+\begin{tabular}{llrrrr}
+Country &  Cross-validation & Enviro &   ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
+\hline 
+ Colombia & Random &  \textbf{0.59} &0.55 & 0.58 \\
+ Colombia &  Spatial &  0.12 &  0.12 & \textbf{0.33}\\
+ Indonesia &  Random &  \textbf{0.52} & 0.32 & 0.46 \\
+ Indonesia &  Spatial &  0.45 & 0.41 & 0.45 \\
+ Madagascar &  Random &  \textbf{0.70} &  0.67 & 0.68 \\
+ Madagascar &  Spatial &  0.22 &  0.51 & \textbf{0.55}\\
+ Senegal &  Random &  \textbf{0.58} &  0.50 & 0.49 \\
+ Senegal &  Spatial &   \textbf{0.63} & 0.55 & 0.52 \\
+
+\end{tabular}
+\label{t:results2}
+\end{table}
 
@@
  Colombia & Random & \textbf{0.28} & \textbf{0.28} & \textbf{0.29} & \textbf{0.28} & \textbf{0.30} \\
+ Colombia &  Spatial & \textbf{0.30} & \textbf{0.33}  & \textbf{0.33} & \textbf{0.33} & \textbf{0.34}  \\
+ Indonesia & Random &0.80 & 0.81& 0.78& 0.79& 0.77  \\
+ Indonesia & Spatial & 0.80 & 0.78& 0.78& 0.76& 0.75  \\
  Madagascar &  Random & 0.80 & 0.77& 0.75& 0.77& 0.76 \\
+ Madagascar & Spatial & \textbf{0.65} & 0.74& 0.70& 0.75 & 0.76   \\
  Senegal & Random &0.79 & 0.79& 0.79& 0.79& 0.82 \\
- Indonesia & Random &0.80 & 0.81& 0.78& 0.79& 0.77  \\
- Colombia &  Spatial & \textbf{0.30} & \textbf{0.33}  & \textbf{0.33} & \textbf{0.33} & \textbf{0.34}  \\
- Madagascar & Spatial & \textbf{0.65} & 0.74& 0.70& 0.75 & 0.76   \\
  Senegal & Spatial & 0.85 & \textbf{0.91}& 0.85& \textbf{0.94} & 0.85  \\
-  Indonesia & Spatial & 0.80 & 0.78& 0.78& 0.76& 0.75  \\
 \end{tabular}
@@
      Country          & Model &      RMSE\textsubscript{l} & $\beta_l$ & RMSE\textsubscript{g} & $\beta_g$ \\ \hline
+Colombia & RF & \textbf{0.068} & 0.625 &  \textbf{0.169} & 0.180\\
+Colombia & GBM & 0.073 & \textbf{0.952} & 0.178& -0.218  \\
+Colombia & enet & 0.070 & 0.219 &0.233 & 0.183 \\
+Colombia & nnet & 0.070 & 0.129 &0.220 & 0.527 \\
+Colombia & ppr & 0.070 & 0.667 & 0.205 &  \textbf{0.546}\vspace{0.3cm}\\
+Indonesia & RF& \textbf{0.081} & 0.447 & \textbf{0.169} & 0.178\\
+Indonesia & GBM & 0.085 & 0.357 & 0.178& 0.289 \\
+Indonesia & enet & 0.091 & 0.303 &0.233 & \textbf{0.526} \\
+Indonesia & nnet & 0.089 & \textbf{0.506} &0.220 & 0.316 \\
+Indonesia & ppr & 0.089 & 0.364 & 0.205 &  0.089\vspace{0.3cm}\\
 Madagascar & RF & \textbf{0.100} & 0.538 &  \textbf{0.169} & \textbf{0.529}\\
@@
 Madagascar & ppr & 0.109 & 0.469 & 0.205 &  0.403\vspace{0.3cm}\\ 
-Colombia & RF & \textbf{0.068} & 0.625 &  \textbf{0.169} & 0.180\\
-Colombia & GBM & 0.073 & \textbf{0.952} & 0.178& -0.218  \\
-Colombia & enet & 0.070 & 0.219 &0.233 & 0.183 \\
-Colombia & nnet & 0.070 & 0.129 &0.220 & 0.527 \\
-Colombia & ppr & 0.070 & 0.667 & 0.205 &  \textbf{0.546}\vspace{0.3cm}\\
 Senegal & RF & \textbf{0.092} & \textbf{0.339} & \textbf{0.169} & \textbf{0.425} \\
@@
 Senegal & nnet & 0.099 & 0.254 &0.220 & 0.190 \\
-Senegal & ppr & 0.098 & 0.268& 0.205 &  0.126\vspace{0.3cm}\\
-Indonesia & RF& \textbf{0.081} & 0.447 & \textbf{0.169} & 0.178\\
-Indonesia & GBM & 0.085 & 0.357 & 0.178& 0.289 \\
-Indonesia & enet & 0.091 & 0.303 &0.233 & \textbf{0.526} \\
-Indonesia & nnet & 0.089 & \textbf{0.506} &0.220 & 0.316 \\
-Indonesia & ppr & 0.089 & 0.364 & 0.205 &  0.089\vspace{0.3cm}\\
+Senegal & ppr & 0.098 & 0.268& 0.205 &  0.126\\
 
@@
 
-Overall, our experiments suggest that using predictions from machine learning models trained on prevalence points provides more accurate predictions than using environmental covariates when fitting disaggregation models of malaria incidence.
+Overall, experiment \emph{i} suggest that using predictions from machine learning models trained on local prevalence points provides more accurate predictions than using environmental covariates when fitting disaggregation models of malaria incidence.
 This increased performance comes despite the data being on different scales, the data being measurements of different aspects of malaria transmission and despite the imperfect model we have used to translate between the two scales.
@@
 However, in a few cases such as Colombia and Madagascar under spatial cross-validation the performance boost is large.
-For example, using just environmental covariates in Colombia under spatial cross-validation gives a correlation between observed and predicted data of 0.12.
+For example, using just environmental covariates in Madagascar under spatial cross-validation gives a correlation between observed and predicted data of 0.22.
 Such a model would be unusable for any applied or policy work.
-In contrast, using the predictions from the machine learning models trained on local data gives a correlation value of 0.33, which while still poor is possibly useable.
+In contrast, using the predictions from the machine learning models trained on local data gives a correlation value of 0.55, which while still relatively poor is possibly a useful model.
 
@@
 
-Overall, the models fitted using predictions from machine learning models trained on the global database of prevalence point-surveys were less good than those using predictions from machine learning models trained on local data.
+Overall, the models fitted using predictions from machine learning models trained on the global database of prevalence point-surveys were less good than those using either environmental data alone or than those using predictions from machine learning models trained on local data.
 This in itself is not particularly surprising.

### 2019-11-08 tim-lucas-model-ensembles.tex
More edits referring back to Dans comments.

@@
 
+\usepackage{amssymb}
+\usepackage{array}
+
 \usepackage{lineno,hyperref}
 \modulolinenumbers[5]
-\bibliographystyle{elsarticle-num}
+
 %\usepackage{numcompress}\bibliographystyle{model3-num-names}
@@
 %\setcitestyle{author[oxford]year}
+\bibliographystyle{elsarticle-num}
 
@@
 As reliable routine surveillance data of malaria incidence becomes more widely available disaggregation regression is becoming an important model framework.
-However, the aggregation of incidence over large, heterogeneous areas means that they are often underpowered for learning complex, non-linear relationships between the environment and malaria risk.
-In contrast, prevalence point-surveys are directly linked to loval environmental conditions but are not common in many areas of the world.
+However, the aggregation of incidence over large, heterogeneous areas means that they are underpowered for learning complex, non-linear relationships between the environment and malaria risk.
+In contrast, prevalence point-surveys are directly linked to local environmental conditions but are not common in many areas of the world.
 A model that combines point surveys and aggregated surveillance data could have the benefits of both but must be able to account for the fact that these two data types measure different malariometric units.
-Here, we train multiple non-linear, machine learning models on prevalence point-surveys and then ensemble the predictions from these with a geostatistical disaggregation model that uses routine surveillance data.
+Here, we train multiple non-linear, machine learning models on prevalence point-surveys.
+We then ensemble the predictions from these machine learning models with a geostatistical disaggregation regression model that uses malaria incidence, aggregated to admnistrative units, as response data.
 We find that, in four national case studies, using a disaggregation regression model to combine predictions from machine learning models trained on prevalence point-surveys improves model accuracy relative to a baseline disaggregation regression model that uses linear effects of the environmental covariates only.
@@
 High-resolution maps of malaria risk are vital for control and elimination \citep{weiss2019mapping, battle2019mapping}.
-However, mapping malaria in lower burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{weiss2019mapping, bhatt2017improved, battle2019mapping, bhatt2015effect} is often not effective for two reasons.
+However, mapping malaria in lower burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{weiss2019mapping,battle2019mapping, bhatt2017improved, bhatt2015effect} is often not effective for two reasons.
 Firstly, so few individuals are infected that most surveys will detect zero cases \citep{sturrock2016mapping}.
@@
 Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more reliable and more widely available \citep{sturrock2016mapping} and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
-However, the aggregation of cases over space means that the data may be relatively uninformative, especially if the case counts are aggregated over large or heterogeneous areas, because it is unclear where within the polygon, and in which environments, the cases occurred.
+However, the aggregation of cases over space means that the data may be spatially uninformative, especially if the case counts are aggregated over large or heterogeneous areas, because it is unclear where within the polygon, and in which environments, the cases occurred.
 This data is therefore often under-powered for fitting flexible, non-linear models as is required for accurate malaria maps \citep{bhatt2017improved, bhatt2015effect}. 
@@
 
-Here we propose training a suite of level-zero machine learning models on point-level, binomial prevalence data and combining these predictions by using them as covariates in a level-one, polygon-level, Poisson incidence model.
+Here we propose training a suite of machine learning models on point-level, binomial prevalence data and combining these predictions by using them as covariates in a polygon-level, disaggregation regression model that uses  malaria incidence, aggregated to administrative units, as the response.
 This modelling scheme has similarities to previous stacking methods used for malaria mapping \citep{bhatt2017improved} and elsewhere \citep{wolpert1992stacked, sill2009feature, hao2019review, breiman1996stacked}.
@@
 
-Model stacking \citep{wolpert1992stacked} has proven effective in many realms \citep{breiman1996stacked, sill2009feature, bhatt2017improved, hao2019review}. 
+Model stacking \citep{wolpert1992stacked} has proven effective in many realms \citep{bhatt2017improved, sill2009feature, hao2019review, breiman1996stacked}. 
 Stacking improves predictions by controlling bias and variance; as long as suitably diverse models are averaged, they will have different biases while high variance in models should be averaged out.
@@
 The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling, pfeffer2018ma}.
-While the data cover a large time period, we did not model time explicitely but future work could include this component.
+While the data covered a large time period, we did not model time explicitely as we are here focussed on spatial, rather than temporal modelling.
 The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
@@
 
-\subsection{Covariates}
+\subsection{Raw Covariates}
 
@@
 The covariates were standardised to have a mean of zero and a standard deviation of one. 
-These covariates were used for both the machine learning models and the polygon-level models.
-Raster surfaces of population for the years 2005, 2010 and 2015, were created using data from WorldPop \citep{tatem2017worldpop} and from GPWv4 \citep{gpw4} where WorldPop did not have values. 
+These covariates were used in the machine learning models as well as being used directly in the baseline dissaggregation regression models.
+Raster surfaces of population for the years 2005, 2010 and 2015, were created using data from WorldPop \citep{linard2012population,gaughan2013high,  sorichetta2015high} and from GPWv4 \citep{gpw4} where WorldPop did not have values.
 Population rasters for the remaining years were created by linear interpolation.
@@
 Our response variable was prevalence and we weighted the data by sample size (i.e.\thinspace the number of people tested for malaria in each survey).
+We used the raw covariates described above.
 For each model we ran five-fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
@@
 Predictions from these models were then made across each country respectively.
-These predictions were finally empirical logit transformed so that they were on the linear predictor scale of the top level model.
+These predictions were empirical logit transformed so that they were on the linear predictor scale of the top level model.
 An empirical logit was used rather than a standard logit as there were many predictions of exactly zero.
+These predicted surfaces were subsequently used as covariates in the dissaggregation regression models.
+This process can therefore be seen as creating non-linear transformations of the raw covariates that are hopefully better correlated with malaria incidence than the raw covariates are.
 
@@
 The top level model was a disaggregation regression model \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
-The models were implemented and fitted using Template Model Builder \citep{TMB} in R \citep{R}.
+The models were implemented and fitted using Template Model Builder \citep{TMB} in R \citep{R} and we note that these models cannot be fitted using INLA \citep{INLA} as we are not using a linear link function.
 This model is defined by a likelihood at the level of the polygon with covariates and a spatial random field at the pixel-level. 
@@
 The fact that the model passes through prevalence space ensures that the predictions from the machine learning models can be appropriately scaled.
-The linear predictor of the model is related to prevalence by a typical logit link function and includes an intercept, $\beta_0$, covariates, $X$, with regression parameters $\beta$, a spatial, Gaussian, random field, $u_s(\rho, \sigma_u)$, and an \emph{iid} random effect, $v_j(\sigma_v)$.
-$$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u_s(\rho, \sigma_u) + v_j(\sigma_v)\right)$$
+The linear predictor of the model is related to prevalence by a typical logit link function and includes: an intercept, $\beta_0$; covariates, $X$, with a vector of regression parameters, $\beta$; a spatial, Gaussian, random field, $u_s(\rho, \sigma_u)$; and an \emph{iid} random effect, $v_j(\sigma_v)$.
+$$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u_s\left(\rho, \sigma_u\right) + v_j\left(\sigma_v\right)\right)$$
 The Gaussian spatial effect has a Mat\'ern covariance function and two hyperparameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
@@
 
-We used the environmental and anthropogenic covariates, centered and standardised, as a baseline model (subsequantly called Enviro).
+We used the raw environmental covariates, centered and standardised, as a baseline model (subsequantly called Enviro).
 We then performed two experiments.
@@
 
+\begin{table}[t!]
+\caption{Summary of models.
+Environmental covariates includes the full set of eight variables.
+Local ML covariates are the predictions from machine learning models, trained with local prevalence data and the eight environmental covariates as inputs.
+Global ML covariates are the predictions from machine learning models, trained with global prevalence data and the eight environmental covariates as inputs.}
+\centering
+\begin{tabular}{l >{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{1.55cm}>{\centering\arraybackslash}p{1.7cm}}
+Model &  Environmental covariates & Local ML covariates &  Global ML covariates \\
+\hline 
+ \textbf{Experiment \emph{i}} & & &  \\
+ Enviro &  \checkmark & & \\
+ Enviro + ML\textsubscript{l} & \checkmark & \checkmark &\\\vspace{0.4cm}
+ ML\textsubscript{l} &  & \checkmark & \\
+ \textbf{Experiment \emph{ii}} & & &  \\
+ Enviro &  \checkmark & & \\
+ ML\textsubscript{g} &  &  & \checkmark\\
+ ML\textsubscript{l} + ML\textsubscript{g} &  & \checkmark &  \checkmark
+\end{tabular}
+\label{t:results}
+\end{table}
+
 In experiment \emph{i} we compared the baseline to two models that used locally trained machine learning models.
-First we used only the predictions from the machine learning models trained on local data (subsequantly called ML\textsubscript{l}).
-Secondly, we combined the predictions from the machine learning models and the environmental coviarates in one model (Enviro + ML\textsubscript{l}).
+Firstly, we combined the predictions from the machine learning models and the environmental coviarates in one model (Enviro + ML\textsubscript{l}).
+In this model therefore, the environmental covariates are effectively used twice, once in their raw form and once in transformed form (i.e. the predictions from the machine learning models).
+Secondly, we used only the predictions from the machine learning models trained on local data (subsequantly called ML\textsubscript{l}).
+As the environmental covariates are used as covariates in the machine learning models, this model is still ultimately driven by the values of the raw enviroenmental covariates.
+However, in this model we are only using the environmental covariates in the transformed space learned by the machine learning models.
 
@@
 We used only predictions from the machine learning models trained on the global data (ML\textsubscript{g}).
-We also combined predictions from the machine learning models trained on regional data and global data (ML\textsubscript{l} + ML\textsubscript{g}).
+We also combined predictions from the machine learning models trained on regional data and predictions from the machine learning models trained on global data (ML\textsubscript{l} + ML\textsubscript{g}).
 
@@
 Due to the small sample size, we set k as 5 for Senegal.
-This spatial cross-validation scheme is testing the models' ability to make predictions far from data where the spatial random field is not informative.
+This spatial cross-validation scheme is testing the ability of the different models to make predictions far from data where the spatial random field is not informative.
 
@@
 Of the remaining five cases, in two cases ML\textsubscript{l} performs best and in three cases Envir + ML\textsubscript{l} performs best.
-The greatest benefits in predictive performance occur under spatial cross-validation and in the cases when the Enviro model is particularly poor.
+The greatest benefits to using prediction from machine learning models instead of, or in combination with, the raw environmental variables occur under spatial cross-validation and in the cases when the Enviro model is particularly poor.
 
@@
 The coverage in Colombia is very poor with no models achieving a coverage above 0.4.
-Otherwise the coverage is generally acceptable with most values lying between 0.7 and 0.9.
+The coverage is acceptable in the other three countries with most values lying between 0.7 and 0.9.
 Overall there is a general tendancy for models to be slightly over confident.
@@
 
-Overall we find that including predictions from machine learning models trained on prevalence point-surveys typically improves the performance of disaggregation regression models for malaria incidence.
-Training the machine learning models on local data gave much better performance, but how local the data should be remains an open question.
+Overall we find that including predictions from machine learning models trained on prevalence point-surveys can improve the performance of disaggregation regression models for malaria incidence relative to using raw environmental covariates.
+This extra modelling step can be seen as finding useful, non-linear transformations of the raw environmental covariates.
+This view is important for understanding how the model will predict in areas with no incidence data; in this situation the Gaussian random field has no effect and predictions are driven by this non-linear transformation of the raw environmental covariates.
+Training the machine learning models on local data (i.e.
+prevalence data from the same country or region as the incidence data) gave much better performance than when training the machine learning models on the full, global, dataset of prevalence point-surveys.
 As more countries start providing timely and accurate routine surveillance data of malaria cases disaggregation regression may become more popular.
-We have here presented a method for improving these models by using ancillary data.
+We have here presented a method for improving these models by using ancillary, prevalence point-survey data.
 

### 2019-11-15 tim-lucas-model-ensembles.tex
Bunch more edits.

@@
 Here, we train multiple non-linear, machine learning models on prevalence point-surveys.
-We then ensemble the predictions from these machine learning models with a geostatistical disaggregation regression model that uses malaria incidence, aggregated to admnistrative units, as response data.
-We find that, in four national case studies, using a disaggregation regression model to combine predictions from machine learning models trained on prevalence point-surveys improves model accuracy relative to a baseline disaggregation regression model that uses linear effects of the environmental covariates only.
+We then ensemble the predictions from these machine learning models with a geostatistical disaggregation regression model that uses malaria incidence, aggregated to administrative units, as response data.
+We find that, in four national case studies, using a disaggregation regression model to combine predictions from machine learning models trained on prevalence point-surveys improves model accuracy relative to a baseline disaggregation regression model that uses linear effects of the raw environmental covariates only.
 
@@
 
-Here we propose training a suite of machine learning models on point-level, binomial prevalence data and combining these predictions by using them as covariates in a polygon-level, disaggregation regression model that uses  malaria incidence, aggregated to administrative units, as the response.
+Here we propose training a suite of machine learning models, using point-level, binomial prevalence data and environmental covariates, and combining predictions from these models by using them as covariates in a polygon-level, disaggregation regression model that uses  malaria incidence, aggregated to administrative units, as the response.
 This modelling scheme has similarities to previous stacking methods used for malaria mapping \citep{bhatt2017improved} and elsewhere \citep{wolpert1992stacked, sill2009feature, hao2019review, breiman1996stacked}.
-However, as the response data in the level-zero models and the level-one models are on different scales (prevalence is a proportion while incidence is a rate) there are a number of important differences.
-Notably, we cannot simply take a weighted average of the predictions from the level-zero models as in a standard stacking scheme \citep{sill2009feature, hao2019review}.
+However, as the response data in the machine learning models and the disaggregation regression models are on different scales (prevalence is a proportion while incidence is a rate) there are a number of important differences.
+Notably, we cannot simply take a weighted average of the predictions from the machine learning models as in a standard stacking scheme \citep{sill2009feature, hao2019review}.
 Instead the predictions  need to be transformed to the incidence scale with a seperately fitted model \citep{cameron2015defining}.
-Applications in other disease contexts have used a similar scheme where data from vector or wild-animal host species are used to train models, the predictions from which are then used in a final model \citep{pigott2014mapping, shearer2016estimating}.
+Applications in other disease contexts have used a similar scheme where data from vector or wild-animal host species are used to train models, the predictions from which are then used as covariates in a final model \citep{pigott2014mapping, shearer2016estimating}.
 In such applications we would always expect to need additional covariates in additional to the modelled distributions of hosts or vectors.
-However, in the case examined here, both sets of data are direct measures of some aspect of malaria transmission rate, and therefore it is possible that we would need no further covariates.
+However, in the case examined here, both sets of data are direct measures of some aspect of malaria transmission rate, and therefore it is possible, though not guaranteed, that we would not need any further covariates.
 
@@
 This understanding of how stacking improves model performance indicates that diversity in models is important for stacking to be effective.
-Diversity in models is typically created in two ways: by using diverse training datasets\citep{breiman1996bagging} (as in Random Forests for example) and by using functionally different models (for example by averaging tree based models and neural networks) \citep{breiman1996stacked}.
+Diversity in models is typically created in two ways: by using diverse training datasets \citep{breiman1996bagging} (as in Random Forests for example) and by using functionally different models (for example by averaging tree based models and neural networks) \citep{breiman1996stacked}.
 One important trade-off in spatial modelling is whether to use local data (with a smaller sample size but that is likely to be representative of the area of study) or global data that have a larger sample size but a less close association with the areas of study.
@@
 To test the effectiveness of the proposed approach we used data from four countries: Madagascar, Colombia, Indonesia and Senegal.
-We first examined whether using predictions from machine learning models trained on local prevalence data could improve predictions.
-We then tested whether using predictions from machine learning models trained on global data could improve predictions either on their own or when combined with predictions from machine learning models trained on local data.
-While there was not one consistently best model, we found that in nearly all cases, using predictions from local machine learning models improved model performance.
+We first examined whether using predictions from machine learning models, trained on local prevalence data and raw environmental covariates, could improve predictions when used as covariates in disaggregation regression models.
+We then tested whether covariates made using predictions from machine learning models trained on a global prevalence dataset could improve predictions either on their own or when combined with predictions from machine learning models trained on local data.
+While there was not one consistently best model, we found that in most cases, using predictions from local machine learning models as covariates improved model performance.
 In contrast, using predictions from machine learning models trained on the global prevalence dataset rarely improved predictive performance.
@@
 
-\subsection{Malaria Data}
+\subsection{Epidemiological Data}
 
@@
 For reports where cases were not reported at the species level, national estimates of the ratio between \emph{P.
-falciparum} and \emph{P.
-vivax} cases were used to calculate \emph{P.
+falciparum} and \emph{Plasmodium vivax} cases were used to calculate \emph{P.
 falciparum} only cases.
@@
 
-\subsection{Raw Covariates}
-
-We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \citep{weiss2015re, weiss2018global}.
-Tasseled cap brightness and urban land cover were subsequently removed as they were highly correlated with other variables. 
-The covariates were standardised to have a mean of zero and a standard deviation of one. 
-These covariates were used in the machine learning models as well as being used directly in the baseline dissaggregation regression models.
 Raster surfaces of population for the years 2005, 2010 and 2015, were created using data from WorldPop \citep{linard2012population,gaughan2013high,  sorichetta2015high} and from GPWv4 \citep{gpw4} where WorldPop did not have values.
@@
 
+\subsection{Raw Environmental Covariates}
+
+We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \citep{weiss2014air, weiss2015re, weiss2018global}.
+Tasseled cap brightness and urban land cover were subsequently removed as they were highly correlated with other variables. 
+The covariates were standardised to have a mean of zero and a standard deviation of one. 
+We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created from predictions from machine learning models.
+These variables were used as covariates in the machine learning models as well as being used directly as covariates the baseline dissaggregation regression models.
+
 \subsection{Machine learning models}
-For each country specific dataset and for the global dataset we fitted 5 models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees \citep{gbm}.
+For each country specific dataset and for the global dataset we fitted 5 models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees (gradient boosted models, subsequently GBM) \citep{gbm}.
 These models were fitted to both the full malaria prevalence dataset and to the regional subsets of the data.
 Our response variable was prevalence and we weighted the data by sample size (i.e.\thinspace the number of people tested for malaria in each survey).
-We used the raw covariates described above.
+We used the raw environmental covariates described above.
 For each model we ran five-fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
@@
 These predicted surfaces were subsequently used as covariates in the dissaggregation regression models.
-This process can therefore be seen as creating non-linear transformations of the raw covariates that are hopefully better correlated with malaria incidence than the raw covariates are.
+This process can therefore be seen as creating non-linear transformations of the raw covariates that are hopefully better correlated with malaria incidence than the raw environmental covariates are.
 
@@
 The \emph{iid} random effect models both missing covariates and extra-Poisson sampling error.
+As described in more detail below, the matrix of covariates $X$ varies and is made up of various combinations of raw environmental covariates and predictions from the machine learning models.
 
@@
 In the second (\emph{ii}) we tested whether using predictions from machine learning models trained on global prevalence data improved predictions.
+These combinations of covariates are summarised in Table~\ref{t:models}.
 Our primary performance metric was correlation between observed and predicted data.
@@
 \end{tabular}
-\label{t:results}
+\label{t:models}
 \end{table}
@@
 In experiment \emph{i} we compared the baseline to two models that used locally trained machine learning models.
-Firstly, we combined the predictions from the machine learning models and the environmental coviarates in one model (Enviro + ML\textsubscript{l}).
+Firstly, we combined the predictions from the machine learning models and the environmental coviarates in one model (subsequantly called Enviro + ML\textsubscript{l}).
 In this model therefore, the environmental covariates are effectively used twice, once in their raw form and once in transformed form (i.e. the predictions from the machine learning models).
@@
 
-In experiment \emph{ii} we compared the baseline to two models that used globally trained machine learning models.
-We used only predictions from the machine learning models trained on the global data (ML\textsubscript{g}).
-We also combined predictions from the machine learning models trained on regional data and predictions from the machine learning models trained on global data (ML\textsubscript{l} + ML\textsubscript{g}).
+In experiment \emph{ii} we compared the baseline model to two models that use predictions from machine learning models trained on the global dataset of prevalence surveys.
+In the first, we used only predictions from the machine learning models trained on the global data (ML\textsubscript{g}).
+In the second, we combined predictions from the machine learning models trained on regional data and predictions from the machine learning models trained on global data (ML\textsubscript{l} + ML\textsubscript{g}).
 
@@
 This spatial cross-validation scheme is testing the ability of the different models to make predictions far from data where the spatial random field is not informative.
+
+\section{Results}
+
+Table \ref{t:results} gives the correlation between observed and held out data (under random and spatially cross-validation) for experiment \emph{i} (models Enviro, ML\textsubscript{l} and Envir + ML\textsubscript{l}).
+Many of the differences in performance are rather marginal.
+Enviro was the best performing model in two cases (random cross-validation in Madagascar and Senegal).
+In one case, Enviro perfomed equally well as another model; under spatial cross-validation in Senegal Enviro and Enviro + ML\textsubscript{l} performed equally well.
+Of the remaining five cases, in two cases ML\textsubscript{l} performed best and in three cases Envir + ML\textsubscript{l} performed best.
+The greatest benefits to using prediction from machine learning models instead of, or in combination with, the raw environmental variables occured under spatial cross-validation and in the cases when the Enviro model was particularly poor.
 
@@
 \end{table}
-
-\section{Results}
-
-Table \ref{t:results} gives the correlation between observed and held out data (under random and spatially cross-validation) for experiment \emph{i} (models Enviro, ML\textsubscript{l} and Envir + ML\textsubscript{l}).
-Many of the differences in performance are rather marginal.
-Enviro is the best performing model in two cases (random cross-validation in Madagascar and Senegal).
-In one case, Enviro perfoms equally well as another model; under spatial cross-validation in Senegal Enviro and Enviro + ML\textsubscript{l} perform equally well.
-Of the remaining five cases, in two cases ML\textsubscript{l} performs best and in three cases Envir + ML\textsubscript{l} performs best.
-The greatest benefits to using prediction from machine learning models instead of, or in combination with, the raw environmental variables occur under spatial cross-validation and in the cases when the Enviro model is particularly poor.
 
@@
 Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatially cross-validation) for experiment \emph{ii} (models Enviro, ML\textsubscript{g} and ML\textsubscript{l} + ML\textsubscript{g}).
-In six cases, Enviro is the best or tied best performing model.
-The ML\textsubscript{g} model is never the best performing model and only outpeforms Enviro in one case (spatial cross-validation in Madagascar).
-In two cases ML\textsubscript{l} + ML\textsubscript{g} is the best performing model (spatial Colombia and spatial Madagascar).
+In six cases, Enviro was the best or tied best performing model.
+The ML\textsubscript{g} model was never the best performing model and only outpeforms Enviro in one case (spatial cross-validation in Madagascar).
+In two cases ML\textsubscript{l} + ML\textsubscript{g} was the best performing model (spatial Colombia and spatial Madagascar).
 Comparing across Tables \ref{t:results} and \ref{t:results2} we can see that ML\textsubscript{g} outpeforms ML\textsubscript{l} once (spatial cross-validation in Senegal).
-In only two cases (Spatial Senegal and Spatial Indonesia) does ML\textsubscript{l} + ML\textsubscript{g} outperform ML\textsubscript{l}.
+In only two cases (Spatial Senegal and Spatial Indonesia) did ML\textsubscript{l} + ML\textsubscript{g} outperform ML\textsubscript{l}.
 
@@
 Table~\ref{t:coverage} shows the out-of-sample coverage of the 80\% credible intervals for all models, countries and cross-validation schemes.
-The coverage in Colombia is very poor with no models achieving a coverage above 0.4.
-The coverage is acceptable in the other three countries with most values lying between 0.7 and 0.9.
-Overall there is a general tendancy for models to be slightly over confident.
+The coverage in Colombia was very poor with no models achieving a coverage above 0.4.
+The coverage was acceptable in the other three countries with most values lying between 0.7 and 0.9.
+Overall there was a general tendancy for models to be slightly over confident.
 
@@
 These values are given in Table~\ref{t:mlresults}.
-In all five sets of machine learning models (four sets trained on local data and one set trained on local data), Random Forest performs the best.
+In all five sets of machine learning models (four sets trained on local data and one set trained on global data), Random Forest performs the best.
 We have not forced $\beta_i > 0$ but we have set the priors for these coefficients with a positive mean.
@@
 
-Overall, experiment \emph{i} suggest that using predictions from machine learning models trained on local prevalence points provides more accurate predictions than using environmental covariates when fitting disaggregation models of malaria incidence.
-This increased performance comes despite the data being on different scales, the data being measurements of different aspects of malaria transmission and despite the imperfect model we have used to translate between the two scales.
-However, there is no clear best model between ML\textsubscript{l} and Enviro + ML\textsubscript{l}.
+Overall, experiment \emph{i} suggest that using predictions from machine learning models trained on local prevalence points provides more accurate predictions than using raw environmental covariates when fitting disaggregation models of malaria incidence.
+This increased performance comes despite the prevalence data being on a different scale and being measurements of a different aspects of malaria transmission as well as the fact that the model we have used to translate between the two scales is imperfect.
+However, there was no clear best model between ML\textsubscript{l} and Enviro + ML\textsubscript{l}.
 Therefore, when using these methods, both of these models should be fitted and the most predictive model selected for a given dataset.
@@
 prevalence data from the same country or region as the incidence data) gave much better performance than when training the machine learning models on the full, global, dataset of prevalence point-surveys.
-As more countries start providing timely and accurate routine surveillance data of malaria cases disaggregation regression may become more popular.
+As more countries, particularly those with medium or low malaria burdens, start providing timely and accurate routine surveillance data of malaria cases disaggregation regression may become more popular.
 We have here presented a method for improving these models by using ancillary, prevalence point-survey data.

### 2019-11-19 tim-lucas-model-ensembles.tex
Lisa and suzanne edits.

@@
 We used two data sources that reflect \emph{Plasmodium falciparum} malaria transmission; point-prevalence surveys and polygon-level, aggregated incidence data. 
-We selected Madagascar, Colombia, Indonesia and Senegal as case examples as they all have fairly complete, publicly available, surveillance data at a finer geographical resolution than admin 1.
+We selected Madagascar, Colombia, Indonesia and Senegal as case examples as they all have fairly complete, publicly available, surveillance data at a finer geographical resolution than admin 1 (i.e.
+higher resolution than state or province).
 The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling, pfeffer2018ma}.
-While the data covered a large time period, we did not model time explicitely as we are here focussed on spatial, rather than temporal modelling.
+While the data covered a large time period, we did not model time explicitly as we are here focussed on spatial, rather than temporal modelling.
 The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
@@
 
-The polygon incidence data were collected from government reports and standardised using methods defined in \cite{cibulskis2011worldwide}.
+The polygon incidence data (i.e. malaria incidence aggregated to administrative units) were collected from government reports and standardised using methods defined in \cite{cibulskis2011worldwide}.
 This standardisation step accounts for missed cases due to lack of treatment seeking, missing case reports, and cases that sought medical attention outside the public health systems \citep{battle2016treatment}.
@@
 To minimise temporal effects we selected, for each country, one year of surveillance data. 
-We used annual surveillance data from data from 2013 for Madagascar (110 districts), 2015 for Colombia (952 municipalities), 2013 for Indonesia (244 regencies and cities) and 2009 for Senegal (34 departments).
-These years were selected as they had the most data in each case.
+We used annual surveillance data from 2013 for Madagascar (110 districts), 2015 for Colombia (952 municipalities), 2013 for Indonesia (244 regencies and cities) and 2009 for Senegal (34 departments).
+These years were selected as they had the most complete data in each case.
 
-Raster surfaces of population for the years 2005, 2010 and 2015, were created using data from WorldPop \citep{linard2012population,gaughan2013high,  sorichetta2015high} and from GPWv4 \citep{gpw4} where WorldPop did not have values.
+Raster surfaces (i.e.
+population gridded to 5$ \times$ 5 km pixles) of population for the years 2005, 2010 and 2015, were created using data from WorldPop \citep{linard2012population,gaughan2013high,  sorichetta2015high} and from GPWv4 \citep{gpw4} where WorldPop did not have values.
 Population rasters for the remaining years were created by linear interpolation.
@@
 Tasseled cap brightness and urban land cover were subsequently removed as they were highly correlated with other variables. 
-The covariates were standardised to have a mean of zero and a standard deviation of one. 
+The covariates were standardised and centered to have a mean of zero and a standard deviation of one. 
 We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created from predictions from machine learning models.
@@
 Our response variable was prevalence and we weighted the data by sample size (i.e.\thinspace the number of people tested for malaria in each survey).
-We used the raw environmental covariates described above.
+We used the raw environmental covariates described above as covariates in these machine learning models.
+This process can therefore be seen as creating non-linear transformations of the raw covariates that are hopefully better correlated with malaria incidence than the raw environmental covariates are.
 For each model we ran five-fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
@@
 These predicted surfaces were subsequently used as covariates in the dissaggregation regression models.
-This process can therefore be seen as creating non-linear transformations of the raw covariates that are hopefully better correlated with malaria incidence than the raw environmental covariates are.
 
 \subsection{Disaggregation regression}
-The top level model was a disaggregation regression model \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
+The model fitted to aggregated incidence data was a disaggregation regression model \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
 The models were implemented and fitted using Template Model Builder \citep{TMB} in R \citep{R} and we note that these models cannot be fitted using INLA \citep{INLA} as we are not using a linear link function.
@@
 
-We used the raw environmental covariates, centered and standardised, as a baseline model (subsequantly called Enviro).
+We used the raw environmental covariates, centered and standardised, as a baseline model (subsequently called Enviro).
 We then performed two experiments.
@@
 Secondly, we used only the predictions from the machine learning models trained on local data (subsequantly called ML\textsubscript{l}).
-As the environmental covariates are used as covariates in the machine learning models, this model is still ultimately driven by the values of the raw enviroenmental covariates.
+As the environmental covariates are used as covariates in the machine learning models, this model is still ultimately driven by the values of the raw environmental covariates.
 However, in this model we are only using the environmental covariates in the transformed space learned by the machine learning models.

### 2019-11-19 tim-lucas-model-ensembles.tex
Final suzanne comments.

@@
 The covariates were standardised and centered to have a mean of zero and a standard deviation of one. 
-We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created from predictions from machine learning models.
-These variables were used as covariates in the machine learning models as well as being used directly as covariates the baseline dissaggregation regression models.
+We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created using predictions from machine learning models.
+These variables were used as covariates in the machine learning models as well as being used directly as covariates in the baseline dissaggregation regression models.
 

### 2019-11-20 tim-lucas-model-ensembles.tex
More minor edits.

@@
 \author[oxford]{Suzanne H. Keddie}
-\author[oxford]{Katherine A. Twohig}
+%\author[oxford]{Katherine A. Twohig}
 \author[oxford]{Elisabeth G. Chestnutt}
 \author[oxford]{Rosalind E. Howes}
-\author[oxford]{Michele Nguyen}
+%\author[oxford]{Michele Nguyen}
 \author[oxford]{Susan F. Rumisha}
 \author[oxford]{Andre Python}
+\author[oxford]{Tasmin L. Symons}
+\author[oxford]{Justin J. Millar}
 \author[oxford]{Rohan Arambepola}
@@
 As reliable routine surveillance data of malaria incidence becomes more widely available disaggregation regression is becoming an important model framework.
-However, the aggregation of incidence over large, heterogeneous areas means that they are underpowered for learning complex, non-linear relationships between the environment and malaria risk.
+However, the aggregation of incidence over large, heterogeneous areas means that these data are underpowered for learning complex, non-linear relationships between the environment and malaria risk.
 In contrast, prevalence point-surveys are directly linked to local environmental conditions but are not common in many areas of the world.
@@
 We then ensemble the predictions from these machine learning models with a geostatistical disaggregation regression model that uses malaria incidence, aggregated to administrative units, as response data.
-We find that, in four national case studies, using a disaggregation regression model to combine predictions from machine learning models trained on prevalence point-surveys improves model accuracy relative to a baseline disaggregation regression model that uses linear effects of the raw environmental covariates only.
+We find that, in four national case studies, using a disaggregation regression model to combine predictions from machine learning models trained on prevalence point-surveys improves model accuracy relative to a baseline disaggregation regression model that only uses linear effects of the raw environmental covariates.
 
@@
 Instead the predictions  need to be transformed to the incidence scale with a seperately fitted model \citep{cameron2015defining}.
-Applications in other disease contexts have used a similar scheme where data from vector or wild-animal host species are used to train models, the predictions from which are then used as covariates in a final model \citep{pigott2014mapping, shearer2016estimating}.
-In such applications we would always expect to need additional covariates in additional to the modelled distributions of hosts or vectors.
+
+Applications in other disease contexts have used a similar stacking scheme where data from vector or wild-animal host species are used to train models, the predictions from which are then used as covariates in a final model \citep{pigott2014mapping, shearer2016estimating}.
+In such applications we would always expect to need additional covariates in addition to the modelled distributions of hosts or vectors.
 However, in the case examined here, both sets of data are direct measures of some aspect of malaria transmission rate, and therefore it is possible, though not guaranteed, that we would not need any further covariates.
@@
 falciparum} only cases.
-To minimise temporal effects we selected, for each country, one year of surveillance data. 
+To minimise temporal effects we selected one year of surveillance data for each country. 
 We used annual surveillance data from 2013 for Madagascar (110 districts), 2015 for Colombia (952 municipalities), 2013 for Indonesia (244 regencies and cities) and 2009 for Senegal (34 departments).
@@
 where $i_a$ is the estimated polygon incidence rate and $\mathrm{pop_a}$ is the observed polygon population-at-risk. 
-This polygon-level likelihood is linked to the pixel level prevalence 
+This polygon-level likelihood is linked to the pixel-level incidence and prevalence by
 $$i_a = \frac{ \sum(i_b \mathrm{pop}_b)}{\sum  \mathrm{pop}_b} $$
@@
 Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$ and $\sigma_v$. 
-We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \citep{fuglstad2018constructing} such that $P(\rho < 1) = 0.0001$ (except for Indonesia where we set $P(\rho < 3) = 0.0001$ due to it's much larger size) and $P(\sigma_u > 1) = 0.0001$.
+We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \citep{fuglstad2018constructing} such that $P(\rho < 1) = 0.0001$ (except for Indonesia where we set $P(\rho < 3) = 0.0001$ due to its much larger size) and $P(\sigma_u > 1) = 0.0001$.
 This prior encoded our \emph{a priori} preference for a simpler, smoother random field.
@@
 Finally, we set different priors on the regression coefficients depending on which covariates were used.
-When environmental covariates or a mix of environmental covariates and predictions from machine learning models were used we set the prior to be weakly regularising, $\beta_i \sim \operatorname{Norm}(0, 0.4)$, such that it is unlikely that any single covariate could explain the full range of the response data.
+When environmental covariates or a mix of environmental covariates and predictions from machine learning models were used we set the prior to be weakly regularising, $\beta_i \sim \operatorname{Norm}(0, 0.4)$, such that it was unlikely that any single covariate could explain the full range of the response data.
 When only machine learning model predictions were used we set $\beta_i \sim \operatorname{ Norm}(\frac{1}{M}, 0.4)$ where $M$ is the number of machine learning models being used. 
@@
 In a preliminary analysis we tested the case where we force $\beta_i > 0$ which in practice is largely the same as the $\sum_i \beta_i = 1$  \citep{breiman1996stacked} but allows a small amount of flexibility to handle mispecification in $\mathrm{p2i}$.
-This analysis did not find any benefits to this approach so it was not considered further.
+This analysis did not show any benefits to this approach so it was not considered further.
 
@@
 
-Table \ref{t:results} gives the correlation between observed and held out data (under random and spatially cross-validation) for experiment \emph{i} (models Enviro, ML\textsubscript{l} and Envir + ML\textsubscript{l}).
+Table \ref{t:results} gives the correlation between observed and held out data (under random and spatial cross-validation) for experiment \emph{i} (models Enviro, ML\textsubscript{l} and Envir + ML\textsubscript{l}).
 Many of the differences in performance are rather marginal.
@@
 
-Figures~\ref{f:scatter1} shows scatter plots of the model performance under random cross-validation for experiment \emph{i}.
-Figures~\ref{f:scatter2} shows scatter plots of the model performance under spatial cross-validation for experiment \emph{i}.
+Figure~\ref{f:scatter1} shows scatter plots of the model performance under random cross-validation for experiment \emph{i} while Figure~\ref{f:scatter2} shows scatter plots of the model performance under spatial cross-validation.
 It can be seen that without environmental covariates, the models in Madagascar fail to predict very high or very low values correctly.
-Figure~\ref{f:map} shows the input data and spatially out-of-sample predictions of the Enviro model and ML\textsubscript{l} model in Colombia.
+Figure~\ref{f:map} shows the input data and spatial out-of-sample predictions of the Enviro model and ML\textsubscript{l} model in Colombia.
 
-Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatially cross-validation) for experiment \emph{ii} (models Enviro, ML\textsubscript{g} and ML\textsubscript{l} + ML\textsubscript{g}).
+Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatial cross-validation) for experiment \emph{ii} (models Enviro, ML\textsubscript{g} and ML\textsubscript{l} + ML\textsubscript{g}).
 In six cases, Enviro was the best or tied best performing model.
@@
 However, there was no clear best model between ML\textsubscript{l} and Enviro + ML\textsubscript{l}.
-Therefore, when using these methods, both of these models should be fitted and the most predictive model selected for a given dataset.
+Therefore, when using these methods, both of these models should be fitted and the model with the best predictive performace for a given dataset selected.
 Furthermore, many of the performance improvements are rather marginal. 
@@
 However, it does indicate that continental scale models should consider using a mosaic of locally trained machine learning models for example.
-What is more surprising is that using both the ML\textsubscript{l} and ML\textsubscript{g} covariates together did not improve performance given that this approach of creating diverse predictions is a core element of stacking methodology.
+What is more surprising is that using both the ML\textsubscript{l} and ML\textsubscript{g} covariates together did not improve performance.
+Given that this approach of creating diverse predictions is a core element of stacking methodology we would expect this set of covariates to perform as well or better than just the ML\textsubscript{l} covariates.
 

### 2019-12-02 tim-lucas-model-ensembles.tex
Finish anita comments

@@
 Secondly, there is a lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}. 
-Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more reliable and more widely available \citep{sturrock2016mapping} and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
+Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more reliable and more widely available \citep{sturrock2016mapping} and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log, johnson2019spatially}.
 However, the aggregation of cases over space means that the data may be spatially uninformative, especially if the case counts are aggregated over large or heterogeneous areas, because it is unclear where within the polygon, and in which environments, the cases occurred.
@@
 Applications in other disease contexts have used a similar stacking scheme where data from vector or wild-animal host species are used to train models, the predictions from which are then used as covariates in a final model \citep{pigott2014mapping, shearer2016estimating}.
-In such applications we would always expect to need additional covariates in addition to the modelled distributions of hosts or vectors.
+In such applications we would always expect to need additional covariates as well as the modelled distributions of hosts or vectors.
 However, in the case examined here, both sets of data are direct measures of some aspect of malaria transmission rate, and therefore it is possible, though not guaranteed, that we would not need any further covariates.
@@
 To test the effectiveness of the proposed approach we used data from four countries: Madagascar, Colombia, Indonesia and Senegal.
-We first examined whether using predictions from machine learning models, trained on local prevalence data and raw environmental covariates, could improve predictions when used as covariates in disaggregation regression models.
-We then tested whether covariates made using predictions from machine learning models trained on a global prevalence dataset could improve predictions either on their own or when combined with predictions from machine learning models trained on local data.
-While there was not one consistently best model, we found that in most cases, using predictions from local machine learning models as covariates improved model performance.
+We focused our analysis on comparing the predictive performance of disaggregation regression when given different sets of covariates.
+In each country we fitted machine learning models trained on prevalence data and raw environmental covariates.
+We made new covariates using predictions from these models.
+We then tested whether disaggregation regression models with these new covariates performed better than a baseline model.
+We tested this approach using machine learning models trained on local prevalence data as well as models trained on a global prevalence dataset.
+While there was no consistently best model we found that, in most cases, using predictions from local machine learning models as covariates improved model performance.
 In contrast, using predictions from machine learning models trained on the global prevalence dataset rarely improved predictive performance.
@@
 We used two data sources that reflect \emph{Plasmodium falciparum} malaria transmission; point-prevalence surveys and polygon-level, aggregated incidence data. 
-We selected Madagascar, Colombia, Indonesia and Senegal as case examples as they all have fairly complete, publicly available, surveillance data at a finer geographical resolution than admin 1 (i.e.
+We selected Madagascar, Colombia, Indonesia and Senegal as case examples as they all have fairly complete, publicly available, surveillance data at a finer geographical resolution than administrative level one (i.e.
 higher resolution than state or province).
@@
 While the data covered a large time period, we did not model time explicitly as we are here focussed on spatial, rather than temporal modelling.
+While we do expect there to be variation in time, as long as the spatial biases in sampling do not change through time, the temporal effects will be minimised and the model will still be learning relationships based on the relative changes along environmental variables.
 The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
@@
 The covariates were standardised and centered to have a mean of zero and a standard deviation of one. 
-We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created using predictions from machine learning models.
-These variables were used as covariates in the machine learning models as well as being used directly as covariates in the baseline dissaggregation regression models.
+We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created from predictions from machine learning models.
+The raw environmental variables were used as covariates in the machine learning models as well as being used directly as covariates in the baseline dissaggregation regression models.
 
@@
 Predictions from these models were then made across each country respectively.
-These predictions were empirical logit transformed so that they were on the linear predictor scale of the top level model.
+These predictions were empirical logit transformed so that they were on the linear predictor scale of the disaggregation regression model.
 An empirical logit was used rather than a standard logit as there were many predictions of exactly zero.
@@
 Finally, we set different priors on the regression coefficients depending on which covariates were used.
-When environmental covariates or a mix of environmental covariates and predictions from machine learning models were used we set the prior to be weakly regularising, $\beta_i \sim \operatorname{Norm}(0, 0.4)$, such that it was unlikely that any single covariate could explain the full range of the response data.
+When raw environmental covariates or a mix of raw environmental covariates and predictions from machine learning models were used we set the prior to be weakly regularising, $\beta_i \sim \operatorname{Norm}(0, 0.4)$, such that it was unlikely that any single covariate could explain the full range of the response data.
 When only machine learning model predictions were used we set $\beta_i \sim \operatorname{ Norm}(\frac{1}{M}, 0.4)$ where $M$ is the number of machine learning models being used. 
@@
 
+All experiments involve comparing predictive performance of disaggregation regression models when different combinations of covariates are used.
 We used the raw environmental covariates, centered and standardised, as a baseline model (subsequently called Enviro).
@@
 
-Overall, experiment \emph{i} suggest that using predictions from machine learning models trained on local prevalence points provides more accurate predictions than using raw environmental covariates when fitting disaggregation models of malaria incidence.
+We have studied the predictive performance of disaggregation regression of malaria incidence when provided with different sets of covariates.
+Overall, experiment \emph{i} suggest that the predictions from the disaggregation models were better when using covariates created using predictions from machine learning models (trained on local prevalence points) than when using raw environmental covariates.
 This increased performance comes despite the prevalence data being on a different scale and being measurements of a different aspects of malaria transmission as well as the fact that the model we have used to translate between the two scales is imperfect.
@@
 However, as models such as Random Forest and elastic net can robustly handle high dimensional data, future work could include many more covariates in the machine learning models, potentially increasing predictive performance.
-However, to make comparisons more easy, here we have used the same environmental covariates in the machine learning models and in the Enviro disaggregation regression models.
+However, to make comparisons easier, we have used the same environmental covariates in the machine learning models as in the Enviro disaggregation regression models.
 

### 2019-12-09 tim-lucas-model-ensembles.tex
Another round of dan edits.

@@
 High-resolution maps of malaria risk are vital for control and elimination \citep{weiss2019mapping, battle2019mapping}.
-However, mapping malaria in lower burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \citep{weiss2019mapping,battle2019mapping, bhatt2017improved, bhatt2015effect} is often not effective for two reasons.
+However, mapping malaria in lower burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \cite{weiss2019mapping,battle2019mapping, bhatt2017improved, bhatt2015effect} is often not effective for two reasons.
 Firstly, so few individuals are infected that most surveys will detect zero cases \citep{sturrock2016mapping}.
@@
 This data is therefore often under-powered for fitting flexible, non-linear models as is required for accurate malaria maps \citep{bhatt2017improved, bhatt2015effect}. 
-A model that combines prevalence point-surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.
+A method that combines prevalence point-surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.
 
-Here we propose training a suite of machine learning models, using point-level, binomial prevalence data and environmental covariates, and combining predictions from these models by using them as covariates in a polygon-level, disaggregation regression model that uses  malaria incidence, aggregated to administrative units, as the response.
+Here we propose a two-stage method.
+In the first stage we train a suite of machine learning models, using point-level, binomial prevalence data and environmental covariates.
+In the second stage we combine predictions from these models by using them as covariates in a polygon-level, disaggregation regression model that uses  malaria incidence (aggregated to administrative units) as the response.
+Unlike joint likelihood models \citep{wang2018generalized}, this method does not combine both prevalence and incidence data within one model.
+Instead the aim is to use the prevalence data to find useful non-linear transformations of the environmental covariates which are then subsequently used in the disaggregation regression models.
+
 This modelling scheme has similarities to previous stacking methods used for malaria mapping \citep{bhatt2017improved} and elsewhere \citep{wolpert1992stacked, sill2009feature, hao2019review, breiman1996stacked}.
-However, as the response data in the machine learning models and the disaggregation regression models are on different scales (prevalence is a proportion while incidence is a rate) there are a number of important differences.
-Notably, we cannot simply take a weighted average of the predictions from the machine learning models as in a standard stacking scheme \citep{sill2009feature, hao2019review}.
+However, as the response data in the machine learning models and the disaggregation regression models are on different scales (prevalence is a proportion while incidence is a rate) we cannot simply take a weighted average of the predictions from the machine learning models as in a standard stacking scheme \citep{sill2009feature, hao2019review}.
 Instead the predictions need to be transformed to the incidence scale with a seperately fitted model \citep{cameron2015defining}.
-
 Applications in other disease contexts have used a similar stacking scheme where data from vector or wild-animal host species are used to train models, the predictions from which are then used as covariates in a final model \citep{pigott2014mapping, shearer2016estimating}.
@@
 We focused our analysis on comparing the predictive performance of disaggregation regression when given different sets of covariates.
-In each country we fitted machine learning models trained on prevalence data and raw environmental covariates.
+Therefore we keep the structure of the disaggregation regression model the same and only vary the covariates provided to the model.
+In each country we fitted stage 1 machine learning models trained on prevalence data and raw environmental covariates.
 We made new covariates using predictions from these models.
-We then tested whether disaggregation regression models with these new covariates performed better than a baseline model.
+We then tested whether stage 2 disaggregation regression models with these new covariates performed better than a baseline disaggregation regression model that directly used the raw environmental covariates.
 We tested this approach using machine learning models trained on local prevalence data as well as models trained on a global prevalence dataset.
-While there was no consistently best model we found that, in most cases, using predictions from local machine learning models as covariates improved model performance.
+While there was no consistently best model we found that, in most cases, the two stage method worked better than the single stage baseline disaggregation regression models.
+Using predictions from machine learning models, trained on local prevalence data as covariates improved the performance of disaggregation regression models relative to the disaggregation regression models that only used the raw environmental covariates.
 In contrast, using predictions from machine learning models trained on the global prevalence dataset rarely improved predictive performance.
@@
 The covariates were standardised and centered to have a mean of zero and a standard deviation of one. 
-We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created from predictions from machine learning models.
-The raw environmental variables were used as covariates in the machine learning models as well as being used directly as covariates in the baseline dissaggregation regression models.
+We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created from predictions from stage one machine learning models.
+The raw environmental variables were used as covariates in the stage one machine learning models as well as being used directly as covariates in the baseline stage two dissaggregation regression models.
 
-\subsection{Machine learning models}
-For each country specific dataset and for the global dataset we fitted 5 models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees (gradient boosted models, subsequently GBM) \citep{gbm}.
+\subsection{Stage 1 machine learning models}
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=0.8\textwidth]{figs/schematic2.pdf}
+\caption{
+  Schematic of the baseline disaggregation regression model and the two stage method. 
+  Models are shown in yellow ovals, malaria data is shown in purple rectangles and covariates are shown in green rectangles.
+  The baseline model uses aggregated incidence data and raw environmental covarates in a disaggregation regression model.
+  In the two stage method, new covariates are created in stage 1 by training machine learning models on prevalence data.
+  Predictions from these machine learning models are used as covariates in the stage 2 disaggregation regression.
+  Only one of the two stage models (ML\textsubscript{l}) is shown for simplicity.
+}
+\label{f:schem}
+\end{figure}
+For each country specific dataset and for the global dataset we fitted 5 stage 1 models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees (gradient boosted models, subsequently GBM) \citep{gbm}.
 These models were fitted to both the full malaria prevalence dataset and to the regional subsets of the data.
@@
 An empirical logit was used rather than a standard logit as there were many predictions of exactly zero.
-These predicted surfaces were subsequently used as covariates in the dissaggregation regression models.
+These predicted surfaces were subsequently used as covariates in the stage 2 dissaggregation regression models (Figure~\ref{f:schem}).
+The experiments that follow in this paper assess the improved performance of disaggregation regression models when using the predictions from these machine learning models as covariates as compared to a baseline disaggregation regression model using only the raw environmental covariates.
 
-\subsection{Disaggregation regression}
+\subsection{Stage 2 disaggregation regression}
 The model fitted to aggregated incidence data was a disaggregation regression model \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
@@
 The \emph{iid} random effect models both missing covariates and extra-Poisson sampling error.
-As described in more detail below, the matrix of covariates $X$ varies and is made up of various combinations of raw environmental covariates and predictions from the machine learning models.
+As described in more detail below we do not vary the structure of this model in the methodological comparison.
+The matrix of covariates, $X$, it the only component that varies and is made up of various combinations of raw environmental covariates and predictions from the machine learning models.
 
@@
 
-All experiments involve comparing predictive performance of disaggregation regression models when different combinations of covariates are used.
-We used the raw environmental covariates, centered and standardised, as a baseline model (subsequently called Enviro).
-We then performed two experiments.
-In the first (\emph{i}) we tested whether using predictions from machine learning models trained on local prevalence data improved predictions.
-In the second (\emph{ii}) we tested whether using predictions from machine learning models trained on global prevalence data improved predictions.
-These combinations of covariates are summarised in Table~\ref{t:models}.
-Our primary performance metric was correlation between observed and predicted data.
-We also examined the calibration of models by calculating the proportion of held out data that were within their 80\% credible intervals.
+\begin{table}[t!]
 
-\begin{table}[t!]
-\caption{Summary of models.
-Environmental covariates includes the full set of eight variables.
-Local ML covariates are the predictions from machine learning models, trained with local prevalence data and the eight environmental covariates as inputs.
-Global ML covariates are the predictions from machine learning models, trained with global prevalence data and the eight environmental covariates as inputs.}
 \centering
@@
 \end{tabular}
+\caption{Summary of stage two models.
+All models are disaggregation regression models fitted to aggregated incidence data.
+The only difference between the models is which covariates are being used.
+Environmental covariates includes the full set of eight variables.
+Local ML covariates are the predictions from stage one machine learning models, trained with local prevalence data and the eight environmental covariates as inputs.
+The local prevalence datasets are data from within each country except for Colombia where local prevalence refers to South American data.
+Global ML covariates are the predictions from stage one machine learning models, trained with global prevalence data (i.e.
+the full prevalence database) and the eight environmental covariates as inputs.}
 \label{t:models}
 \end{table}
+
+All experiments involve comparing predictive performance of stage two disaggregation regression models when different combinations of covariates are used.
+We used the raw environmental covariates, centered and standardised, as a baseline stage two model (Figure~\ref{f:schem}).
+We subsequently refer to this model as Enviro.
+We then performed two experiments.
+In the first (\emph{i}) we tested whether using predictions from stage one machine learning models trained on local (i.e. within-country) prevalence data improved predictions (Figure~\ref{f:schem}).
+In the second (\emph{ii}) we tested whether using predictions from stage one machine learning models trained on global prevalence data improved predictions.
+These combinations of covariates are summarised in Table~\ref{t:models}.
+Our primary performance metric was correlation between observed and predicted aggregated  data.
+We also examined the calibration of models by calculating the proportion of held out data that were within their 80\% credible intervals.
 
@@
 In this model therefore, the environmental covariates are effectively used twice, once in their raw form and once in transformed form (i.e. the predictions from the machine learning models).
-Secondly, we used only the predictions from the machine learning models trained on local data (subsequantly called ML\textsubscript{l}).
+Secondly, we used only the predictions from the machine learning models (Figure~\ref{f:schem}) trained on local data (subsequantly called ML\textsubscript{l}).
 As the environmental covariates are used as covariates in the machine learning models, this model is still ultimately driven by the values of the raw environmental covariates.
@@
 \begin{table}[t!]
-\caption{Pearson correlations between observed and predicted values. }
+
 \centering
@@
 \end{tabular}
+\caption{Pearson correlations between observed and predicted values. }
 \label{t:results}
+
 \end{table}
@@
 \end{tabular}
+\caption{Pearson correlations between observed and predicted values. }
 \label{t:results2}
+
 \end{table}
@@
 \begin{table}[t!]
-\caption{Covarage of 80\% credible intervals. Values outside 0.7--0.9 are shown in bold.}
+
 \centering
@@
 \end{tabular}
+\caption{Covarage of 80\% credible intervals. Values outside 0.7--0.9 are shown in bold.}
 \label{t:coverage}
@@
 \begin{table}[t!]
-\caption{Machine learning model results and fitted parameters (i.e.\thinspace model weights) of the machine learning predictions only models (local only and global only). }
+
 \centering
@@
 \end{tabular}
+\caption{Machine learning model results and fitted parameters (i.e.\thinspace model weights) of the machine learning predictions only models (local only and global only). }
 \label{t:mlresults}
@@
 
+\section*{Citations}
 \bibliography{Malaria} 

### 2019-12-12 tim-lucas-model-ensembles-SI.tex
Add SI

\documentclass[review]{elsarticle}

\usepackage{geometry}
 \geometry{
 a4paper,
 left=30mm,
 right=30mm,
 top=10mm,
 bottom=10mm
 }

\usepackage{graphicx}
\usepackage{amsmath}

\usepackage{subcaption}

\usepackage{amssymb}
\usepackage{array}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

%\usepackage{numcompress}\bibliographystyle{model3-num-names}

\usepackage{natbib}
%\usepackage[numbers, sort&compress]{natbib}
%\setcitestyle{author[oxford]year}
\bibliographystyle{elsarticle-num}

%\geometry{verbose,tmargin=30mm,bmargin=25mm,lmargin=25mm,rmargin=25mm}
%\input{figures.tex}
%\pagestyle{empty}
\begin{document}
%\templatefigures{}

% Model ensembles with different response variables for base and meta models:

% Spatial and Spatio-temporal Epidemiology
% Special issue: Disease Surveillance and Infectious disease Modeling

\begin{frontmatter}

\title{Supporting information: Improving disaggregation models of malaria incidence by ensembling non-linear models of prevalence}

\author[oxford]{Tim C. D. Lucas\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{timcdlucas@gmail.com}
\author[oxford]{Anita K. Nandi}
\author[oxford]{Suzanne H. Keddie}
%\author[oxford]{Katherine A. Twohig}
\author[oxford]{Elisabeth G. Chestnutt}
\author[oxford]{Rosalind E. Howes}
%\author[oxford]{Michele Nguyen}
\author[oxford]{Susan F. Rumisha}
\author[oxford]{Rohan Arambepola}
\author[oxford,idm]{Amelia Bertozzi-Villa}
\author[oxford]{Andre Python}
\author[oxford]{Tasmin L. Symons}
\author[oxford]{Justin J. Millar}
\author[oxford]{Punam Amratia}
\author[oxford]{Penelope Hancock}
\author[oxford]{Katherine E. Battle}
\author[oxford]{Ewan Cameron}
\author[oxford,telethon,curtin]{Peter W. Gething}
\author[oxford]{Daniel J. Weiss}

\address[oxford]{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK}
\address[idm]{Institute for Disease Modeling, Bellevue, WA, USA}
\address[telethon]{Telethon Kids Institute, Perth Children’s Hospital, Perth, Australia}
\address[curtin]{Curtin University, Perth, Australia}

\end{frontmatter}

\clearpage
\tableofcontents

\clearpage
\section{Global dataset Machine Learning}

\clearpage
\subsection{Out-of-sample scatter plots}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnet_obspred_global.png}
\caption{
  Scatter plot of predictions and held out observed data for the neural network trained on the global dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enet_obspred_global.png}
\caption{
  Scatter plot of predictions and held out observed data for the elastic net trained on the global dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppr_obspred_global.png}
\caption{
  Scatter plot of predictions and held out observed data for the PPR trained on the global dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ranger_obspred_global.png}
\caption{
  Scatter plot of predictions and held out observed data for the Random Forest trained on the global dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/gbm_obspred_global.png}
\caption{
  Scatter plot of predictions and held out observed data for the GBM trained on the global dataset.
}

\end{figure}

\clearpage
\subsection{Hyperparameter optimisation}

As ranger and GBM were tuned with random hyperparameter search, the plots become difficult and are not included.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnetopt_global.png}
\caption{
  Optimisation for neural network hyperparameters trained on the global dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enetopt_global.png}
\caption{
  Optimisation for elastic net hyperparameters trained on the global dataset.
}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppropt.png}
\caption{
  Optimisation for PPR hyperparameters trained on the global dataset.
}

\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Colombia (South America) prevalence dataset Machine Learning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Predictions}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/SA_all_ml.png}
\caption{
  Predictions from machine learning models trained on South American prevalence data.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/COL_all_globalml.png}
\caption{
  Predictions over Colombia from machine learning models trained on global prevalence data.
}

\end{figure}

\clearpage
\subsection{Out-of-sample scatter plots}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnet_obspred_sa.png}
\caption{
  Scatter plot of predictions and held out observed data for the neural network trained on the South America dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enet_obspred_sa.png}
\caption{
  Scatter plot of predictions and held out observed data for the elastic net trained on the South America dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppr_obspred_sa.png}
\caption{
  Scatter plot of predictions and held out observed data for the PPR trained on the South America dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ranger_obspred_sa.png}
\caption{
  Scatter plot of predictions and held out observed data for the Random Forest trained on the South America dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/gbm_obspred_sa.png}
\caption{
  Scatter plot of predictions and held out observed data for the GBM trained on the South America dataset.
}

\end{figure}

\clearpage
\subsection{Hyperparameter optimisation}

As ranger and GBM were tuned with random hyperparameter search, the plots become difficult and are not included.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnetopt_sa.png}
\caption{
  Optimisation for neural network hyperparameters trained on the South America dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enetopt_sa.png}
\caption{
  Optimisation for elastic net hyperparameters trained on the South America dataset.
}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppropt_sa.png}
\caption{
  Optimisation for PPR hyperparameters trained on the South America dataset.
}

\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Indonesia dataset Machine Learning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Predictions}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/idn_all_ml.png}
\caption{
  Predictions from machine learning models trained on Indonesian prevalence data.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/idn_all_globalml.png}
\caption{
  Predictions over Indonesia from machine learning models trained on global prevalence data.
}

\end{figure}

\clearpage
\subsection{Out-of-sample scatter plots}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnet_obspred_idn.png}
\caption{
  Scatter plot of predictions and held out observed data for the neural network trained on the Indonesia dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enet_obspred_idn.png}
\caption{
  Scatter plot of predictions and held out observed data for the elastic net trained on the Indonesia dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppr_obspred_idn.png}
\caption{
  Scatter plot of predictions and held out observed data for the PPR trained on the Indonesia dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ranger_obspred_idn.png}
\caption{
  Scatter plot of predictions and held out observed data for the Random Forest trained on the Indonesia dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/xgboost_obspred_idn.png}
\caption{
  Scatter plot of predictions and held out observed data for the GBM trained on the Indonesia dataset.
}

\end{figure}

\clearpage
\subsection{Hyperparameter optimisation}

As ranger and GBM were tuned with random hyperparameter search, the plots become difficult and are not included.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enetopt_idn.png}
\caption{
  Optimisation for elastic net hyperparameters trained on the Indonesia dataset.
}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppropt_idn.png}
\caption{
  Optimisation for PPR hyperparameters trained on the Indonesia dataset.
}

\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Madagascar dataset Machine Learning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Predictions}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/MDG_all_ml.png}
\caption{
  Predictions from machine learning models trained on Malagasy prevalence data.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/MDG_all_globalml.png}
\caption{
  Predictions over Madagascar from machine learning models trained on global prevalence data.
}

\end{figure}

\clearpage
\subsection{Out-of-MDGmple scatter plots}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnet_obspred_mdg.png}
\caption{
  Scatter plot of predictions and held out observed data for the neural network trained on the Madagascar dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enet_obspred_mdg.png}
\caption{
  Scatter plot of predictions and held out observed data for the elastic net trained on the Madagascar dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppr_obspred_mdg.png}
\caption{
  Scatter plot of predictions and held out observed data for the PPR trained on the Madagascar dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ranger_obspred_mdg.png}
\caption{
  Scatter plot of predictions and held out observed data for the Random Forest trained on the Madagascar dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/xgboost_obspred_mdg.png}
\caption{
  Scatter plot of predictions and held out observed data for the GBM trained on the Madagascar dataset.
}

\end{figure}

\clearpage
\subsection{Hyperparameter optimiMDGtion}

As ranger and GBM were tuned with random hyperparameter search, the plots become difficult and are not included.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enetopt_mdg.png}
\caption{
  OptimiMDGtion for elastic net hyperparameters trained on the Madagascar dataset.
}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppropt_mdg.png}
\caption{
  OptimiMDGtion for PPR hyperparameters trained on the Madagascar dataset.
}

\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Senegal dataset Machine Learning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Predictions}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/sen_all_ml.png}
\caption{
  Predictions from machine learning models trained on Senegal prevalence data.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/sen_all_globalml.png}
\caption{
  Predictions from machine learning models trained on Senegal prevalence data.
}

\end{figure}

\clearpage
\subsection{Out-of-sample scatter plots}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnet_obspred_sen.png}
\caption{
  Scatter plot of predictions and held out observed data for the neural network trained on the Senegal dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enet_obspred_sen.png}
\caption{
  Scatter plot of predictions and held out observed data for the elastic net trained on the Senegal dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppr_obspred_sen.png}
\caption{
  Scatter plot of predictions and held out observed data for the PPR trained on the Senegal dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ranger_obspred_sen.png}
\caption{
  Scatter plot of predictions and held out observed data for the Random Forest trained on the Senegal dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/xgboost_obspred_sen.png}
\caption{
  Scatter plot of predictions and held out observed data for the GBM trained on the Senegal dataset.
}

\end{figure}

\clearpage
\subsection{Hyperparameter optimisation}

As ranger and GBM were tuned with random hyperparameter search, the plots become difficult and are not included.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enetopt_sen.png}
\caption{
  Optimisation for elastic net hyperparameters trained on the Senegal dataset.
}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppropt_sen.png}
\caption{
  Optimisation for PPR hyperparameters trained on the Senegal dataset.
}

\end{figure}

%\section*{Citations}
%\bibliography{Malaria} 
%\bibliographystyle{unsrt}
%\bibliographystyle{apalike}

\end{document}

### 2019-12-12 tim-lucas-model-ensembles.tex
Minor author edits.
Tweaks.

@@
 \author[oxford]{Anita K. Nandi}
-\author[oxford]{Emma L. Collins}
 \author[oxford]{Suzanne H. Keddie}
@@
 \author[oxford]{Susan F. Rumisha}
+\author[oxford]{Rohan Arambepola}
+\author[oxford,idm]{Amelia Bertozzi-Villa}
 \author[oxford]{Andre Python}
@@
 \author[oxford]{Justin J. Millar}
-\author[oxford]{Rohan Arambepola}
-\author[oxford,idm]{Amelia Bertozzi-Villa}
+\author[oxford]{Punam Amratia}
 \author[oxford]{Penelope Hancock}
-\author[oxford]{Punam Amratia}
 \author[oxford]{Katherine E. Battle}
@@
 \end{figure}
+
 For each country specific dataset and for the global dataset we fitted 5 stage 1 models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees (gradient boosted models, subsequently GBM) \citep{gbm}.
@@
 
+All experiments involve comparing predictive performance of stage two disaggregation regression models when different combinations of covariates are used.
+We used the raw environmental covariates, centered and standardised, as a baseline stage two model (Figure~\ref{f:schem}).
+We subsequently refer to this model as Enviro.
+We then performed two experiments.
+In the first (\emph{i}) we tested whether using predictions from stage one machine learning models trained on local (i.e. within-country) prevalence data improved predictions (Figure~\ref{f:schem}).
+In the second (\emph{ii}) we tested whether using predictions from stage one machine learning models trained on global prevalence data improved predictions.
+Our primary performance metric was correlation between observed and predicted aggregated  data.
+We also examined the calibration of models by calculating the proportion of held out data that were within their 80\% credible intervals.
+
+In experiment \emph{i} we compared the baseline to two models that used locally trained machine learning models (Table~\ref{t:models}).
+Firstly, we combined the predictions from the machine learning models and the environmental coviarates in one model (subsequently called Enviro + ML\textsubscript{l}).
+In this model therefore, the environmental covariates are effectively used twice, once in their raw form and once in transformed form (i.e. the predictions from the machine learning models).
+Secondly, we used only the predictions from the machine learning models (Figure~\ref{f:schem}) trained on local data (subsequantly called ML\textsubscript{l}).
+As the environmental covariates are used as covariates in the machine learning models, this model is still ultimately driven by the values of the raw environmental covariates.
+However, in this model we are only using the environmental covariates in the transformed space learned by the machine learning models.
+
 \begin{table}[t!]
@@
 
-All experiments involve comparing predictive performance of stage two disaggregation regression models when different combinations of covariates are used.
-We used the raw environmental covariates, centered and standardised, as a baseline stage two model (Figure~\ref{f:schem}).
-We subsequently refer to this model as Enviro.
-We then performed two experiments.
-In the first (\emph{i}) we tested whether using predictions from stage one machine learning models trained on local (i.e. within-country) prevalence data improved predictions (Figure~\ref{f:schem}).
-In the second (\emph{ii}) we tested whether using predictions from stage one machine learning models trained on global prevalence data improved predictions.
-These combinations of covariates are summarised in Table~\ref{t:models}.
-Our primary performance metric was correlation between observed and predicted aggregated  data.
-We also examined the calibration of models by calculating the proportion of held out data that were within their 80\% credible intervals.
-
-In experiment \emph{i} we compared the baseline to two models that used locally trained machine learning models.
-Firstly, we combined the predictions from the machine learning models and the environmental coviarates in one model (subsequantly called Enviro + ML\textsubscript{l}).
-In this model therefore, the environmental covariates are effectively used twice, once in their raw form and once in transformed form (i.e. the predictions from the machine learning models).
-Secondly, we used only the predictions from the machine learning models (Figure~\ref{f:schem}) trained on local data (subsequantly called ML\textsubscript{l}).
-As the environmental covariates are used as covariates in the machine learning models, this model is still ultimately driven by the values of the raw environmental covariates.
-However, in this model we are only using the environmental covariates in the transformed space learned by the machine learning models.
-
-In experiment \emph{ii} we compared the baseline model to two models that use predictions from machine learning models trained on the global dataset of prevalence surveys.
+In experiment \emph{ii} we compared the baseline model to two models that use predictions from machine learning models trained on the global dataset of prevalence surveys (Table~\ref{t:models}).
 In the first, we used only predictions from the machine learning models trained on the global data (ML\textsubscript{g}).

### 2019-12-13 tim-lucas-model-ensembles.tex
Add api refs.

@@
 
-The polygon incidence data (i.e. malaria incidence aggregated to administrative units) were collected from government reports and standardised using methods defined in \cite{cibulskis2011worldwide}.
+The polygon incidence data (i.e.
+malaria incidence aggregated to administrative units) were collected from government reports \citep{mdg, sen, col, idn} and standardised using methods defined in \cite{cibulskis2011worldwide}.
 This standardisation step accounts for missed cases due to lack of treatment seeking, missing case reports, and cases that sought medical attention outside the public health systems \citep{battle2016treatment}.
@@
 The experiments that follow in this paper assess the improved performance of disaggregation regression models when using the predictions from these machine learning models as covariates as compared to a baseline disaggregation regression model using only the raw environmental covariates.
+See the supplementary material for plots of the grid search hyperparameter performance, out-of-sample scatter plots and plots of the predicted surfaces.
 
@@
 
-In experiment \emph{i} we compared the baseline to two models that used locally trained machine learning models (Table~\ref{t:models}).
-Firstly, we combined the predictions from the machine learning models and the environmental coviarates in one model (subsequently called Enviro + ML\textsubscript{l}).
-In this model therefore, the environmental covariates are effectively used twice, once in their raw form and once in transformed form (i.e. the predictions from the machine learning models).
-Secondly, we used only the predictions from the machine learning models (Figure~\ref{f:schem}) trained on local data (subsequantly called ML\textsubscript{l}).
-As the environmental covariates are used as covariates in the machine learning models, this model is still ultimately driven by the values of the raw environmental covariates.
-However, in this model we are only using the environmental covariates in the transformed space learned by the machine learning models.
-
 \begin{table}[t!]
@@
 \end{table}
+
+In experiment \emph{i} we compared the baseline to two models that used locally trained machine learning models (Table~\ref{t:models}).
+Firstly, we combined the predictions from the machine learning models and the environmental coviarates in one model (subsequently called Enviro + ML\textsubscript{l}).
+In this model therefore, the environmental covariates are effectively used twice, once in their raw form and once in transformed form (i.e. the predictions from the machine learning models).
+Secondly, we used only the predictions from the machine learning models (Figure~\ref{f:schem}) trained on local data (subsequantly called ML\textsubscript{l}).
+As the environmental covariates are used as covariates in the machine learning models, this model is still ultimately driven by the values of the raw environmental covariates.
+However, in this model we are only using the environmental covariates in the transformed space learned by the machine learning models.
 
@@
 \end{tabular}
-\caption{Pearson correlations between observed and predicted values. }
+\caption{Pearson correlations between observed and predicted values for experiment \emph{i}. }
 \label{t:results}
@@
 \end{tabular}
-\caption{Pearson correlations between observed and predicted values. }
+\caption{Pearson correlations between observed and predicted values for experiment \emph{ii}. }
 \label{t:results2}

### 2020-01-09 tim-lucas-model-ensembles.tex
Suzanne 2nd edits.

@@
 falciparum} only cases.
+For incidence rates we divide by 1000 to give the Annual Parasite Index (API).
 To minimise temporal effects we selected one year of surveillance data for each country. 
@@
 In six cases, Enviro was the best or tied best performing model.
-The ML\textsubscript{g} model was never the best performing model and only outpeforms Enviro in one case (spatial cross-validation in Madagascar).
+The ML\textsubscript{g} model was never the best performing model and only outperforms Enviro in one case (spatial cross-validation in Madagascar).
 In two cases ML\textsubscript{l} + ML\textsubscript{g} was the best performing model (spatial Colombia and spatial Madagascar).
-Comparing across Tables \ref{t:results} and \ref{t:results2} we can see that ML\textsubscript{g} outpeforms ML\textsubscript{l} once (spatial cross-validation in Senegal).
+Comparing across Tables \ref{t:results} and \ref{t:results2} we can see that ML\textsubscript{g} outperforms ML\textsubscript{l} once (spatial cross-validation in Senegal).
 In only two cases (Spatial Senegal and Spatial Indonesia) did ML\textsubscript{l} + ML\textsubscript{g} outperform ML\textsubscript{l}.
@@
 We have studied the predictive performance of disaggregation regression of malaria incidence when provided with different sets of covariates.
-Overall, experiment \emph{i} suggest that the predictions from the disaggregation models were better when using covariates created using predictions from machine learning models (trained on local prevalence points) than when using raw environmental covariates.
+Overall, experiment \emph{i} suggests that the predictions from the disaggregation models were better when using covariates created using predictions from machine learning models (trained on local prevalence points) than when using raw environmental covariates.
 This increased performance comes despite the prevalence data being on a different scale and being measurements of a different aspects of malaria transmission as well as the fact that the model we have used to translate between the two scales is imperfect.
@@
 While the approach presented here is related to stacking, it differs in that we have not constrained the regression parameters to be positive nor included a sum-to-one constraint i.e.\thinspace the result is not simply a weighted average of the level zero model predictions.
-We did not include these constraints because the base models and the meta model are trained on response data on different scales.
+We did not include these constraints because the first stage and second stage models were trained on response data on different scales.
 However, given our priors, nearly all the fitted regression coefficients in models with only machine learning predictions were positive.

### 2020-01-09 tim-lucas-model-ensembles.tex
Andre comments.

@@
 Maps of infection rate are one of the core tools needed for the control and elimination of malaria.
-As reliable routine surveillance data of malaria incidence becomes more widely available disaggregation regression is becoming an important model framework.
+Reliable routine surveillance data of malaria incidence is becoming more widely available.
+Disaggregation regression is becoming an important model framework for estimating high resolution risk maps from these aggregated data.
 However, the aggregation of incidence over large, heterogeneous areas means that these data are underpowered for learning complex, non-linear relationships between the environment and malaria risk.
@@
 A model that combines point surveys and aggregated surveillance data could have the benefits of both but must be able to account for the fact that these two data types measure different malariometric units.
-Here, we train multiple non-linear, machine learning models on prevalence point-surveys.
+Here, we train multiple non-linear, machine learning models on \emph{Plasmodium falciparum} prevalence point-surveys.
 We then ensemble the predictions from these machine learning models with a geostatistical disaggregation regression model that uses malaria incidence, aggregated to administrative units, as response data.
@@
 
-We used two data sources that reflect \emph{Plasmodium falciparum} malaria transmission; point-prevalence surveys and polygon-level, aggregated incidence data. 
+We used two data sources that reflect \emph{P. falciparum} malaria transmission; point-prevalence surveys and polygon-level, aggregated incidence data. 
 We selected Madagascar, Colombia, Indonesia and Senegal as case examples as they all have fairly complete, publicly available, surveillance data at a finer geographical resolution than administrative level one (i.e.
@@
 While the data covered a large time period, we did not model time explicitly as we are here focussed on spatial, rather than temporal modelling.
-While we do expect there to be variation in time, as long as the spatial biases in sampling do not change through time, the temporal effects will be minimised and the model will still be learning relationships based on the relative changes along environmental variables.
+While we do expect there to be variation in time, as long as the spatial biases in sampling do not change through time, the temporal effects will not bias learned relationships as the model will still be learning relationships based on the relative changes along environmental variables.
 The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
 This data was used as both a global dataset and as regional subsets.
-The global dataset contained 55,914 surveys in 44,842 distinct locations and represented samples from 5,687,304 individuals.
-For Colombia we used all points from South America (7,719 individuals from 522 locations) while for the other countries we used only data from that country (Madagascar: 89,381 individuals from 1,505 locations.
+The global dataset contains 55,914 surveys in 44,842 distinct locations and represents samples from 5,687,304 individuals.
+As there were relatively few surveys in Colombia we used all points from South America (7,719 individuals from 522 locations) while for the other countries we used only data from that country (Madagascar: 89,381 individuals from 1,505 locations.
 Indonesia: 1,512,888 individuals from 4,778 locations.
 Senegal: 80,896 individuals from 1,762 locations).
-We chose these geographic regions based on a trade-off between wanting a large sample size but wanting data from geographically relevant areas.
 
@@
 For incidence rates we divide by 1000 to give the Annual Parasite Index (API).
-To minimise temporal effects we selected one year of surveillance data for each country. 
+To keep the analysis focused on spatial estimates we selected one year of surveillance data for each country. 
 We used annual surveillance data from 2013 for Madagascar (110 districts), 2015 for Colombia (952 municipalities), 2013 for Indonesia (244 regencies and cities) and 2009 for Senegal (34 departments).
@@
 \caption{
-  Schematic of the baseline disaggregation regression model and the two stage method. 
+  Schematic of the baseline disaggregation regression model (Enviro) and the two stage method (ML\textsubscript{l}). 
   Models are shown in yellow ovals, malaria data is shown in purple rectangles and covariates are shown in green rectangles.
-  The baseline model uses aggregated incidence data and raw environmental covarates in a disaggregation regression model.
-  In the two stage method, new covariates are created in stage 1 by training machine learning models on prevalence data.
+  The baseline model (Enviro) uses aggregated incidence data and raw environmental covarates in a disaggregation regression model.
+  In the two stage method (ML\textsubscript{l}), new covariates are created in stage 1 by training machine learning models on prevalence data.
   Predictions from these machine learning models are used as covariates in the stage 2 disaggregation regression.
@@
 These predicted surfaces were subsequently used as covariates in the stage 2 dissaggregation regression models (Figure~\ref{f:schem}).
-The experiments that follow in this paper assess the improved performance of disaggregation regression models when using the predictions from these machine learning models as covariates as compared to a baseline disaggregation regression model using only the raw environmental covariates.
+The experiments that follow in this paper assess the performance of disaggregation regression models when using the predictions from these machine learning models as covariates as compared to a baseline disaggregation regression model using only the raw environmental covariates.
 See the supplementary material for plots of the grid search hyperparameter performance, out-of-sample scatter plots and plots of the predicted surfaces.
@@
 $$i_a = \frac{ \sum(i_b \mathrm{pop}_b)}{\sum  \mathrm{pop}_b} $$
-$$i_b = \mathrm{p2i}(p_b)$$
-where $\mathrm{p2i}$ is from a model that was published previously \citep{cameron2015defining} which defines a function
-$$\mathrm{p2i}: f(P) = 2.616P - 3.596P^2 + 1.594P^3.$$
+$$i_b = 2.616p_b - 3.596p_b^2 + 1.594p_b^3$$
+where the polynomial is  a function from a previously published model \citep{cameron2015defining}.
 The fact that the model passes through prevalence space ensures that the predictions from the machine learning models can be appropriately scaled.
-The linear predictor of the model is related to prevalence by a typical logit link function and includes: an intercept, $\beta_0$; covariates, $X$, with a vector of regression parameters, $\beta$; a spatial, Gaussian, random field, $u_s(\rho, \sigma_u)$; and an \emph{iid} random effect, $v_j(\sigma_v)$.
-$$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u_s\left(\rho, \sigma_u\right) + v_j\left(\sigma_v\right)\right)$$
+The linear predictor of the model is related to prevalence by a typical logit link function and includes: an intercept, $\beta_0$; covariates, $X$, with a vector of regression parameters, $\beta$; a spatial, Gaussian, random field, $u_b(\rho, \sigma_u)$; and an \emph{iid} random effect, $v_j(\sigma_v)$.
+$$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u_b\left(\rho, \sigma_u\right) + v_j\left(\sigma_v\right)\right)$$
 The Gaussian spatial effect has a Mat\'ern covariance function and two hyperparameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
-The \emph{iid} random effect models both missing covariates and extra-Poisson sampling error.
+The \emph{iid} random effect models both unobserved explanatory factors and extra-Poisson sampling error.
 As described in more detail below we do not vary the structure of this model in the methodological comparison.
@@
 \end{tabular}
-\caption{Summary of stage two models.
+\caption{Summary of stage two models and the experiments they are grouped into.
 All models are disaggregation regression models fitted to aggregated incidence data.
@@
 Global ML covariates are the predictions from stage one machine learning models, trained with global prevalence data (i.e.
-the full prevalence database) and the eight environmental covariates as inputs.}
+the full prevalence database) and the eight environmental covariates as inputs.
+Experiment \emph{i} tests whether including Local ML covariates improves predictive performance while Experiment \emph{ii} tests whether including Global ML covariates improves performance.}
 \label{t:models}
@@
   Observed data against predictions for random cross-validation hold-out samples on a square root transformed scale. 
+  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Envir +  ML\textsubsecript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubsecript{l}: Machine learning models trained on local prevalence data only.
 }
@@
   Observed data against predictions for spatial cross-validation hold-out samples on a square root transformed scale.
+  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Envir +  ML\textsubsecript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubsecript{l}: Machine learning models trained on local prevalence data only.
 }
@@
 \end{tabular}
-\caption{Covarage of 80\% credible intervals. Values outside 0.7--0.9 are shown in bold.}
+\caption{Coverage of 80\% credible intervals. Values outside 0.7--0.9 are shown in bold.}
 \label{t:coverage}
@@
 \end{tabular}
-\caption{Machine learning model results and fitted parameters (i.e.\thinspace model weights) of the machine learning predictions only models (local only and global only). }
+\caption{Machine learning model results and fitted parameters (i.e.\thinspace model weights) of the machine learning predictions only models (i.e.
+ML\textsubscript{l} local only and ML\textsubscript{g} global only).
+  For each dataset (the country and whether the data was local or global) the best model (lowest RMSE) is shown in bold. 
+  Similary, the largest coefficient within each disaggregation model is shown in bold.}
 \label{t:mlresults}
@@
 Furthermore, the difference between random and spatial cross-validation is usually bigger than the difference between different models.
-This suggests that better data coverage is more important than which specific model is used.
+This suggests that better data coverage is more important than which specific model is used and that the models are still relying quite strongly on the random field.
 
@@
 Therefore, in practice these models are working in a way similar to standard stacking.
-However, the coefficients certainly do not sum to one, and fitted intercepts are very small to account for this.
+However, the coefficients certainly do not sum to one, and fitted intercepts are negative to account for this.
 
-Due to the low power of typical aggregated incidence datasets, previous analyses using disaggregation regression used a small number of covariates \citep{sturrock2014fine}.
+Due to the low statistical power of typical aggregated incidence datasets, previous analyses using disaggregation regression used a small number of covariates \citep{sturrock2014fine}.
 However, as models such as Random Forest and elastic net can robustly handle high dimensional data, future work could include many more covariates in the machine learning models, potentially increasing predictive performance.
-However, to make comparisons easier, we have used the same environmental covariates in the machine learning models as in the Enviro disaggregation regression models.
+However, for the sake of consistency, we have used the same environmental covariates in the machine learning models as in the Enviro disaggregation regression models.
 
@@
 This extra modelling step can be seen as finding useful, non-linear transformations of the raw environmental covariates.
-This view is important for understanding how the model will predict in areas with no incidence data; in this situation the Gaussian random field has no effect and predictions are driven by this non-linear transformation of the raw environmental covariates.
+This view is important for understanding how the model will predict in areas with no incidence data; in this situation the data cannot inform the Gaussian random field and predictions are driven by this non-linear transformation of the raw environmental covariates.
 Training the machine learning models on local data (i.e.
-prevalence data from the same country or region as the incidence data) gave much better performance than when training the machine learning models on the full, global, dataset of prevalence point-surveys.
-As more countries, particularly those with medium or low malaria burdens, start providing timely and accurate routine surveillance data of malaria cases disaggregation regression may become more popular.
-We have here presented a method for improving these models by using ancillary, prevalence point-survey data.
+prevalence data from the same country or region as the incidence data) shows much better performance than when training the machine learning models on the full, global, dataset of prevalence point-surveys.
+More countries, particularly those with medium or low malaria burdens, are providing timely and accurate routine surveillance data of malaria cases.
+Therefore it may be expected that disaggregation regression may become more popular.
+We have here presented a method for improving the predictive performance of these models by using ancillary, prevalence point-survey data.
 

### 2020-01-09 tim-lucas-model-ensembles.tex
Fix subsecipt macro names.;

@@
 The covariates were standardised and centered to have a mean of zero and a standard deviation of one. 
-We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created from predictions from stage one machine learning models.
-The raw environmental variables were used as covariates in the stage one machine learning models as well as being used directly as covariates in the baseline stage two dissaggregation regression models.
+We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created from predictions from stage 1 machine learning models.
+The raw environmental variables were used as covariates in the stage 1 machine learning models as well as being used directly as covariates in the baseline stage 2 dissaggregation regression models.
 
@@
 The fact that the model passes through prevalence space ensures that the predictions from the machine learning models can be appropriately scaled.
-The linear predictor of the model is related to prevalence by a typical logit link function and includes: an intercept, $\beta_0$; covariates, $X$, with a vector of regression parameters, $\beta$; a spatial, Gaussian, random field, $u_b(\rho, \sigma_u)$; and an \emph{iid} random effect, $v_j(\sigma_v)$.
-$$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u_b\left(\rho, \sigma_u\right) + v_j\left(\sigma_v\right)\right)$$
+The linear predictor of the model is related to prevalence by a typical logit link function and includes: an intercept, $\beta_0$; covariates, $X$, with a vector of regression parameters, $\beta$; a spatial, Gaussian, random field, $u_b(\rho, \sigma_u)$; and an \emph{iid} random effect, $v_a(\sigma_v)$.
+$$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u_b\left(\rho, \sigma_u\right) + v_a\left(\sigma_v\right)\right)$$
 The Gaussian spatial effect has a Mat\'ern covariance function and two hyperparameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
@@
 
-All experiments involve comparing predictive performance of stage two disaggregation regression models when different combinations of covariates are used.
-We used the raw environmental covariates, centered and standardised, as a baseline stage two model (Figure~\ref{f:schem}).
+All experiments involve comparing predictive performance of stage 2 disaggregation regression models when different combinations of covariates are used.
+We used the raw environmental covariates, centered and standardised, as a baseline stage 2 model (Figure~\ref{f:schem}).
 We subsequently refer to this model as Enviro.
 We then performed two experiments.
-In the first (\emph{i}) we tested whether using predictions from stage one machine learning models trained on local (i.e. within-country) prevalence data improved predictions (Figure~\ref{f:schem}).
-In the second (\emph{ii}) we tested whether using predictions from stage one machine learning models trained on global prevalence data improved predictions.
+In the first (\emph{i}) we tested whether using predictions from stage 1 machine learning models trained on local (i.e. within-country) prevalence data improved predictions (Figure~\ref{f:schem}).
+In the second (\emph{ii}) we tested whether using predictions from stage 1 machine learning models trained on global prevalence data improved predictions.
 Our primary performance metric was correlation between observed and predicted aggregated  data.
@@
 \end{tabular}
-\caption{Summary of stage two models and the experiments they are grouped into.
+\caption{Summary of stage 2 models and the experiments they are grouped into.
 All models are disaggregation regression models fitted to aggregated incidence data.
@@
 Environmental covariates includes the full set of eight variables.
-Local ML covariates are the predictions from stage one machine learning models, trained with local prevalence data and the eight environmental covariates as inputs.
+Local ML covariates are the predictions from stage 1 machine learning models, trained with local prevalence data and the eight environmental covariates as inputs.
 The local prevalence datasets are data from within each country except for Colombia where local prevalence refers to South American data.
-Global ML covariates are the predictions from stage one machine learning models, trained with global prevalence data (i.e.
+Global ML covariates are the predictions from stage 1 machine learning models, trained with global prevalence data (i.e.
 the full prevalence database) and the eight environmental covariates as inputs.
@@
   Observed data against predictions for random cross-validation hold-out samples on a square root transformed scale. 
-  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Envir +  ML\textsubsecript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubsecript{l}: Machine learning models trained on local prevalence data only.
+  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Envir +  ML\textsubscript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubscript{l}: Machine learning models trained on local prevalence data only.
 }
@@
   Observed data against predictions for spatial cross-validation hold-out samples on a square root transformed scale.
-  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Envir +  ML\textsubsecript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubsecript{l}: Machine learning models trained on local prevalence data only.
+  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Envir +  ML\textsubscript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubscript{l}: Machine learning models trained on local prevalence data only.
 }

### 2020-01-09 tim-lucas-model-ensembles.tex
Cite WMR.

@@
 
-Maps of infection rate are one of the core tools needed for the control and elimination of malaria.
-Reliable routine surveillance data of malaria incidence is becoming more widely available.
+Maps of disease burden are one of the core tools needed for the control and elimination of malaria.
+Reliable routine surveillance data of malaria incidence, typically aggregated to administrative units, is becoming more widely available.
 Disaggregation regression is becoming an important model framework for estimating high resolution risk maps from these aggregated data.
@@
 However, mapping malaria in lower burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \cite{weiss2019mapping,battle2019mapping, bhatt2017improved, bhatt2015effect} is often not effective for two reasons.
-Firstly, so few individuals are infected that most surveys will detect zero cases \citep{sturrock2016mapping}.
+Firstly, so few individuals are infected that most surveys will detect zero positives \citep{sturrock2016mapping}.
 Secondly, there is a lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}. 
@@
 
-To test the effectiveness of the proposed approach we used data from four countries: Madagascar, Colombia, Indonesia and Senegal.
+To test the effectiveness of the proposed approach we used data from four countries with relatively complete surveillance data: Madagascar, Colombia, Indonesia and Senegal.
 We focused our analysis on comparing the predictive performance of disaggregation regression when given different sets of covariates.
@@
 For reports where cases were not reported at the species level, national estimates of the ratio between \emph{P.
-falciparum} and \emph{Plasmodium vivax} cases were used to calculate \emph{P.
-falciparum} only cases.
+falciparum} and \emph{Plasmodium vivax} cases from the World Malaria Report were used to calculate \emph{P.
+falciparum} only cases \citep{world2016world}.
 For incidence rates we divide by 1000 to give the Annual Parasite Index (API).
@@
 where the polynomial is  a function from a previously published model \citep{cameron2015defining}.
-The fact that the model passes through prevalence space ensures that the predictions from the machine learning models can be appropriately scaled.
+The fact that the model is explicit about the relationship between prevalence and incidence has two advantages.
+Firstly, predictions of prevalence can be easily made directly from the linear predictor of the model.
+Secondly, it means that the logit-transformed predictions from the machine learning models are correctly scaled.
 The linear predictor of the model is related to prevalence by a typical logit link function and includes: an intercept, $\beta_0$; covariates, $X$, with a vector of regression parameters, $\beta$; a spatial, Gaussian, random field, $u_b(\rho, \sigma_u)$; and an \emph{iid} random effect, $v_a(\sigma_v)$.
@@
 
-Table \ref{t:results} gives the correlation between observed and held out data (under random and spatial cross-validation) for experiment \emph{i} (models Enviro, ML\textsubscript{l} and Envir + ML\textsubscript{l}).
+Table \ref{t:results} gives the correlation between observed and held out data (under random and spatial cross-validation) for experiment \emph{i} (models Enviro, ML\textsubscript{l} and Enviro + ML\textsubscript{l}).
 Many of the differences in performance are rather marginal.
@@
 In one case, Enviro perfomed equally well as another model; under spatial cross-validation in Senegal Enviro and Enviro + ML\textsubscript{l} performed equally well.
-Of the remaining five cases, in two cases ML\textsubscript{l} performed best and in three cases Envir + ML\textsubscript{l} performed best.
+Of the remaining five cases, in two cases ML\textsubscript{l} performed best and in three cases Enviro + ML\textsubscript{l} performed best.
 The greatest benefits to using prediction from machine learning models instead of, or in combination with, the raw environmental variables occured under spatial cross-validation and in the cases when the Enviro model was particularly poor.
@@
   Observed data against predictions for random cross-validation hold-out samples on a square root transformed scale. 
-  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Envir +  ML\textsubscript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubscript{l}: Machine learning models trained on local prevalence data only.
+  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Enviro +  ML\textsubscript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubscript{l}: Machine learning models trained on local prevalence data only.
 }
@@
   Observed data against predictions for spatial cross-validation hold-out samples on a square root transformed scale.
-  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Envir +  ML\textsubscript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubscript{l}: Machine learning models trained on local prevalence data only.
+  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Enviro +  ML\textsubscript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubscript{l}: Machine learning models trained on local prevalence data only.
 }
@@
 The coverage was acceptable in the other three countries with most values lying between 0.7 and 0.9.
-Overall there was a general tendancy for models to be slightly over confident.
+Overall there was a general tendency for models to be slightly over confident.
 
@@
 
-Overall, the models fitted using predictions from machine learning models trained on the global database of prevalence point-surveys were less good than those using either environmental data alone or than those using predictions from machine learning models trained on local data.
+Overall, the models fitted using predictions from machine learning models trained on the global database of prevalence point-surveys were worse than those using either environmental data alone or than those using predictions from machine learning models trained on local data.
 This in itself is not particularly surprising.

### 2020-01-09 tim-lucas-model-ensembles.tex
kb edits.

@@
 
-We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \citep{weiss2014air, weiss2015re, weiss2018global}.
+We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \citep{weiss2014air, weiss2015re, weiss2018global, gething2011modelling}.
 Tasseled cap brightness and urban land cover were subsequently removed as they were highly correlated with other variables. 
@@
 All experiments involve comparing predictive performance of stage 2 disaggregation regression models when different combinations of covariates are used.
+After model fitting we made predictions over the study areas and reaggregated back to the administrative level of the surveillance data.
+Our primary performance metric was correlation between observed and aggregated predictions.
+We also examined the calibration of models by calculating the proportion of held out data that were within their 80\% credible intervals.
 We used the raw environmental covariates, centered and standardised, as a baseline stage 2 model (Figure~\ref{f:schem}).
@@
 We then performed two experiments.
-In the first (\emph{i}) we tested whether using predictions from stage 1 machine learning models trained on local (i.e. within-country) prevalence data improved predictions (Figure~\ref{f:schem}).
-In the second (\emph{ii}) we tested whether using predictions from stage 1 machine learning models trained on global prevalence data improved predictions.
-Our primary performance metric was correlation between observed and predicted aggregated  data.
-We also examined the calibration of models by calculating the proportion of held out data that were within their 80\% credible intervals.
+In the first (experiment \emph{i}) we tested whether using predictions from stage 1 machine learning models trained on local (i.e.
+within-country) prevalence data improved predictions (Figure~\ref{f:schem}).
+In the second (experiment \emph{ii}) we tested whether using predictions from stage 1 machine learning models trained on global prevalence data improved predictions.
 
@@
 We have studied the predictive performance of disaggregation regression of malaria incidence when provided with different sets of covariates.
+In experiment \emph{i} we compared a baseline model that used only raw environmental covariates (Enviro) to two models that used predictions from machine learning models trained on local prevalence points (ML\textsubscript{l} and Enviro + ML\textsubscript{l}).
 Overall, experiment \emph{i} suggests that the predictions from the disaggregation models were better when using covariates created using predictions from machine learning models (trained on local prevalence points) than when using raw environmental covariates.
-This increased performance comes despite the prevalence data being on a different scale and being measurements of a different aspects of malaria transmission as well as the fact that the model we have used to translate between the two scales is imperfect.
+This increased performance comes despite the prevalence data being a different scale (a proportion instead of a rate) and being measurements of a different aspects of malaria transmission (prevalence rather than incidence) as well as the fact that the model we have used to translate between the two scales is imperfect.
 However, there was no clear best model between ML\textsubscript{l} and Enviro + ML\textsubscript{l}.
@@
 
-While the approach presented here is related to stacking, it differs in that we have not constrained the regression parameters to be positive nor included a sum-to-one constraint i.e.\thinspace the result is not simply a weighted average of the level zero model predictions.
+While the approach presented here is related to stacking, it differs in that we have not constrained the regression parameters to be positive nor included a sum-to-one constraint, i.e.\thinspace the result is not simply a weighted average of the level zero model predictions.
 We did not include these constraints because the first stage and second stage models were trained on response data on different scales.
@@
 More countries, particularly those with medium or low malaria burdens, are providing timely and accurate routine surveillance data of malaria cases.
-Therefore it may be expected that disaggregation regression may become more popular.
+Therefore it may be expected that disaggregation regression may become more popular and operationally relevant.
 We have here presented a method for improving the predictive performance of these models by using ancillary, prevalence point-survey data.

### 2020-04-07 tim-lucas-model-ensembles-SI.tex
Update to state at first submission.

@@
  top=10mm,
- bottom=10mm
+ bottom=20mm
  }
@@
 \usepackage{amsmath}
+
+\renewcommand\thefigure{S\arabic{figure}} 
+\renewcommand\thetable{S\arabic{table}} 
+%% END MACROS SECTION
 
@@
 \title{Supporting information: Improving disaggregation models of malaria incidence by ensembling non-linear models of prevalence}
-
-\author[oxford]{Tim C. D. Lucas\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
-\ead{timcdlucas@gmail.com}
-\author[oxford]{Anita K. Nandi}
-\author[oxford]{Suzanne H. Keddie}
-%\author[oxford]{Katherine A. Twohig}
-\author[oxford]{Elisabeth G. Chestnutt}
-\author[oxford]{Rosalind E. Howes}
-%\author[oxford]{Michele Nguyen}
-\author[oxford]{Susan F. Rumisha}
-\author[oxford]{Rohan Arambepola}
-\author[oxford,idm]{Amelia Bertozzi-Villa}
-\author[oxford]{Andre Python}
-\author[oxford]{Tasmin L. Symons}
-\author[oxford]{Justin J. Millar}
-\author[oxford]{Punam Amratia}
-\author[oxford]{Penelope Hancock}
-\author[oxford]{Katherine E. Battle}
-\author[oxford]{Ewan Cameron}
-\author[oxford,telethon,curtin]{Peter W. Gething}
-\author[oxford]{Daniel J. Weiss}
-
-\address[oxford]{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK}
-\address[idm]{Institute for Disease Modeling, Bellevue, WA, USA}
-\address[telethon]{Telethon Kids Institute, Perth Children’s Hospital, Perth, Australia}
-\address[curtin]{Curtin University, Perth, Australia}
 
@@
 \clearpage
-\subsection{Hyperparameter optimiMDGtion}
+\subsection{Hyperparameter optimisation}
 
@@
 \caption{
-  OptimiMDGtion for elastic net hyperparameters trained on the Madagascar dataset.
+  Optimisation for elastic net hyperparameters trained on the Madagascar dataset.
 }
@@
 \caption{
-  OptimiMDGtion for PPR hyperparameters trained on the Madagascar dataset.
+  Optimisation for PPR hyperparameters trained on the Madagascar dataset.
 }

### 2020-04-07 tim-lucas-model-ensembles.tex
Update to state at first submission.

@@
 
-\author[oxford]{Tim C. D. Lucas\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
-\ead{timcdlucas@gmail.com}
-\author[oxford]{Anita K. Nandi}
-\author[oxford]{Suzanne H. Keddie}
-%\author[oxford]{Katherine A. Twohig}
-\author[oxford]{Elisabeth G. Chestnutt}
-\author[oxford]{Rosalind E. Howes}
-%\author[oxford]{Michele Nguyen}
-\author[oxford]{Susan F. Rumisha}
-\author[oxford]{Rohan Arambepola}
-\author[oxford,idm]{Amelia Bertozzi-Villa}
-\author[oxford]{Andre Python}
-\author[oxford]{Tasmin L. Symons}
-\author[oxford]{Justin J. Millar}
-\author[oxford]{Punam Amratia}
-\author[oxford]{Penelope Hancock}
-\author[oxford]{Katherine E. Battle}
-\author[oxford]{Ewan Cameron}
-\author[oxford,telethon,curtin]{Peter W. Gething}
-\author[oxford]{Daniel J. Weiss}
-
-\address[oxford]{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK}
-\address[idm]{Institute for Disease Modeling, Bellevue, WA, USA}
-\address[telethon]{Telethon Kids Institute, Perth Children’s Hospital, Perth, Australia}
-\address[curtin]{Curtin University, Perth, Australia}
-
 \begin{abstract}
 
-Maps of disease burden are one of the core tools needed for the control and elimination of malaria.
+Maps of disease burden are a core tool needed for the control and elimination of malaria.
 Reliable routine surveillance data of malaria incidence, typically aggregated to administrative units, is becoming more widely available.
-Disaggregation regression is becoming an important model framework for estimating high resolution risk maps from these aggregated data.
-However, the aggregation of incidence over large, heterogeneous areas means that these data are underpowered for learning complex, non-linear relationships between the environment and malaria risk.
+Disaggregation regression is an important model framework for estimating high resolution risk maps from aggregated data.
+However, the aggregation of incidence over large, heterogeneous areas means that these data are underpowered for estimating complex, non-linear relationships models.
 In contrast, prevalence point-surveys are directly linked to local environmental conditions but are not common in many areas of the world.
-A model that combines point surveys and aggregated surveillance data could have the benefits of both but must be able to account for the fact that these two data types measure different malariometric units.
 Here, we train multiple non-linear, machine learning models on \emph{Plasmodium falciparum} prevalence point-surveys.
-We then ensemble the predictions from these machine learning models with a geostatistical disaggregation regression model that uses malaria incidence, aggregated to administrative units, as response data.
-We find that, in four national case studies, using a disaggregation regression model to combine predictions from machine learning models trained on prevalence point-surveys improves model accuracy relative to a baseline disaggregation regression model that only uses linear effects of the raw environmental covariates.
+We then ensemble the predictions from these machine learning models with a disaggregation regression model that uses aggregated malaria incidences as response data.
+We find that using a disaggregation regression model to combine predictions from machine learning models improves model accuracy relative to a baseline model.
 
@@
 See the supplementary material for plots of the grid search hyperparameter performance, out-of-sample scatter plots and plots of the predicted surfaces.
+All code is available on \href{https://github.com/timcdlucas/polygon_ml}{github}.
 
@@
 The authors acknowledge the National Malaria Control Programme of Madagascar for sharing their routine case data for this analysis.
+This work was funded by the Bill and Melinda Gates Foundation.
 

### 2020-04-07 tim-lucas-model-ensembles-title-page.tex
add the title page

--- 2020-04-07 tim-lucas-model-ensembles.tex
@@
 
+\author[oxford]{Tim C. D. Lucas\corref{mycorrespondingauthor}}
+\cortext[mycorrespondingauthor]{Corresponding author}
+\ead{timcdlucas@gmail.com}
+\author[oxford]{Anita K. Nandi}
+\author[oxford]{Suzanne H. Keddie}
+%\author[oxford]{Katherine A. Twohig}
+\author[oxford]{Elisabeth G. Chestnutt}
+\author[oxford]{Rosalind E. Howes}
+%\author[oxford]{Michele Nguyen}
+\author[oxford]{Susan F. Rumisha}
+\author[oxford]{Rohan Arambepola}
+\author[oxford,idm]{Amelia Bertozzi-Villa}
+\author[oxford]{Andre Python}
+\author[oxford]{Tasmin L. Symons}
+\author[oxford]{Justin J. Millar}
+\author[oxford]{Punam Amratia}
+\author[oxford]{Penelope Hancock}
+\author[oxford]{Katherine E. Battle}
+\author[oxford]{Ewan Cameron}
+\author[oxford,telethon,curtin]{Peter W. Gething}
+\author[oxford]{Daniel J. Weiss}
+
+\address[oxford]{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK}
+\address[idm]{Institute for Disease Modeling, Bellevue, WA, USA}
+\address[telethon]{Telethon Kids Institute, Perth Children’s Hospital, Perth, Australia}
+\address[curtin]{Curtin University, Perth, Australia}
+
 \begin{abstract}
@@
 We find that using a disaggregation regression model to combine predictions from machine learning models improves model accuracy relative to a baseline model.
-
 \end{abstract}
@@
 
-\linenumbers
-
-\section{Introduction}
-
-% New plan
-
-% We need to do disaggregation but how do we do nonlinear
-
-% We can do machine learning on prevalence and then combine...
-
-% Stacking is a thing and has some nice theory.
-
-% We found that it worked ok.
-
-High-resolution maps of malaria risk are vital for control and elimination \citep{weiss2019mapping, battle2019mapping}.
-However, mapping malaria in lower burden countries presents new challenges as traditional mapping of prevalence from cluster-level surveys \cite{weiss2019mapping,battle2019mapping, bhatt2017improved, bhatt2015effect} is often not effective for two reasons.
-Firstly, so few individuals are infected that most surveys will detect zero positives \citep{sturrock2016mapping}.
-Secondly, there is a lack of nationally representative prevalence surveys in low burden countries \citep{sturrock2016mapping, sturrock2014fine}. 
-Routine surveillance data of malaria case counts, often aggregated over administrative regions defined by geographic polygons, is becoming more reliable and more widely available \citep{sturrock2016mapping} and recent work has focussed on methods for estimating high-resolution malaria risk from these data \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log, johnson2019spatially}.
-However, the aggregation of cases over space means that the data may be spatially uninformative, especially if the case counts are aggregated over large or heterogeneous areas, because it is unclear where within the polygon, and in which environments, the cases occurred.
-This data is therefore often under-powered for fitting flexible, non-linear models as is required for accurate malaria maps \citep{bhatt2017improved, bhatt2015effect}. 
-A method that combines prevalence point-surveys and aggregated surveillance data, and therefore leverages the strength of both, has great potential.
-
-Here we propose a two-stage method.
-In the first stage we train a suite of machine learning models, using point-level, binomial prevalence data and environmental covariates.
-In the second stage we combine predictions from these models by using them as covariates in a polygon-level, disaggregation regression model that uses  malaria incidence (aggregated to administrative units) as the response.
-Unlike joint likelihood models \citep{wang2018generalized}, this method does not combine both prevalence and incidence data within one model.
-Instead the aim is to use the prevalence data to find useful non-linear transformations of the environmental covariates which are then subsequently used in the disaggregation regression models.
-
-This modelling scheme has similarities to previous stacking methods used for malaria mapping \citep{bhatt2017improved} and elsewhere \citep{wolpert1992stacked, sill2009feature, hao2019review, breiman1996stacked}.
-However, as the response data in the machine learning models and the disaggregation regression models are on different scales (prevalence is a proportion while incidence is a rate) we cannot simply take a weighted average of the predictions from the machine learning models as in a standard stacking scheme \citep{sill2009feature, hao2019review}.
-Instead the predictions need to be transformed to the incidence scale with a seperately fitted model \citep{cameron2015defining}.
-Applications in other disease contexts have used a similar stacking scheme where data from vector or wild-animal host species are used to train models, the predictions from which are then used as covariates in a final model \citep{pigott2014mapping, shearer2016estimating}.
-In such applications we would always expect to need additional covariates as well as the modelled distributions of hosts or vectors.
-However, in the case examined here, both sets of data are direct measures of some aspect of malaria transmission rate, and therefore it is possible, though not guaranteed, that we would not need any further covariates.
-
-Model stacking \citep{wolpert1992stacked} has proven effective in many realms \citep{bhatt2017improved, sill2009feature, hao2019review, breiman1996stacked}. 
-Stacking improves predictions by controlling bias and variance; as long as suitably diverse models are averaged, they will have different biases while high variance in models should be averaged out.
-This understanding of how stacking improves model performance indicates that diversity in models is important for stacking to be effective.
-Diversity in models is typically created in two ways: by using diverse training datasets \citep{breiman1996bagging} (as in Random Forests for example) and by using functionally different models (for example by averaging tree based models and neural networks) \citep{breiman1996stacked}.
-One important trade-off in spatial modelling is whether to use local data (with a smaller sample size but that is likely to be representative of the area of study) or global data that have a larger sample size but a less close association with the areas of study.
-For the application of malaria mapping, we can think about diversity of training data in this context and expect that stacking seperate models trained on local and global data will also increase the diversity of predictions in a useful way.
-
-To test the effectiveness of the proposed approach we used data from four countries with relatively complete surveillance data: Madagascar, Colombia, Indonesia and Senegal.
-We focused our analysis on comparing the predictive performance of disaggregation regression when given different sets of covariates.
-Therefore we keep the structure of the disaggregation regression model the same and only vary the covariates provided to the model.
-In each country we fitted stage 1 machine learning models trained on prevalence data and raw environmental covariates.
-We made new covariates using predictions from these models.
-We then tested whether stage 2 disaggregation regression models with these new covariates performed better than a baseline disaggregation regression model that directly used the raw environmental covariates.
-We tested this approach using machine learning models trained on local prevalence data as well as models trained on a global prevalence dataset.
-While there was no consistently best model we found that, in most cases, the two stage method worked better than the single stage baseline disaggregation regression models.
-Using predictions from machine learning models, trained on local prevalence data as covariates improved the performance of disaggregation regression models relative to the disaggregation regression models that only used the raw environmental covariates.
-In contrast, using predictions from machine learning models trained on the global prevalence dataset rarely improved predictive performance.
-
-\section{Methods}
-
-\subsection{Epidemiological Data}
-
-We used two data sources that reflect \emph{P. falciparum} malaria transmission; point-prevalence surveys and polygon-level, aggregated incidence data. 
-We selected Madagascar, Colombia, Indonesia and Senegal as case examples as they all have fairly complete, publicly available, surveillance data at a finer geographical resolution than administrative level one (i.e.
-higher resolution than state or province).
-The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from 1990 onwards \citep{bhatt2015effect, guerra2007assembling, pfeffer2018ma}.
-While the data covered a large time period, we did not model time explicitly as we are here focussed on spatial, rather than temporal modelling.
-While we do expect there to be variation in time, as long as the spatial biases in sampling do not change through time, the temporal effects will not bias learned relationships as the model will still be learning relationships based on the relative changes along environmental variables.
-The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
-This data was used as both a global dataset and as regional subsets.
-The global dataset contains 55,914 surveys in 44,842 distinct locations and represents samples from 5,687,304 individuals.
-As there were relatively few surveys in Colombia we used all points from South America (7,719 individuals from 522 locations) while for the other countries we used only data from that country (Madagascar: 89,381 individuals from 1,505 locations.
-Indonesia: 1,512,888 individuals from 4,778 locations.
-Senegal: 80,896 individuals from 1,762 locations).
-
-The polygon incidence data (i.e.
-malaria incidence aggregated to administrative units) were collected from government reports \citep{mdg, sen, col, idn} and standardised using methods defined in \cite{cibulskis2011worldwide}.
-This standardisation step accounts for missed cases due to lack of treatment seeking, missing case reports, and cases that sought medical attention outside the public health systems \citep{battle2016treatment}.
-For reports where cases were not reported at the species level, national estimates of the ratio between \emph{P.
-falciparum} and \emph{Plasmodium vivax} cases from the World Malaria Report were used to calculate \emph{P.
-falciparum} only cases \citep{world2016world}.
-For incidence rates we divide by 1000 to give the Annual Parasite Index (API).
-To keep the analysis focused on spatial estimates we selected one year of surveillance data for each country. 
-We used annual surveillance data from 2013 for Madagascar (110 districts), 2015 for Colombia (952 municipalities), 2013 for Indonesia (244 regencies and cities) and 2009 for Senegal (34 departments).
-These years were selected as they had the most complete data in each case.
-
-Raster surfaces (i.e.
-population gridded to 5$ \times$ 5 km pixles) of population for the years 2005, 2010 and 2015, were created using data from WorldPop \citep{linard2012population,gaughan2013high,  sorichetta2015high} and from GPWv4 \citep{gpw4} where WorldPop did not have values.
-Population rasters for the remaining years were created by linear interpolation.
-
-\subsection{Raw Environmental Covariates}
-
-We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \citep{weiss2014air, weiss2015re, weiss2018global, gething2011modelling}.
-Tasseled cap brightness and urban land cover were subsequently removed as they were highly correlated with other variables. 
-The covariates were standardised and centered to have a mean of zero and a standard deviation of one. 
-We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created from predictions from stage 1 machine learning models.
-The raw environmental variables were used as covariates in the stage 1 machine learning models as well as being used directly as covariates in the baseline stage 2 dissaggregation regression models.
-
-\subsection{Stage 1 machine learning models}
-
-\begin{figure}[t!]
-  \centering
-  \includegraphics[width=0.8\textwidth]{figs/schematic2.pdf}
-\caption{
-  Schematic of the baseline disaggregation regression model (Enviro) and the two stage method (ML\textsubscript{l}). 
-  Models are shown in yellow ovals, malaria data is shown in purple rectangles and covariates are shown in green rectangles.
-  The baseline model (Enviro) uses aggregated incidence data and raw environmental covarates in a disaggregation regression model.
-  In the two stage method (ML\textsubscript{l}), new covariates are created in stage 1 by training machine learning models on prevalence data.
-  Predictions from these machine learning models are used as covariates in the stage 2 disaggregation regression.
-  Only one of the two stage models (ML\textsubscript{l}) is shown for simplicity.
-}
-\label{f:schem}
-\end{figure}
-
-For each country specific dataset and for the global dataset we fitted 5 stage 1 models via \emph{caret} \citep{caret}: elastic net \citep{enet}, Random Forest \citep{wright2015ranger}, projection pursuit regression \citep{friedman1981projection}, neural networks \citep{nnet} and boosted regression trees (gradient boosted models, subsequently GBM) \citep{gbm}.
-These models were fitted to both the full malaria prevalence dataset and to the regional subsets of the data.
-Our response variable was prevalence and we weighted the data by sample size (i.e.\thinspace the number of people tested for malaria in each survey).
-We used the raw environmental covariates described above as covariates in these machine learning models.
-This process can therefore be seen as creating non-linear transformations of the raw covariates that are hopefully better correlated with malaria incidence than the raw environmental covariates are.
-For each model we ran five-fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
-Root mean square error (RMSE) was used to select the best performing model.
-Predictions from these models were then made across each country respectively.
-These predictions were empirical logit transformed so that they were on the linear predictor scale of the disaggregation regression model.
-An empirical logit was used rather than a standard logit as there were many predictions of exactly zero.
-These predicted surfaces were subsequently used as covariates in the stage 2 dissaggregation regression models (Figure~\ref{f:schem}).
-The experiments that follow in this paper assess the performance of disaggregation regression models when using the predictions from these machine learning models as covariates as compared to a baseline disaggregation regression model using only the raw environmental covariates.
-See the supplementary material for plots of the grid search hyperparameter performance, out-of-sample scatter plots and plots of the predicted surfaces.
-All code is available on \href{https://github.com/timcdlucas/polygon_ml}{github}.
-
-\subsection{Stage 2 disaggregation regression}
-The model fitted to aggregated incidence data was a disaggregation regression model \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
-The models were implemented and fitted using Template Model Builder \citep{TMB} in R \citep{R} and we note that these models cannot be fitted using INLA \citep{INLA} as we are not using a linear link function.
-This model is defined by a likelihood at the level of the polygon with covariates and a spatial random field at the pixel-level. 
-Values at the polygon-level are given the subscript $a$ while pixel level values are indexed with $b$.
-
-The polygon case count data, $y_a$ is given a Poisson likelihood
-$$y_a \sim \operatorname{Pois}(i_a\mathrm{pop_a})$$
-where $i_a$ is the estimated polygon incidence rate and $\mathrm{pop_a}$ is the observed polygon population-at-risk. 
-This polygon-level likelihood is linked to the pixel-level incidence and prevalence by
-$$i_a = \frac{ \sum(i_b \mathrm{pop}_b)}{\sum  \mathrm{pop}_b} $$
-$$i_b = 2.616p_b - 3.596p_b^2 + 1.594p_b^3$$
-where the polynomial is  a function from a previously published model \citep{cameron2015defining}.
-The fact that the model is explicit about the relationship between prevalence and incidence has two advantages.
-Firstly, predictions of prevalence can be easily made directly from the linear predictor of the model.
-Secondly, it means that the logit-transformed predictions from the machine learning models are correctly scaled.
-The linear predictor of the model is related to prevalence by a typical logit link function and includes: an intercept, $\beta_0$; covariates, $X$, with a vector of regression parameters, $\beta$; a spatial, Gaussian, random field, $u_b(\rho, \sigma_u)$; and an \emph{iid} random effect, $v_a(\sigma_v)$.
-$$p_b = \operatorname{logit}^{-1}\left(\beta_0 + \beta X  + u_b\left(\rho, \sigma_u\right) + v_a\left(\sigma_v\right)\right)$$
-The Gaussian spatial effect has a Mat\'ern covariance function and two hyperparameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
-The \emph{iid} random effect models both unobserved explanatory factors and extra-Poisson sampling error.
-As described in more detail below we do not vary the structure of this model in the methodological comparison.
-The matrix of covariates, $X$, it the only component that varies and is made up of various combinations of raw environmental covariates and predictions from the machine learning models.
-
-Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$ and $\sigma_v$. 
-We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \citep{fuglstad2018constructing} such that $P(\rho < 1) = 0.0001$ (except for Indonesia where we set $P(\rho < 3) = 0.0001$ due to its much larger size) and $P(\sigma_u > 1) = 0.0001$.
-This prior encoded our \emph{a priori} preference for a simpler, smoother random field.
-We set this prior such that the random field could explain most of the range of the data if required.
-
-We assigned $\sigma_v$ a penalised complexity prior \citep{simpson2017penalising} such that $P(\sigma_v > 0.05) =  0.0001$. 
-This was based on a comparison of the variance of Poisson random variables, with rates given by the number of polygon-level cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \citep{cibulskis2011worldwide}.
-We found that an \emph{iid} effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
-
-Finally, we set different priors on the regression coefficients depending on which covariates were used.
-When raw environmental covariates or a mix of raw environmental covariates and predictions from machine learning models were used we set the prior to be weakly regularising, $\beta_i \sim \operatorname{Norm}(0, 0.4)$, such that it was unlikely that any single covariate could explain the full range of the response data.
-When only machine learning model predictions were used we set $\beta_i \sim \operatorname{ Norm}(\frac{1}{M}, 0.4)$ where $M$ is the number of machine learning models being used. 
-This prior  sets our \emph{a priori} expectation that all the machine learning prediction models are positively and equally correlated with incidence i.e. this prior encodes standard model averaging.
-It is important to note that this setup does not constitute true stacking in which we would enforce $\sum_i \beta_i = 1$ \citep{bhatt2017improved}.
-In a preliminary analysis we tested the case where we force $\beta_i > 0$ which in practice is largely the same as the $\sum_i \beta_i = 1$  \citep{breiman1996stacked} but allows a small amount of flexibility to handle mispecification in $\mathrm{p2i}$.
-This analysis did not show any benefits to this approach so it was not considered further.
-
-\subsection{Experiments}
-
-All experiments involve comparing predictive performance of stage 2 disaggregation regression models when different combinations of covariates are used.
-After model fitting we made predictions over the study areas and reaggregated back to the administrative level of the surveillance data.
-Our primary performance metric was correlation between observed and aggregated predictions.
-We also examined the calibration of models by calculating the proportion of held out data that were within their 80\% credible intervals.
-We used the raw environmental covariates, centered and standardised, as a baseline stage 2 model (Figure~\ref{f:schem}).
-We subsequently refer to this model as Enviro.
-We then performed two experiments.
-In the first (experiment \emph{i}) we tested whether using predictions from stage 1 machine learning models trained on local (i.e.
-within-country) prevalence data improved predictions (Figure~\ref{f:schem}).
-In the second (experiment \emph{ii}) we tested whether using predictions from stage 1 machine learning models trained on global prevalence data improved predictions.
-
-\begin{table}[t!]
-
-\centering
-\begin{tabular}{l >{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{1.55cm}>{\centering\arraybackslash}p{1.7cm}}
-Model &  Environmental covariates & Local ML covariates &  Global ML covariates \\
-\hline 
- \textbf{Experiment \emph{i}} & & &  \\
- Enviro &  \checkmark & & \\
- Enviro + ML\textsubscript{l} & \checkmark & \checkmark &\\\vspace{0.4cm}
- ML\textsubscript{l} &  & \checkmark & \\
- \textbf{Experiment \emph{ii}} & & &  \\
- Enviro &  \checkmark & & \\
- ML\textsubscript{g} &  &  & \checkmark\\
- ML\textsubscript{l} + ML\textsubscript{g} &  & \checkmark &  \checkmark
-\end{tabular}
-\caption{Summary of stage 2 models and the experiments they are grouped into.
-All models are disaggregation regression models fitted to aggregated incidence data.
-The only difference between the models is which covariates are being used.
-Environmental covariates includes the full set of eight variables.
-Local ML covariates are the predictions from stage 1 machine learning models, trained with local prevalence data and the eight environmental covariates as inputs.
-The local prevalence datasets are data from within each country except for Colombia where local prevalence refers to South American data.
-Global ML covariates are the predictions from stage 1 machine learning models, trained with global prevalence data (i.e.
-the full prevalence database) and the eight environmental covariates as inputs.
-Experiment \emph{i} tests whether including Local ML covariates improves predictive performance while Experiment \emph{ii} tests whether including Global ML covariates improves performance.}
-\label{t:models}
-\end{table}
-
-In experiment \emph{i} we compared the baseline to two models that used locally trained machine learning models (Table~\ref{t:models}).
-Firstly, we combined the predictions from the machine learning models and the environmental coviarates in one model (subsequently called Enviro + ML\textsubscript{l}).
-In this model therefore, the environmental covariates are effectively used twice, once in their raw form and once in transformed form (i.e. the predictions from the machine learning models).
-Secondly, we used only the predictions from the machine learning models (Figure~\ref{f:schem}) trained on local data (subsequantly called ML\textsubscript{l}).
-As the environmental covariates are used as covariates in the machine learning models, this model is still ultimately driven by the values of the raw environmental covariates.
-However, in this model we are only using the environmental covariates in the transformed space learned by the machine learning models.
-
-In experiment \emph{ii} we compared the baseline model to two models that use predictions from machine learning models trained on the global dataset of prevalence surveys (Table~\ref{t:models}).
-In the first, we used only predictions from the machine learning models trained on the global data (ML\textsubscript{g}).
-In the second, we combined predictions from the machine learning models trained on regional data and predictions from the machine learning models trained on global data (ML\textsubscript{l} + ML\textsubscript{g}).
-
-In each experiment we used two cross-validation schemes. 
-In the first, polygon incidence data was randomly split into six cross-validation folds.
-In the second, polygon incidence data was split spatially into k folds (via k-means clustering on the polygon centroids).
-We set k as 3 for Madagascar and Colombia.
-Due to its large size we set k as 7 for Indonesia.
-Due to the small sample size, we set k as 5 for Senegal.
-This spatial cross-validation scheme is testing the ability of the different models to make predictions far from data where the spatial random field is not informative.
-
-\section{Results}
-
-Table \ref{t:results} gives the correlation between observed and held out data (under random and spatial cross-validation) for experiment \emph{i} (models Enviro, ML\textsubscript{l} and Enviro + ML\textsubscript{l}).
-Many of the differences in performance are rather marginal.
-Enviro was the best performing model in two cases (random cross-validation in Madagascar and Senegal).
-In one case, Enviro perfomed equally well as another model; under spatial cross-validation in Senegal Enviro and Enviro + ML\textsubscript{l} performed equally well.
-Of the remaining five cases, in two cases ML\textsubscript{l} performed best and in three cases Enviro + ML\textsubscript{l} performed best.
-The greatest benefits to using prediction from machine learning models instead of, or in combination with, the raw environmental variables occured under spatial cross-validation and in the cases when the Enviro model was particularly poor.
-
-\begin{table}[t!]
-
-\centering
-\begin{tabular}{llrrrr}
-Country &  Cross-validation & Enviro &  Enviro + ML\textsubscript{l} &  ML\textsubscript{l} \\
-\hline 
- Colombia & Random &  0.59 & \textbf{0.61}& 0.59  \\
- Colombia &  Spatial &  0.12 &  0.25&  \textbf{0.33} \\
- Indonesia &  Random &  0.52 &  \textbf{0.59}&  0.48  \\
- Indonesia &  Spatial &  0.45 &  \textbf{0.51}&  0.44  \\
- Madagascar & Random &   \textbf{0.70}& 0.69 &  0.68  \\
- Madagascar &  Spatial &  0.22 & 0.18&  \textbf{0.55} \\
- Senegal &  Random &  \textbf{0.58} & 0.57&  0.51  \\
- Senegal &  Spatial &  0.63 &  0.63&  0.51  \\
-\end{tabular}
-\caption{Pearson correlations between observed and predicted values for experiment \emph{i}. }
-\label{t:results}
-
-\end{table}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{figs/cv1_l_scatter.png}
-\caption{
-  Observed data against predictions for random cross-validation hold-out samples on a square root transformed scale. 
-  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Enviro +  ML\textsubscript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubscript{l}: Machine learning models trained on local prevalence data only.
-}
-\label{f:scatter1}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{figs/cv2_l_scatter.png}
-\caption{
-  Observed data against predictions for spatial cross-validation hold-out samples on a square root transformed scale.
-  There are 12 cases composed of 4 countries (COL:, Colombia, IDN: Indonesia, MDG: Madagascar, SEN: Senegal) and three sets of covariates (Envir: raw environmental covariates only, Enviro +  ML\textsubscript{l}: raw environmental covariates and machine learning covariates trained on local prevalence data combined, ML\textsubscript{l}: Machine learning models trained on local prevalence data only.
-}
-\label{f:scatter2}
-\end{figure}
-
-Figure~\ref{f:scatter1} shows scatter plots of the model performance under random cross-validation for experiment \emph{i} while Figure~\ref{f:scatter2} shows scatter plots of the model performance under spatial cross-validation.
-It can be seen that without environmental covariates, the models in Madagascar fail to predict very high or very low values correctly.
-Figure~\ref{f:map} shows the input data and spatial out-of-sample predictions of the Enviro model and ML\textsubscript{l} model in Colombia.
-
-Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatial cross-validation) for experiment \emph{ii} (models Enviro, ML\textsubscript{g} and ML\textsubscript{l} + ML\textsubscript{g}).
-In six cases, Enviro was the best or tied best performing model.
-The ML\textsubscript{g} model was never the best performing model and only outperforms Enviro in one case (spatial cross-validation in Madagascar).
-In two cases ML\textsubscript{l} + ML\textsubscript{g} was the best performing model (spatial Colombia and spatial Madagascar).
-Comparing across Tables \ref{t:results} and \ref{t:results2} we can see that ML\textsubscript{g} outperforms ML\textsubscript{l} once (spatial cross-validation in Senegal).
-In only two cases (Spatial Senegal and Spatial Indonesia) did ML\textsubscript{l} + ML\textsubscript{g} outperform ML\textsubscript{l}.
-
-\begin{figure}[t!]
-\centering
-\includegraphics[width = 1\textwidth]{figs/col_comparison_map.png} %\caption{Indonesia spatial crossvalidation} 
-\caption{
-  A) Observed data for Colombia (grey for zero incidence).
-B) Out-of-sample predictions for the spatial cross-validation, environmental covariates only model.
-C) Out-of-sample predictions for the spatial cross-validation, local machine learning only model.
-For each cross-validation fold, predictions are made for the held out data which are then combined to make a single surface.
-}
-\label{f:map}
-\end{figure}
-
-Table~\ref{t:coverage} shows the out-of-sample coverage of the 80\% credible intervals for all models, countries and cross-validation schemes.
-The coverage in Colombia was very poor with no models achieving a coverage above 0.4.
-The coverage was acceptable in the other three countries with most values lying between 0.7 and 0.9.
-Overall there was a general tendency for models to be slightly over confident.
-
-We can examine the relationship between the RMSE of the machine learning models to their fitted regression coefficients (weights).
-These values are given in Table~\ref{t:mlresults}.
-In all five sets of machine learning models (four sets trained on local data and one set trained on global data), Random Forest performs the best.
-We have not forced $\beta_i > 0$ but we have set the priors for these coefficients with a positive mean.
-In nearly all cases the fitted values are positive.
-If the prevalence data and incidence data are not biased relative to each other we would expect the models with the lowest RMSE to also have the biggest regression coefficient.
-This occurs in three cases where Random Forest has the lowest RMSE and the biggest coefficient.
-In a further two cases, Random Forest has the lowest RMSE but GBM has the highest coefficient.
-The predictions from Random Forest and GBM are highly correlated in part because they both perform well and in part because they are both tree based models.
-Finally, it can be seen that the relationship between RMSE and regression coefficients was much weaker in Indonesia.
-For the models trained on local data a neural network has the highest fitted coefficient while for the models trained on global data an elastic net has the highest fitted coefficient.
-
-%\begin{table}[h!]
-%\caption{Pearson correlations between observed and predicted values. }
-%\centering
-%\begin{tabular}{llrrrr}
-%Cross-validation scheme & Country &  Covariates &  ML\textsubscript{l} &  ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
-%\hline 
-% Random &  Colombia & \textbf{0.59} & \textbf{0.59} &  0.58 & 0.56 \\
-% Random &  Madagascar &  0.70 &  \textbf{0.73} &   0.70 & 0.71 \\
-% Random &  Senegal &  \textbf{0.58} &  0.51 &  0.30 & 0.08 \\
-% Random &  Indonesia &  \textbf{0.52} &  0.48 &  0.46 & 0.32 \\
-% Spatial &  Colombia &  0.12 &  0.33 &  0.18 & \textbf{0.34}\\
-% Spatial &  Madagascar &  0.22 &  \textbf{0.66} &  0.63 & 0.60\\
-% Spatial &  Senegal &  \textbf{0.63} &  0.27 &  0.36 & 0.35 \\
-% Spatial &  Indonesia &  \textbf{0.45} &  0.44 &  0.41 & \textbf{0.45} \\
-%\end{tabular}
-%\label{t:results}
-%\end{table}
-
-% to do table of coverage.
-
-\begin{table}[t!]
-
-\centering
-\begin{tabular}{llrrrr}
-Country &  Cross-validation & Enviro &   ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
-\hline 
- Colombia & Random &  \textbf{0.59} &0.55 & 0.58 \\
- Colombia &  Spatial &  0.12 &  0.12 & \textbf{0.33}\\
- Indonesia &  Random &  \textbf{0.52} & 0.32 & 0.46 \\
- Indonesia &  Spatial &  0.45 & 0.41 & 0.45 \\
- Madagascar &  Random &  \textbf{0.70} &  0.67 & 0.68 \\
- Madagascar &  Spatial &  0.22 &  0.51 & \textbf{0.55}\\
- Senegal &  Random &  \textbf{0.58} &  0.50 & 0.49 \\
- Senegal &  Spatial &   \textbf{0.63} & 0.55 & 0.52 \\
-
-\end{tabular}
-\caption{Pearson correlations between observed and predicted values for experiment \emph{ii}. }
-\label{t:results2}
-
-\end{table}
-
-\begin{table}[t!]
-
-\centering
-\begin{tabular}{llrrrrr}
-Country &  CV & Enviro & ML\textsubscript{l} &  Enviro + ML\textsubscript{l} & ML\textsubscript{g} & ML\textsubscript{l} + ML\textsubscript{g} \\
-\hline 
- Colombia & Random & \textbf{0.28} & \textbf{0.28} & \textbf{0.29} & \textbf{0.28} & \textbf{0.30} \\
- Colombia &  Spatial & \textbf{0.30} & \textbf{0.33}  & \textbf{0.33} & \textbf{0.33} & \textbf{0.34}  \\
- Indonesia & Random &0.80 & 0.81& 0.78& 0.79& 0.77  \\
- Indonesia & Spatial & 0.80 & 0.78& 0.78& 0.76& 0.75  \\
- Madagascar &  Random & 0.80 & 0.77& 0.75& 0.77& 0.76 \\
- Madagascar & Spatial & \textbf{0.65} & 0.74& 0.70& 0.75 & 0.76   \\
- Senegal & Random &0.79 & 0.79& 0.79& 0.79& 0.82 \\
- Senegal & Spatial & 0.85 & \textbf{0.91}& 0.85& \textbf{0.94} & 0.85  \\
-\end{tabular}
-\caption{Coverage of 80\% credible intervals. Values outside 0.7--0.9 are shown in bold.}
-\label{t:coverage}
-\end{table}
-
-\begin{table}[t!]
-
-\centering
-\small
-\begin{tabular}{ll|rr|rr}
-     Country          & Model &      RMSE\textsubscript{l} & $\beta_l$ & RMSE\textsubscript{g} & $\beta_g$ \\ \hline
-Colombia & RF & \textbf{0.068} & 0.625 &  \textbf{0.169} & 0.180\\
-Colombia & GBM & 0.073 & \textbf{0.952} & 0.178& -0.218  \\
-Colombia & enet & 0.070 & 0.219 &0.233 & 0.183 \\
-Colombia & nnet & 0.070 & 0.129 &0.220 & 0.527 \\
-Colombia & ppr & 0.070 & 0.667 & 0.205 &  \textbf{0.546}\vspace{0.3cm}\\
-Indonesia & RF& \textbf{0.081} & 0.447 & \textbf{0.169} & 0.178\\
-Indonesia & GBM & 0.085 & 0.357 & 0.178& 0.289 \\
-Indonesia & enet & 0.091 & 0.303 &0.233 & \textbf{0.526} \\
-Indonesia & nnet & 0.089 & \textbf{0.506} &0.220 & 0.316 \\
-Indonesia & ppr & 0.089 & 0.364 & 0.205 &  0.089\vspace{0.3cm}\\
-Madagascar & RF & \textbf{0.100} & 0.538 &  \textbf{0.169} & \textbf{0.529}\\
-Madagascar & GBM & 0.105 & \textbf{0.570} & 0.178& 0.432 \\
-Madagascar & enet & 0.116 & 0.301 &0.233 & 0.262 \\
-Madagascar & nnet & 0.113 & 0.033 &0.220 & 0.364 \\
-Madagascar & ppr & 0.109 & 0.469 & 0.205 &  0.403\vspace{0.3cm}\\ 
-Senegal & RF & \textbf{0.092} & \textbf{0.339} & \textbf{0.169} & \textbf{0.425} \\
-Senegal & GBM & 0.099 & 0.261& 0.178& 0.408 \\
-Senegal & enet& 0.103 & 0.344  &0.233 & 0.205 \\
-Senegal & nnet & 0.099 & 0.254 &0.220 & 0.190 \\
-Senegal & ppr & 0.098 & 0.268& 0.205 &  0.126\\
-
-\end{tabular}
-\caption{Machine learning model results and fitted parameters (i.e.\thinspace model weights) of the machine learning predictions only models (i.e.
-ML\textsubscript{l} local only and ML\textsubscript{g} global only).
-  For each dataset (the country and whether the data was local or global) the best model (lowest RMSE) is shown in bold. 
-  Similary, the largest coefficient within each disaggregation model is shown in bold.}
-\label{t:mlresults}
-\end{table}
-
-%\begin{table}
-%\caption{Fitted regression parameters of the environmental and machine learning covariates for thee models: just environmental variables, just local machine learning models and the combination of the two.}
-%\centering
-%\tiny
-%\begin{tabular}{ll|rrr}
-%Country    & Variable &  $\beta_{enviro}$   & $\beta_l$ & $\beta_{enviro+l}$\\ \hline
-%Madagascar & LST mean & 0.17 &  &  -0.01\\
-%Madagascar & EVI & 0.32 &  & -0.04 \\
-%Madagascar & TSI & 0.42 &  &0.06  \\
-%Madagascar & access & 0.34 &  &0.00  \\
-%Madagascar & Elev & -0.14 &  & 0.00\\ 
-%Madagascar & LST sd & -0.78 &  &-0.04  \\
-%Madagascar & VIIRS & -0.1 &  &-0.52  \\
-%Madagascar & TCW & -0.22 &  & -0.06\\ 
-%Madagascar & RF &  & 0.29 &  0.01\\
-%Madagascar & GBM &  & 0.52 & 0.25 \\
-%Madagascar & enet &  & 0.31 &0.09  \\
-%Madagascar & nnet &  & 0.43 &0.14  \\
-%Madagascar & ppr &  & 0.49 & 0.22\vspace{0.2cm}\\ 
-%Colombia & LST mean &-0.12  &  & -0.22\\
-%Colombia & EVI &  -0.08&  & 0.17\\
-%Colombia & TSI &  0.54&  & -0.06\\
-%Colombia & access &  0.35&  &0.00 \\
-%Colombia & Elev &  0.47&  & 0.00\\
-%Colombia & LST sd & -0.11 &  & -0.10\\
-%Colombia & VIIRS & 0.24 &  & 0.01\\
-%Colombia & TCW &  0.21&  & 0.00\\
-%Colombia & RF &  & 0.36 & -0.18\\
-%Colombia & GBM &  & 0.54 & 0.26\\
-%Colombia & enet &  & 0.32 & -0.34\\
-%Colombia & nnet &  &0.33  & -0.09\\
-%Colombia & ppr &  & 0.54 & 0.01\vspace{0.2cm}\\
-%Senegal & LST mean &-0.04  &  &-0.02\\
-%Senegal & EVI &  0.33&  & 0.30\\
-%Senegal & TSI &  -0.58&  &  -0.59\\
-%Senegal & access & -0.03 &  &  -0.05\\
-%Senegal & Elev &  0.42&  &0.37\\ 
-%Senegal & LST sd &  0.21&  & 0.10 \\
-%Senegal & VIIRS &  0.13&  &  0.13\\
-%Senegal & TCW & 0.08 &  &-0.01\\ 
-%Senegal & RF &  & 0.34 &  0.20\\
-%Senegal & GBM &  &  0.28& 0.05\\
-%Senegal & enet &  &  0.27&  -0.05\\
-%Senegal & nnet &  & 0.29 &-0.09 \\
-%Senegal & ppr &  &  0.35& 0.25\vspace{0.2cm}\\
-%Indonesia & LST mean & -0.32 &  &-0.34\\
-%Indonesia & EVI &  0.49&  & 0.36\\
-%Indonesia & TSI &  0.58&  &0.52\\
-%Indonesia & access &0.33 &  & 0.06 \\
-%Indonesia & Elev& -0.06&  &-0.13\\ 
-%Indonesia & LST sd & -0.23 &  & -0.33\\
-%Indonesia & VIIRS &  -0.12&  & 0.03\\
-%Indonesia & TCW& 0.12 &  &0.10\\ 
-%Indonesia & RF &  & 0.32 & 0.06\\
-%Indonesia & GBM &  &  0.38&0.03\\
-%Indonesia & enet &  &  0.49&  0.56\\
-%Indonesia & nnet &  &  0.36&  0.18\\
-%Indonesia & ppr &  &  0.45& 0.32\\
-%
-%
-%
-%\end{tabular}
-%\label{t:mlresults2}
-%\end{table}
-%
-
-\section{Discussion}
-
-We have studied the predictive performance of disaggregation regression of malaria incidence when provided with different sets of covariates.
-In experiment \emph{i} we compared a baseline model that used only raw environmental covariates (Enviro) to two models that used predictions from machine learning models trained on local prevalence points (ML\textsubscript{l} and Enviro + ML\textsubscript{l}).
-Overall, experiment \emph{i} suggests that the predictions from the disaggregation models were better when using covariates created using predictions from machine learning models (trained on local prevalence points) than when using raw environmental covariates.
-This increased performance comes despite the prevalence data being a different scale (a proportion instead of a rate) and being measurements of a different aspects of malaria transmission (prevalence rather than incidence) as well as the fact that the model we have used to translate between the two scales is imperfect.
-However, there was no clear best model between ML\textsubscript{l} and Enviro + ML\textsubscript{l}.
-Therefore, when using these methods, both of these models should be fitted and the model with the best predictive performace for a given dataset selected.
-Furthermore, many of the performance improvements are rather marginal. 
-However, in a few cases such as Colombia and Madagascar under spatial cross-validation the performance boost is large.
-For example, using just environmental covariates in Madagascar under spatial cross-validation gives a correlation between observed and predicted data of 0.22.
-Such a model would be unusable for any applied or policy work.
-In contrast, using the predictions from the machine learning models trained on local data gives a correlation value of 0.55, which while still relatively poor is possibly a useful model.
-
-As expected, the model performance was generally worse under spatial cross-validation than under random cross-validation.
-This implies that the models are still relying heavily on the spatial Gaussian random field.
-Furthermore, the difference between random and spatial cross-validation is usually bigger than the difference between different models.
-This suggests that better data coverage is more important than which specific model is used and that the models are still relying quite strongly on the random field.
-
-% this method is an improvement
-% still reliant on spatial field
-
-% allows broad suite of covariates
-
-% sum to one stacking
-
-% global ml models, local experts
-
-Overall, the models fitted using predictions from machine learning models trained on the global database of prevalence point-surveys were worse than those using either environmental data alone or than those using predictions from machine learning models trained on local data.
-This in itself is not particularly surprising.
-However, it does indicate that continental scale models should consider using a mosaic of locally trained machine learning models for example.
-What is more surprising is that using both the ML\textsubscript{l} and ML\textsubscript{g} covariates together did not improve performance.
-Given that this approach of creating diverse predictions is a core element of stacking methodology we would expect this set of covariates to perform as well or better than just the ML\textsubscript{l} covariates.
-
-While the approach presented here is related to stacking, it differs in that we have not constrained the regression parameters to be positive nor included a sum-to-one constraint, i.e.\thinspace the result is not simply a weighted average of the level zero model predictions.
-We did not include these constraints because the first stage and second stage models were trained on response data on different scales.
-However, given our priors, nearly all the fitted regression coefficients in models with only machine learning predictions were positive.
-Therefore, in practice these models are working in a way similar to standard stacking.
-However, the coefficients certainly do not sum to one, and fitted intercepts are negative to account for this.
-
-Due to the low statistical power of typical aggregated incidence datasets, previous analyses using disaggregation regression used a small number of covariates \citep{sturrock2014fine}.
-However, as models such as Random Forest and elastic net can robustly handle high dimensional data, future work could include many more covariates in the machine learning models, potentially increasing predictive performance.
-However, for the sake of consistency, we have used the same environmental covariates in the machine learning models as in the Enviro disaggregation regression models.
-
-\section{Conclusions}
-
-Overall we find that including predictions from machine learning models trained on prevalence point-surveys can improve the performance of disaggregation regression models for malaria incidence relative to using raw environmental covariates.
-This extra modelling step can be seen as finding useful, non-linear transformations of the raw environmental covariates.
-This view is important for understanding how the model will predict in areas with no incidence data; in this situation the data cannot inform the Gaussian random field and predictions are driven by this non-linear transformation of the raw environmental covariates.
-Training the machine learning models on local data (i.e.
-prevalence data from the same country or region as the incidence data) shows much better performance than when training the machine learning models on the full, global, dataset of prevalence point-surveys.
-More countries, particularly those with medium or low malaria burdens, are providing timely and accurate routine surveillance data of malaria cases.
-Therefore it may be expected that disaggregation regression may become more popular and operationally relevant.
-We have here presented a method for improving the predictive performance of these models by using ancillary, prevalence point-survey data.
-
-\section*{Acknowledgments}
-The authors acknowledge the National Malaria Control Programme of Madagascar for sharing their routine case data for this analysis.
-This work was funded by the Bill and Melinda Gates Foundation.
-
-\section*{Citations}
-\bibliography{Malaria} 
-%\bibliographystyle{unsrt}
-%\bibliographystyle{apalike}
-
 \end{document}

### 2020-04-08 tim-lucas-model-ensembles-SI.tex
Add enet details.

@@
 \renewcommand\thetable{S\arabic{table}}
+
+ \DeclareMathOperator*{\argmin}{\arg\!\min}
 %% END MACROS SECTION
@@
 \tableofcontents
+
+\section{Supplementary methods}
+
+\subsection{Extended methods for machine learning models}
+
+We trained five machine learning models to each prevalence dataset.
+These models were elastic net, projection pursuit regression, random forest, boosted regression trees, neural networks.
+All models were trained on untransformed prevalence data with any predictions below zero being set at zero.
+For tree based models, transforming the data makes almost no difference.
+For the other models, training on empirical logit transformed data was considered but as the final performance metric (correlation between predicted and observed polygon incidence data) was on an absolute scale it was deemed better to train the models on absolute scales.
+For all models, the sample size of each survey was used as a weight.
+
+Elastic net models are penalised, linear regression models.
+They find the regression parameters that optimise
+$$\hat{\beta} = \argmin_\beta \sum_{i=1}^N w_i(y_i - X_i\beta) ^ 2 + \lambda\left((1-\alpha)\beta^2 + \alpha|\beta|\right)$$
+where $\beta$ is a vector of regression parameters, $y_i$ is the prevalence response data, $X_i$ is a vector of environmental covariates and $w_i$ is the weight given by the sample size. 
+$\lambda$ is the penalty term and $\alpha$ controls the relative contribution of the l1 and l2 penalties.
+With $\alpha = 0$ the model collapses to ridge regression while with $\alpha = 1$ the model is equivalent to the LASSO.
+$\alpha$ and $\lambda$ are selected by a cross validation with candidate values given by a grid search.
+We examined 15 values for $\alpha$ taken evenly between 0.05 and 1.
+We examined 15 values for $\lambda$ taken as 14 values spread evenly on a log10 scale between -4 and -1 and additionally considering $\lambda = 0$ which gives the least squares estimate.
 
@@
 \clearpage
-\subsection{Out-of-MDGmple scatter plots}
+\subsection{Out-of-sample scatter plots}
 

### 2020-04-08 tim-lucas-model-ensembles.tex
Perhaps line endings?

@@
 
-This modelling scheme has similarities to previous stacking methods used for malaria mapping \citep{bhatt2017improved} and elsewhere \citep{wolpert1992stacked, sill2009feature, hao2019review, breiman1996stacked}.
+Stacking, or stacked generalization, uses a second-stage model to combine predictions from a number of modelsby training the second-stage model using out-of-samples predictions from the first-stage models  \citep{wolpert1992stacked} .
+The modelling scheme proposed here has similarities to stacking methods used for malaria mapping \citep{bhatt2017improved} and elsewhere \citep{wolpert1992stacked, sill2009feature, hao2019review, breiman1996stacked}.
 However, as the response data in the machine learning models and the disaggregation regression models are on different scales (prevalence is a proportion while incidence is a rate) we cannot simply take a weighted average of the predictions from the machine learning models as in a standard stacking scheme \citep{sill2009feature, hao2019review}.
@@
 Raster surfaces (i.e.
-population gridded to 5$ \times$ 5 km pixles) of population for the years 2005, 2010 and 2015, were created using data from WorldPop \citep{linard2012population,gaughan2013high,  sorichetta2015high} and from GPWv4 \citep{gpw4} where WorldPop did not have values.
+population gridded to 5$ \times$ 5 km pixels) of population for the years 2005, 2010 and 2015, were created using data from WorldPop \citep{linard2012population,gaughan2013high,  sorichetta2015high} and from GPWv4 \citep{gpw4} where WorldPop did not have values.
 Population rasters for the remaining years were created by linear interpolation.
@@
 Overall, experiment \emph{i} suggests that the predictions from the disaggregation models were better when using covariates created using predictions from machine learning models (trained on local prevalence points) than when using raw environmental covariates.
-This increased performance comes despite the prevalence data being a different scale (a proportion instead of a rate) and being measurements of a different aspects of malaria transmission (prevalence rather than incidence) as well as the fact that the model we have used to translate between the two scales is imperfect.
+This increased performance comes despite the prevalence data being on a different scale (a proportion instead of a rate) and being measurements of a different aspects of malaria transmission (prevalence rather than incidence) as well as the fact that the model we have used to translate between the two scales is imperfect.
 However, there was no clear best model between ML\textsubscript{l} and Enviro + ML\textsubscript{l}.
-Therefore, when using these methods, both of these models should be fitted and the model with the best predictive performace for a given dataset selected.
+Therefore, when using these methods, both of these models should be fitted and the model with the best predictive performance for a given dataset selected.
 Furthermore, many of the performance improvements are rather marginal. 
@@
 
+One drawback of using predictions from machine learnings models as covariates is that the uncertainties in the predictions are not propagated properly through to the final predictions.
+This could be handled with an appropriate error model \citep{richardson1993conditional}.
+Where the machine learning models explicitly provide estimates of uncertainty, these could be used to inform such models.
+In the absence of such individual-prediction estimates of uncertainty, the cross-validation error could instead be used to inform priors for error models.
+It is however worth noting that the environmental covariates used here are also modelled and therefore similar care would ideally be taken in characterising the uncertainty in their values.
+
 Due to the low statistical power of typical aggregated incidence datasets, previous analyses using disaggregation regression used a small number of covariates \citep{sturrock2014fine}.

### 2020-04-08 responses/reviewer_response_plan.md
Add plan for now.

General comments
1. The double use of the environmental variables (first for the ML prediction and second for estimating together with the ML predictions the incidence) implies some correlation between the covariates used in the prevalence model on page 8. In addition, certain ML methods perform similar (see gradient boosting tree and random forest) expecting very similar results, and thus high correlation. I would thus suggest to add a few sentences commenting on these two sources of correlation and how do you expect this to affect your results.

2. I would also suggest to add (on a supplemental file) more information about the use of the ML methods on the supplement together with the cross validation, the tuning parameters used and the model specifics.

3. A potential drawback of the modelling approach you followed is that the uncertainty from the machine learning algorithms is not propagated in the prevalence model. I would suggest to acknowledge this and make suggestions on how to solve this problem as avenues for further research.

4. I have also missed some definition of methodology used. Please provide the definition of stacking and also provide in the supplement a brief explanation on how the ML methods you selected work.

will add

5. I would suggest to add a few sentences of the potential spatial misalignment between the environmental covariates used.

don't understand

6. I find it slightly unnatural the fact that although the main goal of disaggregation methods is to assess incidence (or any other metric) on higher geographical precision, both metrics used assess the coarse geographical resolution (polygons). Of course this is fair, since the "truth" is available on the polygon level. It would be interesting to have some metrics that perform on this higher geographical resolution of interest (and the truth can be the same, assuming identical on the pixels that belong to each polygon).

note to reviewer saying we're working on it

7. The study, as the title suggests is motivated by the malaria incidence. I would suggest to have a small section or paragraph in the paper, interpreting the maps you created and argue about their usefulness in the Malaria example and in the countries selected.

8. The distinction between global and local model was not clear on the first read of the paper. Please specify explicitly that the global model refer to a single model combining all 4 countries with a single spatial component, whereas the local are 4 different country specific models with different spatial components. Given this, it is not unexpected the local model to perform better, since it have a region specific Gaussian Field.

not correct

9. I would have expected that the predictive ability of a model with $n$ covariates (purely talking about bias and not variance) would have been lower compared to the predictive ability of a model with $n+k$ covariates. How do you explain the fact that there are instances where Enviro perfoms better? Can you add some sentences regarding this?

yes

10. The Malaria atlas project also presents prevalence estimates (2nd stage of your model). I would suggest to add some sentences clarifying: 1) how your methods differ from the Malaria continuous prevalence/incidence atlas 2) How  your estimates differ from the estimates published in the malaria atlas 3) Talk about the possibility or not of using the malaria atlas continuous estimates as an additional covariate.

explain to reviewer that this is weird.

11. Initially, the reviewer got the impression that incidence data on aggregated scale will be used to inform the modelling framework (page 2 lines 17-18). The model description on page 8 implies that the incidence on aggregated scale is calculated using the disaggregation model and the observed incidence is used only for cross validation purposes. If you use the observed incidence to inform the predicted values from the disaggregation model, please provide more information. Otherwise, I would remove any claim that the proposed model combines both incidence and prevalence data.

not correct

12. On page 10, the use of $i$ to define different experiments can confuse, since $i$ was also used to denote incidence.

will change

13. Please provide some additional information about how the spatial cross validation was performed using the global models. Did you take all the possible combinations between the country specific clusters results by the k-means?

huh

Reviewer #2: I think this paper is well written, employs interesting methodology, and addresses an important problem. I have several minor questions and requests for clarification, but generally think the manuscript is in good shape.

Page 2 - More contrast with/discussion about true joint models would be useful

add in discussion. cite preprint?

Page 4- "While we do expect there to be variation in time, as long as the spatial biases in sampling do not change through time, the temporal effects will not bias learned relationships as the model will still be learning relationships based on the relative changes along environmental variables."

This claim is not completely clear to me

not sure

> Page 7: "For each model we ran five-fold cross-validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. Root 150 mean square error (RMSE) was used to select the best performing model."

While it may not be problematic given the intended use, some discussion here about why k-fold cross validation is appropriate in a spatial setting would be helpful. Unlike in most "off-the-shelf" ML analyses, we can't assume that the observations are independent, and the error rates estimated under cross validation can fail to account for this fact. Later in the manuscript a spatial approach is mentioned - making that clear here would help.

agree. perhaps should have run spatial

> Page 7: "An empirical logit was used rather than a standard logit as there were many predictions of exactly zero."

Is there any reason to prefer ML-regression + Empirical Logit vs. using  ML-binomial-classification models directly? It seems that the techniques you're using are well suited to directly estimate the probability surface under an appropriate link function rather than doing so indirectly using a linear  model. More discussion of this issue would be helpful.

> Page 7-8:  "these models cannot be fitted using INLA [47] as we are not using a linear link function" - this is unclear to me - the R-INLA software does support alternate link functions, so it seems like it should be possible. The choice of software does not seem crucial to the authors methods, however.

http://www.r-inla.org/faq#TOC-Does-INLA-support-the-use-of-different-link-functions-

explain

> Page 8: "the polynomial is a function from a previously published model" - somewhere, it would be helpful to have a brief note of any limitations, assumptions, or problems associated with this approach. One issue which comes to mind is that, as written, it's not clear if any uncertainty associated with the given parameters is being propagated to the use of the model in the current work. This seems especially important as the authors note later that coverage of posterior credible intervals was poor.

we agree but difficult. just in review response?

Generally, I think additional limitations are required when discussing things like posterior credible intervals. Given that the models are derived using nonparametric components, there is no guarantee that that the posterior distribution will be consistent. Great caution in interpreting inferential measures seems in order.

agree? not sure where to write this

### 2020-04-08 tim-lucas-model-ensembles-SI.tex
more si methods for ml.

@@
 We trained five machine learning models to each prevalence dataset.
-These models were elastic net, projection pursuit regression, random forest, boosted regression trees, neural networks.
+These models were elastic net, projection pursuit regression,  neural networks, random forest and boosted regression trees.
 All models were trained on untransformed prevalence data with any predictions below zero being set at zero.
@@
 We examined 15 values for $\lambda$ taken as 14 values spread evenly on a log10 scale between -4 and -1 and additionally considering $\lambda = 0$ which gives the least squares estimate.
+
+Projection pursuit regression fits a number of smooth functions to linear combinations of all the covariates. 
+It involves finding regression parameters $\beta$ in the model
+$$y_i = \beta_0 + \sum_{j=1}^r f_j(\beta_j X_i) $$
+where the $f_j$'s are smoothing functions, $X_i$ is the vector of covariates and $r$ is the number of smooth terms to be fitted (a hyperparameter selected by cross-validation). 
+The model is fitted iteratively $r$ times with the model being fitted to the residuals of the previous models at each iteration.
+
+We fitted neural networks with a single hidden layer, logistic activation functions for the hidden units and a linear activation function for the output.
+This architecture, with $h$ hidden units and $c$ covariates, can be written as
+$$y_i = \beta_0 + \sum_h \beta_{ho} \operatorname{logit}^{-1}\left(\beta_{0h} + \sum_c \beta_{ih}X_{ic}\right)$$
+with $\beta_0$ being the intercept, $\beta_{0h}$ being the intercept for each hidden unit, $\beta_{ih}$ being the weights between the inputs and hidden units and $\beta_{ho}$ being the weights between the hidden units and the outputs.
 

### 2020-04-08 tim-lucas-model-ensembles-SI.tex
Finish ml description.

@@
 The model is fitted iteratively $r$ times with the model being fitted to the residuals of the previous models at each iteration.
+We considered  $r$ß in the range 1 to 15.
 
@@
 with $\beta_0$ being the intercept, $\beta_{0h}$ being the intercept for each hidden unit, $\beta_{ih}$ being the weights between the inputs and hidden units and $\beta_{ho}$ being the weights between the hidden units and the outputs.
+The nnet package optimises this model using back-propogation and includes an additional hyperparameter, 'decay', that is a penalty on the sum of the squares of the weights.
+We considered a grid of hyperparameters.
+For the number of hidden units we considered 15 equally spaced values between 1 and 29.
+We examined 15 values for $decay$ taken as 14 values spread evenly on a log10 scale between -4 and -1 and additionally considering $\lambda = 0$.
+
+The Random Forest algorithm creates an ensemble of regression trees.
+To control overfitting each tree is fitted to a bootstrap sample of the data.
+To further control overfitting, at each node of each tree, mtry random selected covariates are considered instead of the full number of covariates.
+For each node in a tree the covariate and split point is found which minimises the sum of the variances in the nodes.
+However, we consider two further variants of this rule.
+We consider "extremely randomised trees" in which the split point for each covariate is drawn uniformly randomly.
+We then consider "conditional inference trees" in which after a split point is chosen, a significance test is performed and the split point is only used if the data on each side of the split are significantly different (with a significance threshold of 0.5).
+The trees are gown until all nodes reach a minimum node size.
+The  split rule and the parameters mtry and the minimal node size are chosen by cross-validation.
+We considered a random search of 15 parameter sets.
+
+Finally, boosted regression trees also makes an ensemble of trees.
+The trees are built sequentially and for each successive tree, the data are weighted by the size of the residual between the observed data and the prediction from all the previous trees.
+The XGBoost implementation of boosted regression trees contains a number of hyperparameters: nrounds is the number of additive trees, max depth is the maximum depth of each tree, eta reduces the additional weight given to poorly predicted data points, gamma sets the minimum loss reduction needed for a split to be allowed, for each tree, a proportion colsamplebytree of covariates are considered, min_child_weight causes the growing of the tree to stop when a node has this many data points and finally subsample is a proportion of the data to randomly sample to use in training of each tree.
+Again we used a random search and considered 15 parameter sets.
 

### 2020-04-11 tim-lucas-model-ensembles-SI.tex
More responses to reviewers.

@@
 The trees are built sequentially and for each successive tree, the data are weighted by the size of the residual between the observed data and the prediction from all the previous trees.
-The XGBoost implementation of boosted regression trees contains a number of hyperparameters: nrounds is the number of additive trees, max depth is the maximum depth of each tree, eta reduces the additional weight given to poorly predicted data points, gamma sets the minimum loss reduction needed for a split to be allowed, for each tree, a proportion colsamplebytree of covariates are considered, min_child_weight causes the growing of the tree to stop when a node has this many data points and finally subsample is a proportion of the data to randomly sample to use in training of each tree.
+The XGBoost implementation of boosted regression trees contains a number of hyperparameters: nrounds is the number of additive trees, max depth is the maximum depth of each tree, eta reduces the additional weight given to poorly predicted data points, gamma sets the minimum loss reduction needed for a split to be allowed, for each tree, a proportion colsamplebytree of covariates are considered, min child weight causes the growing of the tree to stop when a node has this many data points and finally subsample is a proportion of the data to randomly sample to use in training of each tree.
 Again we used a random search and considered 15 parameter sets.

### 2020-04-11 tim-lucas-model-ensembles.tex
More responses to reviewers.

@@
 One important trade-off in spatial modelling is whether to use local data (with a smaller sample size but that is likely to be representative of the area of study) or global data that have a larger sample size but a less close association with the areas of study.
-For the application of malaria mapping, we can think about diversity of training data in this context and expect that stacking seperate models trained on local and global data will also increase the diversity of predictions in a useful way.
+For the application of malaria mapping, we can think about diversity of training data in this context and expect that stacking separate models trained on local and global data will also increase the diversity of predictions in a useful way.
 
@@
   Predictions from these machine learning models are used as covariates in the stage 2 disaggregation regression.
-  Only one of the two stage models (ML\textsubscript{l}) is shown for simplicity.
+  Only one of the two stage models (ML\textsubscript{l}) is shown for simplicity .
+  If ML\textsubscript{g} was included as well for example, it would look the same as ML\textsubscript{l} except that the prevalence data (pink box in stage 1) would have the global database of prevalence surveys.
 }
@@
 Root mean square error (RMSE) was used to select the best performing model.
+We note that spatial or random cross-validation could have been used in this step.
+The choice is less critical than for the cross-validation scheme used to test model performance but using random cross-validation might select for hyperparameters giving more complex or flexible models.
+To an extent, when ensembling models as we are here, high variance is better than high bias as the variance gets averaged out \citep{brieman??}.
+
 Predictions from these models were then made across each country respectively.
@@
 
-The polygon case count data, $y_a$ is given a Poisson likelihood
+The aggregated incidence count data, $y_a$ is given a Poisson likelihood
 $$y_a \sim \operatorname{Pois}(i_a\mathrm{pop_a})$$
@@
 We then performed two experiments.
-In the first (experiment \emph{i}) we tested whether using predictions from stage 1 machine learning models trained on local (i.e.
-within-country) prevalence data improved predictions (Figure~\ref{f:schem}).
-In the second (experiment \emph{ii}) we tested whether using predictions from stage 1 machine learning models trained on global prevalence data improved predictions.
+In the first (experiment 1) we tested whether using predictions from stage 1 machine learning models trained on local (i.e. within-country) prevalence data improved predictions (Figure~\ref{f:schem}).
+In the second (experiment 2) we tested whether using predictions from stage 1 machine learning models trained on global prevalence data improved predictions.
 
@@
 \hline 
- \textbf{Experiment \emph{i}} & & &  \\
+ \textbf{Experiment 1} & & &  \\
  Enviro &  \checkmark & & \\
@@
  ML\textsubscript{l} &  & \checkmark & \\
- \textbf{Experiment \emph{ii}} & & &  \\
+ \textbf{Experiment 2} & & &  \\
  Enviro &  \checkmark & & \\
@@
 the full prevalence database) and the eight environmental covariates as inputs.
-Experiment \emph{i} tests whether including Local ML covariates improves predictive performance while Experiment \emph{ii} tests whether including Global ML covariates improves performance.}
+Experiment 1 tests whether including Local ML covariates improves predictive performance while Experiment 2 tests whether including Global ML covariates improves performance.}
 \label{t:models}
@@
 
-In experiment \emph{i} we compared the baseline to two models that used locally trained machine learning models (Table~\ref{t:models}).
-Firstly, we combined the predictions from the machine learning models and the environmental coviarates in one model (subsequently called Enviro + ML\textsubscript{l}).
+In experiment 1we compared the baseline to two models that used locally trained machine learning models (Table~\ref{t:models}).
+Firstly, we combined the predictions from the machine learning models and the environmental covariates in one model (subsequently called Enviro + ML\textsubscript{l}).
 In this model therefore, the environmental covariates are effectively used twice, once in their raw form and once in transformed form (i.e. the predictions from the machine learning models).
@@
 
-In experiment \emph{ii} we compared the baseline model to two models that use predictions from machine learning models trained on the global dataset of prevalence surveys (Table~\ref{t:models}).
+In experiment 2 we compared the baseline model to two models that use predictions from machine learning models trained on the global dataset of prevalence surveys (Table~\ref{t:models}).
 In the first, we used only predictions from the machine learning models trained on the global data (ML\textsubscript{g}).
@@
 
-Table \ref{t:results} gives the correlation between observed and held out data (under random and spatial cross-validation) for experiment \emph{i} (models Enviro, ML\textsubscript{l} and Enviro + ML\textsubscript{l}).
+Table \ref{t:results} gives the correlation between observed and held out data (under random and spatial cross-validation) for experiment 1 (models Enviro, ML\textsubscript{l} and Enviro + ML\textsubscript{l}).
 Many of the differences in performance are rather marginal.
@@
 \end{tabular}
-\caption{Pearson correlations between observed and predicted values for experiment \emph{i}. }
+\caption{Pearson correlations between observed and predicted values for experiment 1. }
 \label{t:results}
@@
 
-Figure~\ref{f:scatter1} shows scatter plots of the model performance under random cross-validation for experiment \emph{i} while Figure~\ref{f:scatter2} shows scatter plots of the model performance under spatial cross-validation.
+Figure~\ref{f:scatter1} shows scatter plots of the model performance under random cross-validation for experiment 1 while Figure~\ref{f:scatter2} shows scatter plots of the model performance under spatial cross-validation.
 It can be seen that without environmental covariates, the models in Madagascar fail to predict very high or very low values correctly.
@@
 
-Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatial cross-validation) for experiment \emph{ii} (models Enviro, ML\textsubscript{g} and ML\textsubscript{l} + ML\textsubscript{g}).
+Table \ref{t:results2} gives the correlation between observed and held out data (under random and spatial cross-validation) for experiment 2 (models Enviro, ML\textsubscript{g} and ML\textsubscript{l} + ML\textsubscript{g}).
 In six cases, Enviro was the best or tied best performing model.
@@
 \end{tabular}
-\caption{Pearson correlations between observed and predicted values for experiment \emph{ii}. }
+\caption{Pearson correlations between observed and predicted values for experiment 2. }
 \label{t:results2}
@@
 We have studied the predictive performance of disaggregation regression of malaria incidence when provided with different sets of covariates.
-In experiment \emph{i} we compared a baseline model that used only raw environmental covariates (Enviro) to two models that used predictions from machine learning models trained on local prevalence points (ML\textsubscript{l} and Enviro + ML\textsubscript{l}).
-Overall, experiment \emph{i} suggests that the predictions from the disaggregation models were better when using covariates created using predictions from machine learning models (trained on local prevalence points) than when using raw environmental covariates.
+In experiment 1 we compared a baseline model that used only raw environmental covariates (Enviro) to two models that used predictions from machine learning models trained on local prevalence points (ML\textsubscript{l} and Enviro + ML\textsubscript{l}).
+Overall, experiment 1 suggests that the predictions from the disaggregation models were better when using covariates created using predictions from machine learning models (trained on local prevalence points) than when using raw environmental covariates.
 This increased performance comes despite the prevalence data being on a different scale (a proportion instead of a rate) and being measurements of a different aspects of malaria transmission (prevalence rather than incidence) as well as the fact that the model we have used to translate between the two scales is imperfect.
@@
 Therefore, when using these methods, both of these models should be fitted and the model with the best predictive performance for a given dataset selected.
+
 Furthermore, many of the performance improvements are rather marginal. 
@@
 In contrast, using the predictions from the machine learning models trained on local data gives a correlation value of 0.55, which while still relatively poor is possibly a useful model.
+
+It is also of note that in two cases, the simplest model (Enviro) performed the best.
+Given that Enviro + ML\textsubscript{l} includes the same covariates (with some extra) this is likely explained by increased variance.
+The model is trying to estimate a number of regression parameters, a random field and hyperparameters so removing extraneous covariates should help parameter estimation.
+This is particularly true in Senegal which only has 34 datapoints.
+Future work should consider joint models which use the prevalence and incidence data as response data so that the disaggregation has more degrees of freedom.
+This would hopefully allow the accurate estimation of more regression parameters
+The prevalence to incidence relationship used in this paper could be used to link the two data types.
+Furthermore, if the parameters in the prevalence incidence relationship were treated as unknown parameters, with informative priors, this would allow three data to inform the relationship and for the uncertainty in the original model to be propagated properly.
 
@@
 Given that this approach of creating diverse predictions is a core element of stacking methodology we would expect this set of covariates to perform as well or better than just the ML\textsubscript{l} covariates.
+
+It is important to note that we are using out-of-sample prediction of aggregated polygon data as our performance metric while the true target for prediction is the high resolution risk surface.
+In rare cases there is both aggregate and unaggregated data available so that the high resolution accuracy can be tested as well as the aggregate level performance \citep{sturrock2014fine}.
+However, this case is rare.
+Some simulation studies have been performed \citep{wilson2017pointless, law2018variational, johnson2019spatially} but these are often with few covariates and on correctly specified models.
+This is an important area for future research.
 
@@
 Therefore it may be expected that disaggregation regression may become more popular and operationally relevant.
+Indeed, methods similar to ML\textsubscript{g} are already being used for global mapping in malaria \citep{weiss2019mapping, battle2019mapping} though the results here suggest that future global mapping efforts should allow local prevalence data to have a stronger influence on the estimates.
 We have here presented a method for improving the predictive performance of these models by using ancillary, prevalence point-survey data.

### 2020-04-13 tim-lucas-model-ensembles-SI.tex
Revised ms.

@@
 
+\clearpage
+\section{Colinearity between covariates}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.7\textwidth]{figs/SI/col_correlation_plot.png}
+\caption{
+  Correlation matrix for all covariates (raw environmental and machine learning predictions) for Colombia}
+
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.7\textwidth]{figs/SI/idn_correlation_plot.png}
+\caption{
+  Correlation matrix for all covariates (raw environmental and machine learning predictions) for Indonesia}
+
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.7\textwidth]{figs/SI/mdg_correlation_plot.png}
+\caption{
+  Correlation matrix for all covariates (raw environmental and machine learning predictions) for Madagascar}
+
+\end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.7\textwidth]{figs/SI/sen_correlation_plot.png}
+\caption{
+  Correlation matrix for all covariates (raw environmental and machine learning predictions) for Senegal}
+
+\end{figure}
+
+\clearpage
+\section{Model consistency}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.9\textwidth]{figs/SI/pix_consistency_cis.png}
+\caption{80\% credible interval widths for a some pixels when fitting models in Madagascar with increasing dataset sizes. The blue line is a simple linear model fit.}
+
+\end{figure}
+
 %\section*{Citations}

### 2020-04-13 tim-lucas-model-ensembles.tex
Revised ms.

@@
 While the data covered a large time period, we did not model time explicitly as we are here focussed on spatial, rather than temporal modelling.
-While we do expect there to be variation in time, as long as the spatial biases in sampling do not change through time, the temporal effects will not bias learned relationships as the model will still be learning relationships based on the relative changes along environmental variables.
+Although we have not accounted for time in the models, as long as sampling in space is independent of time, the correct relationships should be recovered. 
 The prevalence points were then standardised to an age range of 2--10 using the model from \citep{smith2007standardizing}.
@@
 
-We considered an initial suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap brightness, tasseled cap wetness, log accessibility to cities, log night lights and proportion of urban land cover \citep{weiss2014air, weiss2015re, weiss2018global, gething2011modelling}.
-Tasseled cap brightness and urban land cover were subsequently removed as they were highly correlated with other variables. 
+We considered a suite of environmental and anthropological covariates, at a resolution of approximately $5 \times 5$ kilometres that included the annual mean and log standard deviation of land surface temperature, enhanced vegetation index, malaria parasite temperature suitability index, elevation, tasseled cap wetness, log accessibility to cities and log night lights \citep{weiss2014air, weiss2015re, weiss2018global, gething2011modelling}.
+All covariates were aligned in their native resolution (500m or 1km) and then aggregated to 5km resolution, therefore there should be minimal effects from spatial misalignment.
 The covariates were standardised and centered to have a mean of zero and a standard deviation of one. 
 We refer to this set of transformed variables as the raw environmental covariates (even though some of the covariates are anthropogenic rather than environmental) to distinguish them from other covariates created from predictions from stage 1 machine learning models.
-The raw environmental variables were used as covariates in the stage 1 machine learning models as well as being used directly as covariates in the baseline stage 2 dissaggregation regression models.
+The raw environmental variables were used as covariates in the stage 1 machine learning models as well as being used directly as covariates in the baseline stage 2 disaggregation regression models.
 
@@
 The choice is less critical than for the cross-validation scheme used to test model performance but using random cross-validation might select for hyperparameters giving more complex or flexible models.
-To an extent, when ensembling models as we are here, high variance is better than high bias as the variance gets averaged out \citep{brieman??}.
+To an extent, when ensembling models, high variance is better than high bias as the variance gets averaged out \citep{breiman1996stacked}.
 
@@
 An empirical logit was used rather than a standard logit as there were many predictions of exactly zero.
-These predicted surfaces were subsequently used as covariates in the stage 2 dissaggregation regression models (Figure~\ref{f:schem}).
+These predicted surfaces were subsequently used as covariates in the stage 2 disaggregation regression models (Figure~\ref{f:schem}).
+Plots of the correlation matrices for all covariates can be seen in S46--S49.
+The correlation between covariates varied from country to country.
+In general there is little correlation between global and local machine learning predictions.
+Accessibility is often strongly correlated with predictions from the machine learning models.
+In Senegal there is a lot of correlation between variables.
+However, we note that collinearity between variables is not as problematic in a predictive context as it is when interpreting regression coefficients is the aim of the analysis.
+Though it is still potentially a waste of degrees of freedom.
 The experiments that follow in this paper assess the performance of disaggregation regression models when using the predictions from these machine learning models as covariates as compared to a baseline disaggregation regression model using only the raw environmental covariates.
 See the supplementary material for plots of the grid search hyperparameter performance, out-of-sample scatter plots and plots of the predicted surfaces.
-All code is available on \href{https://github.com/timcdlucas/polygon_ml}{github}.
 
@@
 The model is trying to estimate a number of regression parameters, a random field and hyperparameters so removing extraneous covariates should help parameter estimation.
-This is particularly true in Senegal which only has 34 datapoints.
+This is particularly true in Senegal which only has 34 datapoints and higher correlation between covariates (S49).
 Future work should consider joint models which use the prevalence and incidence data as response data so that the disaggregation has more degrees of freedom.
@@
 However, it does indicate that continental scale models should consider using a mosaic of locally trained machine learning models for example.
-What is more surprising is that using both the ML\textsubscript{l} and ML\textsubscript{g} covariates together did not improve performance.
+What is more surprising is that using both the ML\textsubscript{l} and ML\textsubscript{g} covariates together did not improve performance, especially considering the relatively low correlation between the local and global predictions (Figure S46-49).
 Given that this approach of creating diverse predictions is a core element of stacking methodology we would expect this set of covariates to perform as well or better than just the ML\textsubscript{l} covariates.
 
-It is important to note that we are using out-of-sample prediction of aggregated polygon data as our performance metric while the true target for prediction is the high resolution risk surface.
+We have demonstrated that using prevalence data alongside aggregated incidence data can improve predictive performance.
+Unlike the maps presented by Weiss \emph{et al.} \citep{weiss2019mapping}, the estimates presented in this paper are intended for methodological research, not directly for policy decisions. 
+However, they do further demonstrate that aggregated incidence data can be used to create high resolution estimates of malaria incidence.
+Furthermore, in countries such as Senegal, the incidence data is much coarser than intervention implementation units, so disaggregated maps like the ones here have clear policy uses.
+It is important, however, to note that we are using out-of-sample prediction of aggregated polygon data as our performance metric while the true target for prediction is the high resolution risk surface.
 In rare cases there is both aggregate and unaggregated data available so that the high resolution accuracy can be tested as well as the aggregate level performance \citep{sturrock2014fine}.
@@
 It is however worth noting that the environmental covariates used here are also modelled and therefore similar care would ideally be taken in characterising the uncertainty in their values.
-
-Due to the low statistical power of typical aggregated incidence datasets, previous analyses using disaggregation regression used a small number of covariates \citep{sturrock2014fine}.
-However, as models such as Random Forest and elastic net can robustly handle high dimensional data, future work could include many more covariates in the machine learning models, potentially increasing predictive performance.
-However, for the sake of consistency, we have used the same environmental covariates in the machine learning models as in the Enviro disaggregation regression models.
 

### 2020-04-13 responses/time_factor.R
Short script for demonstrating that sampling through time is ok.

@@ -1,0 +1,60 @@
+
+# Case 1
+# Sampling is biased to towards the present
+t <- rep(c(1, rep(2, 3)), each = 100)
+
+# locations (and therefore temperature) are 
+#   also sampled with biases, but the bias
+#   is independant of time
+temp <- rep(c(1, rep(2, 3)), times = 100) 
+
+baseline_prev <- t * 0.1
+
+prev <-  baseline_prev * temp + rnorm(400, 0, 0.01)
+
+# 1 unit increase in temperature increases
+#   prevalence by a factor of 2.
+#   This is true despite the baseline
+#   prevalence that is determined by time.
+plot(prev ~ temp)
+
+
+samples <- rbinom(400, 1000, prev)
+
+d <- data.frame(positive = samples, n = 1000, temp = temp)
+
+m1 <- glm(cbind(samples, n) ~ temp, data = d, family = binomial)
+# we recover the factor of 2.
+exp(m1$coefficients[2])
+
+
+
+# Case 2
+# Sampling is still biased to towards the present
+t <- rep(c(1, rep(2, 3)), each = 100)
+
+# *However, if sampling in space changes through time*
+temp <- c(rep(c(1, 1, 1, 2), each = 25),
+          rep(c(1, 2, 2, 2), each = 75))
+
+baseline_prev <- t * 0.1
+
+prev <-  baseline_prev * temp + rnorm(400, 0, 0.01)
+
+# 1 unit increase in temperature increases
+#   prevalence by a factor of 2.
+#   This is true despite the baseline
+#   prevalence that is determined by time.
+plot(prev ~ temp)
+
+
+samples <- rbinom(400, 1000, prev)
+
+d <- data.frame(positive = samples, n = 1000, temp = temp)
+
+m2 <- glm(cbind(samples, n) ~ temp, data = d, family = binomial)
+
+# estimates are now biased.
+exp(m2$coefficients[2])
+
+
